### 2022-05-24 Documents/manuscript.Rmd
work in progress

---
title: "Double Standards in Judging Collective Action"
header: "DOUBLE STANDARDS"
keywords: protest, collective action, double standards, social identity theory, system justification theory, Black Lives Matter
abstract: Collective action is a powerful force driving social change but, at the same time, often sparks contention about what are acceptable forms of collective action.
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: preprint-template.tex

---

```{r setup, include = FALSE}

  # Load packages
  library(tidyverse); library(tidybayes); library(numform); library(glue)

  # knitr options
  knitr::opts_chunk$set(echo = FALSE)

```

<!-- I should cite Colin Leach's work. -->

Protest movements often spark contention about what actions are acceptable means of protest. Reactions to Black American athletes kneeling during the national anthem to protest racist police violence are a case in point. Kneeling during the national anthem is not violent, disruptive, or illegal. And yet, only 29\% of White Americans (66\% of Black Americans) and 11\% of Republicans (59\% of Democrats) considered it appropriate for Black athletes to protest in this way [@yougov_nfl_2017]. Indeed, Black athletes who had participated in the protests faced a backlash in public opinion [@lacina_nfl_2020]. This and other examples suggest that *who* the protesters are and *what* they are protesting influences how acceptable observers judge their actions to be.

In this paper, we examine double standards in judging collective action---that is, whether observers judge the same protest actions as more or less acceptable depending on their own and the protesters' group memberships (*identity-based double standards*) or on the protesters' cause and how it aligns with their own ideological positions (*ideology-based double standards*). In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action. In two preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police in the United States. We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but show no ingroup bias when judging collective action by ingroup and outgroup members.

<!-- Below, ... we situate the distinction ... below we introduce the distinction between normative ... derive predictions about ... from ... and introduce the present research ... -->

## Double Standards in Judging Collective Action

Collective action---that is, any action taken by group members to advance a shared political goal [for similar definitions, see @becker_virtual_2012; @van_zomeren_building_2016]---is a powerful force driving social change. For example, Black Lives Matter protests lastingly shifted public discourse about racial inequity toward antiracist ideas [@dunivin_black_2022].

Collective action can take many forms. Psychologists distinguish between *normative* collective action that seeks to achieve a political goal while conforming to the norms of the existing social system and *non-normative* collective action that violates those norms [@wright_responding_1990]. Recent research has shown that this distinction matters for how people respond to collective action. @shuman_disrupting_2021 showed that non-normative, non-violent collective action is more effective than either normative or violent collective action at gaining concessions from those most resistant to social change. @feinberg_activists_2020 demonstrated, however, that observers are less supportive of social movements that use extreme protest actions. Similarly, @teixeira_is_2020 found that advantaged-group members perceive non-normative collective action by disadvantaged-group members as more damaging to their ingroup's social image and are thus less supportive of such action. This means that activists face the dilemma that non-normative collective action might be most effective at gaining concessions but that it also risks reducing popular support for a cause. Together, these studies show that the distinction between normative and non-normative collective action is psychologically and societally consequential.

Research has, for the most part, relied on ad-hoc distinctions between what researchers themselves considered normative and non-normative collective action. In liberal democracies, researchers tend to consider actions such as signing petitions, voting in elections, or peaceful protest to be normative and actions such as blocking traffic, damaging property, or violent protest to be non-normative. But, as the examples in the first paragraph show, the distinction between normative and non-normative collective action might instead be in the eye of the beholder.

In this research, we move this distinction from the researcher's intuition into the realm of scientific investigation and test for double standards in where people draw the line between acceptable and unacceptable forms of collective action. For that purpose, we define a double standard as judging the same action as more or less acceptable depending on who is performing the action and for what reason [for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.] In this section, we derive predictions from social psychological theories about how identity-based and ideology-based double standards could lead people to judge the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting.

### Identity-Based Double Standards

Social Identity Theory [@tajfel_integrative_1979; for a recent review, see @reimer_identity_2022] argues that just as people show self-serving biases to achieve positive self-esteem, they show ingroup-serving biases to achieve positive social identities. In other words, they think about ingroup and outgroup in "me"--"not me" terms and favor ingroup members ("us") over outgroup members ["not us", @brewer_social_2007].

One domain in which ingroup bias manifests is judgements about other people's actions. @hewstone_ultimate_1990 reviewed research showing that people make ingroup-serving causal attributions when judging actions by ingroup and outgroup members, for example, by attributing negative behavior by outgroup members to internal causes but attributing negative behavior by ingroup members to external causes. @valdesolo_moral_2007 demonstrated that participants judged the same selfish action as less unfair when performed by themselves or an ingroup member than when performed by an outgroup member. Other studies [@abrams_double_2013; @endevelt_everyone_2021; for exceptions, see @mendoza_for_2014; @pinto_membership_2010] provided further evidence for ingroup bias in judging moral transgressions by ingroup and outgroup members.

In the same vain, we propose that ingroup bias results in identity-based double standards in judging collective action by ingroup and outgroup members. That is, we hypothesize that people will judge the same protest actions as more acceptable if the protesters are ingroup members (*Hypothesis 1a*) or if the cause of the protest aligns with the ingroup's interests (*Hypothesis 1b*). This hypothesis could explain why, for example, far more Black than White Americans considered it appropriate for Black athletes to kneel down to protest police violence against Black Americans [@yougov_nfl_2017]. Notably, this hypothesis applies equally to historically disadvantaged groups mobilizing against social injustice and to historically advantaged groups defending their group's position.

### Ideology-Based Double Standards

System Justification Theory [@jost_role_1994; for a recent review, see @jost_theory_2020] argues that people are motivated to defend, justify, and bolster the prevailing social, economic, and political system because doing so serves basic epistemic, existential, and relational needs. Both advantaged and disadvantaged groups are thought to be motivated to justify the existing system, although the strength of this motivation and its expression vary across individuals and situations.

While system justification often leads people to resist social change, its relationship to collective action depends on the goal of the action. @jost_missing_2017 argued that research on collective action needs to distinguish between system-challenging protest aimed at changing an unequal status quo and system-supporting protest aimed at maintaining or defending an unequal status quo. @osborne_protesting_2019 showed that, for members of both advantaged and disadvantaged groups, system justification is associated with supporting system-supporting actions but with opposing system-challenging actions.

Just as system justification shapes support for, and opposition to, collective action, we propose that it results in ideology-based double standards in what actions people consider acceptable forms of collective action. First, we hypothesize that  people will judge the same protest actions as more acceptable if the protest supports, rather than challenges, the system (*Hypothesis 2a*). This hypothesis reflects the assumption that the motivation to justify the system is, to some extent, universal. Second, we hypothesize that people will judge the same protest actions as more acceptable if they endorse system-justifying beliefs and the protest supports the system or if they reject system-justifying beliefs and the protest challenges the system (*Hypothesis 2b*). This hypothesis implies an symmetry in how people with different ideological orientations judge system-supporting and system-challenging collective action.

As system-justifying beliefs are closely related to political conservatism [@jost_political_2003], this hypothesis could explain partisan differences in what actions liberals and conservatives consider acceptable means of protest for progressive and conservative causes. Ideology-based double standards could, for example, explain why far more Democrats than Republicans considered it appropriate to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].

<!-- That is, we hypothesize that, at least to the extent that they hold system-justifying beliefs, people will judge the same protest actions as more acceptable if the protest supports, rather than challenges, the system. -->

## Using Item Response Theory to Detect Double Standards

```{=latex}
\begin{figure}
\centering
\caption{Item response curves for three hypothetical protest actions}
\includegraphics[scale=1]{../Scale Development/figures/figure-1}
\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
\label{fig:f1}
\end{figure}
```

Item response theory is a a conceptual and statistical framework for understanding how the characteristics of both items and respondents shape responses to a set of items [@demars_irt_2010]. 
<!-- As such, it allows us to disentangle how acceptable protest actions are from how accepting respondents are of them and, in so doing, to detect double standards in judging collective action.  -->
When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters: 
$$ \Pr ( y_{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
where $\Pr ( y_{ij} = 1 )$ is the probability that respondent $j$ considers action $i$ an acceptable means of protest, $\theta_j$ estimates how *accepting* respondent $j$ is of various protest actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.

Figure \ref{fig:f1} illustrates those relationships for three hypothetical protest actions. $\theta_{(z)}$ is the *z*-standardized tendency for a participant to be more accepting of various actions so that $\beta_i$ determines the probability that the average respondent ($\theta_{(z)} = 0$) considers an action an acceptable means of protest. Action 2 ($\beta_2 = 1.28$) is more acceptable than Actions 1 and 2 ($\{ \beta_2, \beta_3 \} = 0$) as the average respondent is more likely to consider the former ($\Pr = .78$) than the latter ($\Pr = .50$) acceptable means of protest. Action 3 ($\alpha_3 = 3.60$) is more discriminating than Actions 1 and 2 ($\{ \alpha_1, \alpha_2 \} = 1$) as the slope for the relationship between how accepting a respondent is and how likely they are to consider an action acceptable is steeper for the former action. This means that knowing whether someone considers Action 3 acceptable provides more information about them than knowing whether they consider Actions 1 or 2 acceptable.

In this research, item response theory serves two purposes. First, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters and what they are protesting. In Experiments 1 and 2, we operationalize double standards as the differences in $\theta_{(z)}$ between conditions which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes. Second, item response theory provides a statistical framework for scale development. In Studies 1 and 2, we select the most discriminating and informative actions to build an instrument for capturing double standards in the experiments.

## Purpose of the Present Research

Our research serves two purposes. The first purpose is to develop an instrument to measure where people draw the line between acceptable and unacceptable forms of collective action. In Study 1, we ask participants to generate more and less extreme protest actions and compile a pool of protest actions for scale development. In Study 2, we ask participants to rate how acceptable they consider each of those protest action to be and use item response theory to select the most discriminating and informative protest actions for our instrument. In so doing, we build an instrument that enables both present and future research to capture double standards in judging collective action. In this way, our research situates the distinction between normative and non-normative collective action in the eye of the beholders and opens it to scientific investigation.

The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more or less acceptable depending on their own and the protesters' group memberships (Hypothesis 1a, 1b) or on the protesters' cause and how it aligns with their own ideological positions (Hypotheses 2a, 2b). In Experiment 1, we use the instrument developed in Studies 1 and 2 to test for double standards in how acceptable participants consider collective action for workers' rights in the United Kingdom. By varying both the participants' and protesters' group memberships (working/middle class), Experiment 1 compares the relative evidence for Hypotheses 1a and 2a. Extending Experiment 1, Experiment 2 considers both system-challenging and system-defending collective action in the context of the 2020 Black Lives Matter protests in the United States. By varying the protest's cause (for/against defunding the police) as well as the participants' and protesters' group memberships (Black/White), Experiment 2 compares the relative evidence for Hypotheses 1a, 1b, 2a, and 2b. Together, the preregistered experiments provide a comprehensive test of the hypothesized identity- and ideology-based double standards in judging collective action.

<!-- > **Hypothesis 1.** People will consider *more* controversial actions acceptable means of protest if (a) the protesters are ingroup members *or* if (b) the cause of the protest aligns with the ingroup's interests. -->

<!-- > **Hypothesis 2.** People will consider *more* controversial actions acceptable means of protest if (a) the protest's cause supports, rather than challenges, the system *or* if (b) the protest's cause aligns with their system-justifying beliefs. -->

# Scale Development

<!-- In two studies, we developed an instrument of 25 protest actions to capture double standards in judging collective action. -->

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Examples of protest actions rated in Study 2}
\includegraphics[scale=1]{../Scale Development/figures/figure-2}
\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause. Darker bars mark actions that were included in the final scale.}
\label{fig:f2}
\end{figure*}
```

## Study 1

In Study 1, we compiled protest actions from participants' responses and other sources. We recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students.

Participants first read an accessible definition of what a social group is and that collective action is any action group members take to promote the interests of their social group. Participants were then asked to name between five and ten actions that fit that definition. Participants were encouraged to think of actions that varied in how acceptable they were in their opinion. We recoded responses into a smaller set of unique actions, which we supplemented with protest actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973]. This process resulted in 72 actions that varied in how acceptable we would expect them to be. For details, see Supplemental Online Material (SOM).

## Study 2

In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action. We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students. We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.

Participants again read an accessible definition of collective action. Participants were then asked to think of different causes and circumstances and to rate how often a given action would be an acceptable means for a group to advance one of these causes (1 = *never*, 2 = *rarely*, 3 = *sometimes*, 4 = *often*, 5 = *always*). In addition, participants rated how disruptive, violent, and extreme they considered an action to be (1 = *not at all*, 4 = *very*) and how positive or negative they felt, in general, about an action (1 = *very positive*, 5 = *very negative*). Each participant rated 20 of the 72 actions from Study 1 so that each action was rated by 29--53 participants.

We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997]---an item response theory model for ordinal response variables---for participants' ratings of how often an action would be an acceptable form of collective action.^[Ratings of how often a given action would be an acceptable form of collective action were strongly and negatively correlated with how disruptive, violent, and extreme participants considered the action to be and how negative they felt about it (Table S3).] Based on the results, we selected the 25 most discriminating, informative, and relevant protest actions to form an instrument for measuring double standards in judging collective action in Experiments 1 and 2. We selected controversial, and thus diagnostic, actions (e.g., disrupting traffic) over actions that almost all respondents considered acceptable (e.g., signing petitions) or unacceptable (e.g., political violence; Figure \ref{fig:f2}). For detailed ratings and results, see SOM.

# Experiment 1

\ldots

## Method

We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations.^[[https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b](https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b)] We made all materials, data, and analysis scripts available online.^[[https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)] We followed sample size recommendations for item response theory models [@demars_irt_2010, p. 36], planning to recruit 500 participants.

### Study Design

We used a 2 (quasi-experimental: higher-status/lower-status participants) $\times$ 2 (experimental: higher-status/lower-status protesters) between-subjects design to test our hypotheses.

### Participants

We recruited 515 participants from the Prolific subject pool who were UK citizens, 25 years old or older, and not current students.^[We had preregistered a sample of 500 participants, before exclusions, but included another 15 participants who completed the study without returning an approval code to the Prolific platform. Participants received, on average, \pounds 9.00 (\$10.93) per hour of participation.] As preregistered, we excluded 71 participants who failed an attention check. This resulted in a final sample of 443 participants ($\textit{Mdn} = 41$ years, age range: 25--76 years; 272 women, 171 men, 1 non-binary). Of these, 210 participants considered their past, current, or future jobs to be working-class jobs. Participants in this lower-status group did not have a university degree and placed themselves on the bottom three ranks of the subjective socioeconomic status ladder. Another 233 participants considered their past, current job, or future jobs to be middle-class/professional jobs. Participants in this higher-status group had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socioeconomic status ladder.

### Procedure

We used a screening survey to recruit participants who satisfied our preregistered inclusion criteria for the lower-status and higher-status groups. For the lower-status group, we recruited 475 participants who did not have a university degree and placed themselves on the bottom three ranks of a subjective socioeconomic status ladder. For the higher-status group, we recruited 400 participants who had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socioeconomic status ladder.

In the screening survey, participants read an accessible definition of what working-class and middle-class/professional jobs are (for details, see SOM). Participants then answered, among other questions, whether they considered their current job---or the jobs they had had in the past or expected to have in the future---to be a working-class job or a middle-class/professional job. As preregistered, we excluded participants from the lower-status group who did not respond  "working-class job" and participants from the higher-status group who did not respond "middle-class/professional job".

We recruited 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group. Participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (lower-status protesters) or people in professional jobs (higher-status protesters). Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real. Participants in the lower-status protestprs condition read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in working-class jobs. Working-class jobs, in this case, are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These are jobs that do not usually require a university degree. Other jobs are unlikely to be affected.

\noindent Participants in the higher-status protesters condition instead read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in professional jobs. Professional jobs are administrative, managerial, or other jobs that usually require a university degree. Other jobs are unlikely to be affected.

\noindent Participants in both conditions then read an almost identically worded paragraph:

> This government measure would make it easier for companies to hire workers during economic growth and to lay off workers during an economic crisis. As a consequence, companies would be able to fire employees with little notice and without giving a reason. Trade unions are opposed to the measure. They argue that the bill would compromise job security, and prevent employees from challenging harassment or other abuse without the fear of being fired. People in [working-class/professional] jobs are particularly at risk, and there is a rise in tension and outrage among them.

On the next pages, participants completed all remaining measures. On the final page, we asked participants to recall who the people most affected by the fictitious government bill were. As preregistered, we excluded all participants whose response did not qualitatively match their experimental condition.

### Measures

We measured the outcome variable by asking participants to decide, for each of 25 protest actions presented in a randomized order, whether they thought this action was "an acceptable means for people in [working-class/professional] jobs to protest against the government bill" (1 = *yes*, 0 = *no*; see Table A1).

We assessed reactions to the vignette by asking participants how outraged they would be if the government were to introduce this bill in real life, to what extent it would affect them personally, and to what extent the it would affect people like them (1 = *not at all*, 7 = *very much*). We also asked participants to what extent they identified with people with the kinds of jobs most affected by the proposed bill (1 = *not at all*, 7 = *very much*) and how often, if at all, they had participated in protest actions such as the ones we asked about (1 = *never*, 5 = *very often*).

We measured social dominance orientation with the eight-item $\text{SDO}_{7(s)}$ scale [@ho_nature_2015], for example, "it is unjust to try to make groups equal" (1 = *strongly oppose*, 7 = *strongly favour*; McDonald's $\omega = .89$). A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 108.69; \text{CFI} = 0.92; \text{TLI} = 0.88; \text{RMSEA} = 0.11, [0.09, 0.14]$.

We measured system-justifying beliefs with eight items [adapted from @kay_complementary_2003], for example, "in general, I find society to be fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .89$). A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 77.37; \text{CFI} = 0.94; \text{TLI} = 0.92; \text{RMSEA} = 0.09, [0.07, 0.12]$.

In the screening survey, we recorded participants' gender, age, nationality, student status, and employment status. We also included two three-item scales measuring social identification with people in working-class and middle-class/professional jobs [adapted from @becker_committed_2011] and a one-item semantic differential scale measuring political orientation ("People often describe their political orientation as left-wing or right-wing. On a scale from left to right, where would you position yourself?"; 1 = *left*, 7 = *right*).

## Results

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Results from the preregistered (\textbf{A}) and non-preregistered (\textbf{B}) analyses for Experiment 1}
\includegraphics[scale=1]{../Experiment 1/figures/figure-3}
\caption*{\textit{Note.} HS = Higher Status, LS = Lower Status. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. (\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition. Bars enclose the 95\% most plausible estimates. (\textbf{B}) Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:f3}
\end{figure*}
```

### Reactions to the Manipulation

Participants with professional jobs thought that they would be more outraged if the government were to introduce a bill affecting people with similar jobs ($M = 6.00, SD = 1.05$) than by a bill affecting people with working-class jobs ($M = 5.38, SD = 1.28$; Cohen's $d = 0.50$). Conversely, participants with working-class jobs thought that they would be more outraged by a bill affecting people in similar jobs ($M = 6.26, SD = 1.05$) than by a bill affecting people in professional jobs ($M = 5.61, SD = 1.47$; $d = 0.52$). Participants thought that a bill affecting people with similar jobs would affect them personally ($d = 1.14$) and people like them ($d = 1.33$) more than a bill affecting people with other jobs. Participants identified to a greater extent with people in similar jobs ($M = 6.21, SD = 1.12$) than with people in other jobs ($M = 4.06, SD = 1.65$; $d = 1.22$). Overall, participants' reactions suggested that the experimental design worked as intended and that participants in the lower-status and higher-status groups understood the vignette in ingroup--outgroup terms.

### Preregistered Analyses

To test our hypotheses, we estimated a two-parameter logistic item response theory model [@burkner_bayesian_2021] with participants' responses to the question whether they thought each action was an acceptable means of protest as the outcome variable. For each action $i$, the model estimated how acceptable ($\beta_i$) and discriminating ($\alpha_i$) that protest action was. For each participant $j$, the model estimated their unique propensity to consider protest actions acceptable ($\theta_j$). In addition, the model estimated how accepting participants were of controversial protest actions as a function of three dummy-coded variables that encoded the group status of the protesters and the participant.

We estimated this model in CmdStan [@cmdstanr_2021; @cmdstan_2021] using Bayesian statistical methods. Bayesian inference involves choosing a likelihood function and prior distributions. A likelihood function links the observed data to the model parameters and states how likely the observed data are given different values of said model parameters. Prior distributions state how plausible different values of said model parameters are before considering the observed data. Bayesian inference applies Bayes' theorem to update prior distributions in light of the observed data to produce posterior distributions. In contrast to $p$-values and confidence intervals, posterior distributions have a straight-forward interpretation as stating how plausible different values of the model parameters are given the observed data.

Our model derived the likelihood of participants' responses from a Bernoulli likelihood function with a logistic regression equation linking the two item parameters ($\alpha_i$, $\beta_i$), the one participant parameter ($\theta_j$), and the three regression coefficients to the observed data. To identify the model, we constrained $\theta_j$ to have a mean of zero and constrained $\alpha_i$ to have a fixed mean and to be non-negative. We used partial pooling to estimate $\alpha_i$, $\beta_i$, and $\theta_j$. Our model assigned weakly informative prior distributions [@gelman_prior_2017] to all model parameters.^[Our model assigned $\text{Half-Cauchy} (0, 3)$ prior distributions to the standard deviations of $\alpha_i$, $\beta_i$, and $\theta_j$ and $\text{Normal} (0, 3)$ prior distributions to all other model parameters.] We report point estimates, based on the median of posterior samples, and uncertainty intervals, based on the quantiles of posterior samples, that enclose the 95% most plausible estimates.

Figure \ref{fig:f3} (A) shows estimates for each combination of the protesters' and the participants' group membership. Overall, we found that participants' responses depended on both the protesters' and the participants' group status---but not in the directions predicted by our hypotheses. Contradicting Hypothesis 1, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = 0.00, [-0.20, 0.20]$). Instead, we found that participants from higher-status backgrounds considered protest actions performed by both lower-status ($d = 0.34, [0.07, 0.59$]) and higher-status ($d = 0.33, [0.03, 0.63]$) protesters to be, on average, more acceptable than participants from lower-status backgrounds. Contradicting Hypothesis 2, participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.40, -0.00]$).

### Non-Preregistered Analyses

We explored to what extent participants' ideological orientations influenced how acceptable they considered various protest actions to be. To that end, we estimated another two-parameter logistic item response model that estimated participants' responses as a function of the participant's and the protesters' group status and of the participants' political orientation, social dominance orientation, and system-justifying beliefs. We used factor scores to quantify social dominance orientation and system-justifying beliefs and standardized all new predictor variables. We report standardized regression coefficients that quantify by how many standard deviations a participant's propensity to consider more collective actions acceptable increases for each additional standard deviation of the predictor variable.

Figure \ref{fig:f3} (B) the $z$-standardized propensity to consider more controversial protest actions acceptable as a function of the three ideological orientation variables. We found that participants who reported a more right-wing political orientation ($\beta_{xy} = -0.25, [-0.37, -0.15]$) and who expressed more agreement with system-justifying beliefs ($\beta_{xy} = -0.20, [-0.30, -0.10]$) tended to find fewer collective actions to be acceptable. In contrast, we found that, after controlling for the other two variables, social dominance orientation was not associated with participants' judgements about how acceptable various collective actions are ($\beta_{xy} = -0.05, [-0.15, 0.05]$). Overall, these findings suggest that people who are right-wing and endorse system-justifying beliefs tend to find various collective actions to be less acceptable than people who are left-wing and do not endorse system-justifying beliefs.^[We ran additional analyses to explore how these associations differed across experimental conditions. We found uncertainty intervals for these associations to overlap across conditions, suggesting that we did not have enough data to differentiate these varying effects.]

## Discussion

# Experiment 2

- Explain why defunding the police (public support for BLM, so needed to be more controversial)

In Experiment 2, we tested five hypotheses. We tested two predictions derived from social identity theory. First, we hypothesize that observers are more likely to judge an action as an acceptable means of protest when the protesters are ingroup members than when they are outgroup members (Hypothesis 1a). Second, we hypothesize that observers are more likely to judge an action as an acceptable means of protest when the goal of the protests would benefit the observers' ingroup (Hypothesis 1b). We will test two predictions derived from system justification theory. First, we hypothesize that observers are more likely to judge an action as an acceptable means of protest when the goal of the protest is to support rather than challenge the system (Hypothesis 2a). Second, we hypothesize that observers who endorse system-justifying beliefs are more likely to consider system-supporting actions acceptable means of protest-and less likely to consider system-challenging actions acceptable means of protest-than observers who do not endorse system-justifying beliefs (Hypothesis 2b). In addition, we will test an alternative hypothesis. We hypothesize that observers are more likely to consider actions as acceptable means of protest when they support the cause of the protesters than when they oppose their cause (Hypothesis 3). We note that Hypotheses 1a, 1b, and 2a make competing predictions about differences between conditions while Hypotheses 2b and 3 need not contradict each other or any of the other hypotheses.

## Method

We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations.^[[https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1](https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1)] We made all materials, data, and analysis scripts available online.^[[https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)] As reported in the preregistration, we ran simulations, using data from Experiment 1, to determine that a sample size of $N = 1,600$ ($n = 200$ per condition) was sufficient to detect even small differences between conditions ($0.09 < \text{Cohen's}~d < 0.16$).

### Study Design

We used a 2 (quasi-experimental: Black/White participants) $\times$ 2 (experimental: Black/White protesters) $\times$ 2 (experimental: for/against defunding the police) between-subjects design to test our hypotheses.

### Participants

We recruited 1,773 Black and White American participants from the Prolific subject pool who were 18 years old or older, lived in the US, and were US citizens.^[Data were collected between January 15 and 28, 2021. Participants received, on average, \$12.89 per hour of participation.] As preregistered, we excluded 173 participants who failed at least one of three attention checks or who reported a different ethnic background than they had reported in the Prolific prescreening questionnaire. This resulted in a final sample of 1,600 participants ($\textit{Mdn} = 31$ years, age range: 18--84 years; 864 women, 708 men, 27 sex/gender diverse) of whom 800 identified as Black and 800 identified as White.

### Procedure

Participants read the following paragraphs:

> In 2020, police officers killed Breonna Taylor and George Floyd. Many were outraged that police officers had, once again, killed unarmed Black Americans. Across the United States, people called for changes to prevent future police violence.

> Some argue that, to end police violence, we should take money away from police departments. Reducing police funding would mean fewer police officers on the street. Fewer police officers would mean fewer opportunities for them to turn violent. Reducing police funding would also leave more money for other services. Proponents argue for reallocating police funding to social services, housing, and education. Doing so would keep communities safer with fewer police officers. We refer to this position as "defunding the police". This position differs from "reforming the police" which might mean increasing police funding and also differs from "abolishing the police" which means disbanding police departments altogether.

\noindent Participants were then asked to answer whether the text had been about "reforming the police", "defunding the police", or "abolishing the police". If they selected the wrong answer, they were instructed to reread the text and select the right answer. On the next page, participants stated whether they supported or opposed the proposed solution.

Participants in the system-challenging protest condition then read a text about protesters in support of defunding the police:

> Earlier, you read about defunding the police as a possible solution to end police violence. Some local residents want to protest for defunding the police. They argue that reducing police funding would prevent police violence.

\noindent Participants in the system-supporting protest condition instead read about protesters rallying against defunding the police:

> Some local residents want to protest against defunding the police. They argue that reducing police funding would mean fewer police officers serving their community.

\noindent Participants then read either that "most of the protesters are Black" or that "most of the protesters are White". Participants again had to correctly answer multiple choice questions about the text ("Are the protesters for or against defunding the police?", "Who are the protesters?") before being able to proceed.

On the next pages, participants completed all remaining measures. On the final page, participants responded to three attention checks: "In this study, you first read about a proposed solution to police violence. What was it?" (*reforming the police*, *defunding the police*, *abolishing the police*); "In this study, you then read about protesters. Were these protesters for or against the proposed solution?" (*for*, *against*); and "Were most of the protesters Black or White?" (*Black*, *White*). As preregistered, we excluded participants who gave an answer inconsistent with their assigned experimental condition.

### Measures

We measured the outcome variable by asking participants to think about the protesters they had read about and to decide, for each of 25 protest actions presented in a randomized order, whether they thought this action was "an acceptable means for them to protest [for/against] defunding the police" (1 = *yes*, 0 = *no*). We replaced some actions from Experiment 1 because they either did not fit the study context or had been considered acceptable by almost all participants (see Table B1).

We measured system-justifying beliefs with eight items [adapted from @kay_complementary_2003], for example, "in general, I find society to be fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .91$). A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (28) = 4309.15; \text{CFI} = 0.95; \text{TLI} = 0.93; \text{RMSEA} = 0.11, [0.10, 0.12]$.

We measured support for defunding the police with one item: "Do you support or oppose defunding the police?" (1 = *strongly oppose*, 5 = *strongly support*).

We included additional measures to describe the sample, to describe reactions to the manipulation, or to use in non-preregistered analyses. In addition to demographic questions, we asked participants how outraged they were about recent incidents of police violence against Black Americans, to what extent they identifed with the protesters described in the study, and to what extent they identifed with their racial ingroup (1 = *not at all*, 7 = *very much*). We also asked how often, if at all, participants had participated in protest actions such as the ones we had asked about (1 = *never*, 5 = *very often*) and whether they has participated in protests for reforming, defunding or abolishing the police; against reforming, defunding or abolishing the police; or in neither. We measured political orientation with a one-item semantic differential scale: "People often describe their political orientation as liberal or conservative. On a scale from liberal to conservative, where would you position yourself?"; 1 = *liberal*, 7 = *conservative*).

## Results

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-4}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:f4}
\end{figure*}
```

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Predictions from the preregistered (system justification) and non-preregistered (political orientation) analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-5}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
\label{fig:f5}
\end{figure*}
```

### Reactions to the Manipulation

As expected, Black participants ($M = 6.18, \textit{SD} = 1.31$) reported being more outraged about recent incidents of police violence than White participants ($M = 5.56, \textit{SD} = 1.61$; Cohen's $d = 0.42$). Black participants strongly identified with their ingroup ($M = 6.46, \textit{SD} = 1.09$) and, across experimental conditions, identified more with Black protesters ($M = 4.66, \textit{SD} = 1.99$) than with White protesters ($M = 4.03, \textit{SD} = 2.16$; $d = 0.30$). In contrast, White participants identified less strongly with their ingroup ($M = 4.85, \textit{SD} = 1.53$; $d = 1.21$) and did not identify more with White protesters ($M = 3.58, \textit{SD} = 2.01$) than with Black protesters ($M = 3.81, \textit{SD} = 1.77$; $d = -0.12$). Both White ($d = 0.67$) and Black ($d = 0.53$) participants tended to identify more with protesters protesting *for* defunding the police. On average, participants tended to describe their political orientation as moderately liberal ($M = 2.92, \textit{SD} = 1.73$) but, as expected, were divided in their support for defunding the police ($M = 3.37, \textit{SD} = 1.42$).

### Preregistered Analyses

To test our hypotheses, we estimated a series of two-parameter logistic item response theory models with participants' responses to the question whether they thought each action was an acceptable means of protest as the outcome variable. Our models differed from the models used in Experiment 1 in two ways. First, we estimated the item parameters ($\alpha_i$, $\beta_i$) as *correlated* varying effects. Second, we used varying, rather than fixed, effects to estimate differences between the eight conditions. This resulted in partial pooling---condition-wise estimates were shrunk towards each other---and allowed us to compare conditions without multiple comparison problems [@gelman_why_2012]. Our models assigned weakly informative prior distributions to all model parameters.^[Our model assigned $\text{LKJ} (2)$ prior distributions to the Cholesky-transformed correlation matrices for varying effects and $(\text{Half-})\text{Cauchy} (0, 5)$ prior distributions to all other model parameters.] Figure \ref{fig:f4} shows results from three preregistered models we estimated to test our hypotheses.

```{r, include = FALSE}

  # Import results
  m1 <- read_rds("../Experiment 2/results/m1_post.rds")

  # Define contrasts
  m1 <- m1 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1L,
        Participant != Protesters ~ -1L,
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1L,
        Participant == "White" & Cause == "Against" ~ +1L,
        TRUE ~ -1L,
      ),
      h2a_contrast = case_when(
        Cause == "For" ~ -1L,
        Cause == "Against" ~ +1L
      ),
      nh1_contrast = case_when(
        Participant == "Black" ~ -1L,
        Participant == "White" ~ +1L,
      ),
      nh2_contrast = case_when(
        Participant == "White" & Cause == "Against" ~ -1L,
        Participant == "White" & Cause == "For" ~ +1L,
        Participant == "Black" ~ 0L
      ),
      nh3_contrast = case_when(
        Participant == "Black" & Cause == "Against" ~ -1L,
        Participant == "Black" & Cause == "For" ~ +1L,
        Participant == "White" ~ 0L
      )
    )

  # Calculate contrasts
  m1 <- m1 %>% 
    group_by(.draw) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~mean(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

  # Make text
  m1 <- m1 %>% 
    mutate(
      across(
        c(d, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      d_text = glue(
        "$d = {d}, [{.lower}, {.upper}]$"
      )
    )

```

Model 1 (M1) estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause. Contradicting Hypothesis 1a, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`). Contradicting Hypothesis 1b, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`). Contradicting Hypothesis 2a, participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`). Instead, we found that, on average, White participants considered all protest actions to be more acceptable than Black participants (`r with(m1, d_text[contrast == "nh1_contrast"])`) and both Black (`r with(m1, d_text[contrast == "nh3_contrast"])`) and White (`r with(m1, d_text[contrast == "nh2_contrast"])`) participants considered the same actions to be more acceptable when protests were for, rather than against, defunding the police. As in Experiment 1, we thus found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the directions predicted by our hypotheses.

```{r, include = FALSE}

  # Import results
  m2 <- read_rds("../Experiment 2/results/m2_post.rds")

  # Hypothesis 2
  m2_h2 <- m2 %>% 
    distinct(.draw, b_sjb_for, b_sjb_against) %>% 
    pivot_longer(
      c(b_sjb_for, b_sjb_against),
      names_to = "Cause",
      names_prefix = "b_sjb_",
      values_to = "b"
    ) %>% 
    group_by(Cause) %>% 
    mutate(
      b = case_when(
        Cause == "for" ~ -b,
        Cause == "against" ~ b
      )
    ) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

Model 2 (M2) extended Model 1 by estimating participants' responses as a function of their *z*-standardized endorsement of system-justifying beliefs. As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgements about system-defending protest actions and one estimating their effect on judgements about system-challenging protest actions, and one varying effect estimating its variance across conditions. Supporting Hypothesis 2b, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`). We did not, however, find evidence for the ideological symmetry implied by Hypothesis 2b since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).

Figure \ref{fig:f5} shows the predicted pattern of condition-wise differences underlying those fixed effects. Participants who endorsed system-justifying beliefs tended to consider system-challenging and system-defending protest actions to be equally unacceptable. In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with Hypothesis 2b, they considered the same protest actions to be more acceptable if the protesters challenged the system or, to a lesser extent, if the protesters were from the disadvantaged group.

```{r, include = FALSE}

  # Import results
  m3 <- read_rds("../Experiment 2/results/m3_post.rds")

  # Hypothesis 2
  m3_h3 <- m3 %>% 
    distinct(.draw, b_sup) %>%
    rename(b = b_sup) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{y} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

Model 3 (M3) extended Model 1 by estimating participants' responses as a function of their self-reported support for the cause of the protest. We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police. As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Contradicting Hypothesis 3, participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`). Ergo, our results did not support the alternative hypothesis that ideology-based double standards can be reduced to support for, or opposition to, the cause of a protest.

### Non-Preregistered Analyses

```{r, include = FALSE}

  # Import results
  m4 <- read_rds("../Experiment 2/results/m4_post.rds")

  # Hypothesis 2
  m4_exp <- m4 %>% 
    distinct(.draw, b_pol_for, b_pol_against) %>% 
    pivot_longer(
      c(b_pol_for, b_pol_against),
      names_to = "Cause",
      names_prefix = "b_pol_",
      values_to = "b"
    ) %>% 
    group_by(Cause) %>% 
    mutate(
      b = case_when(
        Cause == "for" ~ -b,
        Cause == "against" ~ b
      )
    ) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

```

We reran Model 3 with political orientation, instead of system justification, as an alternative predictor variable. Our results mirrored results from the preregistered analyses: More liberal participants were more likely to consider protest actions for defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "for"])`) but more conservative participants were not more likely to consider protest actions against defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "against"])`). As Figure 5 shows, conservative participants tended to consider all protest actions to be equally unacceptable while liberal participants considered the same protest actions to be more acceptable if protesters rallied around a progressive cause. Our non-preregistered analyses thus found evidence for ideology-based double standards in judging collective action with a different operationalization of political ideology.

## Discussion

# General Discussion

- Summary: No evidence for identity-based double standards, but some evidence for ideology-based double standards
- Two plausible interpretations: (1) conservatives oppose controversial protest (no matter the end) *or* (2) the end justifies the mean
- Re (1) it's plausible but what about recent events? (see public polling from grant)
- Re (2) why is there no asymmetry?
- Outline the-end-justifies-the-mean idea, highlighting that just supporting a cause is not enough, it needs moralization
- Discuss why high-status group are more accepting of controversial protest actions
- Discuss limitations (two cultures, two issues, liberal samples; system-defending vs. system-supporting actions; what about other conservative causes?)
- Invite researchers to use our paradigm

\refsection

\begingroup

\noindent
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\small

<div id="refs"></div>

\endgroup

\newpage

```{=latex}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\begin{table*}

\section{Appendix A}

\caption{List of protest actions used in Experiment 1}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & participate in a public meeting of representatives and elected officials & 97\%\\
2 & hold meetings to inform the public & 96\%\\
3 & make a public speech & 96\%\\
4 & hold meetings to influence the public & 93\%\\
5 & attend or organise a protest march & 93\%\\
6 & attend or organise a protest rally & 92\%\\
7 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 92\%\\
8 & do not buy goods or services from companies who support the bill (consumers' boycott) & 92\%\\
9 & paste up posters with political messages in places where it is allowed and encouraged & 89\%\\
10 & join or form a group of activists who oppose the bill & 87\%\\
11 & refuse to accept honours or awards in protest & 82\%\\
12 & donate to political parties who oppose the bill & 80\%\\
13 & refuse to work (strike) & 80\%\\
14 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 78\%\\
15 & donate to activist groups who oppose the bill & 77\%\\
16 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 54\%\\
17 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 51\%\\
18 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) until the bill is abandoned & 47\%\\
19 & enter and refuse to leave a building (occupation) & 38\%\\
20 & paste up posters with political messages in places where it is not allowed or encouraged & 35\%\\
21 & disrupt traffic (e.g., blocking roads) & 22\%\\
22 & refuse to cooperate with the police and other government agencies & 19\%\\
23 & mock or insult individuals who support the bill & 14\%\\
24 & spray paint political messages in public places & 14\%\\
25 & deface flags or other national symbols & 12\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 1.}

\end{table*}
```

```{=latex}
\setcounter{table}{0}
\renewcommand{\thetable}{B\arabic{table}}

\begin{table*}

\section{Appendix B}

\caption{List of protest actions used in Experiment 2}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & make a public speech & 94\%\\
2 & \textit{hand out flyers, leaflets, or pamphlets} & 93\%\\
3 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
4 & hold meetings to influence the public & 91\%\\
5 & paste up posters with political messages in places where it is allowed and encouraged & 91\%\\
6 & attend or organise a protest march & 91\%\\
7 & donate to activist groups & 89\%\\
8 & join or form a group of activists & 89\%\\
9 & \textit{wear or display political symbols (e.g., patches, flags, bumper stickers)} & 86\%\\
10 & refuse to buy goods or services from companies that advocate [against/for] defunding the police (boycott) & 85\%\\
11 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 83\%\\
12 & donate to politicians who advocate [for/against] defunding the police & 82\%\\
13 & refuse to accept honors or awards in protest & 80\%\\
14 & refuse to work (strike) & 61\%\\
15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 61\%\\
16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 55\%\\
17 & \textit{attend a protest march even though it might turn violent} & 44\%\\
18 & \textit{attend a protest march even though it might be unlawful} & 42\%\\
19 & \textit{attend a protest march while carrying a firearm (where legal)} & 35\%\\
20 & enter and refuse to leave a building (occupation) & 34\%\\
21 & paste up posters with political messages in places where it is not allowed or encouraged & 29\%\\
22 & \textit{refuse to pay fees, fines, or taxes in protest} & 29\%\\
23 & disrupt traffic (e.g., blocking roads) & 25\%\\
24 & spray paint political messages in public places & 23\%\\
25 & mock or insult individuals who are [against/for] defunding the police & 15\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 2. Actions in \textit{italics} replaced actions used in Experiment 1 that were either redundant or did not fit the context of Experiment 2.}

\end{table*}
```

### 2022-05-24 Documents/manuscript.tex
work in progress

\documentclass[twocolumn, 11pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{MinionPro} 
\input{glyphtounicode}
\pdfgentounicode=1 
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setlength{\columnsep}{18pt}
\setcounter{secnumdepth}{0}
\usepackage[compact]{titlesec}
\titleformat*{\section}{\centering\normalfont\Large\bfseries}
\titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
\titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
\usepackage{authblk}
\renewcommand\Affilfont{\small}
\newcommand{\refsection}{\section{References}}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Links
\usepackage[colorlinks = true, linkcolor = black, urlcolor = black, citecolor = black]{hyperref}

% Figures
\usepackage{graphicx}
\usepackage[labelfont = bf, font = small, labelsep = newline, singlelinecheck = false]{caption}

% Tables
\usepackage{booktabs}
\usepackage{tabularx}

% References
\newenvironment{CSLReferences}[2]{}{}

% Frontmatter
\title{ Double Standards in Judging Collective Action\thanks{} }
\author[1]{Nils K. Reimer}
\affil[1]{University of California, Santa Barbara, United States}
\author[2]{Marija Brankovi\'{c}}
\affil[2]{Singidunum University, Serbia}
\author[3]{Iniobong Essien}
\affil[3]{Leuphana University of L\"{u}neburg, Germany}
\author[4]{Jin X. Goh}
\affil[4]{Colby College, United States}
\author[5]{S\'{e}bastien Goudeau}
\affil[5]{University of Poitiers, France}
\author[6]{N\'{o}ra A. Lantos}
\affil[6]{E\"{o}tv\"{o}s Lor\'{o}nd University, Hungary}
\author[7]{Jenny Veldman}
\affil[7]{KU Leuven, Belgium}

%\date{June 15, 2021}

\begin{document}

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Collective action is a powerful force driving social
change but, at the same time, often sparks contention about what are
acceptable forms of collective action.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

{
  \renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{M.B.--J.V. are listed in alphabetical order. This research was supported by a Seedcorn Research Grant from the European Association of Social Psychology.}
}

\noindent Protest movements often spark contention about what actions
are acceptable means of protest. Reactions to Black American athletes
kneeling during the national anthem to protest racist police violence
are a case in point. Kneeling during the national anthem is not violent,
disruptive, or illegal. And yet, only 29\% of White Americans (66\% of
Black Americans) and 11\% of Republicans (59\% of Democrats) considered
it appropriate for Black athletes to protest in this way (YouGov, 2017).
Indeed, Black athletes who had participated in the protests faced a
backlash in public opinion (Lacina, 2020). This and other examples
suggest that \emph{who} the protesters are and \emph{what} they are
protesting influences how acceptable observers judge their actions to
be.

In this paper, we examine double standards in judging collective
action---that is, whether observers judge the same protest actions as
more or less acceptable depending on their own and the protesters' group
memberships (\emph{identity-based double standards}) or on the
protesters' cause and how it aligns with their own ideological positions
(\emph{ideology-based double standards}). In two studies, we first
develop an instrument of 25 controversial protest actions, based in item
response theory, to capture double standards in judging collective
action. In two preregistered experiments, we then use this instrument to
test for double standards in judging protest actions for workers' rights
in the United Kingdom and for or against defunding the police in the
United States. We find evidence for ideology-based, but not for
identity-based, double standards: People judge the same protest actions
to be more acceptable when the protesters' cause aligns with their own
ideological positions---but show no ingroup bias when judging collective
action by ingroup and outgroup members.

\hypertarget{double-standards-in-judging-collective-action}{%
\subsection{Double Standards in Judging Collective
Action}\label{double-standards-in-judging-collective-action}}

Collective action---that is, any action taken by group members to
advance a shared political goal (for similar definitions, see Becker,
2012; van Zomeren, 2016)---is a powerful force driving social change.
For example, Black Lives Matter protests lastingly shifted public
discourse about racial inequity toward antiracist ideas (Dunivin et al.,
2022).

Collective action can take many forms. Psychologists distinguish between
\emph{normative} collective action that seeks to achieve a political
goal while conforming to the norms of the existing social system and
\emph{non-normative} collective action that violates those norms (Wright
et al., 1990). Recent research has shown that this distinction matters
for how people respond to collective action. Shuman et al. (2021) showed
that non-normative, non-violent collective action is more effective than
either normative or violent collective action at gaining concessions
from those most resistant to social change. Feinberg et al. (2020)
demonstrated, however, that observers are less supportive of social
movements that use extreme protest actions. Similarly, Teixeira et al.
(2020) found that advantaged-group members perceive non-normative
collective action by disadvantaged-group members as more damaging to
their ingroup's social image and are thus less supportive of such
action. This means that activists face the dilemma that non-normative
collective action might be most effective at gaining concessions but
that it also risks reducing popular support for a cause. Together, these
studies show that the distinction between normative and non-normative
collective action is psychologically and societally consequential.

Research has, for the most part, relied on ad-hoc distinctions between
what researchers themselves considered normative and non-normative
collective action. In liberal democracies, researchers tend to consider
actions such as signing petitions, voting in elections, or peaceful
protest to be normative and actions such as blocking traffic, damaging
property, or violent protest to be non-normative. But, as the examples
in the first paragraph show, the distinction between normative and
non-normative collective action might instead be in the eye of the
beholder.

In this research, we move this distinction from the researcher's
intuition into the realm of scientific investigation and test for double
standards in where people draw the line between acceptable and
unacceptable forms of collective action. For that purpose, we define a
double standard as judging the same action as more or less acceptable
depending on who is performing the action and for what reason (for an
analogous definition, see Foschi, 2000).\footnote{Feinberg et al. (2020)
  provided incidental evidence for double standards in judging
  collective action as, in a manipulation check, Black participants
  rated Black Lives Matter protests as less extreme than White
  participants and liberal participants rated protest actions for
  progressive causes as less extreme than conservative participants.} In
this section, we derive predictions from social psychological theories
about how identity-based and ideology-based double standards could lead
people to judge the same protest actions as more or less acceptable
depending on who the protesters are and what they are protesting.

\hypertarget{identity-based-double-standards}{%
\subsubsection{Identity-Based Double
Standards}\label{identity-based-double-standards}}

Social Identity Theory (Tajfel \& Turner, 1979; for a recent review, see
Reimer et al., 2022) argues that just as people show self-serving biases
to achieve positive self-esteem, they show ingroup-serving biases to
achieve positive social identities. In other words, they think about
ingroup and outgroup in ``me''--``not me'' terms and favor ingroup
members (``us'') over outgroup members (``not us,'' Brewer, 2007).

One domain in which ingroup bias manifests is judgements about other
people's actions. Hewstone (1990) reviewed research showing that people
make ingroup-serving causal attributions when judging actions by ingroup
and outgroup members, for example, by attributing negative behavior by
outgroup members to internal causes but attributing negative behavior by
ingroup members to external causes. Valdesolo \& DeSteno (2007)
demonstrated that participants judged the same selfish action as less
unfair when performed by themselves or an ingroup member than when
performed by an outgroup member. Other studies (Abrams et al., 2013;
Endevelt et al., 2021; for exceptions, see Mendoza et al., 2014; Pinto
et al., 2010) provided further evidence for ingroup bias in judging
moral transgressions by ingroup and outgroup members.

In the same vain, we propose that ingroup bias results in identity-based
double standards in judging collective action by ingroup and outgroup
members. That is, we hypothesize that people will judge the same protest
actions as more acceptable if the protesters are ingroup members
(\emph{Hypothesis 1a}) or if the cause of the protest aligns with the
ingroup's interests (\emph{Hypothesis 1b}). This hypothesis could
explain why, for example, far more Black than White Americans considered
it appropriate for Black athletes to kneel down to protest police
violence against Black Americans (YouGov, 2017). Notably, this
hypothesis applies equally to historically disadvantaged groups
mobilizing against social injustice and to historically advantaged
groups defending their group's position.

\hypertarget{ideology-based-double-standards}{%
\subsubsection{Ideology-Based Double
Standards}\label{ideology-based-double-standards}}

System Justification Theory (Jost \& Banaji, 1994; for a recent review,
see Jost, 2020) argues that people are motivated to defend, justify, and
bolster the prevailing social, economic, and political system because
doing so serves basic epistemic, existential, and relational needs. Both
advantaged and disadvantaged groups are thought to be motivated to
justify the existing system, although the strength of this motivation
and its expression vary across individuals and situations.

While system justification often leads people to resist social change,
its relationship to collective action depends on the goal of the action.
Jost et al. (2017) argued that research on collective action needs to
distinguish between system-challenging protest aimed at changing an
unequal status quo and system-supporting protest aimed at maintaining or
defending an unequal status quo. Osborne et al. (2019) showed that, for
members of both advantaged and disadvantaged groups, system
justification is associated with supporting system-supporting actions
but with opposing system-challenging actions.

Just as system justification shapes support for, and opposition to,
collective action, we propose that it results in ideology-based double
standards in what actions people consider acceptable forms of collective
action. First, we hypothesize that people will judge the same protest
actions as more acceptable if the protest supports, rather than
challenges, the system (\emph{Hypothesis 2a}). This hypothesis reflects
the assumption that the motivation to justify the system is, to some
extent, universal. Second, we hypothesize that people will judge the
same protest actions as more acceptable if they endorse
system-justifying beliefs and the protest supports the system or if they
reject system-justifying beliefs and the protest challenges the system
(\emph{Hypothesis 2b}). This hypothesis implies an symmetry in how
people with different ideological orientations judge system-supporting
and system-challenging collective action.

As system-justifying beliefs are closely related to political
conservatism (Jost et al., 2003), this hypothesis could explain partisan
differences in what actions liberals and conservatives consider
acceptable means of protest for progressive and conservative causes.
Ideology-based double standards could, for example, explain why far more
Democrats than Republicans considered it appropriate to kneel during the
national anthem to protest racist police violence (YouGov, 2017).

\hypertarget{using-item-response-theory-to-detect-double-standards}{%
\subsection{Using Item Response Theory to Detect Double
Standards}\label{using-item-response-theory-to-detect-double-standards}}

\begin{figure}
\centering
\caption{Item response curves for three hypothetical protest actions}
\includegraphics[scale=1]{../Scale Development/figures/figure-1}
\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
\label{fig:f1}
\end{figure}

Item response theory is a a conceptual and statistical framework for
understanding how the characteristics of both items and respondents
shape responses to a set of items (DeMars, 2010). When respondents judge
whether each of a set of actions is an acceptable means of protest (1 =
\emph{yes}, 0 = \emph{no}), a two parameter logistic item response
theory model estimates responses as a function of several latent
(unobserved) parameters:
\[ \Pr ( y_{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) \]
where \(\Pr ( y_{ij} = 1 )\) is the probability that respondent \(j\)
considers action \(i\) an acceptable means of protest, \(\theta_j\)
estimates how \emph{accepting} respondent \(j\) is of various protest
actions, \(\beta_i\) estimates how \emph{acceptable} action \(i\) is,
and \(\alpha_i\) estimates how \emph{discriminating} action \(i\) is.

Figure \ref{fig:f1} illustrates those relationships for three
hypothetical protest actions. \(\theta_{(z)}\) is the
\emph{z}-standardized tendency for a participant to be more accepting of
various actions so that \(\beta_i\) determines the probability that the
average respondent (\(\theta_{(z)} = 0\)) considers an action an
acceptable means of protest. Action 2 (\(\beta_2 = 1.28\)) is more
acceptable than Actions 1 and 2 (\(\{ \beta_2, \beta_3 \} = 0\)) as the
average respondent is more likely to consider the former (\(\Pr = .78\))
than the latter (\(\Pr = .50\)) acceptable means of protest. Action 3
(\(\alpha_3 = 3.60\)) is more discriminating than Actions 1 and 2
(\(\{ \alpha_1, \alpha_2 \} = 1\)) as the slope for the relationship
between how accepting a respondent is and how likely they are to
consider an action acceptable is steeper for the former action. This
means that knowing whether someone considers Action 3 acceptable
provides more information about them than knowing whether they consider
Actions 1 or 2 acceptable.

In this research, item response theory serves two purposes. First, item
response theory provides a formal and statistical definition of double
standards in judging collective action as respondents being more
accepting of the same protest actions depending on who the protesters
and what they are protesting. In Experiments 1 and 2, we operationalize
double standards as the differences in \(\theta_{(z)}\) between
conditions which, as \(\theta_{(z)}\) is \(z\)-standardized, correspond
to Cohen's \(d\) effect sizes. Second, item response theory provides a
statistical framework for scale development. In Studies 1 and 2, we
select the most discriminating and informative actions to build an
instrument for capturing double standards in the experiments.

\hypertarget{purpose-of-the-present-research}{%
\subsection{Purpose of the Present
Research}\label{purpose-of-the-present-research}}

Our research serves two purposes. The first purpose is to develop an
instrument to measure where people draw the line between acceptable and
unacceptable forms of collective action. In Study 1, we ask participants
to generate more and less extreme protest actions and compile a pool of
protest actions for scale development. In Study 2, we ask participants
to rate how acceptable they consider each of those protest action to be
and use item response theory to select the most discriminating and
informative protest actions for our instrument. In so doing, we build an
instrument that enables both present and future research to capture
double standards in judging collective action. In this way, our research
situates the distinction between normative and non-normative collective
action in the eye of the beholders and opens it to scientific
investigation.

The second purpose of our research is to test for double standards in
judging collective action---that is, whether people judge the same
protest actions as more or less acceptable depending on their own and
the protesters' group memberships (Hypothesis 1a, 1b) or on the
protesters' cause and how it aligns with their own ideological positions
(Hypotheses 2a, 2b). In Experiment 1, we use the instrument developed in
Studies 1 and 2 to test for double standards in how acceptable
participants consider collective action for workers' rights in the
United Kingdom. By varying both the participants' and protesters' group
memberships (working/middle class), Experiment 1 compares the relative
evidence for Hypotheses 1a and 2a. Extending Experiment 1, Experiment 2
considers both system-challenging and system-defending collective action
in the context of the 2020 Black Lives Matter protests in the United
States. By varying the protest's cause (for/against defunding the
police) as well as the participants' and protesters' group memberships
(Black/White), Experiment 2 compares the relative evidence for
Hypotheses 1a, 1b, 2a, and 2b. Together, the preregistered experiments
provide a comprehensive test of the hypothesized identity- and
ideology-based double standards in judging collective action.

\hypertarget{scale-development}{%
\section{Scale Development}\label{scale-development}}

\begin{figure*}[!t]
\centering
\caption{Examples of protest actions rated in Study 2}
\includegraphics[scale=1]{../Scale Development/figures/figure-2}
\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause. Darker bars mark actions that were included in the final scale.}
\label{fig:f2}
\end{figure*}

\hypertarget{study-1}{%
\subsection{Study 1}\label{study-1}}

In Study 1, we compiled protest actions from participants' responses and
other sources. We recruited 60 participants from the Prolific subject
pool, all of whom were citizens of the UK or the US. To increase the
socioeconomic diversity of our sample, we recruited 30 non-students
without a university degree, 15 non-students with a university degree,
and 15 current university students.

Participants first read an accessible definition of what a social group
is and that collective action is any action group members take to
promote the interests of their social group. Participants were then
asked to name between five and ten actions that fit that definition.
Participants were encouraged to think of actions that varied in how
acceptable they were in their opinion. We recoded responses into a
smaller set of unique actions, which we supplemented with protest
actions from the psychological and political science literature (e.g.,
Sharp, 1973). This process resulted in 72 actions that varied in how
acceptable we would expect them to be. For details, see Supplemental
Online Material (SOM).

\hypertarget{study-2}{%
\subsection{Study 2}\label{study-2}}

In Study 2, we measured how acceptable participants judged the actions
from Study 1 to be and applied item response theory to develop an
instrument to capture double standards in judging collective action. We
recruited 158 participants (\(\textit{Mdn} = 30\) years, age range:
18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the
Prolific subject pool, all of whom were citizens of the UK or the US. To
increase the socioeconomic diversity of our sample, we recruited 80
non-students without a university degree, 37 non-students with a
university degree, and 41 current university students. We excluded 15
participants who failed an attention check, leaving a final sample of
143 participants for our analyses.

Participants again read an accessible definition of collective action.
Participants were then asked to think of different causes and
circumstances and to rate how often a given action would be an
acceptable means for a group to advance one of these causes (1 =
\emph{never}, 2 = \emph{rarely}, 3 = \emph{sometimes}, 4 = \emph{often},
5 = \emph{always}). In addition, participants rated how disruptive,
violent, and extreme they considered an action to be (1 = \emph{not at
all}, 4 = \emph{very}) and how positive or negative they felt, in
general, about an action (1 = \emph{very positive}, 5 = \emph{very
negative}). Each participant rated 20 of the 72 actions from Study 1 so
that each action was rated by 29--53 participants.

We estimated a graded response model (Bürkner, 2021; Samejima,
1997)---an item response theory model for ordinal response
variables---for participants' ratings of how often an action would be an
acceptable form of collective action.\footnote{Ratings of how often a
  given action would be an acceptable form of collective action were
  strongly and negatively correlated with how disruptive, violent, and
  extreme participants considered the action to be and how negative they
  felt about it (Table S3).} Based on the results, we selected the 25
most discriminating, informative, and relevant protest actions to form
an instrument for measuring double standards in judging collective
action in Experiments 1 and 2. We selected controversial, and thus
diagnostic, actions (e.g., disrupting traffic) over actions that almost
all respondents considered acceptable (e.g., signing petitions) or
unacceptable (e.g., political violence; Figure \ref{fig:f2}). For
detailed ratings and results, see SOM.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

\ldots

\hypertarget{method}{%
\subsection{Method}\label{method}}

We preregistered the sample size as well as all hypotheses,
inclusion/exclusion criteria, measures, and manipulations.\footnote{\url{https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b}}
We made all materials, data, and analysis scripts available
online.\footnote{\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}}
We followed sample size recommendations for item response theory models
(DeMars, 2010, p. 36), planning to recruit 500 participants.

\hypertarget{study-design}{%
\subsubsection{Study Design}\label{study-design}}

We used a 2 (quasi-experimental: higher-status/lower-status
participants) \(\times\) 2 (experimental: higher-status/lower-status
protesters) between-subjects design to test our hypotheses.

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

We recruited 515 participants from the Prolific subject pool who were UK
citizens, 25 years old or older, and not current students.\footnote{We
  had preregistered a sample of 500 participants, before exclusions, but
  included another 15 participants who completed the study without
  returning an approval code to the Prolific platform. Participants
  received, on average, \pounds 9.00 (\$10.93) per hour of
  participation.} As preregistered, we excluded 71 participants who
failed an attention check. This resulted in a final sample of 443
participants (\(\textit{Mdn} = 41\) years, age range: 25--76 years; 272
women, 171 men, 1 non-binary). Of these, 210 participants considered
their past, current, or future jobs to be working-class jobs.
Participants in this lower-status group did not have a university degree
and placed themselves on the bottom three ranks of the subjective
socioeconomic status ladder. Another 233 participants considered their
past, current job, or future jobs to be middle-class/professional jobs.
Participants in this higher-status group had at least an undergraduate
degree and placed themselves on the top four ranks of the subjective
socioeconomic status ladder.

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

We used a screening survey to recruit participants who satisfied our
preregistered inclusion criteria for the lower-status and higher-status
groups. For the lower-status group, we recruited 475 participants who
did not have a university degree and placed themselves on the bottom
three ranks of a subjective socioeconomic status ladder. For the
higher-status group, we recruited 400 participants who had at least an
undergraduate degree and placed themselves on the top four ranks of the
subjective socioeconomic status ladder.

In the screening survey, participants read an accessible definition of
what working-class and middle-class/professional jobs are (for details,
see SOM). Participants then answered, among other questions, whether
they considered their current job---or the jobs they had had in the past
or expected to have in the future---to be a working-class job or a
middle-class/professional job. As preregistered, we excluded
participants from the lower-status group who did not respond
``working-class job'' and participants from the higher-status group who
did not respond ``middle-class/professional job''.

We recruited 500 participants from the remaining 687 participants, 250
from the lower-status group and 250 from the higher-status group.
Participants were randomly assigned to read a vignette about a
government bill affecting either people in working-class jobs
(lower-status protesters) or people in professional jobs (higher-status
protesters). Participants in both conditions were instructed to
carefully read the vignette and to try to imagine what it would be like
if this situation was real. Participants in the lower-status protestprs
condition read the following introduction:

\begin{quote}
The government, though not necessarily the current government, is going
to introduce a bill that will mostly affect people in working-class
jobs. Working-class jobs, in this case, are jobs done by skilled,
semi-skilled, unskilled manual workers or by casual workers. These are
jobs that do not usually require a university degree. Other jobs are
unlikely to be affected.
\end{quote}

\noindent Participants in the higher-status protesters condition instead
read the following introduction:

\begin{quote}
The government, though not necessarily the current government, is going
to introduce a bill that will mostly affect people in professional jobs.
Professional jobs are administrative, managerial, or other jobs that
usually require a university degree. Other jobs are unlikely to be
affected.
\end{quote}

\noindent Participants in both conditions then read an almost
identically worded paragraph:

\begin{quote}
This government measure would make it easier for companies to hire
workers during economic growth and to lay off workers during an economic
crisis. As a consequence, companies would be able to fire employees with
little notice and without giving a reason. Trade unions are opposed to
the measure. They argue that the bill would compromise job security, and
prevent employees from challenging harassment or other abuse without the
fear of being fired. People in {[}working-class/professional{]} jobs are
particularly at risk, and there is a rise in tension and outrage among
them.
\end{quote}

On the next pages, participants completed all remaining measures. On the
final page, we asked participants to recall who the people most affected
by the fictitious government bill were. As preregistered, we excluded
all participants whose response did not qualitatively match their
experimental condition.

\hypertarget{measures}{%
\subsubsection{Measures}\label{measures}}

We measured the outcome variable by asking participants to decide, for
each of 25 protest actions presented in a randomized order, whether they
thought this action was ``an acceptable means for people in
{[}working-class/professional{]} jobs to protest against the government
bill'' (1 = \emph{yes}, 0 = \emph{no}; see Table A1).

We assessed reactions to the vignette by asking participants how
outraged they would be if the government were to introduce this bill in
real life, to what extent it would affect them personally, and to what
extent the it would affect people like them (1 = \emph{not at all}, 7 =
\emph{very much}). We also asked participants to what extent they
identified with people with the kinds of jobs most affected by the
proposed bill (1 = \emph{not at all}, 7 = \emph{very much}) and how
often, if at all, they had participated in protest actions such as the
ones we asked about (1 = \emph{never}, 5 = \emph{very often}).

We measured social dominance orientation with the eight-item
\(\text{SDO}_{7(s)}\) scale (Ho et al., 2015), for example, ``it is
unjust to try to make groups equal'' (1 = \emph{strongly oppose}, 7 =
\emph{strongly favour}; McDonald's \(\omega = .89\)). A confirmatory
factor analysis model in which all items loaded onto a single factor
showed acceptable fit,
\(\chi^2 (20) = 108.69; \text{CFI} = 0.92; \text{TLI} = 0.88; \text{RMSEA} = 0.11, [0.09, 0.14]\).

We measured system-justifying beliefs with eight items (adapted from Kay
\& Jost, 2003), for example, ``in general, I find society to be fair''
(1 = \emph{strongly disagree}, 7 = \emph{strongly agree}; McDonald's
\(\omega = .89\)). A confirmatory factor analysis model in which all
items loaded onto a single factor showed acceptable fit,
\(\chi^2 (20) = 77.37; \text{CFI} = 0.94; \text{TLI} = 0.92; \text{RMSEA} = 0.09, [0.07, 0.12]\).

In the screening survey, we recorded participants' gender, age,
nationality, student status, and employment status. We also included two
three-item scales measuring social identification with people in
working-class and middle-class/professional jobs (adapted from Becker et
al., 2011) and a one-item semantic differential scale measuring
political orientation (``People often describe their political
orientation as left-wing or right-wing. On a scale from left to right,
where would you position yourself?''; 1 = \emph{left}, 7 =
\emph{right}).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\begin{figure*}[!t]
\centering
\caption{Results from the preregistered (\textbf{A}) and non-preregistered (\textbf{B}) analyses for Experiment 1}
\includegraphics[scale=1]{../Experiment 1/figures/figure-3}
\caption*{\textit{Note.} HS = Higher Status, LS = Lower Status. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. (\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition. Bars enclose the 95\% most plausible estimates. (\textbf{B}) Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:f3}
\end{figure*}

\hypertarget{reactions-to-the-manipulation}{%
\subsubsection{Reactions to the
Manipulation}\label{reactions-to-the-manipulation}}

Participants with professional jobs thought that they would be more
outraged if the government were to introduce a bill affecting people
with similar jobs (\(M = 6.00, SD = 1.05\)) than by a bill affecting
people with working-class jobs (\(M = 5.38, SD = 1.28\); Cohen's
\(d = 0.50\)). Conversely, participants with working-class jobs thought
that they would be more outraged by a bill affecting people in similar
jobs (\(M = 6.26, SD = 1.05\)) than by a bill affecting people in
professional jobs (\(M = 5.61, SD = 1.47\); \(d = 0.52\)). Participants
thought that a bill affecting people with similar jobs would affect them
personally (\(d = 1.14\)) and people like them (\(d = 1.33\)) more than
a bill affecting people with other jobs. Participants identified to a
greater extent with people in similar jobs (\(M = 6.21, SD = 1.12\))
than with people in other jobs (\(M = 4.06, SD = 1.65\); \(d = 1.22\)).
Overall, participants' reactions suggested that the experimental design
worked as intended and that participants in the lower-status and
higher-status groups understood the vignette in ingroup--outgroup terms.

\hypertarget{preregistered-analyses}{%
\subsubsection{Preregistered Analyses}\label{preregistered-analyses}}

To test our hypotheses, we estimated a two-parameter logistic item
response theory model (Bürkner, 2021) with participants' responses to
the question whether they thought each action was an acceptable means of
protest as the outcome variable. For each action \(i\), the model
estimated how acceptable (\(\beta_i\)) and discriminating (\(\alpha_i\))
that protest action was. For each participant \(j\), the model estimated
their unique propensity to consider protest actions acceptable
(\(\theta_j\)). In addition, the model estimated how accepting
participants were of controversial protest actions as a function of
three dummy-coded variables that encoded the group status of the
protesters and the participant.

We estimated this model in CmdStan (Gabry \& Cesnovar, 2021; Stan
Development Team, 2021) using Bayesian statistical methods. Bayesian
inference involves choosing a likelihood function and prior
distributions. A likelihood function links the observed data to the
model parameters and states how likely the observed data are given
different values of said model parameters. Prior distributions state how
plausible different values of said model parameters are before
considering the observed data. Bayesian inference applies Bayes' theorem
to update prior distributions in light of the observed data to produce
posterior distributions. In contrast to \(p\)-values and confidence
intervals, posterior distributions have a straight-forward
interpretation as stating how plausible different values of the model
parameters are given the observed data.

Our model derived the likelihood of participants' responses from a
Bernoulli likelihood function with a logistic regression equation
linking the two item parameters (\(\alpha_i\), \(\beta_i\)), the one
participant parameter (\(\theta_j\)), and the three regression
coefficients to the observed data. To identify the model, we constrained
\(\theta_j\) to have a mean of zero and constrained \(\alpha_i\) to have
a fixed mean and to be non-negative. We used partial pooling to estimate
\(\alpha_i\), \(\beta_i\), and \(\theta_j\). Our model assigned weakly
informative prior distributions (Gelman et al., 2017) to all model
parameters.\footnote{Our model assigned \(\text{Half-Cauchy} (0, 3)\)
  prior distributions to the standard deviations of \(\alpha_i\),
  \(\beta_i\), and \(\theta_j\) and \(\text{Normal} (0, 3)\) prior
  distributions to all other model parameters.} We report point
estimates, based on the median of posterior samples, and uncertainty
intervals, based on the quantiles of posterior samples, that enclose the
95\% most plausible estimates.

Figure \ref{fig:f3} (A) shows estimates for each combination of the
protesters' and the participants' group membership. Overall, we found
that participants' responses depended on both the protesters' and the
participants' group status---but not in the directions predicted by our
hypotheses. Contradicting Hypothesis 1, participants did not consider
protest actions performed by their ingroup to be more acceptable, on
average, than the same actions performed by the relevant outgroup
(Cohen's \(d = 0.00, [-0.20, 0.20]\)). Instead, we found that
participants from higher-status backgrounds considered protest actions
performed by both lower-status (\(d = 0.34, [0.07, 0.59\){]}) and
higher-status (\(d = 0.33, [0.03, 0.63]\)) protesters to be, on average,
more acceptable than participants from lower-status backgrounds.
Contradicting Hypothesis 2, participants considered protest actions
performed by higher-status protesters to be less acceptable, on average,
than the same actions performed by lower-status protesters
(\(d = -0.20, [-0.40, -0.00]\)).

\hypertarget{non-preregistered-analyses}{%
\subsubsection{Non-Preregistered
Analyses}\label{non-preregistered-analyses}}

We explored to what extent participants' ideological orientations
influenced how acceptable they considered various protest actions to be.
To that end, we estimated another two-parameter logistic item response
model that estimated participants' responses as a function of the
participant's and the protesters' group status and of the participants'
political orientation, social dominance orientation, and
system-justifying beliefs. We used factor scores to quantify social
dominance orientation and system-justifying beliefs and standardized all
new predictor variables. We report standardized regression coefficients
that quantify by how many standard deviations a participant's propensity
to consider more collective actions acceptable increases for each
additional standard deviation of the predictor variable.

Figure \ref{fig:f3} (B) the \(z\)-standardized propensity to consider
more controversial protest actions acceptable as a function of the three
ideological orientation variables. We found that participants who
reported a more right-wing political orientation
(\(\beta_{xy} = -0.25, [-0.37, -0.15]\)) and who expressed more
agreement with system-justifying beliefs
(\(\beta_{xy} = -0.20, [-0.30, -0.10]\)) tended to find fewer collective
actions to be acceptable. In contrast, we found that, after controlling
for the other two variables, social dominance orientation was not
associated with participants' judgements about how acceptable various
collective actions are (\(\beta_{xy} = -0.05, [-0.15, 0.05]\)). Overall,
these findings suggest that people who are right-wing and endorse
system-justifying beliefs tend to find various collective actions to be
less acceptable than people who are left-wing and do not endorse
system-justifying beliefs.\footnote{We ran additional analyses to
  explore how these associations differed across experimental
  conditions. We found uncertainty intervals for these associations to
  overlap across conditions, suggesting that we did not have enough data
  to differentiate these varying effects.}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

\begin{itemize}
\tightlist
\item
  Explain why defunding the police (public support for BLM, so needed to
  be more controversial)
\end{itemize}

In Experiment 2, we tested five hypotheses. We tested two predictions
derived from social identity theory. First, we hypothesize that
observers are more likely to judge an action as an acceptable means of
protest when the protesters are ingroup members than when they are
outgroup members (Hypothesis 1a). Second, we hypothesize that observers
are more likely to judge an action as an acceptable means of protest
when the goal of the protests would benefit the observers' ingroup
(Hypothesis 1b). We will test two predictions derived from system
justification theory. First, we hypothesize that observers are more
likely to judge an action as an acceptable means of protest when the
goal of the protest is to support rather than challenge the system
(Hypothesis 2a). Second, we hypothesize that observers who endorse
system-justifying beliefs are more likely to consider system-supporting
actions acceptable means of protest-and less likely to consider
system-challenging actions acceptable means of protest-than observers
who do not endorse system-justifying beliefs (Hypothesis 2b). In
addition, we will test an alternative hypothesis. We hypothesize that
observers are more likely to consider actions as acceptable means of
protest when they support the cause of the protesters than when they
oppose their cause (Hypothesis 3). We note that Hypotheses 1a, 1b, and
2a make competing predictions about differences between conditions while
Hypotheses 2b and 3 need not contradict each other or any of the other
hypotheses.

\hypertarget{method-1}{%
\subsection{Method}\label{method-1}}

We preregistered the sample size as well as all hypotheses,
inclusion/exclusion criteria, measures, and manipulations.\footnote{\url{https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1}}
We made all materials, data, and analysis scripts available
online.\footnote{\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}}
As reported in the preregistration, we ran simulations, using data from
Experiment 1, to determine that a sample size of \(N = 1,600\)
(\(n = 200\) per condition) was sufficient to detect even small
differences between conditions (\(0.09 < \text{Cohen's}~d < 0.16\)).

\hypertarget{study-design-1}{%
\subsubsection{Study Design}\label{study-design-1}}

We used a 2 (quasi-experimental: Black/White participants) \(\times\) 2
(experimental: Black/White protesters) \(\times\) 2 (experimental:
for/against defunding the police) between-subjects design to test our
hypotheses.

\hypertarget{participants-1}{%
\subsubsection{Participants}\label{participants-1}}

We recruited 1,773 Black and White American participants from the
Prolific subject pool who were 18 years old or older, lived in the US,
and were US citizens.\footnote{Data were collected between January 15
  and 28, 2021. Participants received, on average, \$12.89 per hour of
  participation.} As preregistered, we excluded 173 participants who
failed at least one of three attention checks or who reported a
different ethnic background than they had reported in the Prolific
prescreening questionnaire. This resulted in a final sample of 1,600
participants (\(\textit{Mdn} = 31\) years, age range: 18--84 years; 864
women, 708 men, 27 sex/gender diverse) of whom 800 identified as Black
and 800 identified as White.

\hypertarget{procedure-1}{%
\subsubsection{Procedure}\label{procedure-1}}

Participants read the following paragraphs:

\begin{quote}
In 2020, police officers killed Breonna Taylor and George Floyd. Many
were outraged that police officers had, once again, killed unarmed Black
Americans. Across the United States, people called for changes to
prevent future police violence.
\end{quote}

\begin{quote}
Some argue that, to end police violence, we should take money away from
police departments. Reducing police funding would mean fewer police
officers on the street. Fewer police officers would mean fewer
opportunities for them to turn violent. Reducing police funding would
also leave more money for other services. Proponents argue for
reallocating police funding to social services, housing, and education.
Doing so would keep communities safer with fewer police officers. We
refer to this position as ``defunding the police''. This position
differs from ``reforming the police'' which might mean increasing police
funding and also differs from ``abolishing the police'' which means
disbanding police departments altogether.
\end{quote}

\noindent Participants were then asked to answer whether the text had
been about ``reforming the police'', ``defunding the police'', or
``abolishing the police''. If they selected the wrong answer, they were
instructed to reread the text and select the right answer. On the next
page, participants stated whether they supported or opposed the proposed
solution.

Participants in the system-challenging protest condition then read a
text about protesters in support of defunding the police:

\begin{quote}
Earlier, you read about defunding the police as a possible solution to
end police violence. Some local residents want to protest for defunding
the police. They argue that reducing police funding would prevent police
violence.
\end{quote}

\noindent Participants in the system-supporting protest condition
instead read about protesters rallying against defunding the police:

\begin{quote}
Some local residents want to protest against defunding the police. They
argue that reducing police funding would mean fewer police officers
serving their community.
\end{quote}

\noindent Participants then read either that ``most of the protesters
are Black'' or that ``most of the protesters are White''. Participants
again had to correctly answer multiple choice questions about the text
(``Are the protesters for or against defunding the police?'', ``Who are
the protesters?'') before being able to proceed.

On the next pages, participants completed all remaining measures. On the
final page, participants responded to three attention checks: ``In this
study, you first read about a proposed solution to police violence. What
was it?'' (\emph{reforming the police}, \emph{defunding the police},
\emph{abolishing the police}); ``In this study, you then read about
protesters. Were these protesters for or against the proposed
solution?'' (\emph{for}, \emph{against}); and ``Were most of the
protesters Black or White?'' (\emph{Black}, \emph{White}). As
preregistered, we excluded participants who gave an answer inconsistent
with their assigned experimental condition.

\hypertarget{measures-1}{%
\subsubsection{Measures}\label{measures-1}}

We measured the outcome variable by asking participants to think about
the protesters they had read about and to decide, for each of 25 protest
actions presented in a randomized order, whether they thought this
action was ``an acceptable means for them to protest {[}for/against{]}
defunding the police'' (1 = \emph{yes}, 0 = \emph{no}). We replaced some
actions from Experiment 1 because they either did not fit the study
context or had been considered acceptable by almost all participants
(see Table B1).

We measured system-justifying beliefs with eight items (adapted from Kay
\& Jost, 2003), for example, ``in general, I find society to be fair''
(1 = \emph{strongly disagree}, 7 = \emph{strongly agree}; McDonald's
\(\omega = .91\)). A confirmatory factor analysis model in which all
items loaded onto a single factor showed acceptable fit,
\(\chi^2 (28) = 4309.15; \text{CFI} = 0.95; \text{TLI} = 0.93; \text{RMSEA} = 0.11, [0.10, 0.12]\).

We measured support for defunding the police with one item: ``Do you
support or oppose defunding the police?'' (1 = \emph{strongly oppose}, 5
= \emph{strongly support}).

We included additional measures to describe the sample, to describe
reactions to the manipulation, or to use in non-preregistered analyses.
In addition to demographic questions, we asked participants how outraged
they were about recent incidents of police violence against Black
Americans, to what extent they identifed with the protesters described
in the study, and to what extent they identifed with their racial
ingroup (1 = \emph{not at all}, 7 = \emph{very much}). We also asked how
often, if at all, participants had participated in protest actions such
as the ones we had asked about (1 = \emph{never}, 5 = \emph{very often})
and whether they has participated in protests for reforming, defunding
or abolishing the police; against reforming, defunding or abolishing the
police; or in neither. We measured political orientation with a one-item
semantic differential scale: ``People often describe their political
orientation as liberal or conservative. On a scale from liberal to
conservative, where would you position yourself?''; 1 = \emph{liberal},
7 = \emph{conservative}).

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-4}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest. Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
\label{fig:f4}
\end{figure*}

\begin{figure*}[!t]
\centering
\caption{Predictions from the preregistered (system justification) and non-preregistered (political orientation) analyses for Experiment 2}
\includegraphics[scale=1]{../Experiment 2/figures/figure-5}
\caption*{\textit{Note.} Against = Protesters oppose defunding the police. For = Protesters support defunding the police. $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
\label{fig:f5}
\end{figure*}

\hypertarget{reactions-to-the-manipulation-1}{%
\subsubsection{Reactions to the
Manipulation}\label{reactions-to-the-manipulation-1}}

As expected, Black participants (\(M = 6.18, \textit{SD} = 1.31\))
reported being more outraged about recent incidents of police violence
than White participants (\(M = 5.56, \textit{SD} = 1.61\); Cohen's
\(d = 0.42\)). Black participants strongly identified with their ingroup
(\(M = 6.46, \textit{SD} = 1.09\)) and, across experimental conditions,
identified more with Black protesters (\(M = 4.66, \textit{SD} = 1.99\))
than with White protesters (\(M = 4.03, \textit{SD} = 2.16\);
\(d = 0.30\)). In contrast, White participants identified less strongly
with their ingroup (\(M = 4.85, \textit{SD} = 1.53\); \(d = 1.21\)) and
did not identify more with White protesters
(\(M = 3.58, \textit{SD} = 2.01\)) than with Black protesters
(\(M = 3.81, \textit{SD} = 1.77\); \(d = -0.12\)). Both White
(\(d = 0.67\)) and Black (\(d = 0.53\)) participants tended to identify
more with protesters protesting \emph{for} defunding the police. On
average, participants tended to describe their political orientation as
moderately liberal (\(M = 2.92, \textit{SD} = 1.73\)) but, as expected,
were divided in their support for defunding the police
(\(M = 3.37, \textit{SD} = 1.42\)).

\hypertarget{preregistered-analyses-1}{%
\subsubsection{Preregistered Analyses}\label{preregistered-analyses-1}}

To test our hypotheses, we estimated a series of two-parameter logistic
item response theory models with participants' responses to the question
whether they thought each action was an acceptable means of protest as
the outcome variable. Our models differed from the models used in
Experiment 1 in two ways. First, we estimated the item parameters
(\(\alpha_i\), \(\beta_i\)) as \emph{correlated} varying effects.
Second, we used varying, rather than fixed, effects to estimate
differences between the eight conditions. This resulted in partial
pooling---condition-wise estimates were shrunk towards each other---and
allowed us to compare conditions without multiple comparison problems
(Gelman et al., 2012). Our models assigned weakly informative prior
distributions to all model parameters.\footnote{Our model assigned
  \(\text{LKJ} (2)\) prior distributions to the Cholesky-transformed
  correlation matrices for varying effects and
  \((\text{Half-})\text{Cauchy} (0, 5)\) prior distributions to all
  other model parameters.} Figure \ref{fig:f4} shows results from three
preregistered models we estimated to test our hypotheses.

Model 1 (M1) estimated varying intercepts for the eight conditions to
test whether, in line with Hypotheses 1a, 1b, or 2a, participants'
responses depended on their own group membership, the protesters' group
membership, or the protesters' cause. Contradicting Hypothesis 1a,
participants did not consider protest actions performed by their ingroup
to be more acceptable, on average, than the same actions performed by
the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)). Contradicting
Hypothesis 1b, participants did not consider protest actions for a cause
that was nominally aligned with their ingroup's interests to be more
acceptable, on average, than protest actions for a cause not aligned
with their ingroup's interests (\(d = -0.01, [-0.06, 0.04]\)).
Contradicting Hypothesis 2a, participants did not consider protest
actions for a system-defending cause to be more acceptable, on average,
than protest actions for a system-challenging cause
(\(d = -0.14, [-0.20, -0.09]\)). Instead, we found that, on average,
White participants considered all protest actions to be more acceptable
than Black participants (\(d = 0.09, [0.04, 0.14]\)) and both Black
(\(d = 0.07, [0.03, 0.10]\)) and White (\(d = 0.08, [0.04, 0.11]\))
participants considered the same actions to be more acceptable when
protests were for, rather than against, defunding the police. As in
Experiment 1, we thus found that participants' responses depended on
both the participants' and the protesters' group memberships---but not
in the directions predicted by our hypotheses.

Model 2 (M2) extended Model 1 by estimating participants' responses as a
function of their \emph{z}-standardized endorsement of system-justifying
beliefs. As preregistered, we modeled this relationship with two fixed
effects, one estimating the effect of system-justifying beliefs on
judgements about system-defending protest actions and one estimating
their effect on judgements about system-challenging protest actions, and
one varying effect estimating its variance across conditions. Supporting
Hypothesis 2b, participants who \emph{rejected} system-justifying
beliefs were \emph{more} likely to consider system-challenging protest
actions (for defunding the police) to be acceptable means of protest
(\(\beta_{xy} = 0.38, [0.16, 0.58]\)). We did not, however, find
evidence for the ideological symmetry implied by Hypothesis 2b since
participants who \emph{endorsed} system-justifying beliefs were
\emph{not} more likely to consider system-defending protest actions
(against defunding the police) to be acceptable means of protest
(\(\beta_{xy} = -0.18, [-0.40, 0.02]\)).

Figure \ref{fig:f5} shows the predicted pattern of condition-wise
differences underlying those fixed effects. Participants who endorsed
system-justifying beliefs tended to consider system-challenging and
system-defending protest actions to be equally unacceptable. In
contrast, participants who rejected system-justifying beliefs evinced
ideology-based double standards: In line with Hypothesis 2b, they
considered the same protest actions to be more acceptable if the
protesters challenged the system or, to a lesser extent, if the
protesters were from the disadvantaged group.

Model 3 (M3) extended Model 1 by estimating participants' responses as a
function of their self-reported support for the cause of the protest. We
recoded participants' responses to create a predictor variable that
encoded support for defunding the police when protesters supported
defunding the police and opposition to defunding the police when
protesters opposed defunding the police. As preregistered, we modeled
this relationship as a monotonic effect (Bürkner \& Charpentier, 2020)
that estimated the average change in the outcome variable across
predictor categories as well as how much of this change occurred between
each of the four pairs of adjacent predictor categories.\footnote{Our
  model assigned a Dirichlet prior, \(\alpha = {1, 1, 1, 1}\), to the
  proportions of the overall change that was expected to occur between
  each of the four pairs of predictor categories.} Contradicting
Hypothesis 3, participants who supported the protesters' cause did not,
on average, consider the same protest actions to be more acceptable than
participants who opposed the protesters' cause
(\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our results did not support
the alternative hypothesis that ideology-based double standards can be
reduced to support for, or opposition to, the cause of a protest.

\hypertarget{non-preregistered-analyses-1}{%
\subsubsection{Non-Preregistered
Analyses}\label{non-preregistered-analyses-1}}

We reran Model 3 with political orientation, instead of system
justification, as an alternative predictor variable. Our results
mirrored results from the preregistered analyses: More liberal
participants were more likely to consider protest actions for defunding
the police to be acceptable (\(\beta_{xy} = 0.41, [0.17, 0.62]\)) but
more conservative participants were not more likely to consider protest
actions against defunding the police to be acceptable
(\(\beta_{xy} = -0.16, [-0.41, 0.04]\)). As Figure 5 shows, conservative
participants tended to consider all protest actions to be equally
unacceptable while liberal participants considered the same protest
actions to be more acceptable if protesters rallied around a progressive
cause. Our non-preregistered analyses thus found evidence for
ideology-based double standards in judging collective action with a
different operationalization of political ideology.

\hypertarget{discussion-1}{%
\subsection{Discussion}\label{discussion-1}}

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

\begin{itemize}
\tightlist
\item
  Summary: No evidence for identity-based double standards, but some
  evidence for ideology-based double standards
\item
  Two plausible interpretations: (1) conservatives oppose controversial
  protest (no matter the end) \emph{or} (2) the end justifies the mean
\item
  Re (1) it's plausible but what about recent events? (see public
  polling from grant)
\item
  Re (2) why is there no asymmetry?
\item
  Outline the-end-justifies-the-mean idea, highlighting that just
  supporting a cause is not enough, it needs moralization
\item
  Discuss why high-status group are more accepting of controversial
  protest actions
\item
  Discuss limitations (two cultures, two issues, liberal samples;
  system-defending vs.~system-supporting actions; what about other
  conservative causes?)
\item
  Invite researchers to use our paradigm
\end{itemize}

\refsection

\begingroup

\noindent \setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}
\small

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abrams_double_2013}{}}%
Abrams, D., Randsley de Moura, G., \& Travaglino, G. A. (2013). A double
standard when group members behave badly: {Transgression} credit to
ingroup leaders. \emph{Journal of Personality and Social Psychology},
\emph{105}(5), 799--815. \url{https://doi.org/10.1037/a0033600}

\leavevmode\vadjust pre{\hypertarget{ref-becker_virtual_2012}{}}%
Becker, J. C. (2012). Virtual special issue on theory and research on
collective action in the {European} {Journal} of {Social} {Psychology}
{[}{Editorial}{]}. \emph{European Journal of Social Psychology},
\emph{42}(1), 19--23. \url{https://doi.org/10.1002/ejsp.1839}

\leavevmode\vadjust pre{\hypertarget{ref-becker_committed_2011}{}}%
Becker, J. C., Tausch, N., Spears, R., \& Christ, O. (2011). Committed
dis(s)idents: {Participation} in radical collective action fosters
disidentification with the broader in-group but enhances political
identification. \emph{Personality and Social Psychology Bulletin},
\emph{37}(8), 1104--1116. \url{https://doi.org/10.1177/0146167211407076}

\leavevmode\vadjust pre{\hypertarget{ref-brewer_social_2007}{}}%
Brewer, M. B. (2007). The social psychology of intergroup relations:
Social categorization, ingroup bias, and outgroup prejudice. In A. W.
Kruglanski \& E. T. Higgins (Eds.), \emph{Social psychology: Handbook of
basic principles} (2nd ed., pp. 695--715). Guilford Press.

\leavevmode\vadjust pre{\hypertarget{ref-burkner_bayesian_2021}{}}%
Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with brms
and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5).
\url{https://doi.org/10.18637/jss.v100.i05}

\leavevmode\vadjust pre{\hypertarget{ref-burkner_modelling_2020}{}}%
Bürkner, P.-C., \& Charpentier, E. (2020). Modelling monotonic effects
of ordinal predictors in {Bayesian} regression models. \emph{British
Journal of Mathematical and Statistical Psychology}, \emph{73}(3),
420--451. \url{https://doi.org/10.1111/bmsp.12195}

\leavevmode\vadjust pre{\hypertarget{ref-demars_irt_2010}{}}%
DeMars, C. (2010). \emph{Item response theory}. Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-dunivin_black_2022}{}}%
Dunivin, Z. O., Yan, H. Y., Ince, J., \& Rojas, F. (2022). Black {Lives}
{Matter} protests shift public discourse. \emph{Proceedings of the
National Academy of Sciences}, \emph{119}(10), e2117320119.
\url{https://doi.org/10.1073/pnas.2117320119}

\leavevmode\vadjust pre{\hypertarget{ref-endevelt_everyone_2021}{}}%
Endevelt, K., Schori-Eyal, N., \& Halperin, E. (2021). Everyone should
get the same, but we should get more: Group entitlement and intergroup
moral double standard. \emph{Group Processes \& Intergroup Relations},
\emph{24}(3), 350--370. \url{https://doi.org/10.1177/1368430219896618}

\leavevmode\vadjust pre{\hypertarget{ref-feinberg_activists_2020}{}}%
Feinberg, M., Willer, R., \& Kovacheff, C. (2020). The activist's
dilemma: {Extreme} protest actions reduce popular support for social
movements. \emph{Journal of Personality and Social Psychology},
\emph{119}(5), 1086--1111. \url{https://doi.org/10.1037/pspi0000230}

\leavevmode\vadjust pre{\hypertarget{ref-foschi_double_2000}{}}%
Foschi, M. (2000). Double standards for competence: Theory and research.
\emph{Annual Review of Sociology}, \emph{26}(1), 21--42.
\url{https://doi.org/10.1146/annurev.soc.26.1.21}

\leavevmode\vadjust pre{\hypertarget{ref-cmdstanr_2021}{}}%
Gabry, J., \& Cesnovar, R. (2021). \emph{{cmdstanr}: {R} interface to
{CmdStan}} (Version 0.4.0). \url{https://mc-stan.org/cmdstanr}

\leavevmode\vadjust pre{\hypertarget{ref-gelman_why_2012}{}}%
Gelman, A., Hill, J., \& Yajima, M. (2012). Why we (usually) don't have
to worry about multiple comparisons. \emph{Journal of Research on
Educational Effectiveness}, \emph{5}(2), 189--211.
\url{https://doi.org/10.1080/19345747.2011.618213}

\leavevmode\vadjust pre{\hypertarget{ref-gelman_prior_2017}{}}%
Gelman, A., Simpson, D., \& Betancourt, M. (2017). The prior can often
only be understood in the context of the likelihood. \emph{Entropy},
\emph{19, 10}, 555. \url{https://doi.org/10.3390/e19100555}

\leavevmode\vadjust pre{\hypertarget{ref-hewstone_ultimate_1990}{}}%
Hewstone, M. (1990). The {``ultimate attribution error''}? {A} review of
the literature on intergroup causal attribution. \emph{European Journal
of Social Psychology}, \emph{20}(4), 311--335.
\url{https://doi.org/10.1002/ejsp.2420200404}

\leavevmode\vadjust pre{\hypertarget{ref-ho_nature_2015}{}}%
Ho, A. K., Sidanius, J., Kteily, N., Sheehy-Skeffington, J., Pratto, F.,
Henkel, K. E., Foels, R., \& Stewart, A. L. (2015). The nature of social
dominance orientation: {Theorizing} and measuring preferences for
intergroup inequality using the new {SDO}-7 scale. \emph{Journal of
Personality and Social Psychology}, \emph{109}(6), 1003--1028.
\url{https://doi.org/10.1037/pspi0000033}

\leavevmode\vadjust pre{\hypertarget{ref-jost_theory_2020}{}}%
Jost, J. T. (2020). \emph{A theory of system justification}. Harvard
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-jost_role_1994}{}}%
Jost, J. T., \& Banaji, M. R. (1994). The role of stereotyping in
system-justification and the production of false consciousness.
\emph{British Journal of Social Psychology}, \emph{33}(1), 1--27.
\url{https://doi.org/10.1111/j.2044-8309.1994.tb01008.x}

\leavevmode\vadjust pre{\hypertarget{ref-jost_missing_2017}{}}%
Jost, J. T., Becker, J., Osborne, D., \& Badaan, V. (2017). Missing in
(collective) action: Ideology, system justification, and the
motivational antecedents of two types of protest behavior. \emph{Current
Directions in Psychological Science}, \emph{26}(2), 99--108.
\url{https://doi.org/10.1177/0963721417690633}

\leavevmode\vadjust pre{\hypertarget{ref-jost_political_2003}{}}%
Jost, J. T., Glaser, J., Kruglanski, A. W., \& Sulloway, F. J. (2003).
Political conservatism as motivated social cognition.
\emph{Psychological Bulletin}, \emph{129}(3), 339--375.
\url{https://doi.org/10.1037/0033-2909.129.3.339}

\leavevmode\vadjust pre{\hypertarget{ref-kay_complementary_2003}{}}%
Kay, A. C., \& Jost, J. T. (2003). Complementary justice: Effects of
"poor but happy" and "poor but honest" stereotype exemplars on system
justification and implicit activation of the justice motive.
\emph{Journal of Personality and Social Psychology}, \emph{85}(5),
823--837. \url{https://doi.org/10.1037/0022-3514.85.5.823}

\leavevmode\vadjust pre{\hypertarget{ref-lacina_nfl_2020}{}}%
Lacina, B. (2020). \emph{National anthem protests and whites' views of
black NFL players} {[}Unpublished manuscript{]}. Department of Political
Science, University of Rochester.
\url{http://www.bethanylacina.com/docs/Lacina_NFL_paper.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-mendoza_for_2014}{}}%
Mendoza, S. A., Lane, S. P., \& Amodio, D. M. (2014). For members only:
Ingroup punishment of fairness norm violations in the ultimatum game.
\emph{Social Psychological and Personality Science}, \emph{5}(6),
662--670. \url{https://doi.org/10.1177/1948550614527115}

\leavevmode\vadjust pre{\hypertarget{ref-osborne_protesting_2019}{}}%
Osborne, D., Jost, J. T., Becker, J. C., Badaan, V., \& Sibley, C. G.
(2019). Protesting to challenge or defend the system? {A} system
justification perspective on collective action. \emph{European Journal
of Social Psychology}, \emph{49}(2), 244--269.
\url{https://doi.org/10.1002/ejsp.2522}

\leavevmode\vadjust pre{\hypertarget{ref-pinto_membership_2010}{}}%
Pinto, I. R., Marques, J. M., Levine, J. M., \& Abrams, D. (2010).
Membership status and subjective group dynamics: Who triggers the black
sheep effect? \emph{Journal of Personality and Social Psychology},
\emph{99}(1), 107--119. \url{https://doi.org/10.1037/a0018187}

\leavevmode\vadjust pre{\hypertarget{ref-reimer_identity_2022}{}}%
Reimer, N. K., Schmid, K., Hewstone, M., \& Al Ramiah, A. (2022).
Self-categorization and social identification: Making sense of us and
them. In D. Chadee (Ed.), \emph{Theories in social psychology} (2nd ed.,
pp. 273--295). Wiley-Blackwell.

\leavevmode\vadjust pre{\hypertarget{ref-samejima_graded_1997}{}}%
Samejima, F. (1997). Graded response model. In W. J. van der Linden \&
R. K. Hambleton (Eds.), \emph{Handbook of modern item response theory}
(pp. 85--100). Springer.

\leavevmode\vadjust pre{\hypertarget{ref-sharp_nonviolent_1973}{}}%
Sharp, G. (1973). \emph{The politics of nonviolent action}. Porter
Sargent.

\leavevmode\vadjust pre{\hypertarget{ref-shuman_disrupting_2021}{}}%
Shuman, E., Saguy, T., van Zomeren, M., \& Halperin, E. (2021).
Disrupting the system constructively: {Testing} the effectiveness of
nonnormative nonviolent collective action. \emph{Journal of Personality
and Social Psychology}, \emph{121}(4), 819--841.
\url{https://doi.org/10.1037/pspi0000333}

\leavevmode\vadjust pre{\hypertarget{ref-cmdstan_2021}{}}%
Stan Development Team. (2021). \emph{Stan modeling language users guide
and reference manual} (Version 2.27.0). \url{http://mc-stan.org/}

\leavevmode\vadjust pre{\hypertarget{ref-tajfel_integrative_1979}{}}%
Tajfel, H., \& Turner, J. C. (1979). An integrative theory of intergroup
conflict. In W. G. Austin \& S. Worchel (Eds.), \emph{The psychology of
intergroup relations} (pp. 33--48). Brooks/Cole.

\leavevmode\vadjust pre{\hypertarget{ref-teixeira_is_2020}{}}%
Teixeira, C. P., Spears, R., \& Yzerbyt, V. Y. (2020). Is {Martin}
{Luther} {King} or {Malcolm} {X} the more acceptable face of protest?
{High}-status groups' reactions to low-status groups' collective action.
\emph{Journal of Personality and Social Psychology}, \emph{118}(5),
919--944. \url{https://doi.org/10.1037/pspi0000195}

\leavevmode\vadjust pre{\hypertarget{ref-valdesolo_moral_2007}{}}%
Valdesolo, P., \& DeSteno, D. (2007). Moral hypocrisy: Social groups and
the flexibility of virtue. \emph{Psychological Science}, \emph{18}(8),
689--690. \url{https://doi.org/10.1111/j.1467-9280.2007.01961.x}

\leavevmode\vadjust pre{\hypertarget{ref-van_zomeren_building_2016}{}}%
van Zomeren, M. (2016). Building a tower of babel? {Integrating} core
motivations and features of social structure into the political
psychology of political action. \emph{Political Psychology}, \emph{37},
87--114. \url{https://doi.org/10.1111/pops.12322}

\leavevmode\vadjust pre{\hypertarget{ref-wright_responding_1990}{}}%
Wright, S. C., Taylor, D. M., \& Moghaddam, F. M. (1990). Responding to
membership in a disadvantaged group: From acceptance to collective
protest. \emph{Journal of Personality and Social Psychology},
\emph{58}(6), 994--1003.
\url{https://doi.org/10.1037/0022-3514.58.6.994}

\leavevmode\vadjust pre{\hypertarget{ref-yougov_nfl_2017}{}}%
YouGov. (2017). \emph{{HuffPost: NFL} {(Sept. 25--26, 2017)}} {[}Data
set{]}. \url{https://perma.cc/3Q3V-4GC4}

\end{CSLReferences}

\endgroup

\newpage

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\begin{table*}

\section{Appendix A}

\caption{List of protest actions used in Experiment 1}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & participate in a public meeting of representatives and elected officials & 97\%\\
2 & hold meetings to inform the public & 96\%\\
3 & make a public speech & 96\%\\
4 & hold meetings to influence the public & 93\%\\
5 & attend or organise a protest march & 93\%\\
6 & attend or organise a protest rally & 92\%\\
7 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 92\%\\
8 & do not buy goods or services from companies who support the bill (consumers' boycott) & 92\%\\
9 & paste up posters with political messages in places where it is allowed and encouraged & 89\%\\
10 & join or form a group of activists who oppose the bill & 87\%\\
11 & refuse to accept honours or awards in protest & 82\%\\
12 & donate to political parties who oppose the bill & 80\%\\
13 & refuse to work (strike) & 80\%\\
14 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 78\%\\
15 & donate to activist groups who oppose the bill & 77\%\\
16 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 54\%\\
17 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 51\%\\
18 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) until the bill is abandoned & 47\%\\
19 & enter and refuse to leave a building (occupation) & 38\%\\
20 & paste up posters with political messages in places where it is not allowed or encouraged & 35\%\\
21 & disrupt traffic (e.g., blocking roads) & 22\%\\
22 & refuse to cooperate with the police and other government agencies & 19\%\\
23 & mock or insult individuals who support the bill & 14\%\\
24 & spray paint political messages in public places & 14\%\\
25 & deface flags or other national symbols & 12\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 1.}

\end{table*}

\setcounter{table}{0}
\renewcommand{\thetable}{B\arabic{table}}

\begin{table*}

\section{Appendix B}

\caption{List of protest actions used in Experiment 2}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & make a public speech & 94\%\\
2 & \textit{hand out flyers, leaflets, or pamphlets} & 93\%\\
3 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
4 & hold meetings to influence the public & 91\%\\
5 & paste up posters with political messages in places where it is allowed and encouraged & 91\%\\
6 & attend or organise a protest march & 91\%\\
7 & donate to activist groups & 89\%\\
8 & join or form a group of activists & 89\%\\
9 & \textit{wear or display political symbols (e.g., patches, flags, bumper stickers)} & 86\%\\
10 & refuse to buy goods or services from companies that advocate [against/for] defunding the police (boycott) & 85\%\\
11 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 83\%\\
12 & donate to politicians who advocate [for/against] defunding the police & 82\%\\
13 & refuse to accept honors or awards in protest & 80\%\\
14 & refuse to work (strike) & 61\%\\
15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 61\%\\
16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 55\%\\
17 & \textit{attend a protest march even though it might turn violent} & 44\%\\
18 & \textit{attend a protest march even though it might be unlawful} & 42\%\\
19 & \textit{attend a protest march while carrying a firearm (where legal)} & 35\%\\
20 & enter and refuse to leave a building (occupation) & 34\%\\
21 & paste up posters with political messages in places where it is not allowed or encouraged & 29\%\\
22 & \textit{refuse to pay fees, fines, or taxes in protest} & 29\%\\
23 & disrupt traffic (e.g., blocking roads) & 25\%\\
24 & spray paint political messages in public places & 23\%\\
25 & mock or insult individuals who are [against/for] defunding the police & 15\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 2. Actions in \textit{italics} replaced actions used in Experiment 1 that were either redundant or did not fit the context of Experiment 2.}

\end{table*}

\end{document}

### 2022-05-24 Documents/supplemental-materials.Rmd
work in progress

---
title: Supplemental Online Materials
subtitle: Double Standards in Judging Collective Action
bibliography: references.bib
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: som-template.tex
csl: apa.csl
---

```{r setup, include = FALSE}

  # Load packages
  library(tidyverse); library(tidybayes); library(numform)

  # knitr options
  knitr::opts_chunk$set(echo = FALSE)

```

# Scale Development

## Study 1

In Study 1, we compiled a list of collective actions from participants' responses and other sources. To that end, we recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students. Participants first read an accessible description of collective action:

> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong. Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club,  has a religion, or belongs to any number of other such groups. Individuals often act in ways to promote the interests of the social groups to which they belong.

> For example, the workers in a factory might want to get paid more. They might go on a strike to reach that goal. Students at a university might want to prevent an increase in tuition fees. They might hand out flyers or occupy a building to reach that goal. Other groups might want to protest against police violence. They might block traffic on a road to make others aware of this issue.

\noindent Participants were then asked to name at least five (and up to ten) actions that fit that description and were encouraged to think of actions that vary in how acceptable or unacceptable they are (in their opinion):

> Please think of other actions that members of a social group (or social groups as a whole) might take to promote their group's interests or goals. Please name at least 5 actions that fit the description above. You can name up to 10 actions if you want to. Please try to think of actions that vary in how acceptable or unacceptable they are (in your opinion).

\noindent After naming at least five actions, participants sorted their responses into actions that, across a range of situations, were "always acceptable", "sometimes acceptable", and "never acceptable":

> On the left, you see the actions you have named on the previous page. Think about each action. Can you think of situations in which this action is an acceptable means to advance a group's goals or interests? Can you think of situations in which this action is not an acceptable means to advance a group's interests?

\noindent Participants named between 5 and 10 actions ($\textit{Mdn} = 7$). Participants named more actions that they considered always acceptable ($\textit{Mdn} = 3$) than actions they considered sometimes ($\textit{Mdn} = 2$) or never ($\textit{Mdn} = 1$) acceptable. We recoded participants' responses into a smaller set of unique collective actions. We then supplemented the resulting list of actions with collective actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973]. This process resulted in 72 actions that we expected to vary in how acceptable most people would find them to be.

## Study 2

In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action. We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students. We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.

Participants first read an accessible description of collective action:

> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong. Each person lives in a place, has a job (or not) at a firm or organisation, has a religion (or not), is part of a political party, or belongs to any number of other groups like these.

> Individuals often act in ways to promote their group's interest. For example, individuals might want people like them to get paid more. Members of a group might take various kinds of actions (for example, go on a strike).

\noindent Participants were then asked to "suppose that one or more members of a group took the following action to advance a cause in their group's interest" and were presented with one of the actions from Study 1. Participants answered several questions about the protest action. First, they rated how disruptive, violent, and extreme they considered this action to be (1 = *not at all*, 4 = *very*). Second, they were asked to think of a range of different causes and circumstances, and to rate how often this action would be an acceptable means for a group to advance one of these causes (1 = *never*, 2 = *rarely*, 3 = *sometimes*, 4 = *often*, 5 = *always*). Finally, they rated how positive or negative they felt, in general, about this action (1 = *very positive*, 5 = *very negative*). Each participant answered these questions for 20 of the 72 actions from Study 1 so that each action was rated by 29--53 participants.

We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997], an item response theory model for ordinal response variables, for participants' ratings of how often an action would be an acceptable form of collective action. For each participant $j$, the model estimated their unique propensity ($\theta_j$) to consider collective actions acceptable in more situations. For each item $i$, the model estimated four acceptability thresholds ($\beta_{ik}$), separating the five response options, and one discrimination parameter ($\alpha_i$) indicating how well the item differentiates between participants with different propensities to consider collective actions acceptable. We estimated the model in *CmdStanR* using similar prior distribution as in Experiment. We used an induced Dirichlet prior [@betancourt_ordinal_2019] for the item-specific difficulty thresholds.

Table S1 shows each actions' item information and discrimination as well as proportions for the observed responses. Table S2 shows participants' averaged ratings of how acceptable, disruptive, violent, extreme, and negative they considered each action to be. Table S3 shows that all five ratings were strongy correlated across actions and participants. We considered both estimates from the graded response model and their relevance to the social contexts to select actions for Experiments 1 and 2.

# Experiment 1

## Method

### Procedure

We used a two-step process to recruit participants who satisfied our preregistered inclusion criteria.

First, we recruited a larger pool of 875 participants. Of these, 475 participants did not have a university degree and placed themselves on the bottom three ranks of the subjective socio-economic status ladder (*lower-status group*) and 400 participants had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socio-economic status ladder (*higher-status group*). Participants completed a screening survey in which they, among other questions, answered whether they considered either their current job or the jobs they had had in the past or would have in the future to be a working-class job or a middle-class/professional job (1 = *working-class job*, 2 = *middle-class/professional job*, 3 = *neither*). Before answering this question, participants read the following explanation:

> The workforce is often divided into two kinds of jobs. What we call working-class jobs are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These jobs do not usually require a university degree. What we call middle-class or professional jobs are administrative, managerial, or other jobs that usually require a university degree.

\noindent As preregistered, we excluded participants from the lower-status group who had responded "middle-class/professional job" or "neither" to this question; we excluded participants from the higher-status group who had responded "working-class job" or "neither" to this question. This left 687 participants for the next step of the selection procedure.

Second, we set out to recruit 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group. After giving their informed consent, participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (*lower-status actors*) or people in professional jobs (*higher-status actors*). Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real. Participants in the lower-status actors condition then read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in working-class jobs. Working-class jobs, in this case, are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These are jobs that do not usually require a university degree. Other jobs are unlikely to be affected.
Participants in the higher-status actors condition instead read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in professional jobs. Professional jobs are administrative, managerial, or other jobs that usually require a university degree. Other jobs are unlikely to be affected.
Participants in the two conditions then read this almost identically worded paragraph:

>This government measure would make it easier for companies to hire workers during economic growth and to lay off workers during an economic crisis. As a consequence, companies would be able to fire employees with little notice and without giving a reason. Trade unions are opposed to the measure. They argue that the bill would compromise job security, and prevent employees from challenging harassment or other abuse without the fear of being fired. People in [working-class/professional] jobs are particularly at risk, and there is a rise in tension and outrage among them.

After reading the vignette, participants responded to the other measures (in the order in which they are presented in the main text). On the final page, we reminded participants that they had read about a bill that the government planned to introduce and that some people objected to. As an attention check, we then asked participants to recall who the people most affected by this measure were. As preregistered, we excluded all participants who did not respond with an accurate description of the group they had read about.

\newpage

# References

\begingroup

\noindent
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs"></div>

\endgroup

\newpage

```{=latex}
\begin{xltabular}{\linewidth}{rXrrrrrrr}

\caption{Results from Study 2}\\
\toprule
   &        &     &          & \multicolumn{5}{c}{Response $[\%]$} \\ \cmidrule{5-9}
\# & Action & $I$ & $\alpha$ & 1 & 2 & 3 & 4 & 5\\
\midrule
\endfirsthead

\toprule
   &        &     &          & \multicolumn{5}{c}{Response $[\%]$} \\ \cmidrule{5-9}
\# & Action & $I$ & $\alpha$ & 1 & 2 & 3 & 4 & 5\\
\midrule
\endhead

\bottomrule
\addlinespace
\caption*{\textit{Note.} $I$ = Information; $\alpha$ = Discrimination; Response: 1 = \textit{never}, 2 = \textit{rarely}, 3 = \textit{sometimes}, 4 = \textit{often}, 5 = \textit{always}}
\endfoot

\bottomrule
\addlinespace
\caption*{\textit{Note.} $I$ = Information; $\alpha$ = Discrimination; Response: 1 = \textit{never}, 2 = \textit{rarely}, 3 = \textit{sometimes}, 4 = \textit{often}, 5 = \textit{always}}
\endlastfoot

1 & disrupt traffic (e.g., blocking roads) & 7.98 & 1.31 & 41 & 36 & 18 & 5 & 0\\
2 & attend or organise a protest rally & 7.35 & 1.11 & 0 & 7 & 45 & 39 & 9\\
3 & refuse to work (strike) & 6.96 & 1.11 & 2 & 5 & 53 & 33 & 7\\
4 & enter and refuse to leave a building (occupation) & 6.95 & 1.12 & 6 & 52 & 26 & 13 & 3\\
5 & deface flags or other national symbols & 6.83 & 1.21 & 52 & 20 & 20 & 9 & 0\\

6 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) & 6.42 & 1.19 & 15 & 18 & 40 & 15 & 12\\
7 & paste up posters with political messages in places where it is not allowed or encouraged & 6.37 & 1.01 & 14 & 32 & 35 & 19 & 0\\
8 & refuse to cooperate with the police and other government agencies & 6.37 & 1.11 & 22 & 38 & 32 & 5 & 3\\
9 & refuse to accept honours or awards in protest & 6.32 & 1.14 & 11 & 14 & 39 & 23 & 14\\
10 & disrupt public events (e.g., a sports game) with a political message & 6.30 & 1.02 & 24 & 45 & 21 & 10 & 0\\

11 & spray paint political messages in public places & 6.28 & 1.06 & 31 & 39 & 28 & 3 & 0\\
12 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 6.28 & 1.04 & 7 & 41 & 34 & 14 & 5\\
13 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 6.23 & 0.99 & 0 & 16 & 27 & 39 & 18\\
14 & get involved in the media (e.g., newspapers, radio, television) to influence the public & 6.18 & 0.98 & 0 & 11 & 32 & 36 & 20\\
15 & visit people in their homes to convince them about an issue (canvassing, door knocking) & 6.07 & 1.03 & 9 & 30 & 27 & 27 & 6\\

16 & donate to political parties who support the cause & 6.00 & 1.10 & 6 & 6 & 33 & 31 & 25\\
17 & do not buy goods or services from companies who oppose the cause (consumers' boycott) & 5.99 & 1.06 & 2 & 5 & 39 & 29 & 24\\
18 & refuse to interact with or acknowledge individuals who oppose the cause & 5.89 & 0.97 & 21 & 45 & 29 & 5 & 0\\
19 & attend or organise a protest march & 5.88 & 1.06 & 3 & 3 & 34 & 38 & 22\\
20 & paste up posters with political messages in places where it is allowed and encouraged & 5.85 & 1.00 & 0 & 3 & 20 & 43 & 34\\

21 & refuse to pay rent (rent strike) & 5.82 & 0.97 & 23 & 41 & 28 & 8 & 0\\
22 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 5.82 & 1.01 & 5 & 5 & 30 & 45 & 15\\
23 & wear or display political symbols & 5.79 & 1.00 & 3 & 9 & 38 & 32 & 18\\
24 & mock or insult individuals who oppose the cause & 5.73 & 1.11 & 53 & 34 & 3 & 11 & 0\\
25 & write letters to politicians, representatives and elected officials & 5.69 & 1.07 & 0 & 0 & 11 & 37 & 53\\

26 & use social media (e.g., Facebook, Twitter, Instagram) to inform the public & 5.67 & 1.01 & 0 & 3 & 12 & 41 & 44\\
27 & join or form a group of activists & 5.66 & 1.02 & 2 & 12 & 21 & 38 & 26\\
28 & hold meetings to inform the public & 5.65 & 1.01 & 0 & 3 & 11 & 38 & 49\\
29 & donate to activist groups who support the cause & 5.64 & 1.06 & 2 & 9 & 41 & 14 & 34\\
30 & disrupt public services (e.g., shutting down government websites) & 5.61 & 1.01 & 42 & 42 & 17 & 0 & 0\\

31 & refuse to pay fees, fines, and taxes & 5.56 & 0.96 & 9 & 50 & 30 & 7 & 4\\
32 & hold meetings to influence the public & 5.56 & 1.00 & 0 & 3 & 16 & 35 & 46\\
33 & hand out flyers, leaflets, or pamphlets & 5.55 & 1.01 & 0 & 0 & 13 & 45 & 42\\
34 & visit people in their homes to inform them about an issue (canvassing, door knocking) & 5.39 & 0.95 & 9 & 20 & 30 & 32 & 9\\
35 & make a public speech & 5.35 & 1.00 & 2 & 2 & 14 & 46 & 36\\

36 & join or form a political party & 5.28 & 0.96 & 2 & 7 & 19 & 45 & 26\\
37 & refuse to interact with or acknowledge politicians who oppose the cause & 5.27 & 0.95 & 8 & 35 & 38 & 10 & 10\\
38 & disrupt private life of politicians (e.g., protesting outside their home) & 5.24 & 0.92 & 39 & 36 & 23 & 2 & 0\\
39 & get involved in the media (e.g., newspapers, radio, television) to inform the public & 5.23 & 0.93 & 0 & 2 & 11 & 43 & 43\\
40 & pay for adverts in the media (e.g., newspapers, radio, television) to influence public opinion & 5.23 & 0.97 & 5 & 10 & 22 & 42 & 20\\

41 & damage private property (e.g., cars or houses) & 5.17 & 1.03 & 70 & 25 & 5 & 0 & 0\\
42 & refuse service (e.g., in a restaurant or shop) to politicians who oppose the cause & 5.16 & 0.89 & 29 & 37 & 26 & 8 & 0\\
43 & sign or start a petition & 5.10 & 0.96 & 0 & 0 & 8 & 39 & 53\\
44 & mock or insult politicians who oppose the cause & 5.07 & 0.99 & 38 & 31 & 16 & 11 & 4\\
45 & join or form a trade/labor union & 5.07 & 0.99 & 2 & 2 & 14 & 37 & 44\\

46 & donate to charities who support the cause & 5.03 & 0.96 & 0 & 0 & 8 & 39 & 53\\
47 & refuse to eat (hunger strike) & 4.95 & 0.97 & 38 & 24 & 29 & 6 & 3\\
48 & damage public property (e.g., goverment buildings) & 4.83 & 0.98 & 72 & 20 & 5 & 2 & 0\\
49 & spread rumours about politicians who oppose the cause & 4.77 & 1.00 & 69 & 21 & 10 & 0 & 0\\
50 & participate in a public meeting of representatives and elected officials & 4.70 & 1.00 & 3 & 0 & 7 & 47 & 43\\

51 & physically harm oneself (e.g., setting oneself on fire) & 4.70 & 1.03 & 80 & 20 & 0 & 0 & 0\\
52 & boycott an election by not voting or spoiling one's ballot & 4.70 & 0.91 & 14 & 38 & 26 & 12 & 10\\
53 & stand in an election & 4.69 & 0.90 & 2 & 8 & 18 & 35 & 38\\
54 & soil politicians who oppose the cause (e.g., throwing eggs at them) & 4.62 & 1.02 & 70 & 21 & 8 & 0 & 2\\
55 & attack politicians with the intention of harming them (e.g., punching them) & 4.62 & 1.01 & 78 & 22 & 0 & 0 & 0\\

56 & refuse service (e.g., in a restaurant or shop) to individuals who oppose the cause & 4.60 & 0.91 & 32 & 54 & 11 & 0 & 3\\
57 & blackmail individuals who oppose the cause & 4.53 & 0.96 & 75 & 22 & 3 & 0 & 0\\
58 & threaten politicians who oppose the cause with physical harm & 4.37 & 1.00 & 86 & 11 & 3 & 0 & 0\\
59 & attack individuals who oppose the cause with the intention of harming them (e.g., punching them) & 4.36 & 0.95 & 82 & 12 & 6 & 0 & 0\\
60 & damage commerical property (e.g., shop windows) & 4.36 & 1.03 & 88 & 12 & 0 & 0 & 0\\

61 & attack politicians with the intention of killing them (e.g., stabbing them) & 4.30 & 1.01 & 91 & 7 & 2 & 0 & 0\\
62 & bribe politicians, representatives, and other elected officials & 4.24 & 0.95 & 85 & 13 & 2 & 0 & 0\\
63 & voting for candidates/parties & 4.23 & 0.85 & 0 & 5 & 7 & 21 & 67\\
64 & donate to political parties to make them change their position on an issue & 4.21 & 0.90 & 35 & 24 & 30 & 3 & 8\\
65 & blackmail politicians who oppose the cause & 4.20 & 0.88 & 72 & 26 & 3 & 0 & 0\\

66 & attack police officers and other government agents with the intention of killing them (e.g., stabbing them) & 4.20 & 1.02 & 93 & 7 & 0 & 0 & 0\\
67 & attack police officers and other government agents with the intention of harming them (e.g., punching them) & 4.17 & 1.00 & 92 & 5 & 3 & 0 & 0\\
68 & attack individuals who oppose the cause with the intention of killing them (e.g., stabbing them) & 3.98 & 1.00 & 96 & 2 & 2 & 0 & 0\\
69 & threaten individuals who oppose the cause with physical harm & 3.96 & 0.99 & 94 & 6 & 0 & 0 & 0\\
70 & attack members of the public (e.g., by setting off a bomb in a public place) & 3.72 & 0.99 & 98 & 0 & 2 & 0 & 0\\

71 & spread misinformation to influence public opinion & 3.71 & 0.82 & 78 & 19 & 3 & 0 & 0\\
72 & threaten to attack members of the public (e.g., by making a bomb threat) & 3.51 & 0.97 & 98 & 0 & 0 & 2 & 0

\end{xltabular}
```

\newpage

```{=latex}
\setcounter{table}{1}

\begin{xltabular}{\linewidth}{rXrrrrr}

\caption{Ratings from Study 2}\\
\toprule
   &        & \multicolumn{5}{c}{Rating $[M]$} \\ \cmidrule{3-7}
\# & Action & Acce. & Disr. & Viol. & Extr. & Nega.\\
\midrule
\endfirsthead

\toprule
   &        & \multicolumn{5}{c}{Rating $[M]$} \\ \cmidrule{3-7}
\# & Action & Acce. & Disr. & Viol. & Extr. & Nega.\\
\midrule
\endhead

\bottomrule
\addlinespace
\caption*{\textit{Note.} Acce. = Acceptable (1 = \textit{never}, 5 = \textit{always}); Disr. = Disruptive, Viol. = Violent, Extr. = Extreme (1 = \textit{not at all}, 4 = \textit{very}); Nega. = Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\endfoot

\bottomrule
\addlinespace
\caption*{\textit{Note.} Acce. = Acceptable (1 = \textit{never}, 5 = \textit{always}); Disr. = Disruptive, Viol. = Violent, Extr. = Extreme (1 = \textit{not at all}, 4 = \textit{very}); Nega. = Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\endlastfoot

1 & disrupt traffic (e.g., blocking roads) & 1.86 & 3.82 & 2.05 & 2.98 & 4.16\\
2 & attend or organise a protest rally & 3.50 & 2.50 & 1.36 & 1.48 & 2.64\\
3 & refuse to work (strike) & 3.37 & 3.35 & 1.09 & 2.07 & 2.60\\
4 & enter and refuse to leave a building (occupation) & 2.55 & 3.32 & 1.48 & 2.42 & 3.19\\
5 & deface flags or other national symbols & 1.85 & 2.98 & 2.30 & 2.98 & 4.15\\

6 & refuse to honour national symbols and traditions (e.g., refusing to sing the national anthem) & 2.92 & 1.85 & 1.23 & 1.62 & 3.20\\
7 & paste up posters with political messages in places where it is not allowed or encouraged & 2.59 & 2.84 & 1.22 & 1.92 & 3.38\\
8 & refuse to cooperate with the police and other government agencies & 2.30 & 3.14 & 2.11 & 2.78 & 3.76\\
9 & refuse to accept honours or awards in protest & 3.14 & 1.84 & 1.07 & 1.59 & 2.86\\
10 & disrupt public events (e.g., a sports game) with a political message & 2.17 & 3.40 & 1.60 & 2.31 & 3.79\\

11 & spray paint political messages in public places & 2.03 & 2.72 & 1.53 & 2.44 & 3.97\\
12 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 2.68 & 3.34 & 1.27 & 2.16 & 3.02\\
13 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 3.59 & 1.59 & 1.02 & 1.30 & 2.75\\
14 & get involved in the media (e.g., newspapers, radio, television) to influence the public & 3.66 & 1.61 & 1.02 & 1.23 & 2.41\\
15 & visit people in their homes to convince them about an issue (canvassing, door knocking) & 2.91 & 2.76 & 1.06 & 1.36 & 3.33\\

16 & donate to political parties who support the cause & 3.64 & 1.14 & 1.00 & 1.17 & 2.47\\
17 & do not buy goods or services from companies who oppose the cause (consumers' boycott) & 3.68 & 2.07 & 1.05 & 1.29 & 2.27\\
18 & refuse to interact with or acknowledge individuals who oppose the cause & 2.18 & 2.42 & 1.24 & 2.37 & 3.79\\
19 & attend or organise a protest march & 3.72 & 2.56 & 1.31 & 1.56 & 2.41\\
20 & paste up posters with political messages in places where it is allowed and encouraged & 4.09 & 1.31 & 1.00 & 1.06 & 2.17\\

21 & refuse to pay rent (rent strike) & 2.21 & 2.72 & 1.10 & 2.41 & 3.67\\
22 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 3.60 & 1.40 & 1.00 & 1.15 & 2.42\\
23 & wear or display political symbols & 3.53 & 1.29 & 1.03 & 1.09 & 2.71\\
24 & mock or insult individuals who oppose the cause & 1.71 & 2.79 & 1.95 & 2.58 & 4.18\\
25 & write letters to politicians, representatives and elected officials & 4.42 & 1.39 & 1.05 & 1.03 & 1.53\\

26 & use social media (e.g., Facebook, Twitter, Instagram) to inform the public & 4.25 & 1.31 & 1.00 & 1.06 & 2.03\\
27 & join or form a group of activists & 3.74 & 1.81 & 1.38 & 1.50 & 2.38\\
28 & hold meetings to inform the public & 4.32 & 1.19 & 1.00 & 1.03 & 1.81\\
29 & donate to activist groups who support the cause & 3.68 & 1.39 & 1.11 & 1.25 & 2.39\\
30 & disrupt public services (e.g., shutting down government websites) & 1.75 & 3.75 & 1.61 & 3.06 & 4.28\\

31 & refuse to pay fees, fines, and taxes & 2.48 & 2.61 & 1.04 & 2.09 & 3.46\\
32 & hold meetings to influence the public & 4.24 & 1.14 & 1.00 & 1.03 & 1.86\\
33 & hand out flyers, leaflets, or pamphlets & 4.29 & 1.47 & 1.03 & 1.05 & 2.00\\
34 & visit people in their homes to inform them about an issue (canvassing, door knocking) & 3.11 & 2.45 & 1.07 & 1.50 & 3.16\\
35 & make a public speech & 4.12 & 1.46 & 1.00 & 1.06 & 1.92\\

36 & join or form a political party & 3.86 & 1.40 & 1.00 & 1.24 & 2.10\\
37 & refuse to interact with or acknowledge politicians who oppose the cause & 2.80 & 1.88 & 1.15 & 1.88 & 3.52\\
38 & disrupt private life of politicians (e.g., protesting outside their home) & 1.89 & 3.36 & 2.07 & 2.77 & 4.07\\
39 & get involved in the media (e.g., newspapers, radio, television) to inform the public & 4.27 & 1.36 & 1.02 & 1.14 & 1.82\\
40 & pay for adverts in the media (e.g., newspapers, radio, television) to influence public opinion & 3.62 & 1.60 & 1.10 & 1.35 & 2.50\\

41 & damage private property (e.g., cars or houses) & 1.34 & 3.77 & 3.64 & 3.86 & 4.75\\
42 & refuse service (e.g., in a restaurant or shop) to politicians who oppose the cause & 2.13 & 2.95 & 1.26 & 2.39 & 3.71\\
43 & sign or start a petition & 4.45 & 1.18 & 1.03 & 1.03 & 1.74\\
44 & mock or insult politicians who oppose the cause & 2.13 & 2.31 & 1.58 & 2.02 & 3.84\\
45 & join or form a trade/labor union & 4.19 & 1.40 & 1.00 & 1.07 & 1.70\\

46 & donate to charities who support the cause & 4.44 & 1.03 & 1.00 & 1.03 & 1.47\\
47 & refuse to eat (hunger strike) & 2.12 & 2.26 & 1.38 & 3.32 & 3.74\\
48 & damage public property (e.g., goverment buildings) & 1.38 & 3.67 & 3.55 & 3.65 & 4.68\\
49 & spread rumours about politicians who oppose the cause & 1.41 & 3.14 & 1.72 & 2.76 & 4.31\\
50 & participate in a public meeting of representatives and elected officials & 4.27 & 1.23 & 1.07 & 1.07 & 1.80\\

51 & physically harm oneself (e.g., setting oneself on fire) & 1.20 & 3.68 & 3.76 & 3.95 & 4.83\\
52 & boycott an election by not voting or spoiling one's ballot & 2.64 & 2.24 & 1.19 & 1.90 & 3.45\\
53 & stand in an election & 3.98 & 1.27 & 1.02 & 1.15 & 1.98\\
54 & soil politicians who oppose the cause (e.g., throwing eggs at them) & 1.43 & 3.28 & 3.11 & 3.19 & 4.47\\
55 & attack politicians with the intention of harming them (e.g., punching them) & 1.22 & 3.65 & 3.81 & 3.84 & 4.81\\

56 & refuse service (e.g., in a restaurant or shop) to individuals who oppose the cause & 1.86 & 3.03 & 1.46 & 2.65 & 4.16\\
57 & blackmail individuals who oppose the cause & 1.28 & 3.33 & 2.56 & 3.61 & 4.72\\
58 & threaten politicians who oppose the cause with physical harm & 1.16 & 3.57 & 3.76 & 3.81 & 4.81\\
59 & attack individuals who oppose the cause with the intention of harming them (e.g., punching them) & 1.24 & 3.76 & 3.88 & 3.88 & 4.79\\
60 & damage commerical property (e.g., shop windows) & 1.12 & 3.85 & 3.79 & 3.79 & 4.91\\

61 & attack politicians with the intention of killing them (e.g., stabbing them) & 1.11 & 3.93 & 4.00 & 4.00 & 4.91\\
62 & bribe politicians, representatives, and other elected officials & 1.17 & 3.02 & 1.28 & 3.04 & 4.85\\
63 & voting for candidates/parties & 4.51 & 1.19 & 1.09 & 1.07 & 1.63\\
64 & donate to political parties to make them change their position on an issue & 2.24 & 2.03 & 1.05 & 1.68 & 3.68\\
65 & blackmail politicians who oppose the cause & 1.31 & 3.15 & 2.15 & 3.54 & 4.64\\

66 & attack police officers and other government agents with the intention of killing them (e.g., stabbing them) & 1.07 & 4.00 & 4.00 & 4.00 & 4.98\\
67 & attack police officers and other government agents with the intention of harming them (e.g., punching them) & 1.11 & 3.81 & 3.95 & 3.95 & 4.95\\
68 & attack individuals who oppose the cause with the intention of killing them (e.g., stabbing them) & 1.06 & 3.94 & 4.00 & 4.00 & 4.92\\
69 & threaten individuals who oppose the cause with physical harm & 1.06 & 3.72 & 3.92 & 3.97 & 4.92\\
70 & attack members of the public (e.g., by setting off a bomb in a public place) & 1.05 & 3.90 & 3.92 & 3.92 & 4.92\\

71 & spread misinformation to influence public opinion & 1.24 & 3.41 & 1.41 & 2.95 & 4.76\\
72 & threaten to attack members of the public (e.g., by making a bomb threat) & 1.07 & 3.83 & 3.83 & 3.85 & 4.95

\end{xltabular}
```

\newpage

```{=latex}
\setcounter{table}{2}

\begin{table}[t!]
\caption{Correlations between ratings in Study 2}
\centering
\begin{tabular}{rlrrrrr}
\toprule
\# & Rating & Acceptable & Disruptive & Violent & Extreme & Negative\\
\midrule1 & Acceptable &  & -.69 & -.64 & -.77 & -.89\\
2 & Disruptive & -.69 &  & .66 & .78 & .70\\
3 & Violent & -.64 & .66 &  & .79 & .64\\
4 & Extreme & -.77 & .78 & .79 &  & .77\\
5 & Negative & -.89 & .70 & .64 & .77 & \\
\bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Acceptable (1 = \textit{never}, 5 = \textit{always}); Disruptive, Violent, Extreme (1 = \textit{not at all}, 4 = \textit{very}); Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\end{table}
```

### 2022-05-24 Documents/supplemental-materials.tex
work in progress

--- 2022-05-24 Documents/supplemental-materials.Rmd
@@
----
-title: Supplemental Online Materials
-subtitle: Double Standards in Judging Collective Action
-bibliography: references.bib
-output:
-  pdf_document:
-    latex_engine: pdflatex
-    keep_tex: yes
-    template: som-template.tex
-csl: apa.csl
----
+\documentclass[12pt, letterpaper]{article}
 
-```{r setup, include = FALSE}
+% Font
+\usepackage[utf8]{inputenc}
+\usepackage{MinionPro} 
+\input{glyphtounicode}
+\pdfgentounicode=1 
+\usepackage{microtype}
 
-  # Load packages
-  library(tidyverse); library(tidybayes); library(numform)
+% Format
+\usepackage[letterpaper, margin = 1in]{geometry}
+\setcounter{secnumdepth}{0}
+\usepackage{titlesec}  
+\titleformat*{\section}{\centering\normalfont\Large\bfseries}
 
-  # knitr options
-  knitr::opts_chunk$set(echo = FALSE)
+% Links
+\usepackage[colorlinks = true, linkcolor = black, urlcolor = black, citecolor = black]{hyperref}
 
-```
+% Figures
+\usepackage{graphicx}
+\usepackage[labelfont = bf, font = small, labelsep = newline, singlelinecheck = false]{caption}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
-# Scale Development
+% Tables
+\usepackage{booktabs}
+\usepackage{xltabular}
 
-## Study 1
+% References
+\newenvironment{CSLReferences}[2]{}{}
 
-In Study 1, we compiled a list of collective actions from participants' responses and other sources.
-To that end, we recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US.
-To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students.
+% Frontmatter
+\title{ Supplemental Online Materials:\\\textit{Double Standards in
+Judging Collective Action} }
+\author{  }
+
+\begin{document}
+
+\maketitle
+
+\hypertarget{scale-development}{%
+\section{Scale Development}\label{scale-development}}
+
+\hypertarget{study-1}{%
+\subsection{Study 1}\label{study-1}}
+
+In Study 1, we compiled a list of collective actions from participants'
+responses and other sources. To that end, we recruited 60 participants
+from the Prolific subject pool, all of whom were citizens of the UK or
+the US. To increase the socioeconomic diversity of our sample, we
+recruited 30 non-students without a university degree, 15 non-students
+with a university degree, and 15 current university students.
 Participants first read an accessible description of collective action:
 
-> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong.
-Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club,  has a religion, or belongs to any number of other such groups.
-Individuals often act in ways to promote the interests of the social groups to which they belong.
+\begin{quote}
+Society is not only made up of individuals, but consists of `social
+groups' to which these individuals belong. Each person lives in a place,
+has a job (or not) at an organisation, is a fan of a specific sports
+club, has a religion, or belongs to any number of other such groups.
+Individuals often act in ways to promote the interests of the social
+groups to which they belong.
+\end{quote}
 
-> For example, the workers in a factory might want to get paid more.
-They might go on a strike to reach that goal.
-Students at a university might want to prevent an increase in tuition fees.
-They might hand out flyers or occupy a building to reach that goal.
-Other groups might want to protest against police violence.
-They might block traffic on a road to make others aware of this issue.
+\begin{quote}
+For example, the workers in a factory might want to get paid more. They
+might go on a strike to reach that goal. Students at a university might
+want to prevent an increase in tuition fees. They might hand out flyers
+or occupy a building to reach that goal. Other groups might want to
+protest against police violence. They might block traffic on a road to
+make others aware of this issue.
+\end{quote}
 
-\noindent Participants were then asked to name at least five (and up to ten) actions that fit that description and were encouraged to think of actions that vary in how acceptable or unacceptable they are (in their opinion):
+\noindent Participants were then asked to name at least five (and up to
+ten) actions that fit that description and were encouraged to think of
+actions that vary in how acceptable or unacceptable they are (in their
+opinion):
 
-> Please think of other actions that members of a social group (or social groups as a whole) might take to promote their group's interests or goals.
-Please name at least 5 actions that fit the description above.
-You can name up to 10 actions if you want to.
-Please try to think of actions that vary in how acceptable or unacceptable they are (in your opinion).
+\begin{quote}
+Please think of other actions that members of a social group (or social
+groups as a whole) might take to promote their group's interests or
+goals. Please name at least 5 actions that fit the description above.
+You can name up to 10 actions if you want to. Please try to think of
+actions that vary in how acceptable or unacceptable they are (in your
+opinion).
+\end{quote}
 
-\noindent After naming at least five actions, participants sorted their responses into actions that, across a range of situations, were "always acceptable", "sometimes acceptable", and "never acceptable":
+\noindent After naming at least five actions, participants sorted their
+responses into actions that, across a range of situations, were ``always
+acceptable'', ``sometimes acceptable'', and ``never acceptable'':
 
-> On the left, you see the actions you have named on the previous page.
-Think about each action.
-Can you think of situations in which this action is an acceptable means to advance a group's goals or interests?
-Can you think of situations in which this action is not an acceptable means to advance a group's interests?
+\begin{quote}
+On the left, you see the actions you have named on the previous page.
+Think about each action. Can you think of situations in which this
+action is an acceptable means to advance a group's goals or interests?
+Can you think of situations in which this action is not an acceptable
+means to advance a group's interests?
+\end{quote}
 
-\noindent Participants named between 5 and 10 actions ($\textit{Mdn} = 7$).
-Participants named more actions that they considered always acceptable ($\textit{Mdn} = 3$) than actions they considered sometimes ($\textit{Mdn} = 2$) or never ($\textit{Mdn} = 1$) acceptable.
-We recoded participants' responses into a smaller set of unique collective actions.
-We then supplemented the resulting list of actions with collective actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973].
-This process resulted in 72 actions that we expected to vary in how acceptable most people would find them to be.
+\noindent Participants named between 5 and 10 actions
+(\(\textit{Mdn} = 7\)). Participants named more actions that they
+considered always acceptable (\(\textit{Mdn} = 3\)) than actions they
+considered sometimes (\(\textit{Mdn} = 2\)) or never
+(\(\textit{Mdn} = 1\)) acceptable. We recoded participants' responses
+into a smaller set of unique collective actions. We then supplemented
+the resulting list of actions with collective actions from the
+psychological and political science literature (e.g., Sharp, 1973). This
+process resulted in 72 actions that we expected to vary in how
+acceptable most people would find them to be.
 
-## Study 2
+\hypertarget{study-2}{%
+\subsection{Study 2}\label{study-2}}
 
-In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action.
-We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US.
-To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students.
-We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.
+In Study 2, we measured how acceptable participants judged the actions
+from Study 1 to be and applied item response theory to develop an
+instrument to capture double standards in judging collective action. We
+recruited 158 participants (\(\textit{Mdn} = 30\) years, age range:
+18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the
+Prolific subject pool, all of whom were citizens of the UK or the US. To
+increase the socioeconomic diversity of our sample, we recruited 80
+non-students without a university degree, 37 non-students with a
+university degree, and 41 current university students. We excluded 15
+participants who failed an attention check, leaving a final sample of
+143 participants for our analyses.
 
@@
 
-> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong.
-Each person lives in a place, has a job (or not) at a firm or organisation, has a religion (or not), is part of a political party, or belongs to any number of other groups like these.
+\begin{quote}
+Society is not only made up of individuals, but consists of `social
+groups' to which these individuals belong. Each person lives in a place,
+has a job (or not) at a firm or organisation, has a religion (or not),
+is part of a political party, or belongs to any number of other groups
+like these.
+\end{quote}
 
-> Individuals often act in ways to promote their group's interest.
-For example, individuals might want people like them to get paid more.
-Members of a group might take various kinds of actions (for example, go on a strike).
+\begin{quote}
+Individuals often act in ways to promote their group's interest. For
+example, individuals might want people like them to get paid more.
+Members of a group might take various kinds of actions (for example, go
+on a strike).
+\end{quote}
 
-\noindent Participants were then asked to "suppose that one or more members of a group took the following action to advance a cause in their group's interest" and were presented with one of the actions from Study 1.
-Participants answered several questions about the protest action.
-First, they rated how disruptive, violent, and extreme they considered this action to be (1 = *not at all*, 4 = *very*).
-Second, they were asked to think of a range of different causes and circumstances, and to rate how often this action would be an acceptable means for a group to advance one of these causes (1 = *never*, 2 = *rarely*, 3 = *sometimes*, 4 = *often*, 5 = *always*).
-Finally, they rated how positive or negative they felt, in general, about this action (1 = *very positive*, 5 = *very negative*).
-Each participant answered these questions for 20 of the 72 actions from Study 1 so that each action was rated by 29--53 participants.
+\noindent Participants were then asked to ``suppose that one or more
+members of a group took the following action to advance a cause in their
+group's interest'' and were presented with one of the actions from Study
+1. Participants answered several questions about the protest action.
+First, they rated how disruptive, violent, and extreme they considered
+this action to be (1 = \emph{not at all}, 4 = \emph{very}). Second, they
+were asked to think of a range of different causes and circumstances,
+and to rate how often this action would be an acceptable means for a
+group to advance one of these causes (1 = \emph{never}, 2 =
+\emph{rarely}, 3 = \emph{sometimes}, 4 = \emph{often}, 5 =
+\emph{always}). Finally, they rated how positive or negative they felt,
+in general, about this action (1 = \emph{very positive}, 5 = \emph{very
+negative}). Each participant answered these questions for 20 of the 72
+actions from Study 1 so that each action was rated by 29--53
+participants.
 
-We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997], an item response theory model for ordinal response variables, for participants' ratings of how often an action would be an acceptable form of collective action.
-For each participant $j$, the model estimated their unique propensity ($\theta_j$) to consider collective actions acceptable in more situations.
-For each item $i$, the model estimated four acceptability thresholds ($\beta_{ik}$), separating the five response options, and one discrimination parameter ($\alpha_i$) indicating how well the item differentiates between participants with different propensities to consider collective actions acceptable.
-We estimated the model in *CmdStanR* using similar prior distribution as in Experiment.
-We used an induced Dirichlet prior [@betancourt_ordinal_2019] for the item-specific difficulty thresholds.
+We estimated a graded response model (Bürkner, 2021; Samejima, 1997), an
+item response theory model for ordinal response variables, for
+participants' ratings of how often an action would be an acceptable form
+of collective action. For each participant \(j\), the model estimated
+their unique propensity (\(\theta_j\)) to consider collective actions
+acceptable in more situations. For each item \(i\), the model estimated
+four acceptability thresholds (\(\beta_{ik}\)), separating the five
+response options, and one discrimination parameter (\(\alpha_i\))
+indicating how well the item differentiates between participants with
+different propensities to consider collective actions acceptable. We
+estimated the model in \emph{CmdStanR} using similar prior distribution
+as in Experiment. We used an induced Dirichlet prior (Betancourt, 2019)
+for the item-specific difficulty thresholds.
 
-Table S1 shows each actions' item information and discrimination as well as proportions for the observed responses.
-Table S2 shows participants' averaged ratings of how acceptable, disruptive, violent, extreme, and negative they considered each action to be.
-Table S3 shows that all five ratings were strongy correlated across actions and participants.
-We considered both estimates from the graded response model and their relevance to the social contexts to select actions for Experiments 1 and 2.
+Table S1 shows each actions' item information and discrimination as well
+as proportions for the observed responses. Table S2 shows participants'
+averaged ratings of how acceptable, disruptive, violent, extreme, and
+negative they considered each action to be. Table S3 shows that all five
+ratings were strongy correlated across actions and participants. We
+considered both estimates from the graded response model and their
+relevance to the social contexts to select actions for Experiments 1 and
+2.
 
-# Experiment 1
+\hypertarget{experiment-1}{%
+\section{Experiment 1}\label{experiment-1}}
 
-## Method
+\hypertarget{method}{%
+\subsection{Method}\label{method}}
 
-### Procedure
+\hypertarget{procedure}{%
+\subsubsection{Procedure}\label{procedure}}
 
-We used a two-step process to recruit participants who satisfied our preregistered inclusion criteria.
+We used a two-step process to recruit participants who satisfied our
+preregistered inclusion criteria.
 
-First, we recruited a larger pool of 875 participants.
-Of these, 475 participants did not have a university degree and placed themselves on the bottom three ranks of the subjective socio-economic status ladder (*lower-status group*) and 400 participants had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socio-economic status ladder (*higher-status group*).
-Participants completed a screening survey in which they, among other questions, answered whether they considered either their current job or the jobs they had had in the past or would have in the future to be a working-class job or a middle-class/professional job (1 = *working-class job*, 2 = *middle-class/professional job*, 3 = *neither*).
-Before answering this question, participants read the following explanation:
+First, we recruited a larger pool of 875 participants. Of these, 475
+participants did not have a university degree and placed themselves on
+the bottom three ranks of the subjective socio-economic status ladder
+(lower-status group) and 400 participants had at least an undergraduate
+degree and placed themselves on the top four ranks of the subjective
+socio-economic status ladder (higher-status group). Participants
+completed a screening survey in which they, among other questions,
+answered whether they considered either their current job or the jobs
+they had had in the past or would have in the future to be a
+working-class job or a middle-class/professional job (1 = working-class
+job, 2 = middle-class/professional job, 3 = neither). Before answering
+this question, participants read the following explanation:
 
-> The workforce is often divided into two kinds of jobs.
-What we call working-class jobs are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers.
-These jobs do not usually require a university degree.
-What we call middle-class or professional jobs are administrative, managerial, or other jobs that usually require a university degree.
+\begin{quote}
+The workforce is often divided into two kinds of jobs. What we call
+working-class jobs are jobs done by skilled, semi-skilled, unskilled
+manual workers or by casual workers. These jobs do not usually require a
+university degree. What we call middle-class or professional jobs are
+administrative, managerial, or other jobs that usually require a
+university degree.
+\end{quote}
 
-\noindent As preregistered, we excluded participants from the lower-status group who had responded "middle-class/professional job" or "neither" to this question; we excluded participants from the higher-status group who had responded "working-class job" or "neither" to this question.
+As preregistered, we excluded participants from the lower-status group
+who had responded ``middle-class/professional job'' or ``neither'' to
+this question; we excluded participants from the higher-status group who
+had responded ``working-class job'' or ``neither'' to this question.
 This left 687 participants for the next step of the selection procedure.
 
-Second, we set out to recruit 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group.
-After giving their informed consent, participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (*lower-status actors*) or people in professional jobs (*higher-status actors*).
-Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real.
-Participants in the lower-status actors condition then read the following introduction:
+Second, we set out to recruit 500 participants from the remaining 687
+participants, 250 from the lower-status group and 250 from the
+higher-status group. After giving their informed consent, participants
+were randomly assigned to read a vignette about a government bill
+affecting either people in working-class jobs (lower-status actors) or
+people in professional jobs (higher-status actors). Participants in both
+conditions were instructed to carefully read the vignette and to try to
+imagine what it would be like if this situation was real. Participants
+in the lower-status actors condition then read the following
+introduction:
 
-> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in working-class jobs.
-Working-class jobs, in this case, are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers.
-These are jobs that do not usually require a university degree.
-Other jobs are unlikely to be affected.
-Participants in the higher-status actors condition instead read the following introduction:
+\begin{quote}
+The government, though not necessarily the current government, is going
+to introduce a bill that will mostly affect people in working-class
+jobs. Working-class jobs, in this case, are jobs done by skilled,
+semi-skilled, unskilled manual workers or by casual workers. These are
+jobs that do not usually require a university degree. Other jobs are
+unlikely to be affected. Participants in the higher-status actors
+condition instead read the following introduction:
+\end{quote}
 
-> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in professional jobs.
-Professional jobs are administrative, managerial, or other jobs that usually require a university degree.
-Other jobs are unlikely to be affected.
-Participants in the two conditions then read this almost identically worded paragraph:
+\begin{quote}
+The government, though not necessarily the current government, is going
+to introduce a bill that will mostly affect people in professional jobs.
+Professional jobs are administrative, managerial, or other jobs that
+usually require a university degree. Other jobs are unlikely to be
+affected. Participants in the two conditions then read this almost
+identically worded paragraph:
+\end{quote}
 
->This government measure would make it easier for companies to hire workers during economic growth and to lay off workers during an economic crisis.
-As a consequence, companies would be able to fire employees with little notice and without giving a reason.
-Trade unions are opposed to the measure.
-They argue that the bill would compromise job security, and prevent employees from challenging harassment or other abuse without the fear of being fired.
-People in [working-class/professional] jobs are particularly at risk, and there is a rise in tension and outrage among them.
+\begin{quote}
+This government measure would make it easier for companies to hire
+workers during economic growth and to lay off workers during an economic
+crisis. As a consequence, companies would be able to fire employees with
+little notice and without giving a reason. Trade unions are opposed to
+the measure. They argue that the bill would compromise job security, and
+prevent employees from challenging harassment or other abuse without the
+fear of being fired. People in {[}working-class/professional{]} jobs are
+particularly at risk, and there is a rise in tension and outrage among
+them.
+\end{quote}
 
-After reading the vignette, participants responded to the other measures (in the order in which they are presented in the main text).
-On the final page, we reminded participants that they had read about a bill that the government planned to introduce and that some people objected to.
-As an attention check, we then asked participants to recall who the people most affected by this measure were.
-As preregistered, we excluded all participants who did not respond with an accurate description of the group they had read about.
+After reading the vignette, participants responded to the other measures
+(in the order in which they are presented in the main text). On the
+final page, we reminded participants that they had read about a bill
+that the government planned to introduce and that some people objected
+to. As an attention check, we then asked participants to recall who the
+people most affected by this measure were. As preregistered, we excluded
+all participants who did not respond with an accurate description of the
+group they had read about.
 
@@
 
-# References
+\hypertarget{references}{%
+\section{References}\label{references}}
 
@@
 
-\noindent
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\noindent \setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}
 
-<div id="refs"></div>
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-betancourt_ordinal_2019}{}}%
+Betancourt, M. (2019). \emph{Ordinal regression}.
+\url{https://betanalpha.github.io/assets/case_studies/ordinal_regression.html}
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner_bayesian_2021}{}}%
+Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with brms
+and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5).
+\url{https://doi.org/10.18637/jss.v100.i05}
+
+\leavevmode\vadjust pre{\hypertarget{ref-samejima_graded_1997}{}}%
+Samejima, F. (1997). Graded response model. In W. J. van der Linden \&
+R. K. Hambleton (Eds.), \emph{Handbook of modern item response theory}
+(pp. 85--100). Springer.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sharp_nonviolent_1973}{}}%
+Sharp, G. (1973). \emph{The politics of nonviolent action}. Porter
+Sargent.
+
+\end{CSLReferences}
 
@@
 
-```{=latex}
 \begin{xltabular}{\linewidth}{rXrrrrrrr}
@@
 \end{xltabular}
-```
 
@@
 
-```{=latex}
 \setcounter{table}{1}
@@
 
-```{=latex}
 \setcounter{table}{2}
@@
 \end{table}
-```
+
+\end{document}

### 2022-05-24 Experiment 1/figures/figure-3.R

@@ -1,0 +1,245 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("ggtext")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(ggtext)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import results (preregistered analyses)
+  d_prereg <- read_rds("Experiment 1/results/results_pregistered_analyses.rds")
+  
+  # Import results (exploratory analyses)
+  d_explor <- read_rds("Experiment 1/results/results_exploratory_analyses.rds")
+  
+
+# Figure 3 (top) ----------------------------------------------------------
+
+  # Make Figure 3a (z)
+  f_3a_z <- ggplot(d_prereg, aes(x = z, y = actor)) +
+    stat_pointinterval(
+      aes(shape = ingroup),
+      .width = 0.95,
+      fill = "white",
+      size = 0.455,
+      point_size = 0.455 * 3
+    ) +
+    scale_x_continuous(minor_breaks = NULL) +
+    scale_shape_manual(values = c("Ingroup" = 21, "Outgroup" = 19)) +
+    coord_cartesian(xlim = c(-0.5, 0.5)) +
+    facet_grid(observer ~ .) +
+    theme_grey(base_size = 10) +
+    theme(
+      legend.position = "none",
+      # legend.position = c(0.95, 0.95),
+      # legend.justification = c(1, 1),
+      # legend.title = element_blank(),
+      # legend.spacing = unit(0, "pt"),
+      strip.background = element_blank(),
+      strip.text = element_blank(),
+      strip.placement = "outside",
+      axis.line.x = element_line(size = rel(0.5)),
+      axis.title = element_text(size = rel(0.8)),
+      axis.text = element_text(colour = "black"),
+      axis.text.y = element_text(colour = "black"),
+      plot.margin = margin(0, 5, 0, 5),
+      plot.tag = element_text(face = "bold")
+    ) +
+    labs(
[…]

### 2022-05-24 Experiment 1/figures/table-a1.R

@@ -1,0 +1,43 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("numform")
+  
+  # Load packages
+  library(tidyverse)
+  library(numform)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 1/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  
+  # Summarize 
+  ds <- dl %>% 
+    group_by(action, action_text) %>% 
+    summarize(
+      Pr = mean(response)
+    ) %>% 
+    ungroup() %>% 
+    arrange(desc(Pr)) %>% 
+    transmute(
+      `#` = row_number(),
+      Action = action_text,
+      Pr = f_prop2percent(Pr, digits = 0)
+    )
+    
+
+# Export ------------------------------------------------------------------
+  
+  # Save as .rds
+  write_rds(ds, "Experiment 1/figures/table-a1.rds")

### 2022-05-24 Experiment 1/scripts/prepare_data.R

@@ -1,0 +1,210 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+  #########################################################################
+  # All identifying information (e.g., Prolific IDs) have been removed in #
+  # preprocessing and replaced with newly generated IDs.                  #
+  #########################################################################
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("janitor")
+  # devtools::install_github("ropenscilabs/gendercodeR")
+
+  # Load packages
+  library(tidyverse); library(janitor); library(gendercodeR)
+
+  # Add to gender dictionary (gendercodeR)
+  gender_dictionary <- list(
+    email = "female",
+    `my sex is female` = "female",
+    makle = "male",
+    femalew = "female",
+    `female only` = "female"
+  )
+
+
+# Import ------------------------------------------------------------------
+
+  # Import metadata from Prolific
+  dp <- read_csv("Experiment 1/data/raw/prolific_export.csv")
+
+  # Import data from Qualtrics (Screening Survey)
+  ds <- read_csv("Experiment 1/data/raw/Screening_Survey.csv")[-1:-2,]
+  
+  # Import data from Qualtrics (Experiment 1)
+  dr <- read_csv("Experiment 1/data/raw/Study_1.csv")[-1:-2,]
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Transform metadata (Prolific)
+  dp <- dp %>% 
+    clean_names() %>% 
+    mutate(
+      across(c(age, socioeconomic_status), as.integer),
+      across(where(is.character), ~if_else(. == "CONSENT REVOKED", NA_character_, .))
+    )
+  
+  # Select variables (Prolific)
+  dp <- dp %>% select(id, age, highest_education_level:student_status)
+  
+  # Transform data (Screening Survey)
+  ds <- ds %>% 
+    clean_names() %>% 
+    select(id, sample, q2:q9, q10 = q10_1, q12_1:q13_3) %>%
+    mutate(across(c(q3, q4, q5:q10, q12_1:q13_3), as.integer))
+  
+  # Code gender (Screening Survey)
+  ds <- ds %>% 
+    mutate(
+      gender = recode_gender(
+        q2,
[…]

### 2022-05-24 Experiment 1/scripts/run_exploratory_analyses.R

@@ -1,0 +1,230 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 1/data/dl.rds")
+  dw <- read_rds("Experiment 1/data/dw.rds")
+  ds <- read_rds("Experiment 1/data/ds.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Code predictors (preregistered analyses)
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      x_1 = if_else(target == "PJ", 1L, 0L),
+      x_2 = if_else(target == sample, 1L, 0L),
+      x_3 = if_else(target == "PJ" & sample == "PJ", 1L, 0L),
+      y  = response
+    )
+  
+  # Code predictors (exploratory analyses)
+  dw <- dw %>% 
+    left_join(
+      ds %>% select(id, political_orientation = q10),
+      by = "id"
+    ) %>% 
+    rename_with(~str_replace(., "q11", "sdo")) %>% 
+    rename_with(~str_replace(., "q12", "sjb")) %>% 
+    mutate(
+      across(c(sdo_3, sdo_4, sdo_7, sdo_8, sjb_3, sjb_7), ~8L - .)
+    )
+  
+  # Calculate factor scores
+  source("Experiment 1/scripts/run_factor_analyses.R")
[…]

### 2022-05-24 Experiment 1/scripts/run_factor_analyses.R

@@ -1,0 +1,51 @@
+# Load packages -----------------------------------------------------------
+  library(lavaan)
+
+
+# Social Dominance Orientation (SDO) --------------------------------------
+  
+  # Load model
+  sdo_model <- '
+    sdo =~ sdo_1 + sdo_2 + sdo_3 + sdo_4 + sdo_5 + sdo_6 + sdo_7 + sdo_8
+  '
+
+  # Run model
+  sdo_fit <- cfa(
+    sdo_model, 
+    estimator = "mlr", 
+    missing = "fiml",
+    std.lv = TRUE,
+    data = dw
+  )
+
+  # Export factor scores
+  dw <- predict(sdo_fit) %>%
+    as.data.frame() %>%
+    bind_cols(dw, .)
+
+  
+# System-Justifying Beliefs (SJB) -----------------------------------------
+
+  # Load models
+  sjb_model <- '
+    sjb =~ sjb_1 + sjb_2 + sjb_3 + sjb_4 + sjb_5 + sjb_6 + sjb_7 + sjb_8 
+  '
+  
+  # Run models
+  sjb_fit <- cfa(
+    sjb_model,
+    estimator = "mlr",
+    missing = "fiml",
+    std.lv = TRUE,
+    data = dw
+  )
+  
+  # Export factor scores
+  dw <- predict(sjb_fit) %>%
+    as.data.frame() %>%
+    bind_cols(dw, .)
+
+  
+# Remove ------------------------------------------------------------------
+  detach(package:lavaan)
+  rm("sdo_model", "sdo_fit", "sjb_model", "sjb_fit")

### 2022-05-24 Experiment 1/scripts/run_preregistered_analyses.R

@@ -1,0 +1,173 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+
+  # Load packages
+  library(tidyverse)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 1/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  
+  # Code predictors
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      x_1 = if_else(target == "PJ", 1L, 0L),
+      x_2 = if_else(target == sample, 1L, 0L),
+      x_3 = if_else(target == "PJ" & sample == "PJ", 1L, 0L),
+      y  = response
+    )
+  
+  # Compile data list
+  dlist <- with(dl, list(
+    N = length(ii),
+    I = max(ii),
+    J = max(jj),
+    K = 3L,
+    ii = ii,
+    jj = jj,
+    X = cbind(x_1, x_2, x_3),
+    y = y
+  ))
+
+
+# Estimate ----------------------------------------------------------------
+
+  # Compile model
+  model <- cmdstan_model("Experiment 1/models/s1_2pl_model.stan")
+
+  # Run model
+  fit <- model$sample(
+    data = dlist,
+    seed = 36939726,
+    chains = 8,
+    parallel_chains = parallel::detectCores(),
+    iter_sampling = 1000,
+    iter_warmup = 1000
+  )
+
+  # Save results as .rds
+  fit$save_object("Experiment 1/results/fit_pregistered_analyses.rds")
+
+
[…]

### 2022-05-24 Experiment 2/figures/figure-4.R

@@ -1,0 +1,441 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+  # CHECK FIGURE M1
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = case_when(
+        protest == "for" ~ q5,
+        protest == "against" ~ 6L - q5
+      )
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
[…]

### 2022-05-24 Experiment 2/figures/figure-5.R

@@ -1,0 +1,290 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = (q5 - mean(q5))/sd(q5),
+      x_pol = (q19_1 - mean(q19_1))/sd(q19_1)
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
+        observers == "White" & protesters == "Black" & protest == "for"     ~ 5L, 
[…]

### 2022-05-24 Experiment 2/figures/table-b1.R

@@ -1,0 +1,43 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("numform")
+  
+  # Load packages
+  library(tidyverse)
+  library(numform)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  
+  # Summarize 
+  ds <- dl %>% 
+    group_by(action, action_text) %>% 
+    summarize(
+      Pr = mean(response)
+    ) %>% 
+    ungroup() %>% 
+    arrange(desc(Pr)) %>% 
+    transmute(
+      `#` = row_number(),
+      Action = action_text,
+      Pr = f_prop2percent(Pr, digits = 0)
+    )
+    
+
+# Export ------------------------------------------------------------------
+  
+  # Save as .rds
+  write_rds(ds, "Experiment 2/figures/table-b1.rds")

### 2022-05-24 Experiment 2/scripts/prepare_data.R

@@ -1,0 +1,150 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+  #########################################################################
+  # All identifying information (e.g., Prolific IDs) have been removed in #
+  # preprocessing and replaced with newly generated IDs.                  #
+  #########################################################################
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("janitor")
+  # devtools::install_github("ropenscilabs/gendercodeR")
+
+  # Load packages
+  library(tidyverse); library(janitor); library(gendercoder)
+
+
+# Import ------------------------------------------------------------------
+
+  # Import metadata from Prolific
+  dp <- read_csv("Experiment 2/data/raw/prolific_export.csv")
+
+  # Import data from Qualtrics (Experiment 2)
+  dr <- read_csv("Experiment 2/data/raw/Study_2.csv")[-1:-2,]
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Transform metadata (Prolific)
+  dp <- dp %>% 
+    clean_names() %>% 
+    mutate(
+      across(c(age), as.integer),
+      across(where(is.character), ~if_else(. == "CONSENT REVOKED", NA_character_, .))
+    )
+  
+  # Transform data (Experiment 2)
+  dr <- dr %>% clean_names()
+  
+  # Remove duplicates
+  dr <- dr %>% 
+    mutate(
+      recorded_date = lubridate::as_datetime(recorded_date)
+    ) %>% 
+    group_by(id) %>% 
+    slice_min(recorded_date) %>% 
+    ungroup()
+  
+  # Remove incomplete responses
+  dr <- dr %>% filter(finished == "1")
+  
+  # Code gender (Experiment 2)
+  dr <- dr %>% 
+    mutate(
+      gender = recode_gender(
+        q20_1,
+        dictionary = fewlevels_en,
+        retain_unmatched = TRUE
+      ),
+      gender = case_when(
+        gender == "fermale" ~ "woman",
+        gender == "Genderfluid" ~ "sex and gender diverse",
+        gender == "genderfluid" ~ "sex and gender diverse",
[…]

### 2022-05-24 Experiment 2/scripts/run_exploratory_analyses.R

@@ -1,0 +1,157 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+  
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = (q5 - mean(q5))/sd(q5),
+      x_pol = (q19_1 - mean(q19_1))/sd(q19_1)
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
+        observers == "White" & protesters == "Black" & protest == "for"     ~ 5L, 
+        observers == "White" & protesters == "Black" & protest == "against" ~ 6L, 
[…]

### 2022-05-24 Experiment 2/scripts/run_factor_analyses.R

@@ -1,0 +1,28 @@
+# Load packages -----------------------------------------------------------
+  library(lavaan)
+
+# System-Justifying Beliefs (SJB) -----------------------------------------
+
+  # Load models
+  sjb_model <- '
+      sjb =~ sjb_1 + sjb_2 + sjb_3 + sjb_4 + sjb_5 + sjb_6 + sjb_7 + sjb_8 
+    '
+  
+  # Run models
+  sjb_fit <- cfa(
+    sjb_model,
+    estimator = "mlr",
+    missing = "fiml",
+    std.lv = TRUE,
+    data = dw
+  )
+  
+  # Export factor scores
+  dw <- predict(sjb_fit) %>%
+    as.data.frame() %>%
+    bind_cols(dw, .)
+
+
+# Remove ------------------------------------------------------------------
+  detach(package:lavaan)
+  rm("sjb_model", "sjb_fit")

### 2022-05-24 Experiment 2/scripts/run_preregistered_analyses.R

@@ -1,0 +1,270 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+  
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = case_when(
+        protest == "for" ~ q5,
+        protest == "against" ~ 6L - q5
+      )
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
+        observers == "White" & protesters == "Black" & protest == "for"     ~ 5L, 
[…]

### 2022-05-24 Experiment 2/scripts/run_simulations.R

@@ -1,0 +1,305 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 1/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  
+  # Code indices
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        sample == "WJ" & target == "PJ" ~ 1L,
+        sample == "WJ" & target == "WJ" ~ 2L,
+        sample == "PJ" & target == "PJ" ~ 3L,
+        sample == "PJ" & target == "WJ" ~ 4L
+      ),
+      y  = response
+    )
+  
+  # Compile data list
+  dlist <- with(dl, list(
+    N = length(ii),
+    I = max(ii),
+    J = max(jj),
+    K = max(kk),
+    ii = ii,
+    jj = jj,
+    kk = kk,
+    y = y
+  ))
+
+
+# Estimate ----------------------------------------------------------------
+
+  # Compile model
+  model <- cmdstan_model("Experiment 2/models/s2_2pl_model.stan")
+
+  # Run model
+  fit <- model$sample(
+    data = dlist,
+    seed = 3909872,
+    chains = 8,
+    parallel_chains = parallel::detectCores(),
+    iter_sampling = 500,
+    iter_warmup = 500,
+    adapt_delta = 0.95
+  )
+
+
[…]

### 2022-05-24 Scale Development/figures/figure-1.R

@@ -1,0 +1,141 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("ggtext")
+  
+  # Load packages
+  library(tidyverse)
+  library(ggtext)
+
+
+# Functions ---------------------------------------------------------------
+
+  # Link function
+  inv_logit <- function(x) exp(x) / ( 1 + exp(x) )
+  
+  # IRT equation
+  irt <- function(theta, alpha, beta) {
+    inv_logit(alpha*(theta + beta))
+  }
+  irt <- Vectorize(irt)
+  
+
+# Visualize ---------------------------------------------------------------
+  
+  # Make figure
+  ggplot() +
+    geom_vline(xintercept = 0.0, colour = "grey92", size = 0.455) +
+    geom_hline(yintercept = 0.5, colour = "grey92", size = 0.455) +
+    stat_function(
+      aes(colour = "Action 1"),
+      fun = irt, 
+      args = list(
+        alpha = 1.00,
+        beta  = 0.00
+      ),
+      n = 10001, size = 0.455
+    ) +
+    stat_function(
+      aes(colour = "Action 2"),
+      fun = irt, 
+      args = list(
+        alpha = 1.00,
+        beta  = qnorm(0.9)
+      ),
+      n = 10001, size = 0.455
+    ) +
+    stat_function(
+      aes(colour = "Action 3"),
+      fun = irt, 
+      args = list(
+        alpha = 3.60,
+        beta  = 0.00
+      ),
+      n = 10001, size = 0.455
+    ) +
+    annotate(
+      geom = "rich_text",
+      x = c(rep(qnorm(0.1), 3), rep(qnorm(0.9), 3)),
+      y = c(
+        irt(qnorm(0.1), 1.00, 0.00),
+        irt(qnorm(0.1), 1.00, qnorm(0.9)),
+        irt(qnorm(0.1), 3.60, 0.00),
+        irt(qnorm(0.9), 1.00, 0.00),
+        irt(qnorm(0.9), 1.00, qnorm(0.9)),
+        irt(qnorm(0.9), 3.60, 0.00)
+      ),
+      label = rep(1:3, 2),
+      colour = rep(c("#648FFF", "#DC267F", "#FFB000"), 2),
[…]

### 2022-05-24 Scale Development/figures/figure-2.R

@@ -1,0 +1,101 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("ggtext")
+  
+  # Load packages
+  library(tidyverse)
+  library(ggtext)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Scale Development/data/dl_ps2.rds")
+  
+  # Exclude
+  dl <- dl %>% filter(assigned == 1)
+  
+  # Summarize
+  ds <- dl %>% 
+    group_by(action, action_text) %>% 
+    summarize(p_sometimes = mean(q2 >= 3L)) %>% 
+    ungroup()
+  
+
+# Visualize ---------------------------------------------------------------
+  
+  # Prepare data
+  d_fig <- ds %>% 
+    filter(action %in% c(1L, 23L, 31L, 38L, 46L, 52L, 56L, 60L)) %>% 
+    mutate(
+      action_text = case_when(
+        str_detect(action_text, "\\(e.g.,") ~ str_replace(action_text, " \\(e.g.,", "\n(e.g.,"),
+        TRUE ~ str_wrap(action_text, width = 40)
+      ),
+      included = (action %in% c(23L, 38L, 46L, 56L))
+    ) %>% 
+    arrange(p_sometimes)
+  
+  # Make figure
+  d_fig %>% 
+    mutate(action_text = factor(action_text, levels = d_fig$action_text)) %>% 
+    ggplot(., aes(x = p_sometimes, y = action_text)) +
+    geom_col(
+      aes(fill = included),
+      width = 0.8
+    ) +
+    geom_vline(xintercept = 0, size = 0.455) +
+    geom_richtext(
+      aes(
+        label = numform::f_prop2percent(p_sometimes, digits = 0),
+        hjust = if_else(p_sometimes > 0.25, 1, 0),
+        colour = if_else(p_sometimes > 0.25, "white", "black")
+      ),
+      size = 10*0.8/.pt,
+      fill = NA,
+      label.colour = NA
+    ) +
+    scale_x_continuous(labels = scales::percent) +
+    scale_colour_identity() +
+    scale_fill_manual(values = c("grey70", "black")) +
+    coord_cartesian(xlim = c(-0.008, 1.008), ylim = c(0.5, 8.5), expand = FALSE) +
+    theme_classic(base_size = 10) +
[…]

### 2022-05-24 Scale Development/figures/table-s1.R

@@ -1,0 +1,40 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("numform")
+  
+  # Load packages
+  library(tidyverse)
+  library(numform)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import results 
+  results <- read_rds("Scale Development/results/results_scale_development.rds")
+  
+  # Reformat
+  results <- results %>% 
+    arrange(desc(I)) %>% 
+    transmute(
+      `#` = row_number(),
+      Action = action_text,
+      Information = f_num(I, digits = 2, retain.leading.zero = T),
+      alpha = f_num(alpha, digits = 2, retain.leading.zero = T),
+      across(
+        c(`1`:`5`),
+        ~round(. * 100)
+      )
+    )
+  
+
+# Export ------------------------------------------------------------------
+
+  # Save as .rds
+  write_rds(results, "Scale Development/figures/table-s1.rds")

### 2022-05-24 Scale Development/figures/table-s2.R

@@ -1,0 +1,54 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("numform")
+  
+  # Load packages
+  library(tidyverse)
+  library(numform)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Scale Development/data/dl_ps2.rds")
+  
+  # Import Table S2
+  results <- read_rds("Scale Development/figures/table-s1.rds")
+  
+  # Exclude
+  dl <- dl %>% filter(assigned == 1)
+  
+  # Summarize
+  ds <- dl %>% 
+    group_by(action, action_text) %>% 
+    summarize(
+      `Acce.` = mean(q2, na.rm = TRUE),
+      `Disr.` = mean(q1_1, na.rm = TRUE),
+      `Viol.` = mean(q1_2, na.rm = TRUE),
+      `Extr.` = mean(q1_3, na.rm = TRUE),
+      `Nega.` = mean(q3, na.rm = TRUE)
+    ) %>% 
+    ungroup() %>% 
+    transmute(
+      Action = action_text,
+      across(
+        `Acce.`:`Nega.`, 
+        ~f_num(., digits = 2, retain.leading.zero = T)
+      )
+    ) %>% 
+    left_join(results %>% distinct(`#`, Action), ., by = c("Action"))
+  
+
+  
+
+# Export ------------------------------------------------------------------
+
+  # Save as .rds
+  write_rds(ds, "Scale Development/figures/table-s2.rds")

### 2022-05-24 Scale Development/figures/table-s3.R

@@ -1,0 +1,48 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("numform")
+  # install.packages("corrr")
+  
+  # Load packages
+  library(tidyverse)
+  library(numform)
+  library(corrr)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Scale Development/data/dl_ps2.rds")
+  
+  # Exclude
+  dl <- dl %>% filter(assigned == 1)
+  
+  # Summarize 
+  ds <- dl %>% 
+    select(
+      Acceptable = q2,
+      Disruptive = q1_1,
+      Violent = q1_2,
+      Extreme = q1_3,
+      Negative = q3
+    ) %>% 
+    correlate() %>% 
+    transmute(
+      `#` = row_number(),
+      Rating = term,
+      across(c(-`#`, -Rating, -term), ~f_num(., digits = 2)),
+      across(c(-`#`, -Rating, -term), ~replace_na(., ""))
+    )
+  
+
+# Export ------------------------------------------------------------------
+
+  # Save as .rds
+  write_rds(ds, "Scale Development/figures/table-s3.rds")

### 2022-05-24 Scale Development/scripts/run_analyses.R

@@ -1,0 +1,180 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+  # An ideal item should be (1) ambigious in that neither "never" nor 
+  # "always" gets all the votes, and (2) polarising in that an item should
+  # be prefered if not-"sometimes" responses fall on either side of the 
+  # mid-point.
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+  
+# Prepare -----------------------------------------------------------------
+  
+  # Import data
+  dl <- read_rds("Scale Development/data/dl_ps2.rds")
+  
+  # Exclude observations
+  dl <- dl %>% filter(assigned == 1)
+  
+  # Exclude actions
+  actions <- dl %>%
+    group_by(action, action_text) %>%
+    count(q2) %>%
+    spread(q2, n) %>%
+    ungroup() %>%
+    mutate_all(~replace_na(., 0L)) %>%
+    # filter(`1` + `2` > 0, `4` + `5` > 0) %>%
+    pull(action)
+  
+  # Prepare data frame
+  df <- dl %>%
+    select(id, sample, q2, action, action_text) %>%
+    filter(action %in% actions) %>%
+    transmute(
+      id, sample, action, action_text,
+      ii = as.integer(factor(action)),
+      jj = id,
+      y = q2
+    ) %>%
+    arrange(jj, ii)
+  
+  # Prepare data list
+  dlist <- with(df, list(
+    N  = length(y),
+    I  = max(ii),
+    J  = max(jj),
+    K  = max(y),
+    ii = ii,
+    jj = jj,
+    y  = y
+  ))
+  
+
+# Estimate ----------------------------------------------------------------
+  
+  # Compile model
[…]

### 2022-05-24 Experiment 2/scripts/run_exploratory_analyses.R
remove M5 from Experiment 2

@@ -93,25 +93,6 @@ rm(list = ls())
   # Save results as .rds
   m4_fit$save_object("Experiment 2/results/m4_fit.rds")
   
-  
-# Estimate (Model 5) ------------------------------------------------------
-
-  # Compile model
-  m5_model <- cmdstan_model("Experiment 2/models/s2_m5_2pl_model.stan")
-
-  # Run model
-  m5_fit <- m5_model$sample(
-    data = dlist,
-    seed = 1187052,
-    chains = 8,
-    parallel_chains = parallel::detectCores(),
-    iter_sampling = 500,
-    iter_warmup = 1000
-  )
-
-  # Save results as .rds
-  m5_fit$save_object("Experiment 2/results/m5_fit.rds")
-  
 
 # Process (Model 4) -------------------------------------------------------
   

### 2022-06-16 Documents/manuscript.Rmd
update manuscript

@@
 keywords: protest, collective action, double standards, social identity theory, system justification theory, Black Lives Matter
-abstract: Collective action is a powerful force driving social change but, at the same time, often sparks contention about what are acceptable forms of collective action.
+abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
+In four studies (total *N* = 2,246), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*).
+In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action.
+In two preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police in the United States.
+We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position---but participants showed no ingroup bias when judging collective action by ingroup and outgroup members.
+Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
 bibliography: references.bib
@@
 In two preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police in the United States.
-We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but show no ingroup bias when judging collective action by ingroup and outgroup members.
+We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own (system-challenging) ideological positions---but do not show ingroup bias when judging collective action by ingroup and outgroup members.
 
-<!-- Below, ... we situate the distinction ... below we introduce the distinction between normative ... derive predictions about ... from ... and introduce the present research ... -->
+<!-- In the next sections, we, first, introduce the consequential distinction between normative and nonnormative collective action and define double standards in judging what actions are acceptable and unacceptable means to effect social change.
+We, second, derive predictions from social psychological theories about identity- and ideology-based double standards that might affect judgements about collective action.
+We, third, introduce item response theory as a conceptual and statistical framework for study double standards in judging collective action.
+We, finally, provide an overview of the present research.
+-->
 
@@
 Similarly, @teixeira_is_2020 found that advantaged-group members perceive non-normative collective action by disadvantaged-group members as more damaging to their ingroup's social image and are thus less supportive of such action.
+@verkuyten_intolerance_2022 found that, in general, the public tends to be intolerant of non-normative collective action.
 This means that activists face the dilemma that non-normative collective action might be most effective at gaining concessions but that it also risks reducing popular support for a cause.
@@
 Second, we hypothesize that people will judge the same protest actions as more acceptable if they endorse system-justifying beliefs and the protest supports the system or if they reject system-justifying beliefs and the protest challenges the system (*Hypothesis 2b*).
-This hypothesis implies an symmetry in how people with different ideological orientations judge system-supporting and system-challenging collective action.
+Notably, this hypothesis implies a symmetry in how people with different ideological orientations judge system-supporting and system-challenging collective action.
 
@@
 In this research, item response theory serves two purposes.
-First, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters and what they are protesting.
+First, item response theory provides a statistical framework for scale development.
+In Studies 1 and 2, we select the most discriminating and informative actions to build an instrument for capturing double standards in the experiments.
+Second, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters and what they are protesting.
 In Experiments 1 and 2, we operationalize double standards as the differences in $\theta_{(z)}$ between conditions which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
-Second, item response theory provides a statistical framework for scale development.
-In Studies 1 and 2, we select the most discriminating and informative actions to build an instrument for capturing double standards in the experiments.
 
@@
 
-Our research serves two purposes.
-The first purpose is to develop an instrument to measure where people draw the line between acceptable and unacceptable forms of collective action.
+The first purpose of our research is to develop an instrument to measure where people draw the line between acceptable and unacceptable forms of collective action.
 In Study 1, we ask participants to generate more and less extreme protest actions and compile a pool of protest actions for scale development.
@@
 
-The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more or less acceptable depending on their own and the protesters' group memberships (Hypothesis 1a, 1b) or on the protesters' cause and how it aligns with their own ideological positions (Hypotheses 2a, 2b).
+The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more acceptable if the protesters' are ingroup members or the protesters' cause aligns with the ingroup's interests (Hypothesis 1a, 1b) or if the protesters' cause aligns with their own system-justifying or system-challenging ideological orientation (Hypotheses 2a, 2b).
 In Experiment 1, we use the instrument developed in Studies 1 and 2 to test for double standards in how acceptable participants consider collective action for workers' rights in the United Kingdom.
-By varying both the participants' and protesters' group memberships (working/middle class), Experiment 1 compares the relative evidence for Hypotheses 1a and 2a.
+By varying both the participants' and the protesters' group memberships (working/middle class), Experiment 1 compares the relative evidence for identity-based (Hypotheses 1a, 1b) and ideology-based double standards (Hypothesis 2a).
 Extending Experiment 1, Experiment 2 considers both system-challenging and system-defending collective action in the context of the 2020 Black Lives Matter protests in the United States.
-By varying the protest's cause (for/against defunding the police) as well as the participants' and protesters' group memberships (Black/White), Experiment 2 compares the relative evidence for Hypotheses 1a, 1b, 2a, and 2b.
-Together, the preregistered experiments provide a comprehensive test of the hypothesized identity- and ideology-based double standards in judging collective action.
+By varying the protesters' cause (for/against defunding the police) as well as the participants' and the protesters' group memberships (Black/White), Experiment 2 compares the relative evidence identity-based (Hypotheses 1a, 1b) and ideology-based double standards (Hypothesis 2a, 2b).
+Together, the preregistered experiments provide a comprehensive test of the hypothesized double standards in judging collective action.
 
@@
 
-<!-- In two studies, we developed an instrument of 25 protest actions to capture double standards in judging collective action. -->
+In two studies, we developed an instrument of 25 protest actions to capture double standards in judging collective action.
 
@@
 
-\ldots
+In Experiment 1, we used the newly developed instrument to test for double standards in judging collective action for workers' rights in the United Kingdom.
+That is, we tested whether people with either working-class or professional jobs applied different standards when judging collective action taken by people with either working-class or middle-class jobs to protest against a fictitious government bill threatening their rights.
+
+By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that, in line with Hypothesis 1a and 1b, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights than if they were outgroup members protesting for the outgroup's rights.
+By comparing reactions to protesters with working-class or professional jobs, we tested how observers judge collective action by lower-status (i.e.
+working-class) and higher-status (professional/middle-class) group members.
+Assuming that collective action by lower-status group members is more threatening to the unequal status than collective action by higher-status group members, we tested the preregistered hypothesis that, in line with Hypothesis 2a, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group than if they were from a lower-status group.
+In non-preregistered analyses, we further explored whether, in line with Hypothesis 2b, participants would judge the same protest actions as more or less acceptable depending on their own ideological positions.
+In this way, Experiment 1 provided a first test of the hypothesized identity- and ideology-based double standards in judging collective action.
 
@@
 
+\ldots
+
 # Experiment 2
+
+In Experiment 2, we investigated potential double standards in judging collective action for or against defunding the police in the United States.
+That is, we tested whether Black and White Americans applied different standards when judging collective action taken by either Black or White protesters to protest either for or against police divestment as a possible solution to racist police violence.
 
@@
 
-In Experiment 2, we tested five hypotheses.
+<!-- In Experiment 2, we tested five hypotheses.
 We tested two predictions derived from social identity theory.
@@
 We note that Hypotheses 1a, 1b, and 2a make competing predictions about differences between conditions while Hypotheses 2b and 3 need not contradict each other or any of the other hypotheses.
+-->
 
@@
 We measured political orientation with a one-item semantic differential scale: "People often describe their political orientation as liberal or conservative.
-On a scale from liberal to conservative, where would you position yourself?"; 1 = *liberal*, 7 = *conservative*).
+On a scale from liberal to conservative, where would you position yourself?" (1 = *liberal*, 7 = *conservative*).
 
@@
 
-Figure \ref{fig:f5} shows the predicted pattern of condition-wise differences underlying those fixed effects.
+Figure \ref{fig:f5} shows the estimated pattern of condition-wise differences underlying those fixed effects.
 Participants who endorsed system-justifying beliefs tended to consider system-challenging and system-defending protest actions to be equally unacceptable.
-In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with Hypothesis 2b, they considered the same protest actions to be more acceptable if the protesters challenged the system or, to a lesser extent, if the protesters were from the disadvantaged group.
+In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with Hypothesis 2b, they considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.
 
@@
 
-We reran Model 3 with political orientation, instead of system justification, as an alternative predictor variable.
+We reran Model 2 with political orientation, instead of system justification, as an alternative predictor variable.
 Our results mirrored results from the preregistered analyses: More liberal participants were more likely to consider protest actions for defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "for"])`) but more conservative participants were not more likely to consider protest actions against defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "against"])`).
-As Figure 5 shows, conservative participants tended to consider all protest actions to be equally unacceptable while liberal participants considered the same protest actions to be more acceptable if protesters rallied around a progressive cause.
-Our non-preregistered analyses thus found evidence for ideology-based double standards in judging collective action with a different operationalization of political ideology.
+As Figure 5 shows, conservative participants tended to consider all protest actions to be equally unacceptable while liberal participants considered the same protest actions to be more acceptable when protesters rallied around a progressive cause.
+Our non-preregistered analyses thus replicated the ideology-based double standards from the preregistered analyses with a different operationalization of political ideology.
 
@@
 
+\ldots
+
 # General Discussion
 
+<!-- Our research was motivated by examples showing that different groups judge the same protest action more acceptable depending on who's doing the action.
+That is, in contrast to previous research, not the case.
+-->
+
 - Summary: No evidence for identity-based double standards, but some evidence for ideology-based double standards
+- Reject the alternative hypothesis that there is consensus about what is an acceptable means of protest.
+- Relate our findings to @verkuyten_intolerance_2022 
 - Two plausible interpretations: (1) conservatives oppose controversial protest (no matter the end) *or* (2) the end justifies the mean
@@
 - Discuss why high-status group are more accepting of controversial protest actions
-- Discuss limitations (two cultures, two issues, liberal samples; system-defending vs. system-supporting actions; what about other conservative causes?)
+- Discuss limitations (two cultures, two issues, liberal samples; system-defending vs.
+system-supporting actions; what about other conservative causes?; ingroup bias might need zero-sum situation to succeed)
 - Invite researchers to use our paradigm

### 2022-06-16 Documents/manuscript.tex
update manuscript

@@
       \noindent Collective action is a powerful force driving social
-change but, at the same time, often sparks contention about what are
-acceptable forms of collective action.
+change but often sparks contention about what actions are acceptable
+means to effect social change. In four studies (total \emph{N} = 2,246),
+we investigated double standards in judging collective action---that is,
+whether observers will judge the same protest actions as more acceptable
+if the protesters are ingroup members and their cause aligns with the
+ingroup's interests (\emph{identity-based double standards}) or if the
+protesters' cause aligns with the observers' ideological positions
+(\emph{ideology-based double standards}). In two studies, we developed
+an instrument of 25 controversial protest actions, based in item
+response theory, to measure where people draw the line between
+acceptable and unacceptable forms of collective action. In two
+preregistered experiments, we used this instrument to test for double
+standards in judging collective action for workers' rights in the United
+Kingdom and for or against defunding the police in the United States. We
+found evidence for ideology-based, but not for identity-based, double
+standards: Participants judged the same protest actions to be more
+acceptable when the protesters' cause aligned with their own
+(system-challenging) ideological position---but participants showed no
+ingroup bias when judging collective action by ingroup and outgroup
+members. Our findings have theoretical and practical implications for
+understanding the often divided response to prominent social movements.
     \end{abstract}
@@
 to be more acceptable when the protesters' cause aligns with their own
-ideological positions---but show no ingroup bias when judging collective
-action by ingroup and outgroup members.
+(system-challenging) ideological positions---but do not show ingroup
+bias when judging collective action by ingroup and outgroup members.
 
@@
 their ingroup's social image and are thus less supportive of such
-action. This means that activists face the dilemma that non-normative
-collective action might be most effective at gaining concessions but
-that it also risks reducing popular support for a cause. Together, these
-studies show that the distinction between normative and non-normative
-collective action is psychologically and societally consequential.
+action. Verkuyten et al. (2022) found that, in general, the public tends
+to be intolerant of non-normative collective action. This means that
+activists face the dilemma that non-normative collective action might be
+most effective at gaining concessions but that it also risks reducing
+popular support for a cause. Together, these studies show that the
+distinction between normative and non-normative collective action is
+psychologically and societally consequential.
 
@@
 reject system-justifying beliefs and the protest challenges the system
-(\emph{Hypothesis 2b}). This hypothesis implies an symmetry in how
-people with different ideological orientations judge system-supporting
-and system-challenging collective action.
+(\emph{Hypothesis 2b}). Notably, this hypothesis implies a symmetry in
+how people with different ideological orientations judge
+system-supporting and system-challenging collective action.
 
@@
 In this research, item response theory serves two purposes. First, item
-response theory provides a formal and statistical definition of double
-standards in judging collective action as respondents being more
-accepting of the same protest actions depending on who the protesters
-and what they are protesting. In Experiments 1 and 2, we operationalize
-double standards as the differences in \(\theta_{(z)}\) between
-conditions which, as \(\theta_{(z)}\) is \(z\)-standardized, correspond
-to Cohen's \(d\) effect sizes. Second, item response theory provides a
-statistical framework for scale development. In Studies 1 and 2, we
-select the most discriminating and informative actions to build an
-instrument for capturing double standards in the experiments.
+response theory provides a statistical framework for scale development.
+In Studies 1 and 2, we select the most discriminating and informative
+actions to build an instrument for capturing double standards in the
+experiments. Second, item response theory provides a formal and
+statistical definition of double standards in judging collective action
+as respondents being more accepting of the same protest actions
+depending on who the protesters and what they are protesting. In
+Experiments 1 and 2, we operationalize double standards as the
+differences in \(\theta_{(z)}\) between conditions which, as
+\(\theta_{(z)}\) is \(z\)-standardized, correspond to Cohen's \(d\)
+effect sizes.
 
@@
 
-Our research serves two purposes. The first purpose is to develop an
-instrument to measure where people draw the line between acceptable and
-unacceptable forms of collective action. In Study 1, we ask participants
-to generate more and less extreme protest actions and compile a pool of
-protest actions for scale development. In Study 2, we ask participants
-to rate how acceptable they consider each of those protest action to be
-and use item response theory to select the most discriminating and
-informative protest actions for our instrument. In so doing, we build an
-instrument that enables both present and future research to capture
-double standards in judging collective action. In this way, our research
+The first purpose of our research is to develop an instrument to measure
+where people draw the line between acceptable and unacceptable forms of
+collective action. In Study 1, we ask participants to generate more and
+less extreme protest actions and compile a pool of protest actions for
+scale development. In Study 2, we ask participants to rate how
+acceptable they consider each of those protest action to be and use item
+response theory to select the most discriminating and informative
+protest actions for our instrument. In so doing, we build an instrument
+that enables both present and future research to capture double
+standards in judging collective action. In this way, our research
 situates the distinction between normative and non-normative collective
@@
 judging collective action---that is, whether people judge the same
-protest actions as more or less acceptable depending on their own and
-the protesters' group memberships (Hypothesis 1a, 1b) or on the
-protesters' cause and how it aligns with their own ideological positions
+protest actions as more acceptable if the protesters' are ingroup
+members or the protesters' cause aligns with the ingroup's interests
+(Hypothesis 1a, 1b) or if the protesters' cause aligns with their own
+system-justifying or system-challenging ideological orientation
 (Hypotheses 2a, 2b). In Experiment 1, we use the instrument developed in
@@
 participants consider collective action for workers' rights in the
-United Kingdom. By varying both the participants' and protesters' group
-memberships (working/middle class), Experiment 1 compares the relative
-evidence for Hypotheses 1a and 2a. Extending Experiment 1, Experiment 2
-considers both system-challenging and system-defending collective action
-in the context of the 2020 Black Lives Matter protests in the United
-States. By varying the protest's cause (for/against defunding the
-police) as well as the participants' and protesters' group memberships
-(Black/White), Experiment 2 compares the relative evidence for
-Hypotheses 1a, 1b, 2a, and 2b. Together, the preregistered experiments
-provide a comprehensive test of the hypothesized identity- and
-ideology-based double standards in judging collective action.
+United Kingdom. By varying both the participants' and the protesters'
+group memberships (working/middle class), Experiment 1 compares the
+relative evidence for identity-based (Hypotheses 1a, 1b) and
+ideology-based double standards (Hypothesis 2a). Extending Experiment 1,
+Experiment 2 considers both system-challenging and system-defending
+collective action in the context of the 2020 Black Lives Matter protests
+in the United States. By varying the protesters' cause (for/against
+defunding the police) as well as the participants' and the protesters'
+group memberships (Black/White), Experiment 2 compares the relative
+evidence identity-based (Hypotheses 1a, 1b) and ideology-based double
+standards (Hypothesis 2a, 2b). Together, the preregistered experiments
+provide a comprehensive test of the hypothesized double standards in
+judging collective action.
 
@@
 \section{Scale Development}\label{scale-development}}
+
+In two studies, we developed an instrument of 25 protest actions to
+capture double standards in judging collective action.
 
@@
 
-\ldots
+In Experiment 1, we used the newly developed instrument to test for
+double standards in judging collective action for workers' rights in the
+United Kingdom. That is, we tested whether people with either
+working-class or professional jobs applied different standards when
+judging collective action taken by people with either working-class or
+middle-class jobs to protest against a fictitious government bill
+threatening their rights.
+
+By varying the participants' and the protesters' group memberships, we
+tested the preregistered hypothesis that, in line with Hypothesis 1a and
+1b, participants would judge the same protest actions as more acceptable
+if the protesters were ingroup members protesting for the ingroup's
+rights than if they were outgroup members protesting for the outgroup's
+rights. By comparing reactions to protesters with working-class or
+professional jobs, we tested how observers judge collective action by
+lower-status (i.e.~working-class) and higher-status
+(professional/middle-class) group members. Assuming that collective
+action by lower-status group members is more threatening to the unequal
+status than collective action by higher-status group members, we tested
+the preregistered hypothesis that, in line with Hypothesis 2a,
+participants would judge the same protest actions as more acceptable if
+the protesters were from a higher-status group than if they were from a
+lower-status group. In non-preregistered analyses, we further explored
+whether, in line with Hypothesis 2b, participants would judge the same
+protest actions as more or less acceptable depending on their own
+ideological positions. In this way, Experiment 1 provided a first test
+of the hypothesized identity- and ideology-based double standards in
+judging collective action.
 
@@
 
+\ldots
+
 \hypertarget{experiment-2}{%
 \section{Experiment 2}\label{experiment-2}}
+
+In Experiment 2, we investigated potential double standards in judging
+collective action for or against defunding the police in the United
+States. That is, we tested whether Black and White Americans applied
+different standards when judging collective action taken by either Black
+or White protesters to protest either for or against police divestment
+as a possible solution to racist police violence.
 
@@
 \end{itemize}
-
-In Experiment 2, we tested five hypotheses. We tested two predictions
-derived from social identity theory. First, we hypothesize that
-observers are more likely to judge an action as an acceptable means of
-protest when the protesters are ingroup members than when they are
-outgroup members (Hypothesis 1a). Second, we hypothesize that observers
-are more likely to judge an action as an acceptable means of protest
-when the goal of the protests would benefit the observers' ingroup
-(Hypothesis 1b). We will test two predictions derived from system
-justification theory. First, we hypothesize that observers are more
-likely to judge an action as an acceptable means of protest when the
-goal of the protest is to support rather than challenge the system
-(Hypothesis 2a). Second, we hypothesize that observers who endorse
-system-justifying beliefs are more likely to consider system-supporting
-actions acceptable means of protest-and less likely to consider
-system-challenging actions acceptable means of protest-than observers
-who do not endorse system-justifying beliefs (Hypothesis 2b). In
-addition, we will test an alternative hypothesis. We hypothesize that
-observers are more likely to consider actions as acceptable means of
-protest when they support the cause of the protesters than when they
-oppose their cause (Hypothesis 3). We note that Hypotheses 1a, 1b, and
-2a make competing predictions about differences between conditions while
-Hypotheses 2b and 3 need not contradict each other or any of the other
-hypotheses.
 
@@
 orientation as liberal or conservative. On a scale from liberal to
-conservative, where would you position yourself?''; 1 = \emph{liberal},
+conservative, where would you position yourself?'' (1 = \emph{liberal},
 7 = \emph{conservative}).
@@
 
-Figure \ref{fig:f5} shows the predicted pattern of condition-wise
+Figure \ref{fig:f5} shows the estimated pattern of condition-wise
 differences underlying those fixed effects. Participants who endorsed
@@
 ideology-based double standards: In line with Hypothesis 2b, they
-considered the same protest actions to be more acceptable if the
-protesters challenged the system or, to a lesser extent, if the
+considered the same protest actions to be more acceptable when the
+protesters challenged the system or, to a lesser extent, when the
 protesters were from the disadvantaged group.
@@
 
-We reran Model 3 with political orientation, instead of system
+We reran Model 2 with political orientation, instead of system
 justification, as an alternative predictor variable. Our results
@@
 unacceptable while liberal participants considered the same protest
-actions to be more acceptable if protesters rallied around a progressive
-cause. Our non-preregistered analyses thus found evidence for
-ideology-based double standards in judging collective action with a
+actions to be more acceptable when protesters rallied around a
+progressive cause. Our non-preregistered analyses thus replicated the
+ideology-based double standards from the preregistered analyses with a
 different operationalization of political ideology.
@@
 \subsection{Discussion}\label{discussion-1}}
+
+\ldots
 
@@
   evidence for ideology-based double standards
+\item
+  Reject the alternative hypothesis that there is consensus about what
+  is an acceptable means of protest.
+\item
+  Relate our findings to Verkuyten et al. (2022)
 \item
@@
   system-defending vs.~system-supporting actions; what about other
-  conservative causes?)
+  conservative causes?; ingroup bias might need zero-sum situation to
+  succeed)
 \item
@@
 87--114. \url{https://doi.org/10.1111/pops.12322}
+
+\leavevmode\vadjust pre{\hypertarget{ref-verkuyten_intolerance_2022}{}}%
+Verkuyten, M., Adelman, L., \& Yogeeswaran, K. (2022). Intolerance of
+transgressive protest actions: The differential roles of deontological
+and utilitarian morality. \emph{Personality and Social Psychology
+Bulletin}. \url{https://doi.org/10.1177/01461672221099709}
 

### 2022-06-16 Documents/supplemental-materials.Rmd
update manuscript

@@
 Second, we set out to recruit 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group.
-After giving their informed consent, participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (*lower-status actors*) or people in professional jobs (*higher-status actors*).
+After giving their informed consent, participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (*lower-status protesters*) or people in professional jobs (*higher-status protesters*).
 Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real.
-Participants in the lower-status actors condition then read the following introduction:
+Participants in the lower-status protesters condition then read the following introduction:
 
@@
 Other jobs are unlikely to be affected.
-Participants in the higher-status actors condition instead read the following introduction:
+
+\noindent Participants in the higher-status protesters condition instead read the following introduction:
 
@@
 Other jobs are unlikely to be affected.
-Participants in the two conditions then read this almost identically worded paragraph:
+
+\noindent Participants in the two conditions then read this almost identically worded paragraph:
 

### 2022-06-16 Documents/supplemental-materials.tex
update manuscript

@@
 the bottom three ranks of the subjective socio-economic status ladder
-(lower-status group) and 400 participants had at least an undergraduate
-degree and placed themselves on the top four ranks of the subjective
-socio-economic status ladder (higher-status group). Participants
-completed a screening survey in which they, among other questions,
-answered whether they considered either their current job or the jobs
-they had had in the past or would have in the future to be a
-working-class job or a middle-class/professional job (1 = working-class
-job, 2 = middle-class/professional job, 3 = neither). Before answering
-this question, participants read the following explanation:
+(\emph{lower-status group}) and 400 participants had at least an
+undergraduate degree and placed themselves on the top four ranks of the
+subjective socio-economic status ladder (\emph{higher-status group}).
+Participants completed a screening survey in which they, among other
+questions, answered whether they considered either their current job or
+the jobs they had had in the past or would have in the future to be a
+working-class job or a middle-class/professional job (1 =
+\emph{working-class job}, 2 = \emph{middle-class/professional job}, 3 =
+\emph{neither}). Before answering this question, participants read the
+following explanation:
 
@@
 
-As preregistered, we excluded participants from the lower-status group
-who had responded ``middle-class/professional job'' or ``neither'' to
-this question; we excluded participants from the higher-status group who
-had responded ``working-class job'' or ``neither'' to this question.
-This left 687 participants for the next step of the selection procedure.
+\noindent As preregistered, we excluded participants from the
+lower-status group who had responded ``middle-class/professional job''
+or ``neither'' to this question; we excluded participants from the
+higher-status group who had responded ``working-class job'' or
+``neither'' to this question. This left 687 participants for the next
+step of the selection procedure.
 
@@
 were randomly assigned to read a vignette about a government bill
-affecting either people in working-class jobs (lower-status actors) or
-people in professional jobs (higher-status actors). Participants in both
-conditions were instructed to carefully read the vignette and to try to
-imagine what it would be like if this situation was real. Participants
-in the lower-status actors condition then read the following
-introduction:
+affecting either people in working-class jobs (\emph{lower-status
+protesters}) or people in professional jobs (\emph{higher-status
+protesters}). Participants in both conditions were instructed to
+carefully read the vignette and to try to imagine what it would be like
+if this situation was real. Participants in the lower-status protesters
+condition then read the following introduction:
 
@@
 jobs that do not usually require a university degree. Other jobs are
-unlikely to be affected. Participants in the higher-status actors
-condition instead read the following introduction:
+unlikely to be affected.
 \end{quote}
+
+\noindent Participants in the higher-status protesters condition instead
+read the following introduction:
 
@@
 usually require a university degree. Other jobs are unlikely to be
-affected. Participants in the two conditions then read this almost
+affected.
+\end{quote}
+
+\noindent Participants in the two conditions then read this almost
 identically worded paragraph:
-\end{quote}
 

### 2022-06-16 Experiment 1/scripts/run_preregistered_analyses.R

@@ -142,7 +142,7 @@ rm(list = ls())
 # Compare -----------------------------------------------------------------
 
   # By ingroup
-  d_fig %>%
+  post %>%
     group_by(.draw, .chain, .iteration, ingroup) %>%
     summarize(
       z = mean(z)
@@ -150,10 +150,11 @@ rm(list = ls())
     ungroup() %>%
     pivot_wider(names_from = ingroup, values_from = z) %>%
     # summarize(p = mean(ingroup > outgroup))
-    median_qi(d = ingroup - outgroup)
+    median_qi(d = Ingroup - Outgroup) %>% 
+    mutate(across(where(is.double), round, 2))
 
   # By actor
-  d_fig %>%
+  post %>%
     group_by(.draw, .chain, .iteration, actor) %>%
     summarize(
       z = mean(z)
@@ -161,13 +162,16 @@ rm(list = ls())
     ungroup() %>%
     pivot_wider(names_from = actor, values_from = z) %>%
     # summarize(p = mean(HS > LS))
-    median_qi(d = HS - LS)
+    median_qi(d = HS - LS) %>% 
+    mutate(across(where(is.double), round, 2))
 
   # By actor/observer
-  d_fig %>%
-    select(-sample, -target, -n, -p, -ingroup) %>%
-    pivot_wider(names_from = observer, values_from = z) %>%
-    group_by(actor) %>%
+  post %>%
+    # select(-sample, -target, -n, -p, -ingroup) %>%
+    select(-sample, -target, -ingroup, -kk) %>%
+    pivot_wider(names_from = actor, values_from = z) %>%
+    group_by(observer) %>%
     # summarize(p = mean(HS > LS))
-    median_qi(d = HS - LS)
+    median_qi(d = HS - LS) %>% 
+    mutate(across(where(is.double), round, 2))
   

### 2022-06-16 Experiment 2/scripts/run_exploratory_analyses.R

@@ -39,7 +39,8 @@ rm(list = ls())
       id,
       x_sjb = (sjb - mean(sjb))/sd(sjb),
       x_sup = (q5 - mean(q5))/sd(q5),
-      x_pol = (q19_1 - mean(q19_1))/sd(q19_1)
+      x_pol = (q19_1 - mean(q19_1))/sd(q19_1),
+      x_ide = (q16 - mean(q16, na.rm = T))/sd(q16, na.rm = T)
     ) %>% 
     left_join(dl, ., by = "id") %>% 
     mutate(

### 2022-08-16 Documents/manuscript.Rmd
first complete draft

@@
 In two preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police in the United States.
-We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position---but participants showed no ingroup bias when judging collective action by ingroup and outgroup members.
+We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position---but showed no ingroup bias when judging collective action by ingroup and outgroup members.
 Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
@@
 
-<!-- I should cite Colin Leach's work. -->
-
 Protest movements often spark contention about what actions are acceptable means of protest.
@@
 Indeed, Black athletes who had participated in the protests faced a backlash in public opinion [@lacina_nfl_2020].
+<!-- Likewise, 42\% of Americans thought that, in the 2020 Black Lives Matter protests, most protesters were trying to incite violence or destroy property, even though very few protests involved any violence or property destruction [@acled_year_2021].
+-->
 This and other examples suggest that *who* the protesters are and *what* they are protesting influences how acceptable observers judge their actions to be.
@@
 
-The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more acceptable if the protesters' are ingroup members or the protesters' cause aligns with the ingroup's interests (Hypothesis 1a, 1b) or if the protesters' cause aligns with their own system-justifying or system-challenging ideological orientation (Hypotheses 2a, 2b).
+The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more acceptable if the protesters' are ingroup members or the protesters' cause aligns with the ingroup's interests (*Hypothesis 1a*, *1b*) or if the protesters' cause aligns with their own system-justifying or system-challenging motives (*Hypotheses 2a*, *2b*).
 In Experiment 1, we use the instrument developed in Studies 1 and 2 to test for double standards in how acceptable participants consider collective action for workers' rights in the United Kingdom.
-By varying both the participants' and the protesters' group memberships (working/middle class), Experiment 1 compares the relative evidence for identity-based (Hypotheses 1a, 1b) and ideology-based double standards (Hypothesis 2a).
+By varying both the participants' and the protesters' group memberships (working/middle class), Experiment 1 compares the relative evidence for identity-based (*Hypothesis 1a*, *1b*) and ideology-based double standards (*Hypothesis 2a*).
 Extending Experiment 1, Experiment 2 considers both system-challenging and system-defending collective action in the context of the 2020 Black Lives Matter protests in the United States.
-By varying the protesters' cause (for/against defunding the police) as well as the participants' and the protesters' group memberships (Black/White), Experiment 2 compares the relative evidence identity-based (Hypotheses 1a, 1b) and ideology-based double standards (Hypothesis 2a, 2b).
+By varying the protesters' cause (for/against defunding the police) as well as the participants' and the protesters' group memberships (Black/White), Experiment 2 compares the relative evidence identity-based (*Hypothesis 1a*, *1b*) and ideology-based double standards (*Hypothesis 2a*, *2b*).
 Together, the preregistered experiments provide a comprehensive test of the hypothesized double standards in judging collective action.
@@
 
-By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that, in line with Hypothesis 1a and 1b, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights than if they were outgroup members protesting for the outgroup's rights.
-By comparing reactions to protesters with working-class or professional jobs, we tested how observers judge collective action by lower-status (i.e.
-working-class) and higher-status (professional/middle-class) group members.
-Assuming that collective action by lower-status group members is more threatening to the unequal status than collective action by higher-status group members, we tested the preregistered hypothesis that, in line with Hypothesis 2a, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group than if they were from a lower-status group.
-In non-preregistered analyses, we further explored whether, in line with Hypothesis 2b, participants would judge the same protest actions as more or less acceptable depending on their own ideological positions.
+By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that, in line with *Hypothesis 1a* and *1b*, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights than if they were outgroup members protesting for the outgroup's rights.
+By comparing reactions to working-class and middle-class protesters, we tested how observers judge collective action by lower-status (i.e.
+working-class) and higher-status (middle-class) group members.
+Assuming that collective action by lower-status group members is seen as more threatening to the unequal status quo than collective action by higher-status group members and that people are, to same extent, motivated to justify the system, we tested the preregistered hypothesis that, in line with *Hypothesis 2a*, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group than if they were from a lower-status group.
+In non-preregistered analyses, we further explored whether, in line with *Hypothesis 2b*, participants would judge the same protest actions as more or less acceptable depending on their own ideological positions.
 In this way, Experiment 1 provided a first test of the hypothesized identity- and ideology-based double standards in judging collective action.
@@
 
-We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations.^[[https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b](https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b)] We made all materials, data, and analysis scripts available online.^[[https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)] We followed sample size recommendations for item response theory models [@demars_irt_2010, p.
+We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations ([https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b](https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b)).
+We made all materials, data, and analysis scripts available online ([https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)).
+We followed sample size recommendations for item response theory models [@demars_irt_2010, p.
 36], planning to recruit 500 participants.
@@
 Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real.
-Participants in the lower-status protestprs condition read the following introduction:
+Participants in the lower-status protesters condition read the following introduction:
 
@@
 Overall, we found that participants' responses depended on both the protesters' and the participants' group status---but not in the directions predicted by our hypotheses.
-Contradicting Hypothesis 1, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = 0.00, [-0.20, 0.20]$).
+Contradicting our first hypothesis, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = 0.00, [-0.20, 0.20]$).
 Instead, we found that participants from higher-status backgrounds considered protest actions performed by both lower-status ($d = 0.34, [0.07, 0.59$]) and higher-status ($d = 0.33, [0.03, 0.63]$) protesters to be, on average, more acceptable than participants from lower-status backgrounds.
-Contradicting Hypothesis 2, participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.40, -0.00]$).
+Contradicting our second hypothesis, participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.40, -0.00]$).
 
@@
 
-\ldots
+Experiment 1 provided a first test of the hypothesized identity- and ideology-based double standards in judging collective action.
+That is, we tested whether, in line with *Hypothesis 1a* and *1b*, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights or whether, in line with *Hypothesis 2a*, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group.
+We found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the hypothesized directions.
+Instead, we found that both lower-status and higher-status participants tended to find the same protest actions more acceptable when lower-status protesters protested a bill threatening their rights.
+In non-preregistered analyses, we further found that people who were right-wing and endorsed system-justifying beliefs tended to find all protest actions to be less acceptable.
+
+Experiment 1 did not, however, provide a conclusive test of our hypotheses.
+First, by focusing on fictitious government bills, we might have chosen a scenario too far removed from current affairs to evoke strong ingroup bias in participants' judgements of protest actions.
+That said, participants' reactions to the experimental manipulation suggested that it evoked outrage and was understood in ingroup--outgroup terms.
+Second, by focusing on collective action to defend workers' rights, we arguably studied reactions to collective action for a progressive cause.
+To test for ideology-based double standards, we relied on the assumption that collective action by lower-status group members would be seen as more challenging to the prevailing system than collective action by higher-status group members.
+This assumption, however, was contradicted by our observation that participants who endorsed system-justifying beliefs tended to find actions by lower-status and higher-status protesters to be less acceptable.
+Experiment 2 addressed those limitations by focusing on a scenario we expected to evoke stronger responses from participants and by examining reactions to both system-challenging and system-defending collective action.
 
@@
 That is, we tested whether Black and White Americans applied different standards when judging collective action taken by either Black or White protesters to protest either for or against police divestment as a possible solution to racist police violence.
+We conducted this study in January 2021, a time when, after a year of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement.
+We focused on defunding the police as this position remained controversial among both Black and White Americans even as they broadly agreed on the need for police reform.^[In December 2020, only 34% of Black Americans and 18% of White Americans supported defunding the police while majorities of Black and White Americans supported police reforms such as eliminating qualified immunity (50% and 57%) or banning chokeholds [79% and 61%, @yougov_police_2020].]
 
-- Explain why defunding the police (public support for BLM, so needed to be more controversial)
-
-<!-- In Experiment 2, we tested five hypotheses.
-We tested two predictions derived from social identity theory.
-First, we hypothesize that observers are more likely to judge an action as an acceptable means of protest when the protesters are ingroup members than when they are outgroup members (Hypothesis 1a).
-Second, we hypothesize that observers are more likely to judge an action as an acceptable means of protest when the goal of the protests would benefit the observers' ingroup (Hypothesis 1b).
-We will test two predictions derived from system justification theory.
-First, we hypothesize that observers are more likely to judge an action as an acceptable means of protest when the goal of the protest is to support rather than challenge the system (Hypothesis 2a).
-Second, we hypothesize that observers who endorse system-justifying beliefs are more likely to consider system-supporting actions acceptable means of protest-and less likely to consider system-challenging actions acceptable means of protest-than observers who do not endorse system-justifying beliefs (Hypothesis 2b).
-In addition, we will test an alternative hypothesis.
-We hypothesize that observers are more likely to consider actions as acceptable means of protest when they support the cause of the protesters than when they oppose their cause (Hypothesis 3).
-We note that Hypotheses 1a, 1b, and 2a make competing predictions about differences between conditions while Hypotheses 2b and 3 need not contradict each other or any of the other hypotheses.
--->
+By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the protesters were ingroup members (*Hypothesis 1a*).
+By assuming that defunding the police aligns more closely with the interests of Black than White Americans, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the cause of the protest aligned with their ingroup's interests (*Hypothesis 1b*).
+By varying whether the protesters protested for or against defunding the police, we tested the preregistered hypotheses that participants would judge the same protest actions as more acceptable if the protest supported, rather than challenged, the system (*Hypothesis 2a*) and if they endorsed system-justifying beliefs and the protest supported the system or if they rejected system-justifying beliefs and the protest challenged the system (*Hypothesis 2b*).
+In this way, Experiment 2 provided a complete test of the hypothesized identity- and ideology-based double standards in judging collective action.
+In addition, we tested the simpler, alternative hypothesis that participants, in general, would judge the same protest actions as more acceptable if they supported the protesters' cause (*Hypothesis 3*).
 
@@
 
-We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations.^[[https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1](https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1)] We made all materials, data, and analysis scripts available online.^[[https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)] As reported in the preregistration, we ran simulations, using data from Experiment 1, to determine that a sample size of $N = 1,600$ ($n = 200$ per condition) was sufficient to detect even small differences between conditions ($0.09 < \text{Cohen's}~d < 0.16$).
+We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations ([https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1](https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1)).
+We made all materials, data, and analysis scripts available online ([https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)).
+As reported in the preregistration, we ran simulations, using data from Experiment 1, to determine that a sample size of $N = 1,600$ ($n = 200$ per condition) was sufficient to detect even small differences between conditions ($0.09 < \text{Cohen's}~d < 0.16$).
 
@@
 On average, participants tended to describe their political orientation as moderately liberal ($M = 2.92, \textit{SD} = 1.73$) but, as expected, were divided in their support for defunding the police ($M = 3.37, \textit{SD} = 1.42$).
+That is, 56% supported defunding the police, 30% opposed defunding the police, and 14% remained undecided.
 
@@
 Model 1 (M1) estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause.
-Contradicting Hypothesis 1a, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`).
-Contradicting Hypothesis 1b, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`).
-Contradicting Hypothesis 2a, participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`).
+Contradicting *Hypothesis 1a*, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`).
+Contradicting *Hypothesis 1b*, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`).
+Contradicting *Hypothesis 2a*, participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`).
 Instead, we found that, on average, White participants considered all protest actions to be more acceptable than Black participants (`r with(m1, d_text[contrast == "nh1_contrast"])`) and both Black (`r with(m1, d_text[contrast == "nh3_contrast"])`) and White (`r with(m1, d_text[contrast == "nh2_contrast"])`) participants considered the same actions to be more acceptable when protests were for, rather than against, defunding the police.
@@
 As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgements about system-defending protest actions and one estimating their effect on judgements about system-challenging protest actions, and one varying effect estimating its variance across conditions.
-Supporting Hypothesis 2b, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`).
-We did not, however, find evidence for the ideological symmetry implied by Hypothesis 2b since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).
+Supporting *Hypothesis 2b*, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`).
+We did not, however, find evidence for the ideological symmetry implied by *Hypothesis 2b* since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).
 
@@
 Participants who endorsed system-justifying beliefs tended to consider system-challenging and system-defending protest actions to be equally unacceptable.
-In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with Hypothesis 2b, they considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.
+In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with *Hypothesis 2b*, they considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.
 
@@
 We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police.
-As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Contradicting Hypothesis 3, participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
+As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Contradicting *Hypothesis 3*, participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
 Ergo, our results did not support the alternative hypothesis that ideology-based double standards can be reduced to support for, or opposition to, the cause of a protest.
@@
 ### Non-Preregistered Analyses
+
+```{r, include = FALSE}
+
+  # Import results
+  m5 <- read_rds("../Experiment 2/results/m5_post.rds")
+
+  # Define contrasts
+  m5 <- m5 %>% 
+    mutate(
+      h1a_contrast = case_when(
+        Participant == Protesters ~ +1L,
+        Participant != Protesters ~ -1L,
+      ),
+      h1b_contrast = case_when(
+        Participant == "Black" & Cause == "For" ~ +1L,
+        Participant == "White" & Cause == "Against" ~ +1L,
+        TRUE ~ -1L,
+      )
+    )
+
+  # Calculate contrasts
+  m5 <- m5 %>% 
+    group_by(.draw) %>% 
+    summarize(
+      across(
+        ends_with("_contrast"),
+        ~mean(z * .)
+      )
+    ) %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "d"
+    ) %>% 
+    group_by(contrast) %>% 
+    median_qi(d)
+
+  # Make text
+  m5 <- m5 %>% 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
+```
+
+We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses.
+To that end, we extended Model 1 by estimating participants' responses as a function of their *z*-standardized identification with their racial ingroup.
+As in Model 2, we modeled this relationship with a fixed and a varying effect.
+We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast"])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable  than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast"])`).
+Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgements about collective action by ingroup and outgroup members.
 
@@
 
-We reran Model 2 with political orientation, instead of system justification, as an alternative predictor variable.
-Our results mirrored results from the preregistered analyses: More liberal participants were more likely to consider protest actions for defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "for"])`) but more conservative participants were not more likely to consider protest actions against defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "against"])`).
+We also explored whether we would find ideology-based double standards when operationalizing ideology as political orientation instead of as system justification.
+To that end, we reran Model 2 with political orientation as the *z*-standardized predictor variable.
+Our results mirrored the preregistered analyses: More liberal participants were more likely to consider protest actions for defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "for"])`) but more conservative participants were not more likely to consider protest actions against defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "against"])`).
 As Figure 5 shows, conservative participants tended to consider all protest actions to be equally unacceptable while liberal participants considered the same protest actions to be more acceptable when protesters rallied around a progressive cause.
-Our non-preregistered analyses thus replicated the ideology-based double standards from the preregistered analyses with a different operationalization of political ideology.
+Our non-preregistered analyses thus replicated the ideology-based double standards from the preregistered analyses with a different operationalization of ideology.
 
@@
 
-\ldots
+Experiment 2 provided a complete test of the hypothesized identity- and ideology-based double standards in judging collective action.
+As in Experiment 1, we found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the hypothesized directions.
+We found that, contrary to *Hypothesis 1a*, participants considered protest actions taken by ingroup and outgroup members to be equally acceptable and that, contrary to *Hypothesis 1b*, participants did not consider the same protest actions to be more acceptable when the protesters' cause aligned with their ingroup's interests.
+Like Experiment 1, Experiment 2 thus found no evidence for identity-based double standards---even though it focused on an issue of direct and current relevance to the participants.
+
+Expanding Experiment 1, Experiment 2 considered reactions to both system-challenging and system-defending collective action and, in so doing, provided a stronger test of the hypothesized ideology-based double standards.
+Contrary to *Hypothesis 2a* and the underlying assumption of a universal system-justifying motive, we found that participants considered the same protest actions to be *more* acceptable when judging system-challenging collective action (for defunding the police) than when judging system-defending collective action.
+This finding might reflect our liberal-leaning sample.
+In line with *Hypothesis 2b*, we found that participants who rejected system-justifying beliefs considered the same protest action more acceptable when judging system-challenging collective action (for defunding the police).
+We did not, however, find evidence for ideological symmetry in this relationship as participants who endorsed system-justifying beliefs considered system-challenging and system-defending collective action to be equally unacceptable.
+In addition, we did not find evidence for the alternative, simpler *Hypothesis 3* that that participants would judge the same protest actions to be acceptable when they supported the protesters' cause.
+In this way, Experiment 2 provided evidence for ideology-based, but not for identity-based, double standards in judging collective action.
 
 # General Discussion
+
+Our research was motivated by real-world examples that seemed to suggest that where people draw the line between acceptable and unacceptable protest actions depends on *who* the protesters are and *what* they are protesting---in other words, that there are double standards in judging collective action.
+Our findings, however, showed that it is not so simple.
+Contradicting our hypothesis of identity-based double standards, participants in two preregistered experiments did not show ingroup bias when judging controversial protest actions taken by ingroup and outgroup members to advance their group's interests.
+Instead, supporting our hypothesis of ideology-based double standards, participants judged the same controversial protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position.
+<!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.
+-->
+In the remainder of this General Discussion, we discuss theoretical and practical implications for understanding the often divided response to collective action and consider some limiting conditions on our findings.
+
+## Implications
+
+<!-- Our findings raised the question why we did not find double standards in judging collective action, even though there are well-documents double standards in other domains. -->
+
+<!-- 1. What does it mean that we did not find identity-based double standards? -->
+<!-- We do not find ingroup bias, at least not for broad social categories such as race and class, in contrast to evidence and predictions from social identity theory.
+Protest actions are not the same as moral transgressions, perhaps.
+We might have reached a different conclusion when adopting the framework of opinion-based group memberships and understanding political ideology in partisan identity terms.
+-->
+<!-- 2. What does it mean that we did not find symmetrical ideology-based double standards? -->
+<!-- It could just be that conservatives don't like protest.
+But consider Jan.
+6 --- maybe it's about the cause of the protest?
+We know about the distinct moral values of conservatives and liberals (see MFT, see MFT and political orientation) and how they shape attitudes toward policy issues (see MFT and culture war issues).
+This aligns with Verkuyten and Richardson's finding of morality and tolerance judgements.
+-->
+
+Our results showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
+Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
+In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.
+By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provided a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
+
+After all, this distinction can be consequential.
+If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
+In this way, our findings could explain why 45% of Republicans (2% of Democrats) supported the storming of the U.S.
+Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
+If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
+In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].
+For those reasons, researchers should further investigate when and why people apply different standards when judging collective action for various causes.
+
+## Limitations
+
+Our research applied item response theory to judgements about collective action to investigate whether where people draw the line between acceptable and unacceptable means of protest depends on who the protesters are and what they are protesting.
+In so doing, we moved the distinction between normative and non-normative collective action from the realm of the researcher's intuition into the realm of scientific investigation, allowing us to test for double standards in judging collective action.
+
+Despite its strengths, our research is qualified by several limitations.
+First, our research examined the hypothesized relationships for only two political causes---protecting workers' rights, defunding the police---and thus provided only limited evidence that our findings generalize beyond those causes.
+Relying on few stimuli to establish an effect threatens the validity and replicability of research findings and severely constrains psychological research generalize beyond those few stimuli [@yarkoni_generalizability_2022].
+Future research should sample a wider range of causes to address this pervasive but often ignored problem [@judd_treating_2012].
+This is particularly important when studying collective action as recent theorizing [@jost_missing_2017] and research [@osborne_protesting_2019] highlighted the importance of differentiating between progressive and conservative causes.
+Second, our research was based on two Western, Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al., 2010) samples, which severely limits the cross-cultural generalizability of our findings.
+Future research should include samples from a wider range of cultural context to examine how the observed relationships vary across cultural contexts.
+
+# Conclusion
+
+Opinion polls show deep divides not only about the ends of protests, but also about the means of protest.
+For example, only 11\% of Republicans and 29\% of White Americans---compared to 59\% of Democrats and 66\% of Black Americans---considered kneeling during the national anthem---a non-violent form of collective action---an appropriate means to protest racist police violence [@yougov_nfl_2017].
+Our research showed that, in contrast to well-documented double standards in other domains, people do not apply a different standard when judging collective action by ingroup members or for a cause aligned with the ingroup's interests.
+Instead, our research demonstrated that people consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological positions---in other words, that there are partisan double standards in judging collective action.
+
+<!-- In two preregistered experiments, we showed that, unlike judgements in other domains, judgements what are acceptable means of protest are not subject to double standards based on group identities (such as class and race) but instead depend on whether the protesters' cause aligns with one's ideological orientation.
+-->
+
+<!-- Our research sought to understand the often divided response to protest movements.
+Reactions to Black athletes kneeling to protest racist police violence seem to suggest that Americans are divided not only about the ends of protests, but also about the means of protest.
+-->
+
+<!-- Contradicting research showing that people sometimes apply a more lenient standard when judging moral transgressions by ingroup members [e.g., @abrams_double_2013; @valdesolo_moral_2007], ...
+that who the protesters are really does not matter all that much.
+-->
 
@@
 
-- Summary: No evidence for identity-based double standards, but some evidence for ideology-based double standards
-- Reject the alternative hypothesis that there is consensus about what is an acceptable means of protest.
-- Relate our findings to @verkuyten_intolerance_2022 
-- Two plausible interpretations: (1) conservatives oppose controversial protest (no matter the end) *or* (2) the end justifies the mean
-- Re (1) it's plausible but what about recent events? (see public polling from grant)
-- Re (2) why is there no asymmetry?
-- Outline the-end-justifies-the-mean idea, highlighting that just supporting a cause is not enough, it needs moralization
-- Discuss why high-status group are more accepting of controversial protest actions
-- Discuss limitations (two cultures, two issues, liberal samples; system-defending vs.
-system-supporting actions; what about other conservative causes?; ingroup bias might need zero-sum situation to succeed)
-- Invite researchers to use our paradigm
+<!-- Examples strongly suggest that who's protesting and why shapes where people draw the line ...
+However, public opinion polling cannot tell us whether the underlying psychological mechanism is one of identity or ideology.
+In the present research, we developed a systematic approach to ...
+and, across two experiments, test for ...
+double standards.
+Contrary to prior theorizing and research, we found no evidence for ingroup bias...
+Instead, we found ...
+-->
+
+<!-- I should discuss how our findings could explain why people are so ready to perceive BLM protests as non-normative, in contrast to the evidence, because they applied a stricter standard when judging protests for a progressive cause.
+-->
+
+<!-- 1. Summary of Findings -->
+<!-- 3. What does it mean that we did not find identity-based double standards? -->
+<!-- We do not find ingroup bias, at least not for broad social categories such as race and class, in contrast to evidence and predictions from social identity theory.
+Protest actions are not the same as moral transgressions, perhaps.
+We might have reached a different conclusion when adopting the framework of opinion-based group memberships and understanding political ideology in partisan identity terms.
+-->
+<!-- 4. What does it mean that we did not  find symmetrical ideology-based double standards? -->
+<!-- It could just be that conservatives don't like protest.
+But consider Jan.
+6 --- maybe it's about the cause of the protest?
+We know about the distinct moral values of conservatives and liberals (see MFT, see MFT and political orientation) and how they shape attitudes toward policy issues (see MFT and culture war issues).
+This aligns with Verkuyten and Richardson's finding of morality and tolerance judgements.
+-->
+<!-- 2. Is the distinction between normative and non-normative collective action in the eye of the beholder? -->
+<!-- Relate to "protect framework" (e.g., Leach et al.).
+Question asssumption that there is a universal standard for acceptable protest actions.
+Relate to legislative issues such as protest bans/NFL's decision to ban certain forms of political expression.
+-->
+<!-- 5. What are the limitations of the present research? -->
+<!-- 6. Conclusion -->
+
+<!-- - Summary: No evidence for identity-based double standards, but some evidence for ideology-based double standards -->
+<!-- - Reject the alternative hypothesis that there is consensus about what is an acceptable means of protest. -->
+<!-- - Relate our findings to @verkuyten_intolerance_2022  -->
+<!-- - Two plausible interpretations: (1) conservatives oppose controversial protest (no matter the end) *or* (2) the end justifies the mean -->
+<!-- - Re (1) it's plausible but what about recent events? (see public polling from grant) -->
+<!-- - Re (2) why is there no asymmetry? -->
+<!-- - Outline the-end-justifies-the-mean idea, highlighting that just supporting a cause is not enough, it needs moralization -->
+<!-- - Discuss why high-status group are more accepting of controversial protest actions -->
+<!-- - Discuss limitations (two cultures, two issues, liberal samples; system-defending vs.
+system-supporting actions; what about other conservative causes?; ingroup bias might need zero-sum situation to succeed) -->
+<!-- - Invite researchers to use our paradigm -->
 

### 2022-08-16 Documents/manuscript.tex
first complete draft

@@
 
-%\date{June 15, 2021}
+\date{July 15, 2022}
 
@@
 acceptable when the protesters' cause aligned with their own
-(system-challenging) ideological position---but participants showed no
-ingroup bias when judging collective action by ingroup and outgroup
-members. Our findings have theoretical and practical implications for
-understanding the often divided response to prominent social movements.
+(system-challenging) ideological position---but showed no ingroup bias
+when judging collective action by ingroup and outgroup members. Our
+findings have theoretical and practical implications for understanding
+the often divided response to prominent social movements.
     \end{abstract}
@@
 members or the protesters' cause aligns with the ingroup's interests
-(Hypothesis 1a, 1b) or if the protesters' cause aligns with their own
-system-justifying or system-challenging ideological orientation
-(Hypotheses 2a, 2b). In Experiment 1, we use the instrument developed in
-Studies 1 and 2 to test for double standards in how acceptable
-participants consider collective action for workers' rights in the
-United Kingdom. By varying both the participants' and the protesters'
-group memberships (working/middle class), Experiment 1 compares the
-relative evidence for identity-based (Hypotheses 1a, 1b) and
-ideology-based double standards (Hypothesis 2a). Extending Experiment 1,
-Experiment 2 considers both system-challenging and system-defending
-collective action in the context of the 2020 Black Lives Matter protests
-in the United States. By varying the protesters' cause (for/against
-defunding the police) as well as the participants' and the protesters'
-group memberships (Black/White), Experiment 2 compares the relative
-evidence identity-based (Hypotheses 1a, 1b) and ideology-based double
-standards (Hypothesis 2a, 2b). Together, the preregistered experiments
-provide a comprehensive test of the hypothesized double standards in
-judging collective action.
+(\emph{Hypothesis 1a}, \emph{1b}) or if the protesters' cause aligns
+with their own system-justifying or system-challenging motives
+(\emph{Hypotheses 2a}, \emph{2b}). In Experiment 1, we use the
+instrument developed in Studies 1 and 2 to test for double standards in
+how acceptable participants consider collective action for workers'
+rights in the United Kingdom. By varying both the participants' and the
+protesters' group memberships (working/middle class), Experiment 1
+compares the relative evidence for identity-based (\emph{Hypothesis 1a},
+\emph{1b}) and ideology-based double standards (\emph{Hypothesis 2a}).
+Extending Experiment 1, Experiment 2 considers both system-challenging
+and system-defending collective action in the context of the 2020 Black
+Lives Matter protests in the United States. By varying the protesters'
+cause (for/against defunding the police) as well as the participants'
+and the protesters' group memberships (Black/White), Experiment 2
+compares the relative evidence identity-based (\emph{Hypothesis 1a},
+\emph{1b}) and ideology-based double standards (\emph{Hypothesis 2a},
+\emph{2b}). Together, the preregistered experiments provide a
+comprehensive test of the hypothesized double standards in judging
+collective action.
 
@@
 By varying the participants' and the protesters' group memberships, we
-tested the preregistered hypothesis that, in line with Hypothesis 1a and
-1b, participants would judge the same protest actions as more acceptable
-if the protesters were ingroup members protesting for the ingroup's
-rights than if they were outgroup members protesting for the outgroup's
-rights. By comparing reactions to protesters with working-class or
-professional jobs, we tested how observers judge collective action by
-lower-status (i.e.~working-class) and higher-status
-(professional/middle-class) group members. Assuming that collective
-action by lower-status group members is more threatening to the unequal
-status than collective action by higher-status group members, we tested
-the preregistered hypothesis that, in line with Hypothesis 2a,
+tested the preregistered hypothesis that, in line with \emph{Hypothesis
+1a} and \emph{1b}, participants would judge the same protest actions as
+more acceptable if the protesters were ingroup members protesting for
+the ingroup's rights than if they were outgroup members protesting for
+the outgroup's rights. By comparing reactions to working-class and
+middle-class protesters, we tested how observers judge collective action
+by lower-status (i.e.~working-class) and higher-status (middle-class)
+group members. Assuming that collective action by lower-status group
+members is seen as more threatening to the unequal status quo than
+collective action by higher-status group members and that people are, to
+same extent, motivated to justify the system, we tested the
+preregistered hypothesis that, in line with \emph{Hypothesis 2a},
 participants would judge the same protest actions as more acceptable if
@@
 lower-status group. In non-preregistered analyses, we further explored
-whether, in line with Hypothesis 2b, participants would judge the same
-protest actions as more or less acceptable depending on their own
+whether, in line with \emph{Hypothesis 2b}, participants would judge the
+same protest actions as more or less acceptable depending on their own
 ideological positions. In this way, Experiment 1 provided a first test
@@
 We preregistered the sample size as well as all hypotheses,
-inclusion/exclusion criteria, measures, and manipulations.\footnote{\url{https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b}}
-We made all materials, data, and analysis scripts available
-online.\footnote{\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}}
+inclusion/exclusion criteria, measures, and manipulations
+(\url{https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b}).
+We made all materials, data, and analysis scripts available online
+(\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}).
 We followed sample size recommendations for item response theory models
@@
 carefully read the vignette and to try to imagine what it would be like
-if this situation was real. Participants in the lower-status protestprs
+if this situation was real. Participants in the lower-status protesters
 condition read the following introduction:
@@
 participants' group status---but not in the directions predicted by our
-hypotheses. Contradicting Hypothesis 1, participants did not consider
-protest actions performed by their ingroup to be more acceptable, on
-average, than the same actions performed by the relevant outgroup
-(Cohen's \(d = 0.00, [-0.20, 0.20]\)). Instead, we found that
+hypotheses. Contradicting our first hypothesis, participants did not
+consider protest actions performed by their ingroup to be more
+acceptable, on average, than the same actions performed by the relevant
+outgroup (Cohen's \(d = 0.00, [-0.20, 0.20]\)). Instead, we found that
 participants from higher-status backgrounds considered protest actions
@@
 more acceptable than participants from lower-status backgrounds.
-Contradicting Hypothesis 2, participants considered protest actions
-performed by higher-status protesters to be less acceptable, on average,
-than the same actions performed by lower-status protesters
+Contradicting our second hypothesis, participants considered protest
+actions performed by higher-status protesters to be less acceptable, on
+average, than the same actions performed by lower-status protesters
 (\(d = -0.20, [-0.40, -0.00]\)).
@@
 
-\ldots
+Experiment 1 provided a first test of the hypothesized identity- and
+ideology-based double standards in judging collective action. That is,
+we tested whether, in line with \emph{Hypothesis 1a} and \emph{1b},
+participants would judge the same protest actions as more acceptable if
+the protesters were ingroup members protesting for the ingroup's rights
+or whether, in line with \emph{Hypothesis 2a}, participants would judge
+the same protest actions as more acceptable if the protesters were from
+a higher-status group. We found that participants' responses depended on
+both the participants' and the protesters' group memberships---but not
+in the hypothesized directions. Instead, we found that both lower-status
+and higher-status participants tended to find the same protest actions
+more acceptable when lower-status protesters protested a bill
+threatening their rights. In non-preregistered analyses, we further
+found that people who were right-wing and endorsed system-justifying
+beliefs tended to find all protest actions to be less acceptable.
+
+Experiment 1 did not, however, provide a conclusive test of our
+hypotheses. First, by focusing on fictitious government bills, we might
+have chosen a scenario too far removed from current affairs to evoke
+strong ingroup bias in participants' judgements of protest actions. That
+said, participants' reactions to the experimental manipulation suggested
+that it evoked outrage and was understood in ingroup--outgroup terms.
+Second, by focusing on collective action to defend workers' rights, we
+arguably studied reactions to collective action for a progressive cause.
+To test for ideology-based double standards, we relied on the assumption
+that collective action by lower-status group members would be seen as
+more challenging to the prevailing system than collective action by
+higher-status group members. This assumption, however, was contradicted
+by our observation that participants who endorsed system-justifying
+beliefs tended to find actions by lower-status and higher-status
+protesters to be less acceptable. Experiment 2 addressed those
+limitations by focusing on a scenario we expected to evoke stronger
+responses from participants and by examining reactions to both
+system-challenging and system-defending collective action.
 
@@
 or White protesters to protest either for or against police divestment
-as a possible solution to racist police violence.
-
-\begin{itemize}
-\tightlist
-\item
-  Explain why defunding the police (public support for BLM, so needed to
-  be more controversial)
-\end{itemize}
+as a possible solution to racist police violence. We conducted this
+study in January 2021, a time when, after a year of unprecedented
+protests for racial justice, most Americans could be expected to be
+aware of, and have formed an opinion on, the Black Lives Matter
+movement. We focused on defunding the police as this position remained
+controversial among both Black and White Americans even as they broadly
+agreed on the need for police reform.\footnote{In December 2020, only
+  34\% of Black Americans and 18\% of White Americans supported
+  defunding the police while majorities of Black and White Americans
+  supported police reforms such as eliminating qualified immunity (50\%
+  and 57\%) or banning chokeholds (79\% and 61\%, YouGov, 2020).}
+
+By varying the participants' and the protesters' group memberships, we
+tested the preregistered hypothesis that participants would judge the
+same protest actions as more acceptable if the protesters were ingroup
+members (\emph{Hypothesis 1a}). By assuming that defunding the police
+aligns more closely with the interests of Black than White Americans, we
+tested the preregistered hypothesis that participants would judge the
+same protest actions as more acceptable if the cause of the protest
+aligned with their ingroup's interests (\emph{Hypothesis 1b}). By
+varying whether the protesters protested for or against defunding the
+police, we tested the preregistered hypotheses that participants would
+judge the same protest actions as more acceptable if the protest
+supported, rather than challenged, the system (\emph{Hypothesis 2a}) and
+if they endorsed system-justifying beliefs and the protest supported the
+system or if they rejected system-justifying beliefs and the protest
+challenged the system (\emph{Hypothesis 2b}). In this way, Experiment 2
+provided a complete test of the hypothesized identity- and
+ideology-based double standards in judging collective action. In
+addition, we tested the simpler, alternative hypothesis that
+participants, in general, would judge the same protest actions as more
+acceptable if they supported the protesters' cause (\emph{Hypothesis
+3}).
 
@@
 We preregistered the sample size as well as all hypotheses,
-inclusion/exclusion criteria, measures, and manipulations.\footnote{\url{https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1}}
-We made all materials, data, and analysis scripts available
-online.\footnote{\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}}
+inclusion/exclusion criteria, measures, and manipulations
+(\url{https://osf.io/skxjt/?view_only=d8ce44a700884b5ab36b64ef08f833a1}).
+We made all materials, data, and analysis scripts available online
+(\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}).
 As reported in the preregistration, we ran simulations, using data from
@@
 were divided in their support for defunding the police
-(\(M = 3.37, \textit{SD} = 1.42\)).
+(\(M = 3.37, \textit{SD} = 1.42\)). That is, 56\% supported defunding
+the police, 30\% opposed defunding the police, and 14\% remained
+undecided.
 
@@
 responses depended on their own group membership, the protesters' group
-membership, or the protesters' cause. Contradicting Hypothesis 1a,
-participants did not consider protest actions performed by their ingroup
-to be more acceptable, on average, than the same actions performed by
-the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)). Contradicting
-Hypothesis 1b, participants did not consider protest actions for a cause
-that was nominally aligned with their ingroup's interests to be more
-acceptable, on average, than protest actions for a cause not aligned
-with their ingroup's interests (\(d = -0.01, [-0.06, 0.04]\)).
-Contradicting Hypothesis 2a, participants did not consider protest
-actions for a system-defending cause to be more acceptable, on average,
-than protest actions for a system-challenging cause
-(\(d = -0.14, [-0.20, -0.09]\)). Instead, we found that, on average,
-White participants considered all protest actions to be more acceptable
-than Black participants (\(d = 0.09, [0.04, 0.14]\)) and both Black
-(\(d = 0.07, [0.03, 0.10]\)) and White (\(d = 0.08, [0.04, 0.11]\))
-participants considered the same actions to be more acceptable when
-protests were for, rather than against, defunding the police. As in
-Experiment 1, we thus found that participants' responses depended on
-both the participants' and the protesters' group memberships---but not
-in the directions predicted by our hypotheses.
+membership, or the protesters' cause. Contradicting \emph{Hypothesis
+1a}, participants did not consider protest actions performed by their
+ingroup to be more acceptable, on average, than the same actions
+performed by the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)).
+Contradicting \emph{Hypothesis 1b}, participants did not consider
+protest actions for a cause that was nominally aligned with their
+ingroup's interests to be more acceptable, on average, than protest
+actions for a cause not aligned with their ingroup's interests
+(\(d = -0.01, [-0.06, 0.04]\)). Contradicting \emph{Hypothesis 2a},
+participants did not consider protest actions for a system-defending
+cause to be more acceptable, on average, than protest actions for a
+system-challenging cause (\(d = -0.14, [-0.20, -0.09]\)). Instead, we
+found that, on average, White participants considered all protest
+actions to be more acceptable than Black participants
+(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.07, [0.03, 0.10]\))
+and White (\(d = 0.08, [0.04, 0.11]\)) participants considered the same
+actions to be more acceptable when protests were for, rather than
+against, defunding the police. As in Experiment 1, we thus found that
+participants' responses depended on both the participants' and the
+protesters' group memberships---but not in the directions predicted by
+our hypotheses.
 
@@
 one varying effect estimating its variance across conditions. Supporting
-Hypothesis 2b, participants who \emph{rejected} system-justifying
+\emph{Hypothesis 2b}, participants who \emph{rejected} system-justifying
 beliefs were \emph{more} likely to consider system-challenging protest
@@
 (\(\beta_{xy} = 0.38, [0.16, 0.58]\)). We did not, however, find
-evidence for the ideological symmetry implied by Hypothesis 2b since
-participants who \emph{endorsed} system-justifying beliefs were
+evidence for the ideological symmetry implied by \emph{Hypothesis 2b}
+since participants who \emph{endorsed} system-justifying beliefs were
 \emph{not} more likely to consider system-defending protest actions
@@
 contrast, participants who rejected system-justifying beliefs evinced
-ideology-based double standards: In line with Hypothesis 2b, they
+ideology-based double standards: In line with \emph{Hypothesis 2b}, they
 considered the same protest actions to be more acceptable when the
@@
   each of the four pairs of predictor categories.} Contradicting
-Hypothesis 3, participants who supported the protesters' cause did not,
-on average, consider the same protest actions to be more acceptable than
-participants who opposed the protesters' cause
+\emph{Hypothesis 3}, participants who supported the protesters' cause
+did not, on average, consider the same protest actions to be more
+acceptable than participants who opposed the protesters' cause
 (\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our results did not support
@@
 
-We reran Model 2 with political orientation, instead of system
-justification, as an alternative predictor variable. Our results
-mirrored results from the preregistered analyses: More liberal
-participants were more likely to consider protest actions for defunding
-the police to be acceptable (\(\beta_{xy} = 0.41, [0.17, 0.62]\)) but
-more conservative participants were not more likely to consider protest
-actions against defunding the police to be acceptable
+We explored whether group identification moderated how the participants'
+and the protesters' group memberships affected participants' responses.
+To that end, we extended Model 1 by estimating participants' responses
+as a function of their \emph{z}-standardized identification with their
+racial ingroup. As in Model 2, we modeled this relationship with a fixed
+and a varying effect. We found, however, that even participants who
+strongly identified with their racial ingroup (+1\emph{SD}) did not
+consider protest actions performed by their ingroup to be more
+acceptable than the same actions performed by the relevant outgroup
+(\(d = 0.02, [-0.04, 0.07]\)) and did not consider protest actions for a
+cause that was aligned with their ingroup's interests to be more
+acceptable than protest actions for a cause not aligned with their
+ingroup's interests (\(d = -0.03, [-0.09, 0.02]\)). Our
+non-preregistered analyses thus suggested that group identification did
+not moderate group differences in judgements about collective action by
+ingroup and outgroup members.
+
+We also explored whether we would find ideology-based double standards
+when operationalizing ideology as political orientation instead of as
+system justification. To that end, we reran Model 2 with political
+orientation as the \emph{z}-standardized predictor variable. Our results
+mirrored the preregistered analyses: More liberal participants were more
+likely to consider protest actions for defunding the police to be
+acceptable (\(\beta_{xy} = 0.41, [0.17, 0.62]\)) but more conservative
+participants were not more likely to consider protest actions against
+defunding the police to be acceptable
 (\(\beta_{xy} = -0.16, [-0.41, 0.04]\)). As Figure 5 shows, conservative
@@
 ideology-based double standards from the preregistered analyses with a
-different operationalization of political ideology.
+different operationalization of ideology.
 
@@
 
-\ldots
+Experiment 2 provided a complete test of the hypothesized identity- and
+ideology-based double standards in judging collective action. As in
+Experiment 1, we found that participants' responses depended on both the
+participants' and the protesters' group memberships---but not in the
+hypothesized directions. We found that, contrary to \emph{Hypothesis
+1a}, participants considered protest actions taken by ingroup and
+outgroup members to be equally acceptable and that, contrary to
+\emph{Hypothesis 1b}, participants did not consider the same protest
+actions to be more acceptable when the protesters' cause aligned with
+their ingroup's interests. Like Experiment 1, Experiment 2 thus found no
+evidence for identity-based double standards---even though it focused on
+an issue of direct and current relevance to the participants.
+
+Expanding Experiment 1, Experiment 2 considered reactions to both
+system-challenging and system-defending collective action and, in so
+doing, provided a stronger test of the hypothesized ideology-based
+double standards. Contrary to \emph{Hypothesis 2a} and the underlying
+assumption of a universal system-justifying motive, we found that
+participants considered the same protest actions to be \emph{more}
+acceptable when judging system-challenging collective action (for
+defunding the police) than when judging system-defending collective
+action. This finding might reflect our liberal-leaning sample. In line
+with \emph{Hypothesis 2b}, we found that participants who rejected
+system-justifying beliefs considered the same protest action more
+acceptable when judging system-challenging collective action (for
+defunding the police). We did not, however, find evidence for
+ideological symmetry in this relationship as participants who endorsed
+system-justifying beliefs considered system-challenging and
+system-defending collective action to be equally unacceptable. In
+addition, we did not find evidence for the alternative, simpler
+\emph{Hypothesis 3} that that participants would judge the same protest
+actions to be acceptable when they supported the protesters' cause. In
+this way, Experiment 2 provided evidence for ideology-based, but not for
+identity-based, double standards in judging collective action.
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Summary: No evidence for identity-based double standards, but some
-  evidence for ideology-based double standards
-\item
-  Reject the alternative hypothesis that there is consensus about what
-  is an acceptable means of protest.
-\item
-  Relate our findings to Verkuyten et al. (2022)
-\item
-  Two plausible interpretations: (1) conservatives oppose controversial
-  protest (no matter the end) \emph{or} (2) the end justifies the mean
-\item
-  Re (1) it's plausible but what about recent events? (see public
-  polling from grant)
-\item
-  Re (2) why is there no asymmetry?
-\item
-  Outline the-end-justifies-the-mean idea, highlighting that just
-  supporting a cause is not enough, it needs moralization
-\item
-  Discuss why high-status group are more accepting of controversial
-  protest actions
-\item
-  Discuss limitations (two cultures, two issues, liberal samples;
-  system-defending vs.~system-supporting actions; what about other
-  conservative causes?; ingroup bias might need zero-sum situation to
-  succeed)
-\item
-  Invite researchers to use our paradigm
-\end{itemize}
+Our research was motivated by real-world examples that seemed to suggest
+that where people draw the line between acceptable and unacceptable
+protest actions depends on \emph{who} the protesters are and \emph{what}
+they are protesting---in other words, that there are double standards in
+judging collective action. Our findings, however, showed that it is not
+so simple. Contradicting our hypothesis of identity-based double
+standards, participants in two preregistered experiments did not show
+ingroup bias when judging controversial protest actions taken by ingroup
+and outgroup members to advance their group's interests. Instead,
+supporting our hypothesis of ideology-based double standards,
+participants judged the same controversial protest actions to be more
+acceptable when the protesters' cause aligned with their own
+(system-challenging) ideological position. In the remainder of this
+General Discussion, we discuss theoretical and practical implications
+for understanding the often divided response to collective action and
+consider some limiting conditions on our findings.
+
+\hypertarget{implications}{%
+\subsection{Implications}\label{implications}}
+
+Our results showed that, for many controversial protest actions, there
+is no consensus on what is and is not an acceptable means to advance a
+cause. Instead, this judgement seems to depend on what the cause of a
+protest is and how it aligns with one's beliefs and values. In this way,
+our results confirm our contention that the distinction between
+normative and non-normative forms of collective action should be
+situated in the eye of the beholder. By developing an instrument for
+measuring where people draw the line between acceptable and unacceptable
+forms of collective action, we provided a paradigm that we hope will
+stimulate research on when and why people differ in making this
+distinction.
+
+After all, this distinction can be consequential. If people believe that
+even extreme actions are acceptable means to advance a cause they
+support, especially if they moralize their support for that cause
+(Mooijman et al., 2018), they might support violent extremism. In this
+way, our findings could explain why 45\% of Republicans (2\% of
+Democrats) supported the storming of the U.S. Capitol on January 6, 2021
+(YouGov, 2021) even though it was illegal, destructive, and violent. If
+people believe that even non-violent actions are unacceptable means to
+advance a cause they oppose, they might support stifling dissent. In
+this way, our findings could explain why Republicans responded to Black
+Lives Matter protests by seeking to further criminalize disruptive
+protest (e.g., blocking highways, Quinton, 2021) that might be most
+effective at challenging injustice (Shuman et al., 2022, 2021). For
+those reasons, researchers should further investigate when and why
+people apply different standards when judging collective action for
+various causes.
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+Our research applied item response theory to judgements about collective
+action to investigate whether where people draw the line between
+acceptable and unacceptable means of protest depends on who the
+protesters are and what they are protesting. In so doing, we moved the
+distinction between normative and non-normative collective action from
+the realm of the researcher's intuition into the realm of scientific
+investigation, allowing us to test for double standards in judging
+collective action.
+
+Despite its strengths, our research is qualified by several limitations.
+First, our research examined the hypothesized relationships for only two
+political causes---protecting workers' rights, defunding the
+police---and thus provided only limited evidence that our findings
+generalize beyond those causes. Relying on few stimuli to establish an
+effect threatens the validity and replicability of research findings and
+severely constrains psychological research generalize beyond those few
+stimuli (Yarkoni, 2022). Future research should sample a wider range of
+causes to address this pervasive but often ignored problem (Judd et al.,
+2012). This is particularly important when studying collective action as
+recent theorizing (Jost et al., 2017) and research (Osborne et al.,
+2019) highlighted the importance of differentiating between progressive
+and conservative causes. Second, our research was based on two Western,
+Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al.,
+2010) samples, which severely limits the cross-cultural generalizability
+of our findings. Future research should include samples from a wider
+range of cultural context to examine how the observed relationships vary
+across cultural contexts.
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+Opinion polls show deep divides not only about the ends of protests, but
+also about the means of protest. For example, only 11\% of Republicans
+and 29\% of White Americans---compared to 59\% of Democrats and 66\% of
+Black Americans---considered kneeling during the national anthem---a
+non-violent form of collective action---an appropriate means to protest
+racist police violence (YouGov, 2017). Our research showed that, in
+contrast to well-documented double standards in other domains, people do
+not apply a different standard when judging collective action by ingroup
+members or for a cause aligned with the ingroup's interests. Instead,
+our research demonstrated that people consider the same protest actions
+to be more acceptable when the cause of the protest aligns with their
+own ideological positions---in other words, that there are partisan
+double standards in judging collective action.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-judd_treating_2012}{}}%
+Judd, C. M., Westfall, J., \& Kenny, D. A. (2012). Treating stimuli as a
+random factor in social psychology: {A} new and comprehensive solution
+to a pervasive but largely ignored problem. \emph{Journal of Personality
+and Social Psychology}, \emph{103}(1), 54--69.
+\url{https://doi.org/10.1037/a0028347}
+
 \leavevmode\vadjust pre{\hypertarget{ref-kay_complementary_2003}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mooijman_moralization_2018}{}}%
+Mooijman, M., Hoover, J., Lin, Y., Ji, H., \& Dehghani, M. (2018).
+Moralization in social networks and the emergence of violence during
+protests. \emph{Nature Human Behaviour}, \emph{2}(6), 389--396.
+\url{https://doi.org/10.1038/s41562-018-0353-0}
+
 \leavevmode\vadjust pre{\hypertarget{ref-osborne_protesting_2019}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pew_republicans_2021}{}}%
+Quinton, S. (2021). Republicans respond to {Black Lives Matter} with
+anti-protest bills. \emph{Pew Charitable Trusts}.
+\url{https://www.pewtrusts.org/en/research-and-analysis/blogs/stateline/2021/02/04/republicans-respond-to-black-lives-matter-with-anti-protest-bills}
+
 \leavevmode\vadjust pre{\hypertarget{ref-reimer_identity_2022}{}}%
@@
 Sargent.
+
+\leavevmode\vadjust pre{\hypertarget{ref-shuman_protest_2022}{}}%
+Shuman, E., Hasan-Aslih, S., Zomeren, M. van, Saguy, T., \& Halperin, E.
+(2022). Protest movements involving limited violence can sometimes be
+effective: Evidence from the 2020 {Black Lives Matter} protests.
+\emph{Proceedings of the National Academy of Sciences}, \emph{119}(14),
+e2118990119. \url{https://doi.org/10.1073/pnas.2118990119}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-yarkoni_generalizability_2022}{}}%
+Yarkoni, T. (2022). The generalizability crisis. \emph{Behavioral and
+Brain Sciences}, \emph{45}, e1.
+\url{https://doi.org/10.1017/S0140525X20001685}
+
 \leavevmode\vadjust pre{\hypertarget{ref-yougov_nfl_2017}{}}%
@@
 set{]}. \url{https://perma.cc/3Q3V-4GC4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-yougov_police_2020}{}}%
+YouGov. (2020). \emph{{HuffPost}: Police reform {(Dec. 3--6, 2020)}}
+{[}Data set{]}. \url{https://perma.cc/JH5P-3W3K}
+
+\leavevmode\vadjust pre{\hypertarget{ref-yougov_capitol_2021}{}}%
+YouGov. (2021). \emph{Capitol protest {(Jan. 6, 2021)}} {[}Data set{]}.
+\url{https://perma.cc/YG99-EUYX}
 

### 2022-08-16 Experiment 2/scripts/run_exploratory_analyses.R

@@ -13,6 +13,7 @@ rm(list = ls())
   
   # Load packages
   library(tidyverse)
+  library(tidybayes)
   library(cmdstanr)
   library(posterior)
   
@@ -95,6 +96,41 @@ rm(list = ls())
   m4_fit$save_object("Experiment 2/results/m4_fit.rds")
   
 
+# Estimate (Model 5) ------------------------------------------------------
+  
+  # Compile data list
+  m5_dlist <- dl %>% 
+    drop_na(x_ide) %>% 
+    mutate(jj = as.integer(factor(id))) %>% 
+    with(., list(
+      N = length(ii),
+      I = max(ii),
+      J = max(jj),
+      K = max(kk),
+      ii = ii,
+      jj = jj,
+      kk = kk,
+      y = y,
+      x_ide = x_ide
+    ))
+  
+  # Compile model
+  m5_model <- cmdstan_model("Experiment 2/models/s2_m5_2pl_model.stan")
+  
+  # Run model
+  m5_fit <- m5_model$sample(
+    data = m5_dlist,
+    seed = 2369044,
+    chains = 8,
+    parallel_chains = parallel::detectCores(),
+    iter_sampling = 500,
+    iter_warmup = 1000
+  )
+  
+  # Save results as .rds
+  m5_fit$save_object("Experiment 2/results/m5_fit.rds")
+  
+
 # Process (Model 4) -------------------------------------------------------
   
   # Import results
@@ -137,3 +173,81 @@ rm(list = ls())
   # Export results as .rds
   write_rds(m4_post, "Experiment 2/results/m4_post.rds")
 
+
+# Process (Model 5) -------------------------------------------------------
+
+  # Import results
+  m5_fit <- read_rds("Experiment 2/results/m5_fit.rds")
+  
+  # Extract draws
+  m5_ps <- m5_fit$draws() %>% as_draws_df()
+  
+  # Extract posterior draws
+  m5_post_ii <- m5_ps %>% spread_draws(alpha[ii], beta[ii]) %>% ungroup()
+  m5_post_jj <- m5_ps %>% spread_draws(theta[jj]) %>% ungroup()
+  m5_post_kk <- m5_ps %>% spread_draws(delta[kk], b_ide_kk[kk]) %>% ungroup()
+  
+  # Calculate effect sizes
+  m5_post <- dl %>% 
+    drop_na(x_ide) %>% 
+    mutate(jj = as.integer(factor(id))) %>% 
[…]

### 2022-08-22 Documents/manuscript.Rmd
update manuscript

@@
 abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
-In four studies (total *N* = 2,246), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*).
+In five studies (total *N* = 2,979), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*).
 In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action.
-In two preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police in the United States.
-We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position---but showed no ingroup bias when judging collective action by ingroup and outgroup members.
+In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
+We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no ingroup bias when judging collective action by ingroup and outgroup members.
 Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
@@
 In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action.
-In two preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police in the United States.
-We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own (system-challenging) ideological positions---but do not show ingroup bias when judging collective action by ingroup and outgroup members.
+In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
+We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show ingroup bias when judging collective action by ingroup and outgroup members.
 
@@
       h1a_contrast = case_when(
-        Participant == Protesters ~ +1L,
-        Participant != Protesters ~ -1L,
+        Participant == Protesters ~ +1/8,
+        Participant != Protesters ~ -1/8,
       ),
       h1b_contrast = case_when(
-        Participant == "Black" & Cause == "For" ~ +1L,
-        Participant == "White" & Cause == "Against" ~ +1L,
-        TRUE ~ -1L,
+        Participant == "Black" & Cause == "For" ~ +1/8,
+        Participant == "White" & Cause == "Against" ~ +1/8,
+        TRUE ~ -1/8,
       ),
       h2a_contrast = case_when(
-        Cause == "For" ~ -1L,
-        Cause == "Against" ~ +1L
+        Cause == "For" ~ -1/8,
+        Cause == "Against" ~ +1/8
       ),
       nh1_contrast = case_when(
-        Participant == "Black" ~ -1L,
-        Participant == "White" ~ +1L,
+        Participant == "Black" ~ -1/8,
+        Participant == "White" ~ +1/8
       ),
       nh2_contrast = case_when(
-        Participant == "White" & Cause == "Against" ~ -1L,
-        Participant == "White" & Cause == "For" ~ +1L,
-        Participant == "Black" ~ 0L
+        Participant == "White" & Cause == "Against" ~ -1/4,
+        Participant == "White" & Cause == "For" ~ +1/4,
+        Participant == "Black" ~ 0
       ),
       nh3_contrast = case_when(
-        Participant == "Black" & Cause == "Against" ~ -1L,
-        Participant == "Black" & Cause == "For" ~ +1L,
-        Participant == "White" ~ 0L
+        Participant == "Black" & Cause == "Against" ~ -1/4,
+        Participant == "Black" & Cause == "For" ~ +1/4,
+        Participant == "White" ~ 0
       )
@@
         ends_with("_contrast"),
-        ~mean(z * .)
+        ~sum(z * .)
       )
@@
       h1a_contrast = case_when(
-        Participant == Protesters ~ +1L,
-        Participant != Protesters ~ -1L,
+        Participant == Protesters ~ +1/8,
+        Participant != Protesters ~ -1/8
       ),
       h1b_contrast = case_when(
-        Participant == "Black" & Cause == "For" ~ +1L,
-        Participant == "White" & Cause == "Against" ~ +1L,
-        TRUE ~ -1L,
+        Participant == "Black" & Cause == "For" ~ +1/8,
+        Participant == "White" & Cause == "Against" ~ +1/8,
+        TRUE ~ -1/8,
       )
@@
   m5 <- m5 %>% 
-    group_by(.draw) %>% 
+    group_by(.draw, x_ide) %>% 
     summarize(
@@
     ) %>% 
-    group_by(contrast) %>% 
+    group_by(x_ide, contrast) %>% 
     median_qi(d)
@@
 As in Model 2, we modeled this relationship with a fixed and a varying effect.
-We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast"])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable  than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast"])`).
+We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast" & x_ide == 1L])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast" & x_ide == 1L])`).
 Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgements about collective action by ingroup and outgroup members.
@@
 
+# Experiment 3
+
+## Method
+
+We preregistered the sample size, hypotheses, inclusion/exclusion criteria, measures, and manipulations ([https://osf.io/kphyw/?view_only=59f5792a6d3f4fa48516ea9d5f637822](https://osf.io/kphyw/?view_only=59f5792a6d3f4fa48516ea9d5f637822)).
+We deviated from the preregistration in two ways.
+First, we stopped collecting data before reaching the preregistered sample size.
+We did so as the Supreme Court decision on reproductive rights, announced on June 24, 2022, had rendered the experimental manipulation invalid.
+Second, we do not report results for two further hypotheses---testing moral conviction as a potential mechanism underlying ideology-based double standards---as those findings are beyond the scope of this manuscript and will be reported in a future publication.
+We made all materials, data, and analysis scripts available online ([https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)).
+
+### Study Design
+
+We used a 2 (quasi-experimental: conservative/liberal participants) $\times$ 2 (quasi-experimental: male/female participants) $\times$ 2 (experimental: for/against restricting abortion) between-subjects design to test our hypotheses.
+
+### Participants
+
+We recruited 804 participants from the Prolific subject pool who were 18 years old or older and lived in the United States.^[Data were collected between June 10 and 23, 2022.
+Participants received, on average, \$24.70 per hour of participation.] We balanced the sample's gender and political composition so that half of all participants identified as conservative and the other half identified as liberal and so that roughly half of the participants in each group were women and men.
+As preregistered, we excluded 71 participants who failed at least one of five attention checks or reported a different political orientation than they had reported in the Prolific prescreening questionnaire.
+As noted above, we had to stop collecting data before reaching the preregistered sample size of 800 eligible participants.
+This resulted in a final sample of 733 participants ($\textit{Mdn} = 37$ years, age range: 18--93 years) of whom 181 were conservative men, 190 were conservative women, 185 were liberal men, and 177 were liberal women.
+
+### Procedure
+
+After answering where they would you place themselves along the political spectrum, in line with their response in the Prolific prescreening questionnaire, participants read the following paragraph:
+
+> Recent news suggested that the Supreme Court might soon overturn its *Roe v.
+Wade* decision that, so far, has guaranteed legal abortions in the first 20--24 weeks of pregnancy.
+Overturning *Roe v.
+Wade* would allow states to enact laws to further restrict or ban abortions.
+
+\noindent In the supporting abortion restrictions condition, participants then read a text presenting arguments for restricting or banning abortion:
+
+> Some people welcome this news because they oppose legal abortions.
+They argue that life begins at conception and that, therefore, abortion ends the life of an unborn child.
+As human life is sacred, ending an unborn life is wrong and cannot be justified by a pregnant person's right to choose.
+In this view, restricting or banning abortions is the only way to protect unborn lives.
+We refer to this position as "supporting abortion restrictions".
+
+\noindent In the opposing abortion restrictions condition, participants instead read a text presenting arguments against restricting or banning abortion:
+
+> Some people are alarmed by this news because they support legal abortions.
+They argue that pregnant people have a right to control their own body and, therefore, have a right to safe and legal abortions.
+Restricting or banning abortion would lead to more illegal and unsafe abortions.
+In this view, providing legal abortions is the only way to protect pregnant people's rights and to prevent harm.
+We refer to this position as "opposing abortion restrictions".
+
+\noindent Participants were then asked to select which position the text described and what arguments for that position were presented in the text.
+If they selected the wrong answer, they were instructed to reread the text and select the right answer.
+Participants then rated to what extent they support or oppose restricting abortion and to what extent their feelings about this issue are a moral conviction.
+
+Participants then read about protesters protesting either for (in the supporting abortion restrictions condition) or against (in the opposing abortion restrictions condition) restricting or banning abortion:
+
+> Earlier, you read why some people [support/oppose] further restricting or banning abortion. Some local residents who hold this view want to protest [for/against] restricting or banning abortion.
+
+\noindent Participants rated which of several protest actions they consider acceptable means to protest for or against restricting abortion.
+On the next pages, participants completed the remaining measures.
+On the final page, participants responded to two attention checks that assessed whether they paid attention to the experimental manipulation: "In this study, you read about some people's view on a policy change.
+What was the policy change?" and "In this study, you then read about some protesters.
+Were the protesters for or against restricting abortion?".
+
+### Measures
+
+We measured the outcome variable by asking participants to think about the protesters they had read about and to decide, for each of 25 protest actions presented in a randomized order, whether they thought this action was "an acceptable means for them to protest [for/against] restricting or banning abortion" (1 = *yes*, 0 = *no*; see Table C1).
+
+We included additional measures to describe the sample, to report as descriptive statistics, to use in non-preregistered analyses or future analyses not reported in this paper.
+
+We measured support for restricting abortion with one item: "Do you support or oppose restricting (or banning) abortion?" (1 = *strongly oppose*, 5 = *strongly support*).
+We also included a  three-item measure of moral conviction [@ryan_reconsidering_2014] that assessed to what extent participants' feelings about restricting or banning abortion were, for example, connected to their core moral beliefs or convictions (1 = *not at all*, 5 = *very much*).
+
+We measured gender-related system-justifying beliefs with eight items [adapted from @jost_exposure_2005], for example, "in general, relations between men and women are fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .93$).
+A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]$.
+
+We will measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
+We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
+
+In addition to demographic questions, we asked participants to what extent they identified with their gender (1 = *strongly disagree*, 7 = *strongly agree*) and measured political orientation with a one-item semantic differential scale: "People often describe their political orientation as liberal or conservative.
+On a scale from liberal to conservative, where would you position yourself?" (1 = *liberal*, 7 = *conservative*).
+
+## Results
+
+### Reactions to the Manipulation
+
+As expected, participants who identified as conservative reported being far more conservative ($M = 6.05, SD = 0.79$) than participants who identified as liberal ($M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43$).
+Likewise, conservative participants tended to endorse system-justifying beliefs ($M = 5.10, SD = 0.95$) while liberal participants tended to reject system-justifying beliefs ($M = 2.86, SD = 1.03; d = 2.26$).
+As expected, conservative and liberal participants differed in their reactions to the manipulation as the former tended to support restricting abortion ($M = 3.77, SD = 1.48$) while the latter tended to oppose restricting abortion ($M = 1.28, SD = 0.85; d = 2.06$).
+When comparing politically balanced subsamples of men and women, women ($M = 2.55, SD = 1.76$) did not differ from men ($M = 2.55, SD = 1.72$) in their attitudes toward restricting abortion ($d = -0.02$).
+
+### Preregistered Analyses
+
+```{=latex}
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered analyses for Experiment 3}
+\includegraphics[scale=1]{../Experiment 3/figures/figure-6}
+\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
+\label{fig:f6}
+\end{figure*}
+```
+
+```{r, include = FALSE}
+
+  # THIS MIGHT BE WRONG
+
+  # Import results
+  m1 <- read_rds("../Experiment 3/results/m1_post.rds")
+
+  # Define contrasts
+  m1 <- m1 %>% 
+    mutate(
+      h1_contrast = case_when(
+        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L
+      ),
+      h2a_contrast = case_when(
+        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L
+      ),
+      h2b_contrast = case_when(
+        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L
+      ),
+      nh1_contrast = case_when(
+        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Conservative" & protesters_cause == "Against" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh2_contrast = case_when(
+        participant_ideology == "Liberal" & protesters_cause == "For" ~ -1L,
+        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ NA_integer_
+      ),
+      nh3_contrast = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "Against" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh4_contrast = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "Against" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh5_contrast = case_when(
+        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "For" ~ -1L,
+        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ NA_integer_
+      ),
+      nh6_contrast = case_when(
+        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
+        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ NA_integer_
+      )
+    )
+
+  # Calculate contrasts
+  m1 <- m1 %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "direction",
+      values_drop_na = TRUE
+    ) %>% 
+    group_by(.draw, contrast, direction) %>% 
+    summarize(z = mean(z)) %>% 
+    summarize(d = sum(z * direction)) %>% 
+    pivot_wider(
+      names_from = contrast,
+      values_from = d
+    ) %>% 
+    mutate(nh7_contrast = nh2_contrast - nh1_contrast) %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "d"
+    ) %>% 
+    group_by(contrast) %>% 
+    median_qi(d)
+
+  # Make text
+  m1 <- m1 %>% 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
+```
+
+To test our hypotheses, we used a two-parameter logistic item response theory model, identical to the one used in Experiment 2, to estimate how likely participants were to consider each action an acceptable means of protest as a function of the protesters' cause and of the participants' gender and ideology.
+Figure \ref{fig:f6} shows the results of our preregistered analyses.
+
+As in Experiments 1 and 2, we found no evidence for identity-based double standards: Female participants did not consider protest actions *against* restricting abortion to be any more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
+Instead, we again found evidence for ideology-based double standards: Conservative participants considered the same protest actions to be more acceptable, on average, when protesters *supported* restricting abortion (`r with(m1, d_text[contrast == "nh1_contrast"])`) while liberal participants considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion (`r with(m1, d_text[contrast == "nh2_contrast"])`).
+This pattern was consistent across subsamples as conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and as both liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) and women (`r with(m1, d_text[contrast == "nh6_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
+This double standard was, however, more pronounced among liberal participants (`r with(m1, d_text[contrast == "nh7_contrast"])`).
+In this way, Experiment 3 provided evidence for both ideological symmetry and asymmetry in judging collective action for and against restricting abortion.
+
+### Non-Preregistered Analyses
+
+```{r, include = FALSE}
+
+  # Import results
+  ide <- read_rds("../Experiment 3/results/ide_post.rds")
+
+  # Define contrasts
+  ide <- ide %>% 
+    mutate(
+      h1_contrast = case_when(
+        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L,
+      ),
+      nh1_contrast = case_when(
+        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+        participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh2_contrast = case_when(
+        participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
+        participant_gender == "Man" & protesters_cause == "For" ~ -1L,
+        TRUE ~ NA_integer_
+      )
+    )
+
+  # Calculate contrasts
+  ide <- ide %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "direction",
+      values_drop_na = TRUE
+    ) %>% 
+    group_by(.draw, z_ide, contrast, direction) %>% 
+    summarize(z = mean(z)) %>% 
+    summarize(d = sum(z * direction)) %>% 
+    group_by(z_ide, contrast) %>% 
+    median_qi(d)
+
+  # Make text
+  ide <- ide %>% 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
+```
+
+<!-- We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses.
+We found some evidence that, when considering only strongly identified (+1*SD*) men and women, female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(ide, d_text[contrast == "h1_contrast" & z_ide == 1L])`).
+Our non-preregistered analyses thus suggested that gender identification might moderate gender differences in judgements about collective action concerning reproductive rights.
+-->
+
+```{r, include = FALSE}
+
+  # Import results
+  sjb <- read_rds("../Experiment 3/results/sjb_post.rds")
+
+  # Hypothesis 2
+  sjb <- sjb %>% 
+    distinct(.draw, b_sjb_for, b_sjb_against) %>% 
+    pivot_longer(
+      c(b_sjb_for, b_sjb_against),
+      names_to = "cause",
+      names_prefix = "b_sjb_",
+      values_to = "b"
+    ) %>%
+    group_by(cause) %>% 
+    median_qi(b) %>% 
+    mutate(
+      across(
+        c(b, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      b_text = glue(
+        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
+        .open = "(", .close = ")"
+      )
+    )
+
+```
+
+<!-- We also explored whether gender-related system justification predicted how acceptable participants judged protest actions for and against restricting abortion to be.
+To that end, we added two fixed effects to the model used in the preregistered analyses that estimated participants' responses as a function of their *z*-standardized system-justifying beliefs.
+As in Experiment 2, we found evidence for ideological asymmetry in this relationship: Participants who endorsed system-justifying beliefs considered the same protest action less acceptable when judging protests *against* restricting abortion (`r with(sjb, b_text[cause == "against"])`) but did not differ from participants who rejected system-justifying beliefs when judging protests *for* restricting abortion (`r with(sjb, b_text[cause == "for"])`).
+As in Experiment 2, our non-preregistered analyses thus suggested that system justification affects judgements about protests for a progressive cause (opposing abortion restrictions) but not for a conservative cause (supporting abortion restrictions).
+-->
+
+## Discussion
+
 # General Discussion
@@
 -->
-In the remainder of this General Discussion, we discuss theoretical and practical implications for understanding the often divided response to collective action and consider some limiting conditions on our findings.
+In the remainder of this General Discussion, we consider some limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.
+
+## Limitations
+
+Our research applied item response theory to judgements about collective action to investigate whether where people draw the line between acceptable and unacceptable means of protest depends on who the protesters are and what they are protesting.
+In so doing, we moved the distinction between normative and non-normative collective action from the realm of the researcher's intuition into the realm of scientific investigation, allowing us to test for double standards in judging collective action.
+
+Despite its strengths, our research is qualified by important limitations.
+First, our research examined the hypothesized relationships for three political causes---protecting workers' rights, defunding the police, restricting abortion---and thus provided only limited evidence that our findings generalize beyond those causes.
+Relying on few stimuli to establish an effect threatens the validity and replicability of research findings and constrains the generalizability of psychological research [@yarkoni_generalizability_2022].
+Future research should sample a wider range of causes to address this pervasive but often ignored problem [@judd_treating_2012].
+This is particularly important when studying collective action as recent theorizing [@jost_missing_2017] and research [@osborne_protesting_2019] highlighted the importance of differentiating between progressive and conservative causes.
+Second, our research was based on samples from two Western, Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al., 2010) contexts---the United Kingdom and the United States---which limits the cross-cultural generalizability of our findings.
 
@@
 
-Our results showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
+Most broadly, our results showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
 Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
@@
 If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
-In this way, our findings could explain why 45% of Republicans (2% of Democrats) supported the storming of the U.S.
+In this way, our findings could explain why 45% of Republicans supported the storming of the U.S.
 Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
@@
 In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].
-For those reasons, researchers should further investigate when and why people apply different standards when judging collective action for various causes.
-
-## Limitations
-
-Our research applied item response theory to judgements about collective action to investigate whether where people draw the line between acceptable and unacceptable means of protest depends on who the protesters are and what they are protesting.
-In so doing, we moved the distinction between normative and non-normative collective action from the realm of the researcher's intuition into the realm of scientific investigation, allowing us to test for double standards in judging collective action.
-
-Despite its strengths, our research is qualified by several limitations.
-First, our research examined the hypothesized relationships for only two political causes---protecting workers' rights, defunding the police---and thus provided only limited evidence that our findings generalize beyond those causes.
-Relying on few stimuli to establish an effect threatens the validity and replicability of research findings and severely constrains psychological research generalize beyond those few stimuli [@yarkoni_generalizability_2022].
-Future research should sample a wider range of causes to address this pervasive but often ignored problem [@judd_treating_2012].
-This is particularly important when studying collective action as recent theorizing [@jost_missing_2017] and research [@osborne_protesting_2019] highlighted the importance of differentiating between progressive and conservative causes.
-Second, our research was based on two Western, Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al., 2010) samples, which severely limits the cross-cultural generalizability of our findings.
-Future research should include samples from a wider range of cultural context to examine how the observed relationships vary across cultural contexts.
+For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.
 
@@
 ```
+
+```{=latex}
+\setcounter{table}{0}
+\renewcommand{\thetable}{C\arabic{table}}
+
+\begin{table*}
+
+\section{Appendix C}
+
+\caption{List of protest actions used in Experiment 3}
+\small
+\begin{tabularx}{\linewidth}{rXr}
+\addlinespace
+\toprule
+\# & Action & Pr\\
+\midrule
+1 & make a public speech & 94\%\\
+2 & hand out flyers, leaflets, or pamphlets & 92\%\\
+3 & wear or display political symbols (e.g., patches, flags, bumper stickers) & 92\%\\
+4 & hold meetings to influence the public & 92\%\\
+5 & attend or organise a protest march & 91\%\\
+6 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
+7 & donate to activist groups & 90\%\\
+8 & refuse to buy goods or services from companies that advocate [against/for] restricting or banning abortion (boycott) & 90\%\\
+9 & paste up posters with political messages in places where it is allowed and encouraged & 90\%\\
+10 & join or form a group of activists & 90\%\\
+11 & donate to politicians who advocate [for/against] restricting or banning abortion & 89\%\\
+12 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 84\%\\
+13 & refuse to accept honors or awards in protest & 82\%\\
+14 & refuse to work (strike) & 56\%\\
+15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 50\%\\
+16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 41\%\\
+17 & attend a protest march even though it might be unlawful & 40\%\\
+18 & \textit{protest outside the homes of politicians who advocate [against/for] restricting or banning abortion} & 40\%\\
+19 & attend a protest march even though it might turn violent & 40\%\\
+20 & paste up posters with political messages in places where it is not allowed or encouraged & 26\%\\
+21 & enter and refuse to leave a building (occupation) & 25\%\\
+22 & refuse to pay fees, fines, or taxes in protest & 23\%\\
+23 & mock or insult individuals who are [against/for] restricting or banning abortion & 20\%\\
+24 & spray paint political messages in public places & 15\%\\
+25 & disrupt traffic (e.g., blocking roads) & 12\%\\
+\bottomrule
+\end{tabularx}
+\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 3.
+Action in \textit{italics} replaced an action used in Experiment 2 because it better fit the context of Experiment 3.}
+
+\end{table*}
+```

### 2022-08-22 Documents/manuscript.tex
update manuscript

@@
 change but often sparks contention about what actions are acceptable
-means to effect social change. In four studies (total \emph{N} = 2,246),
+means to effect social change. In five studies (total \emph{N} = 2,979),
 we investigated double standards in judging collective action---that is,
@@
 response theory, to measure where people draw the line between
-acceptable and unacceptable forms of collective action. In two
+acceptable and unacceptable forms of collective action. In three
 preregistered experiments, we used this instrument to test for double
 standards in judging collective action for workers' rights in the United
-Kingdom and for or against defunding the police in the United States. We
-found evidence for ideology-based, but not for identity-based, double
-standards: Participants judged the same protest actions to be more
-acceptable when the protesters' cause aligned with their own
-(system-challenging) ideological position---but showed no ingroup bias
-when judging collective action by ingroup and outgroup members. Our
-findings have theoretical and practical implications for understanding
-the often divided response to prominent social movements.
+Kingdom and for or against defunding the police and for or against
+restricting abortion in the United States. We found evidence for
+ideology-based, but not for identity-based, double standards:
+Participants judged the same protest actions to be more acceptable when
+the protesters' cause aligned with their own ideological position---but
+showed no ingroup bias when judging collective action by ingroup and
+outgroup members. Our findings have theoretical and practical
+implications for understanding the often divided response to prominent
+social movements.
     \end{abstract}
@@
 response theory, to capture double standards in judging collective
-action. In two preregistered experiments, we then use this instrument to
-test for double standards in judging protest actions for workers' rights
-in the United Kingdom and for or against defunding the police in the
-United States. We find evidence for ideology-based, but not for
-identity-based, double standards: People judge the same protest actions
-to be more acceptable when the protesters' cause aligns with their own
-(system-challenging) ideological positions---but do not show ingroup
-bias when judging collective action by ingroup and outgroup members.
+action. In three preregistered experiments, we then use this instrument
+to test for double standards in judging protest actions for workers'
+rights in the United Kingdom and for or against defunding the police and
+for or against restricting abortion in the United States. We find
+evidence for ideology-based, but not for identity-based, double
+standards: People judge the same protest actions to be more acceptable
+when the protesters' cause aligns with their own ideological
+positions---but do not show ingroup bias when judging collective action
+by ingroup and outgroup members.
 
@@
 actions to be more acceptable than Black participants
-(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.07, [0.03, 0.10]\))
-and White (\(d = 0.08, [0.04, 0.11]\)) participants considered the same
+(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.13, [0.06, 0.20]\))
+and White (\(d = 0.16, [0.08, 0.23]\)) participants considered the same
 actions to be more acceptable when protests were for, rather than
@@
 acceptable than the same actions performed by the relevant outgroup
-(\(d = 0.02, [-0.04, 0.07]\)) and did not consider protest actions for a
+(\(d = 0.01, [-0.00, 0.02]\)) and did not consider protest actions for a
 cause that was aligned with their ingroup's interests to be more
 acceptable than protest actions for a cause not aligned with their
-ingroup's interests (\(d = -0.03, [-0.09, 0.02]\)). Our
-non-preregistered analyses thus suggested that group identification did
-not moderate group differences in judgements about collective action by
-ingroup and outgroup members.
+ingroup's interests (\(d = 0.00, [-0.01, 0.01]\)). Our non-preregistered
+analyses thus suggested that group identification did not moderate group
+differences in judgements about collective action by ingroup and
+outgroup members.
 
@@
 
+\hypertarget{experiment-3}{%
+\section{Experiment 3}\label{experiment-3}}
+
+\hypertarget{method-2}{%
+\subsection{Method}\label{method-2}}
+
+We preregistered the sample size, hypotheses, inclusion/exclusion
+criteria, measures, and manipulations
+(\url{https://osf.io/kphyw/?view_only=59f5792a6d3f4fa48516ea9d5f637822}).
+We deviated from the preregistration in two ways. First, we stopped
+collecting data before reaching the preregistered sample size. We did so
+as the Supreme Court decision on reproductive rights, announced on June
+24, 2022, had rendered the experimental manipulation invalid. Second, we
+do not report results for two further hypotheses---testing moral
+conviction as a potential mechanism underlying ideology-based double
+standards---as those findings are beyond the scope of this manuscript
+and will be reported in a future publication. We made all materials,
+data, and analysis scripts available online
+(\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}).
+
+\hypertarget{study-design-2}{%
+\subsubsection{Study Design}\label{study-design-2}}
+
+We used a 2 (quasi-experimental: conservative/liberal participants)
+\(\times\) 2 (quasi-experimental: male/female participants) \(\times\) 2
+(experimental: for/against restricting abortion) between-subjects design
+to test our hypotheses.
+
+\hypertarget{participants-2}{%
+\subsubsection{Participants}\label{participants-2}}
+
+We recruited 804 participants from the Prolific subject pool who were 18
+years old or older and lived in the United States.\footnote{Data were
+  collected between June 10 and 23, 2022. Participants received, on
+  average, \$24.70 per hour of participation.} We balanced the sample's
+gender and political composition so that half of all participants
+identified as conservative and the other half identified as liberal and
+so that roughly half of the participants in each group were women and
+men. As preregistered, we excluded 71 participants who failed at least
+one of five attention checks or reported a different political
+orientation than they had reported in the Prolific prescreening
+questionnaire. As noted above, we had to stop collecting data before
+reaching the preregistered sample size of 800 eligible participants.
+This resulted in a final sample of 733 participants
+(\(\textit{Mdn} = 37\) years, age range: 18--93 years) of whom 181 were
+conservative men, 190 were conservative women, 185 were liberal men, and
+177 were liberal women.
+
+\hypertarget{procedure-2}{%
+\subsubsection{Procedure}\label{procedure-2}}
+
+After answering where they would you place themselves along the
+political spectrum, in line with their response in the Prolific
+prescreening questionnaire, participants read the following paragraph:
+
+\begin{quote}
+Recent news suggested that the Supreme Court might soon overturn its
+\emph{Roe v. Wade} decision that, so far, has guaranteed legal abortions
+in the first 20--24 weeks of pregnancy. Overturning \emph{Roe v. Wade}
+would allow states to enact laws to further restrict or ban abortions.
+\end{quote}
+
+\noindent In the supporting abortion restrictions condition,
+participants then read a text presenting arguments for restricting or
+banning abortion:
+
+\begin{quote}
+Some people welcome this news because they oppose legal abortions. They
+argue that life begins at conception and that, therefore, abortion ends
+the life of an unborn child. As human life is sacred, ending an unborn
+life is wrong and cannot be justified by a pregnant person's right to
+choose. In this view, restricting or banning abortions is the only way
+to protect unborn lives. We refer to this position as ``supporting
+abortion restrictions''.
+\end{quote}
+
+\noindent In the opposing abortion restrictions condition, participants
+instead read a text presenting arguments against restricting or banning
+abortion:
+
+\begin{quote}
+Some people are alarmed by this news because they support legal
+abortions. They argue that pregnant people have a right to control their
+own body and, therefore, have a right to safe and legal abortions.
+Restricting or banning abortion would lead to more illegal and unsafe
+abortions. In this view, providing legal abortions is the only way to
+protect pregnant people's rights and to prevent harm. We refer to this
+position as ``opposing abortion restrictions''.
+\end{quote}
+
+\noindent Participants were then asked to select which position the text
+described and what arguments for that position were presented in the
+text. If they selected the wrong answer, they were instructed to reread
+the text and select the right answer. Participants then rated to what
+extent they support or oppose restricting abortion and to what extent
+their feelings about this issue are a moral conviction.
+
+Participants then read about protesters protesting either for (in the
+supporting abortion restrictions condition) or against (in the opposing
+abortion restrictions condition) restricting or banning abortion:
+
+\begin{quote}
+Earlier, you read why some people {[}support/oppose{]} further
+restricting or banning abortion. Some local residents who hold this view
+want to protest {[}for/against{]} restricting or banning abortion.
+\end{quote}
+
+\noindent Participants rated which of several protest actions they
+consider acceptable means to protest for or against restricting
+abortion. On the next pages, participants completed the remaining
+measures. On the final page, participants responded to two attention
+checks that assessed whether they paid attention to the experimental
+manipulation: ``In this study, you read about some people's view on a
+policy change. What was the policy change?'' and ``In this study, you
+then read about some protesters. Were the protesters for or against
+restricting abortion?''.
+
+\hypertarget{measures-2}{%
+\subsubsection{Measures}\label{measures-2}}
+
+We measured the outcome variable by asking participants to think about
+the protesters they had read about and to decide, for each of 25 protest
+actions presented in a randomized order, whether they thought this
+action was ``an acceptable means for them to protest {[}for/against{]}
+restricting or banning abortion'' (1 = \emph{yes}, 0 = \emph{no}; see
+Table C1).
+
+We included additional measures to describe the sample, to report as
+descriptive statistics, to use in non-preregistered analyses or future
+analyses not reported in this paper.
+
+We measured support for restricting abortion with one item: ``Do you
+support or oppose restricting (or banning) abortion?'' (1 =
+\emph{strongly oppose}, 5 = \emph{strongly support}). We also included a
+three-item measure of moral conviction (Ryan, 2014) that assessed to
+what extent participants' feelings about restricting or banning abortion
+were, for example, connected to their core moral beliefs or convictions
+(1 = \emph{not at all}, 5 = \emph{very much}).
+
+We measured gender-related system-justifying beliefs with eight items
+(adapted from Jost \& Kay, 2005), for example, ``in general, relations
+between men and women are fair'' (1 = \emph{strongly disagree}, 7 =
+\emph{strongly agree}; McDonald's \(\omega = .93\)). A confirmatory
+factor analysis model in which all items loaded onto a single factor
+showed acceptable fit,
+\(\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]\).
+
+We will measured moral concerns with the 36-item moral foundations
+questionnaire (MFQ-2, Atari et al., 2022) which assesses to what extent
+participants endorse concerns about care, equality, proportionality,
+loyalty, authority, and purity (e.g., ``I believe chastity is an
+important virtue.''; 1 = \emph{does not describe me at all}, 5 =
+\emph{describes me extremely well}). We embedded three further attention
+checks within the questionnaire (e.g., ``To show that you are paying
+attention and giving your best effort, please select `moderately
+describes me'.'').
+
+In addition to demographic questions, we asked participants to what
+extent they identified with their gender (1 = \emph{strongly disagree},
+7 = \emph{strongly agree}) and measured political orientation with a
+one-item semantic differential scale: ``People often describe their
+political orientation as liberal or conservative. On a scale from
+liberal to conservative, where would you position yourself?'' (1 =
+\emph{liberal}, 7 = \emph{conservative}).
+
+\hypertarget{results-2}{%
+\subsection{Results}\label{results-2}}
+
+\hypertarget{reactions-to-the-manipulation-2}{%
+\subsubsection{Reactions to the
+Manipulation}\label{reactions-to-the-manipulation-2}}
+
+As expected, participants who identified as conservative reported being
+far more conservative (\(M = 6.05, SD = 0.79\)) than participants who
+identified as liberal
+(\(M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43\)). Likewise,
+conservative participants tended to endorse system-justifying beliefs
+(\(M = 5.10, SD = 0.95\)) while liberal participants tended to reject
+system-justifying beliefs (\(M = 2.86, SD = 1.03; d = 2.26\)). As
+expected, conservative and liberal participants differed in their
+reactions to the manipulation as the former tended to support
+restricting abortion (\(M = 3.77, SD = 1.48\)) while the latter tended
+to oppose restricting abortion (\(M = 1.28, SD = 0.85; d = 2.06\)). When
+comparing politically balanced subsamples of men and women, women
+(\(M = 2.55, SD = 1.76\)) did not differ from men
+(\(M = 2.55, SD = 1.72\)) in their attitudes toward restricting abortion
+(\(d = -0.02\)).
+
+\hypertarget{preregistered-analyses-2}{%
+\subsubsection{Preregistered Analyses}\label{preregistered-analyses-2}}
+
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered analyses for Experiment 3}
+\includegraphics[scale=1]{../Experiment 3/figures/figure-6}
+\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
+\label{fig:f6}
+\end{figure*}
+
+To test our hypotheses, we used a two-parameter logistic item response
+theory model, identical to the one used in Experiment 2, to estimate how
+likely participants were to consider each action an acceptable means of
+protest as a function of the protesters' cause and of the participants'
+gender and ideology. Figure \ref{fig:f6} shows the results of our
+preregistered analyses.
+
+As in Experiments 1 and 2, we found no evidence for identity-based
+double standards: Female participants did not consider protest actions
+\emph{against} restricting abortion to be any more acceptable, on
+average, than male participants or protest actions \emph{for}
+restricting abortion (\(d = 0.14, [0.00, 0.29]\)). Instead, we again
+found evidence for ideology-based double standards: Conservative
+participants considered the same protest actions to be more acceptable,
+on average, when protesters \emph{supported} restricting abortion
+(\(d = 0.24, [0.07, 0.43]\)) while liberal participants considered the
+same protest actions to be more acceptable when protesters
+\emph{opposed} restricting abortion (\(d = 0.90, [0.72, 1.08]\)). This
+pattern was consistent across subsamples as conservative women
+(\(d = 0.30, [0.04, 0.55]\)) and, to a lesser extent, conservative men
+(\(d = 0.19, [-0.07, 0.46]\)) considered the same protest actions to be
+more acceptable when protesters \emph{supported} restricting abortion
+and as both liberal men (\(d = 0.76, [0.49, 1.03]\)) and women
+(\(d = 1.04, [0.78, 1.32]\)) considered the same protest actions to be
+more acceptable when protesters \emph{opposed} restricting abortion.
+This double standard was, however, more pronounced among liberal
+participants (\(d = 0.66, [0.41, 0.92]\)). In this way, Experiment 3
+provided evidence for both ideological symmetry and asymmetry in judging
+collective action for and against restricting abortion.
+
+\hypertarget{non-preregistered-analyses-2}{%
+\subsubsection{Non-Preregistered
+Analyses}\label{non-preregistered-analyses-2}}
+
+\hypertarget{discussion-2}{%
+\subsection{Discussion}\label{discussion-2}}
+
 \hypertarget{general-discussion}{%
@@
 (system-challenging) ideological position. In the remainder of this
-General Discussion, we discuss theoretical and practical implications
-for understanding the often divided response to collective action and
-consider some limiting conditions on our findings.
-
-\hypertarget{implications}{%
-\subsection{Implications}\label{implications}}
-
-Our results showed that, for many controversial protest actions, there
-is no consensus on what is and is not an acceptable means to advance a
-cause. Instead, this judgement seems to depend on what the cause of a
-protest is and how it aligns with one's beliefs and values. In this way,
-our results confirm our contention that the distinction between
-normative and non-normative forms of collective action should be
-situated in the eye of the beholder. By developing an instrument for
-measuring where people draw the line between acceptable and unacceptable
-forms of collective action, we provided a paradigm that we hope will
-stimulate research on when and why people differ in making this
-distinction.
-
-After all, this distinction can be consequential. If people believe that
-even extreme actions are acceptable means to advance a cause they
-support, especially if they moralize their support for that cause
-(Mooijman et al., 2018), they might support violent extremism. In this
-way, our findings could explain why 45\% of Republicans (2\% of
-Democrats) supported the storming of the U.S. Capitol on January 6, 2021
-(YouGov, 2021) even though it was illegal, destructive, and violent. If
-people believe that even non-violent actions are unacceptable means to
-advance a cause they oppose, they might support stifling dissent. In
-this way, our findings could explain why Republicans responded to Black
-Lives Matter protests by seeking to further criminalize disruptive
-protest (e.g., blocking highways, Quinton, 2021) that might be most
-effective at challenging injustice (Shuman et al., 2022, 2021). For
-those reasons, researchers should further investigate when and why
-people apply different standards when judging collective action for
-various causes.
+General Discussion, we consider some limiting conditions on our findings
+and discuss theoretical and practical implications for understanding the
+often divided response to collective action.
 
@@
 
-Despite its strengths, our research is qualified by several limitations.
-First, our research examined the hypothesized relationships for only two
-political causes---protecting workers' rights, defunding the
-police---and thus provided only limited evidence that our findings
-generalize beyond those causes. Relying on few stimuli to establish an
-effect threatens the validity and replicability of research findings and
-severely constrains psychological research generalize beyond those few
-stimuli (Yarkoni, 2022). Future research should sample a wider range of
+Despite its strengths, our research is qualified by important
+limitations. First, our research examined the hypothesized relationships
+for three political causes---protecting workers' rights, defunding the
+police, restricting abortion---and thus provided only limited evidence
+that our findings generalize beyond those causes. Relying on few stimuli
+to establish an effect threatens the validity and replicability of
+research findings and constrains the generalizability of psychological
+research (Yarkoni, 2022). Future research should sample a wider range of
 causes to address this pervasive but often ignored problem (Judd et al.,
@@
 2019) highlighted the importance of differentiating between progressive
-and conservative causes. Second, our research was based on two Western,
-Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al.,
-2010) samples, which severely limits the cross-cultural generalizability
-of our findings. Future research should include samples from a wider
-range of cultural context to examine how the observed relationships vary
-across cultural contexts.
+and conservative causes. Second, our research was based on samples from
+two Western, Educated, Industrialized, Rich, and Democratic (WEIRD;
+Henrich et al., 2010) contexts---the United Kingdom and the United
+States---which limits the cross-cultural generalizability of our
+findings.
+
+\hypertarget{implications}{%
+\subsection{Implications}\label{implications}}
+
+Most broadly, our results showed that, for many controversial protest
+actions, there is no consensus on what is and is not an acceptable means
+to advance a cause. Instead, this judgement seems to depend on what the
+cause of a protest is and how it aligns with one's beliefs and values.
+In this way, our results confirm our contention that the distinction
+between normative and non-normative forms of collective action should be
+situated in the eye of the beholder. By developing an instrument for
+measuring where people draw the line between acceptable and unacceptable
+forms of collective action, we provided a paradigm that we hope will
+stimulate research on when and why people differ in making this
+distinction.
+
+After all, this distinction can be consequential. If people believe that
+even extreme actions are acceptable means to advance a cause they
+support, especially if they moralize their support for that cause
+(Mooijman et al., 2018), they might support violent extremism. In this
+way, our findings could explain why 45\% of Republicans supported the
+storming of the U.S. Capitol on January 6, 2021 (YouGov, 2021) even
+though it was illegal, destructive, and violent. If people believe that
+even non-violent actions are unacceptable means to advance a cause they
+oppose, they might support stifling dissent. In this way, our findings
+could explain why Republicans responded to Black Lives Matter protests
+by seeking to further criminalize disruptive protest (e.g., blocking
+highways, Quinton, 2021) that might be most effective at challenging
+injustice (Shuman et al., 2022, 2021). For those reasons, we encourage
+researchers to use our paradigm to further investigate when and why
+people apply different standards when judging collective action for
+different causes.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2022}{}}%
+Atari, M., Haidt, J., Graham, J., Koleva, S., Stevens, S. T., \&
+Dehghani, M. (2022). Morality beyond the {WEIRD}: How the nomological
+network of morality varies across cultures. \emph{PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/q6c9r}
+
 \leavevmode\vadjust pre{\hypertarget{ref-becker_virtual_2012}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-jost_exposure_2005}{}}%
+Jost, J. T., \& Kay, A. C. (2005). Exposure to benevolent sexism and
+complementary gender stereotypes: Consequences for specific and diffuse
+forms of system justification. \emph{Journal of Personality and Social
+Psychology}, \emph{88}(3), 498--509.
+\url{https://doi.org/10.1037/0022-3514.88.3.498}
+
 \leavevmode\vadjust pre{\hypertarget{ref-judd_treating_2012}{}}%
@@
 pp. 273--295). Wiley-Blackwell.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ryan_reconsidering_2014}{}}%
+Ryan, T. J. (2014). Reconsidering moral issues in politics. \emph{The
+Journal of Politics}, \emph{76}(2), 380--397.
+\url{https://doi.org/10.1017/S0022381613001357}
 
@@
 
+\setcounter{table}{0}
+\renewcommand{\thetable}{C\arabic{table}}
+
+\begin{table*}
+
+\section{Appendix C}
+
+\caption{List of protest actions used in Experiment 3}
+\small
+\begin{tabularx}{\linewidth}{rXr}
+\addlinespace
+\toprule
+\# & Action & Pr\\
+\midrule
+1 & make a public speech & 94\%\\
+2 & hand out flyers, leaflets, or pamphlets & 92\%\\
+3 & wear or display political symbols (e.g., patches, flags, bumper stickers) & 92\%\\
+4 & hold meetings to influence the public & 92\%\\
+5 & attend or organise a protest march & 91\%\\
+6 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
+7 & donate to activist groups & 90\%\\
+8 & refuse to buy goods or services from companies that advocate [against/for] restricting or banning abortion (boycott) & 90\%\\
+9 & paste up posters with political messages in places where it is allowed and encouraged & 90\%\\
+10 & join or form a group of activists & 90\%\\
+11 & donate to politicians who advocate [for/against] restricting or banning abortion & 89\%\\
+12 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 84\%\\
+13 & refuse to accept honors or awards in protest & 82\%\\
+14 & refuse to work (strike) & 56\%\\
+15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 50\%\\
+16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 41\%\\
+17 & attend a protest march even though it might be unlawful & 40\%\\
+18 & \textit{protest outside the homes of politicians who advocate [against/for] restricting or banning abortion} & 40\%\\
+19 & attend a protest march even though it might turn violent & 40\%\\
+20 & paste up posters with political messages in places where it is not allowed or encouraged & 26\%\\
+21 & enter and refuse to leave a building (occupation) & 25\%\\
+22 & refuse to pay fees, fines, or taxes in protest & 23\%\\
+23 & mock or insult individuals who are [against/for] restricting or banning abortion & 20\%\\
+24 & spray paint political messages in public places & 15\%\\
+25 & disrupt traffic (e.g., blocking roads) & 12\%\\
+\bottomrule
+\end{tabularx}
+\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 3.
+Action in \textit{italics} replaced an action used in Experiment 2 because it better fit the context of Experiment 3.}
+
+\end{table*}
+
 \end{document}

### 2022-08-22 Experiment 2/figures/figure-5.R

@@ -223,15 +223,16 @@ rm(list = ls())
   # Make plot
   f_5_p <- ggplot(d_fig, aes(x = z, y = fct_rev(factor(order)), group = x)) + 
     geom_ribbon(
-      data = d_fig %>% group_by(order, x, predictor) %>% median_qi(z),
-      aes(xmin = .lower, xmax = .upper, fill = x),
+      data = d_fig %>% group_by(order, x, predictor, Participant) %>% median_qi(z),
+      aes(xmin = .lower, xmax = .upper, fill = x, group = interaction(Participant, x)),
       alpha = 0.4
     ) +
     stat_pointinterval(
       aes(shape = x, colour = x, group = x),
       fill = "white",
       size = 0.455,
-      point_size = 3*0.455
+      point_size = 3*0.455,
+      .width = 0.95
     ) +
     geom_hline(yintercept = c(0.5, 8.5), size = 0.455) +
     geom_hline(yintercept = c(4.5), size = 3*0.455, colour = "white") +

### 2022-08-22 Experiment 3/figures/figure-6.R

@@ -1,0 +1,131 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Load results
+  m1_post <- read_rds("Experiment 3/results/m1_post.rds")
+  
+  # Prepare data for figure
+  d_fig <- m1_post %>% mutate(order = kk)
+  
+
+# Figure 6 ----------------------------------------------------------------
+
+  # Make table
+  f_6_t <- d_fig %>% 
+    distinct(order, participant_ideology, participant_gender, protesters_cause) %>% 
+    rename(Gender = participant_gender, Ideology = participant_ideology, Cause = protesters_cause) %>% 
+    pivot_longer(
+      c(Gender, Ideology, Cause),
+      names_to = "column",
+      values_to = "text"
+    ) %>% 
+    mutate(
+      column = factor(column, levels = c("Ideology", "Gender", "Cause"))
+    ) %>% 
+    ggplot(., aes(x = 0, y = order)) +
+    geom_text(
+      aes(label = text),
+      size = 10/.pt,
+      hjust = 0
+    ) +
+    geom_hline(yintercept = c(0.5, 8.5), size = 0.455) +
+    scale_y_reverse(expand = c(0, 0)) +
+    xlim(0, 1) +
+    facet_grid(. ~ column) +
+    theme_grey(base_size = 10) +
+    theme(
+      legend.position = "none",
+      strip.background = element_blank(),
+      strip.text.x = element_text(size = 10, colour = "black", face = "bold"),
+      strip.text.y = element_blank(),
+      axis.text.x = element_text(colour = NA),
+      axis.ticks.x = element_line(colour = NA),
+      axis.title.x = element_text(colour = NA),
+      axis.text.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      axis.title.y = element_blank(),
+      panel.background = element_blank(),
+      plot.margin = margin(2, 1, 2, 0)
+    )
+  
+  # Make plot
+  f_6_p <- d_fig %>% 
[…]

### 2022-08-22 Experiment 3/figures/table-c1.R

@@ -1,0 +1,43 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("numform")
+  
+  # Load packages
+  library(tidyverse)
+  library(numform)
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 3/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  
+  # Summarize 
+  ds <- dl %>% 
+    group_by(action, action_text) %>% 
+    summarize(
+      Pr = mean(response)
+    ) %>% 
+    ungroup() %>% 
+    arrange(desc(Pr)) %>% 
+    transmute(
+      `#` = row_number(),
+      Action = action_text,
+      Pr = f_prop2percent(Pr, digits = 0)
+    )
+    
+
+# Export ------------------------------------------------------------------
+  
+  # Save as .rds
+  write_rds(ds, "Experiment 3/figures/table-c1.rds")

### 2022-08-22 Experiment 3/scripts/prepare_data.R

@@ -1,0 +1,309 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+  #########################################################################
+  # All identifying information (e.g., Prolific IDs) have been removed in #
+  # preprocessing and replaced with newly generated IDs.                  #
+  #########################################################################
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("janitor")
+  # devtools::install_github("ropenscilabs/gendercoder")
+  
+  # Load packages
+  library(tidyverse); library(janitor); library(gendercoder)
+
+
+# Import ------------------------------------------------------------------
+
+  # Import metadata from Prolific
+  dp <- read_csv("Experiment 3/data/raw/prolific_export.csv")
+  
+  # Import data from Qualtrics (Study 2)
+  dr <- read_csv("Experiment 3/data/raw/qualtrics_export.csv")[-1:-2,]
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Transform metadata (Prolific)
+  dp <- dp %>% 
+    clean_names() %>% 
+    mutate(
+      across(c(age), as.integer),
+      across(
+        where(is.character), 
+        ~if_else(. %in% c("CONSENT REVOKED", "DATA EXPIRED"), NA_character_, .)
+      )
+    )
+
+  # Transform data
+  dr <- dr %>% clean_names()
+
+  # Remove incomplete responses
+  dr <- dr %>% filter(finished == "1")
+  
+  # Code demographics
+  dr <- dr %>% 
+    mutate(
+      gender = recode_gender(
+        q64_1,
+        dictionary = fewlevels_en,
+        retain_unmatched = FALSE
+      ),
+      gender = case_when(
+        !is.na(gender) ~ gender,
+        q64_1 == "Female non-binary" ~ "sex and gender diverse",
+        q64_1 == "female/nonbinary" ~ "sex and gender diverse",
[…]

### 2022-08-22 Experiment 3/scripts/run_exploratory_analyses.R

@@ -1,0 +1,201 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+  
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 3/data/dl.rds")
+  dw <- read_rds("Experiment 3/data/dw.rds")
+  
+  # Exclude participants
+  dl <- dl %>% 
+    filter(!exclude, participant_gender %in% c("Man", "Woman")) %>% 
+    select(-exclude)
+  dw <- dw %>% 
+    filter(!exclude, participant_gender %in% c("Man", "Woman")) %>% 
+    select(-exclude)
+  
+  # Calculate factor scores
+  source("Experiment 3/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    mutate(
+      ide = case_when(
+        participant_gender == "Woman" ~ ide_women,
+        participant_gender == "Man" ~ ide_men
+      )
+    ) %>% 
+    transmute(
+      id,
+      across(
+        c(ide, sjb),
+        ~(. - mean(., na.rm = T))/sd(., na.rm = T),
+        .names = "z_{.col}"
+      )
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    drop_na() %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Man"   & protesters_cause == "For"     ~ 1L, 
[…]

### 2022-08-22 Experiment 3/scripts/run_factor_analyses.R

@@ -1,0 +1,28 @@
+# Load packages -----------------------------------------------------------
+library(lavaan)
+
+# System-Justifying Beliefs (SJB) -----------------------------------------
+
+  # Load model
+  sjb_model <- '
+        sjb =~ sjb_1 + sjb_2 + sjb_3 + sjb_4 + sjb_5 + sjb_6 + sjb_7 + sjb_8 
+      '
+  
+  # Run models
+  sjb_fit <- cfa(
+    sjb_model,
+    estimator = "mlr",
+    missing = "fiml",
+    std.lv = TRUE,
+    data = dw
+  )
+  
+  # Export factor scores
+  dw <- predict(sjb_fit) %>%
+    as.data.frame() %>%
+    bind_cols(dw, .)
+
+
+# Remove ------------------------------------------------------------------
+detach(package:lavaan)
+rm("sjb_model", "sjb_fit")

### 2022-08-22 Experiment 3/scripts/run_preregistered_analyses.R

@@ -1,0 +1,133 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+  
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 3/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% 
+    filter(!exclude, participant_gender %in% c("Man", "Woman")) %>% 
+    select(-exclude)
+  
+  # Code predictors
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Man"   & protesters_cause == "For"     ~ 1L, 
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For"     ~ 2L, 
+        participant_ideology == "Conservative" & participant_gender == "Man"   & protesters_cause == "Against" ~ 3L, 
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "Against" ~ 4L, 
+        participant_ideology == "Liberal"      & participant_gender == "Man"   & protesters_cause == "For"     ~ 5L, 
+        participant_ideology == "Liberal"      & participant_gender == "Woman" & protesters_cause == "For"     ~ 6L, 
+        participant_ideology == "Liberal"      & participant_gender == "Man"   & protesters_cause == "Against" ~ 7L, 
+        participant_ideology == "Liberal"      & participant_gender == "Woman" & protesters_cause == "Against" ~ 8L
+      ),
[…]

### 2022-08-26 Documents/manuscript.Rmd
first complete draft

@@
 In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
-We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no ingroup bias when judging collective action by ingroup and outgroup members.
+We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no consistent ingroup bias when judging collective action by ingroup and outgroup members.
 Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
@@
 In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
-We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show ingroup bias when judging collective action by ingroup and outgroup members.
+We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show consistent ingroup bias when judging collective action by ingroup and outgroup members.
 
@@
 Second, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters and what they are protesting.
-In Experiments 1 and 2, we operationalize double standards as the differences in $\theta_{(z)}$ between conditions which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
+In Experiments 1--3, we operationalize double standards as the differences in $\theta_{(z)}$ between conditions which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
 
@@
 In so doing, we build an instrument that enables both present and future research to capture double standards in judging collective action.
-In this way, our research situates the distinction between normative and non-normative collective action in the eye of the beholders and opens it to scientific investigation.
+<!-- In this way, our research situates the distinction between normative and non-normative collective action in the eye of the beholders and opens it to scientific investigation. -->
 
@@
 Extending Experiment 1, Experiment 2 considers both system-challenging and system-defending collective action in the context of the 2020 Black Lives Matter protests in the United States.
-By varying the protesters' cause (for/against defunding the police) as well as the participants' and the protesters' group memberships (Black/White), Experiment 2 compares the relative evidence identity-based (*Hypothesis 1a*, *1b*) and ideology-based double standards (*Hypothesis 2a*, *2b*).
+By varying the protesters' cause (for/against defunding the police) as well as the participants' and the protesters' group memberships (Black/White), Experiment 2 compares the relative evidence for identity-based (*Hypothesis 1a*, *1b*) and ideology-based double standards (*Hypothesis 2a*, *2b*).
+Building on Experiments 1--2, Experiment 3 examines how liberal and conservative men and women judge collective action for or against restricting abortion in anticipation of the Supreme Court overturning its *Roe v.
+Wade* decision on reproductive rights in the United States.
+By varying the protesters' cause as well as the participants' ideology and gender, Experiment 3 compares the relative evidence for identity-based (*Hypothesis 1b*) and ideology-based (*Hypothesis 2b*) double standards in judging collective action concerning a conservative cause in a sample balanced by gender and ideology.
 Together, the preregistered experiments provide a comprehensive test of the hypothesized double standards in judging collective action.
@@
 This finding might reflect our liberal-leaning sample.
+
 In line with *Hypothesis 2b*, we found that participants who rejected system-justifying beliefs considered the same protest action more acceptable when judging system-challenging collective action (for defunding the police).
@@
 In addition, we did not find evidence for the alternative, simpler *Hypothesis 3* that that participants would judge the same protest actions to be acceptable when they supported the protesters' cause.
-In this way, Experiment 2 provided evidence for ideology-based, but not for identity-based, double standards in judging collective action.
+
+One reason why we did not find evidence for ideological symmetry might be that participants judged protests for or against a progressive cause and that *opposing* a progressive cause is not equivalent to *supporting* a conservative cause.
+Another reason might be that our sample skewed liberal and lacked committed conservatives who might accept more extreme means to oppose a progressive cause.
+Experiment 3 addressed those limitations by testing for ideology-based double standards in judging collective action for or against a conservative cause in a balanced sample of conservatives and progressives.
 
 # Experiment 3
+
+In Experiment 3, we investigated potential double standards in judging collective action for or against restricting abortion in the United States.
+That is, we tested whether liberal and conservative women and men applied different standards when judging collective action opposing or supporting further restricting or banning abortion.
+We conducted this study in June 2022 when many expected the Supreme Court to overturn its *Roe v.
+Wade* decision that hitherto had guaranteed access to legal abortions in the first 20--24 weeks of pregnancy in the United States.
+We focused on restricting abortion as it is a conservative issue, as it primarily affects women's reproductive rights, and as the anticipated Supreme Court decision would prompt several states to restrict or ban abortion.
+
+By varying the participants' political orientation and the protesters' cause, we tested the preregistered hypothesis that, in line with *symmetrical* ideology-based double standards, conservatives would judge the same protest actions to be more acceptable if the protesters supported restricting abortion and that liberals would judge the same protest actions to be more acceptable if the protesters opposed restricting abortion.
+We tested the alternative hypothesis that, in line with *asymmetrical* ideology-based double standards, conservatives would judge all protest actions to be equally unacceptable while liberals would judge the same protest actions to be more acceptable if the protesters opposed restricting abortion.
+By varying the participants' gender, we further tested the preregistered hypothesis that, in line with identity-based double standards, female participants would judge protest actions against restricting abortion to be more acceptable, on average, than male participants or protest actions for restricting abortion.
 
@@
 
-We will measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
+We measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
 We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
@@
 ```{r, include = FALSE}
-
-  # THIS MIGHT BE WRONG
 
@@
 
-As in Experiments 1 and 2, we found no evidence for identity-based double standards: Female participants did not consider protest actions *against* restricting abortion to be any more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
-Instead, we again found evidence for ideology-based double standards: Conservative participants considered the same protest actions to be more acceptable, on average, when protesters *supported* restricting abortion (`r with(m1, d_text[contrast == "nh1_contrast"])`) while liberal participants considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion (`r with(m1, d_text[contrast == "nh2_contrast"])`).
-This pattern was consistent across subsamples as conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and as both liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) and women (`r with(m1, d_text[contrast == "nh6_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
-This double standard was, however, more pronounced among liberal participants (`r with(m1, d_text[contrast == "nh7_contrast"])`).
+As in Experiments 1 and 2, we found evidence for ideology-based double standards: Conservative participants considered the same protest actions to be more acceptable, on average, when protesters *supported* restricting abortion (`r with(m1, d_text[contrast == "nh1_contrast"])`) while liberal participants considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion (`r with(m1, d_text[contrast == "nh2_contrast"])`).
+This double standard was more pronounced among liberal participants (`r with(m1, d_text[contrast == "nh7_contrast"])`).
 In this way, Experiment 3 provided evidence for both ideological symmetry and asymmetry in judging collective action for and against restricting abortion.
+
+We found some evidence for identity-based double standards: Female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
+This pattern was, however, overshadowed by stronger ideology-based double standards that were consistent across subsamples: Conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and both liberal women (`r with(m1, d_text[contrast == "nh6_contrast"])`) and men (`r with(m1, d_text[contrast == "nh5_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
+In this way, Experiment 3 provided more consistent evidence for ideology-based than for identity-based double standards in judging collective action.
 
@@
 
-  # Define contrasts
-  ide <- ide %>% 
-    mutate(
-      h1_contrast = case_when(
-        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
-        TRUE ~ -1L,
-      ),
-      nh1_contrast = case_when(
-        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
-        participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
-        TRUE ~ NA_integer_
-      ),
-      nh2_contrast = case_when(
-        participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
-        participant_gender == "Man" & protesters_cause == "For" ~ -1L,
-        TRUE ~ NA_integer_
-      )
-    )
-
-  # Calculate contrasts
-  ide <- ide %>% 
-    pivot_longer(
-      ends_with("_contrast"),
-      names_to = "contrast",
-      values_to = "direction",
-      values_drop_na = TRUE
-    ) %>% 
-    group_by(.draw, z_ide, contrast, direction) %>% 
-    summarize(z = mean(z)) %>% 
-    summarize(d = sum(z * direction)) %>% 
-    group_by(z_ide, contrast) %>% 
-    median_qi(d)
-
   # Make text
   ide <- ide %>%
+    group_by(.draw, participant_gender, protesters_cause) %>% 
+    summarize(
+      b = mean(b_ide_kk)
+    ) %>% 
+    group_by(participant_gender, protesters_cause) %>%
+    median_qi(b) %>% 
     mutate(
       across(
-        c(d, .lower, .upper), 
+        c(b, .lower, .upper),
         f_num, digits = 2, retain.leading.zero = T
       ),
-      d_text = glue(
-        "$d = {d}, [{.lower}, {.upper}]$"
-      )
-    )
-
-```
-
-<!-- We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses.
-We found some evidence that, when considering only strongly identified (+1*SD*) men and women, female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(ide, d_text[contrast == "h1_contrast" & z_ide == 1L])`).
-Our non-preregistered analyses thus suggested that gender identification might moderate gender differences in judgements about collective action concerning reproductive rights.
--->
+      b_text = glue(
+        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
+        .open = "(", .close = ")"
+      )
+    )
+
+  # # Define contrasts
+  # ide <- ide %>% 
+  #   mutate(
+  #     h1_contrast = case_when(
+  #       participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+  #       TRUE ~ -1L,
+  #     ),
+  #     nh1_contrast = case_when(
+  #       participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+  #       participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
+  #       TRUE ~ NA_integer_
+  #     ),
+  #     nh2_contrast = case_when(
+  #       participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
+  #       participant_gender == "Man" & protesters_cause == "For" ~ -1L,
+  #       TRUE ~ NA_integer_
+  #     )
+  #   )
+  # 
+  # # Calculate contrasts
+  # ide <- ide %>% 
+  #   pivot_longer(
+  #     ends_with("_contrast"),
+  #     names_to = "contrast",
+  #     values_to = "direction",
+  #     values_drop_na = TRUE
+  #   ) %>% 
+  #   group_by(.draw, z_ide, contrast, direction) %>% 
+  #   summarize(z = mean(z)) %>% 
+  #   summarize(d = sum(z * direction)) %>% 
+  #   group_by(z_ide, contrast) %>% 
+  #   median_qi(d)
+  # 
+  # # Make text
+  # ide <- ide %>% 
+  #   mutate(
+  #     across(
+  #       c(d, .lower, .upper), 
+  #       f_num, digits = 2, retain.leading.zero = T
+  #     ),
+  #     d_text = glue(
+  #       "$d = {d}, [{.lower}, {.upper}]$"
+  #     )
+  #   )
+
+```
+
+We explored whether group identification moderated participants' judgements.
+We found, however, that gender identification neither affected how women judged protesters opposing (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "For"])`) abortion restrictions nor how men judged protesters opposing (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "For"])`) abortion restrictions.
+Our non-preregistered analyses thus suggested that gender identification did not moderate identity-based double standards in judging collective action concerning reproductive rights.
 
@@
 
+Experiment 3 tested for double standards in judging collective action for and against restricting abortion in a sample balanced by gender and political orientation.
+As in Experiments 1--2, participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their ideological orientation.
+While participants in Experiment 2 judged protests opposing or supporting a progressive cause (defunding police), participants in Experiment 3 considered protests opposing or supporting a conservative cause (restricting abortion).
+As both conservative and liberal participants judged protests for a cause aligned with their ideological orientation to be more acceptable and as this double standard was stronger among liberal participants, Experiment 3 provided evidence for both symmetry and asymmetry in ideology-based double standards in judging collective action.
+In contrast to Experiments 1--2, Experiment 3 provided some evidence for identity-based double standards, although any gender differences in judging protests for or against restricting abortion were overshadowed by more consistent ideology-based double standards.
+
 # General Discussion
@@
 Our findings, however, showed that it is not so simple.
-Contradicting our hypothesis of identity-based double standards, participants in two preregistered experiments did not show ingroup bias when judging controversial protest actions taken by ingroup and outgroup members to advance their group's interests.
-Instead, supporting our hypothesis of ideology-based double standards, participants judged the same controversial protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position.
+Contradicting our hypothesis of identity-based double standards, participants in three preregistered experiments did not show consistent ingroup bias---in terms of class, race, and gender---when judging controversial protest actions taken by ingroup and outgroup members to advance their group's interests.
+Instead, supporting our hypothesis of ideology-based double standards, participants judged the same controversial protest actions to be more acceptable when the protesters' cause---protecting workers' rights, defunding the police, restricting abortion---aligned with their own ideological position.
 <!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.
@@
 
-Despite its strengths, our research is qualified by important limitations.
+Despite its strengths, our research is qualified by several limitations.
 First, our research examined the hypothesized relationships for three political causes---protecting workers' rights, defunding the police, restricting abortion---and thus provided only limited evidence that our findings generalize beyond those causes.
@@
 This is particularly important when studying collective action as recent theorizing [@jost_missing_2017] and research [@osborne_protesting_2019] highlighted the importance of differentiating between progressive and conservative causes.
-Second, our research was based on samples from two Western, Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al., 2010) contexts---the United Kingdom and the United States---which limits the cross-cultural generalizability of our findings.
+Second, our research was based on samples from two Western, Educated, Industrialized, Rich, and Democratic [WEIRD, @henrich_weirdest_2010] contexts---the United Kingdom and the United States---which limits the cross-cultural generalizability of our findings.
 
 ## Implications
+
+Growing evidence shows that, in many situations, observers apply a more lenient standards when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
+And yet, our findings suggest that, unlike judgements in other domains, judgements about collective action are not subject to double standards based on group identities such as class, race, or gender---not even among highly identified group members.
+One reason might be that collective action, even in its controversial forms, is not judged as a moral transgression.
+Instead, observers might understand collective action primarily as a means to an end and judge how acceptable a form of collective action is in relation to its end.
+This aligns with our finding that how acceptable participants judged protest actions to be depended on the protesters' cause, not their group membership.
+
+Our findings showed that progressives and, to a lesser extent, conservatives consider the same controversial protest actions to be more acceptable when the protesters' cause aligns with their ideological position.
+So far, we explained this ideology-based double standard as resulting from conservatives being motivated to defend the system, progressives being motivated to challenge the system, and both being motivated to support collective action to those effects [@jost_missing_2017; @osborne_protesting_2019].
+An alternative explanation is to understand those double standards as an expression of ingroup bias based on an opinion-based [@mcgarty_collective_2009] or partisan [@finkel_political_2020] identity.
+This aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of the same transgressive protest actions when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
+Another explanation is that their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
+This aligns with Richardson & Conway's [-@richardson_standing_2022] finding that their different moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives and vice versa.
+Future research should seek to establish which of those plausible mechansims best explains partisan differences in judging collective action.
+
+Our findings raised the question why double standards in judging collective action were more pronounced among liberal than among conservative participants.
+One reason for this ideological asymmetry might be that collective action that violates prevailing norms or laws is seen as inherently threatening to the status quo, even if its aim is to bolster and defend the system.
+Political conservatism can be understood as having two psychologically related but distinguishable dimensions, resistance to change and justification of inequality [@jost_political_2003].
+In this way, conservatives might be expected to condemn disruptive protest even if the protest's cause is to defend inequality.
+Future research should seek to establish how consistent this ideological asymmetry is across different political causes and cultural contexts.
+
+<!-- Why do we find ideology-based double standards?
+We argued for system justification as the relevant distinction.
+One alternative explanation is opinion-based groups, see Verkuhyten et al.
+Another alternative explanation is moral concerns, see Richardson et al.
+Ultimately, ideology-based double standards are probably multiply determined.
+-->
+
+<!-- 1.
+There are well-documented identity-based double standards in other domains.
+And yet, we did not find consistent evidence that judgements about collective action are subject to double standards based on social categories such as class, race, and gender---not even among highly identified group members.
+Why is that?
+(1) Collective action is not a moral transgression.
+(2) Opinion-based groups is one way to look at ideology.
+-->
+
+<!-- 2.
+Our findings provide strong evidence for ideology-based double standards such that ...
+This aligns with other recent research that provides converging evidence.
+This all suggest that the right cause can justify the means of protest.
+-->
+
+<!-- 3.
+Why asymmetry?
+System Justification != Social Change == Collective Action.
+However, that we do find effects for a conservative cause ...
+One explanation might lie in the different moral values by ...
+moral conviction ...
+-->
 
@@
 -->
-
-Most broadly, our results showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
-Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
-In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.
-By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provided a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
-
-After all, this distinction can be consequential.
-If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
-In this way, our findings could explain why 45% of Republicans supported the storming of the U.S.
-Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
-If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
-In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].
-For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.
-
-# Conclusion
-
-Opinion polls show deep divides not only about the ends of protests, but also about the means of protest.
-For example, only 11\% of Republicans and 29\% of White Americans---compared to 59\% of Democrats and 66\% of Black Americans---considered kneeling during the national anthem---a non-violent form of collective action---an appropriate means to protest racist police violence [@yougov_nfl_2017].
-Our research showed that, in contrast to well-documented double standards in other domains, people do not apply a different standard when judging collective action by ingroup members or for a cause aligned with the ingroup's interests.
-Instead, our research demonstrated that people consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological positions---in other words, that there are partisan double standards in judging collective action.
 
@@
 <!-- - Invite researchers to use our paradigm -->
+
+Most broadly, our findings showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
+Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
+In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.
+By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provide a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
+
+After all, this distinction is often consequential.
+If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
+In this way, our findings could explain why 45% of Republicans supported the storming of the U.S.
+Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
+If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
+In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].
+For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.
+
+# Conclusion
+
+Opinion polls show deep divides not only about the ends of protests, but also about the means of protest.
+For example, only 11\% of Republicans and 29\% of White Americans---compared to 59\% of Democrats and 66\% of Black Americans---considered kneeling during the national anthem---a non-violent form of collective action---an appropriate means to protest racist police violence [@yougov_nfl_2017].
+Our research showed that, in contrast to well-documented double standards in other domains, people do not apply a different standard when judging collective action by members of the same class, race or gender or for a cause aligned with their own group's interests.
+Instead, our research demonstrated that people consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological positions---in other words, that there are partisan double standards in judging collective action.
 

### 2022-08-26 Documents/manuscript.tex
first complete draft

@@
 \affil[7]{KU Leuven, Belgium}
-
-\date{July 15, 2022}
 
@@
 the protesters' cause aligned with their own ideological position---but
-showed no ingroup bias when judging collective action by ingroup and
-outgroup members. Our findings have theoretical and practical
-implications for understanding the often divided response to prominent
-social movements.
+showed no consistent ingroup bias when judging collective action by
+ingroup and outgroup members. Our findings have theoretical and
+practical implications for understanding the often divided response to
+prominent social movements.
     \end{abstract}
@@
 when the protesters' cause aligns with their own ideological
-positions---but do not show ingroup bias when judging collective action
-by ingroup and outgroup members.
+positions---but do not show consistent ingroup bias when judging
+collective action by ingroup and outgroup members.
 
@@
 depending on who the protesters and what they are protesting. In
-Experiments 1 and 2, we operationalize double standards as the
-differences in \(\theta_{(z)}\) between conditions which, as
-\(\theta_{(z)}\) is \(z\)-standardized, correspond to Cohen's \(d\)
-effect sizes.
+Experiments 1--3, we operationalize double standards as the differences
+in \(\theta_{(z)}\) between conditions which, as \(\theta_{(z)}\) is
+\(z\)-standardized, correspond to Cohen's \(d\) effect sizes.
 
@@
 that enables both present and future research to capture double
-standards in judging collective action. In this way, our research
-situates the distinction between normative and non-normative collective
-action in the eye of the beholders and opens it to scientific
-investigation.
+standards in judging collective action.
 
@@
 and the protesters' group memberships (Black/White), Experiment 2
-compares the relative evidence identity-based (\emph{Hypothesis 1a},
+compares the relative evidence for identity-based (\emph{Hypothesis 1a},
 \emph{1b}) and ideology-based double standards (\emph{Hypothesis 2a},
-\emph{2b}). Together, the preregistered experiments provide a
+\emph{2b}). Building on Experiments 1--2, Experiment 3 examines how
+liberal and conservative men and women judge collective action for or
+against restricting abortion in anticipation of the Supreme Court
+overturning its \emph{Roe v. Wade} decision on reproductive rights in
+the United States. By varying the protesters' cause as well as the
+participants' ideology and gender, Experiment 3 compares the relative
+evidence for identity-based (\emph{Hypothesis 1b}) and ideology-based
+(\emph{Hypothesis 2b}) double standards in judging collective action
+concerning a conservative cause in a sample balanced by gender and
+ideology. Together, the preregistered experiments provide a
 comprehensive test of the hypothesized double standards in judging
@@
 defunding the police) than when judging system-defending collective
-action. This finding might reflect our liberal-leaning sample. In line
-with \emph{Hypothesis 2b}, we found that participants who rejected
-system-justifying beliefs considered the same protest action more
-acceptable when judging system-challenging collective action (for
+action. This finding might reflect our liberal-leaning sample.
+
+In line with \emph{Hypothesis 2b}, we found that participants who
+rejected system-justifying beliefs considered the same protest action
+more acceptable when judging system-challenging collective action (for
 defunding the police). We did not, however, find evidence for
@@
 \emph{Hypothesis 3} that that participants would judge the same protest
-actions to be acceptable when they supported the protesters' cause. In
-this way, Experiment 2 provided evidence for ideology-based, but not for
-identity-based, double standards in judging collective action.
+actions to be acceptable when they supported the protesters' cause.
+
+One reason why we did not find evidence for ideological symmetry might
+be that participants judged protests for or against a progressive cause
+and that \emph{opposing} a progressive cause is not equivalent to
+\emph{supporting} a conservative cause. Another reason might be that our
+sample skewed liberal and lacked committed conservatives who might
+accept more extreme means to oppose a progressive cause. Experiment 3
+addressed those limitations by testing for ideology-based double
+standards in judging collective action for or against a conservative
+cause in a balanced sample of conservatives and progressives.
 
@@
 \section{Experiment 3}\label{experiment-3}}
+
+In Experiment 3, we investigated potential double standards in judging
+collective action for or against restricting abortion in the United
+States. That is, we tested whether liberal and conservative women and
+men applied different standards when judging collective action opposing
+or supporting further restricting or banning abortion. We conducted this
+study in June 2022 when many expected the Supreme Court to overturn its
+\emph{Roe v. Wade} decision that hitherto had guaranteed access to legal
+abortions in the first 20--24 weeks of pregnancy in the United States.
+We focused on restricting abortion as it is a conservative issue, as it
+primarily affects women's reproductive rights, and as the anticipated
+Supreme Court decision would prompt several states to restrict or ban
+abortion.
+
+By varying the participants' political orientation and the protesters'
+cause, we tested the preregistered hypothesis that, in line with
+\emph{symmetrical} ideology-based double standards, conservatives would
+judge the same protest actions to be more acceptable if the protesters
+supported restricting abortion and that liberals would judge the same
+protest actions to be more acceptable if the protesters opposed
+restricting abortion. We tested the alternative hypothesis that, in line
+with \emph{asymmetrical} ideology-based double standards, conservatives
+would judge all protest actions to be equally unacceptable while
+liberals would judge the same protest actions to be more acceptable if
+the protesters opposed restricting abortion. By varying the
+participants' gender, we further tested the preregistered hypothesis
+that, in line with identity-based double standards, female participants
+would judge protest actions against restricting abortion to be more
+acceptable, on average, than male participants or protest actions for
+restricting abortion.
 
@@
 
-We will measured moral concerns with the 36-item moral foundations
+We measured moral concerns with the 36-item moral foundations
 questionnaire (MFQ-2, Atari et al., 2022) which assesses to what extent
@@
 
-As in Experiments 1 and 2, we found no evidence for identity-based
-double standards: Female participants did not consider protest actions
-\emph{against} restricting abortion to be any more acceptable, on
-average, than male participants or protest actions \emph{for}
-restricting abortion (\(d = 0.14, [0.00, 0.29]\)). Instead, we again
-found evidence for ideology-based double standards: Conservative
-participants considered the same protest actions to be more acceptable,
-on average, when protesters \emph{supported} restricting abortion
-(\(d = 0.24, [0.07, 0.43]\)) while liberal participants considered the
-same protest actions to be more acceptable when protesters
-\emph{opposed} restricting abortion (\(d = 0.90, [0.72, 1.08]\)). This
-pattern was consistent across subsamples as conservative women
-(\(d = 0.30, [0.04, 0.55]\)) and, to a lesser extent, conservative men
-(\(d = 0.19, [-0.07, 0.46]\)) considered the same protest actions to be
-more acceptable when protesters \emph{supported} restricting abortion
-and as both liberal men (\(d = 0.76, [0.49, 1.03]\)) and women
-(\(d = 1.04, [0.78, 1.32]\)) considered the same protest actions to be
-more acceptable when protesters \emph{opposed} restricting abortion.
-This double standard was, however, more pronounced among liberal
-participants (\(d = 0.66, [0.41, 0.92]\)). In this way, Experiment 3
-provided evidence for both ideological symmetry and asymmetry in judging
-collective action for and against restricting abortion.
+As in Experiments 1 and 2, we found evidence for ideology-based double
+standards: Conservative participants considered the same protest actions
+to be more acceptable, on average, when protesters \emph{supported}
+restricting abortion (\(d = 0.24, [0.07, 0.43]\)) while liberal
+participants considered the same protest actions to be more acceptable
+when protesters \emph{opposed} restricting abortion
+(\(d = 0.90, [0.72, 1.08]\)). This double standard was more pronounced
+among liberal participants (\(d = 0.66, [0.41, 0.92]\)). In this way,
+Experiment 3 provided evidence for both ideological symmetry and
+asymmetry in judging collective action for and against restricting
+abortion.
+
+We found some evidence for identity-based double standards: Female
+participants considered protest actions \emph{against} restricting
+abortion to be more acceptable, on average, than male participants or
+protest actions \emph{for} restricting abortion
+(\(d = 0.14, [0.00, 0.29]\)). This pattern was, however, overshadowed by
+stronger ideology-based double standards that were consistent across
+subsamples: Conservative women (\(d = 0.30, [0.04, 0.55]\)) and, to a
+lesser extent, conservative men (\(d = 0.19, [-0.07, 0.46]\)) considered
+the same protest actions to be more acceptable when protesters
+\emph{supported} restricting abortion and both liberal women
+(\(d = 1.04, [0.78, 1.32]\)) and men (\(d = 0.76, [0.49, 1.03]\))
+considered the same protest actions to be more acceptable when
+protesters \emph{opposed} restricting abortion. In this way, Experiment
+3 provided more consistent evidence for ideology-based than for
+identity-based double standards in judging collective action.
 
@@
 
+We explored whether group identification moderated participants'
+judgements. We found, however, that gender identification neither
+affected how women judged protesters opposing
+(\(\beta_{xy} = -0.04, [-0.14, 0.07]\)) or supporting
+(\(\beta_{xy} = -0.06, [-0.17, 0.03]\)) abortion restrictions nor how
+men judged protesters opposing (\(\beta_{xy} = -0.05, [-0.14, 0.04]\))
+or supporting (\(\beta_{xy} = -0.04, [-0.12, 0.06]\)) abortion
+restrictions. Our non-preregistered analyses thus suggested that gender
+identification did not moderate identity-based double standards in
+judging collective action concerning reproductive rights.
+
 \hypertarget{discussion-2}{%
 \subsection{Discussion}\label{discussion-2}}
+
+Experiment 3 tested for double standards in judging collective action
+for and against restricting abortion in a sample balanced by gender and
+political orientation. As in Experiments 1--2, participants judged the
+same protest actions to be more acceptable when the protesters' cause
+aligned with their ideological orientation. While participants in
+Experiment 2 judged protests opposing or supporting a progressive cause
+(defunding police), participants in Experiment 3 considered protests
+opposing or supporting a conservative cause (restricting abortion). As
+both conservative and liberal participants judged protests for a cause
+aligned with their ideological orientation to be more acceptable and as
+this double standard was stronger among liberal participants, Experiment
+3 provided evidence for both symmetry and asymmetry in ideology-based
+double standards in judging collective action. In contrast to
+Experiments 1--2, Experiment 3 provided some evidence for identity-based
+double standards, although any gender differences in judging protests
+for or against restricting abortion were overshadowed by more consistent
+ideology-based double standards.
 
@@
 so simple. Contradicting our hypothesis of identity-based double
-standards, participants in two preregistered experiments did not show
-ingroup bias when judging controversial protest actions taken by ingroup
-and outgroup members to advance their group's interests. Instead,
-supporting our hypothesis of ideology-based double standards,
-participants judged the same controversial protest actions to be more
-acceptable when the protesters' cause aligned with their own
-(system-challenging) ideological position. In the remainder of this
-General Discussion, we consider some limiting conditions on our findings
-and discuss theoretical and practical implications for understanding the
-often divided response to collective action.
+standards, participants in three preregistered experiments did not show
+consistent ingroup bias---in terms of class, race, and gender---when
+judging controversial protest actions taken by ingroup and outgroup
+members to advance their group's interests. Instead, supporting our
+hypothesis of ideology-based double standards, participants judged the
+same controversial protest actions to be more acceptable when the
+protesters' cause---protecting workers' rights, defunding the police,
+restricting abortion---aligned with their own ideological position. In
+the remainder of this General Discussion, we consider some limiting
+conditions on our findings and discuss theoretical and practical
+implications for understanding the often divided response to collective
+action.
 
@@
 
-Despite its strengths, our research is qualified by important
-limitations. First, our research examined the hypothesized relationships
-for three political causes---protecting workers' rights, defunding the
-police, restricting abortion---and thus provided only limited evidence
-that our findings generalize beyond those causes. Relying on few stimuli
-to establish an effect threatens the validity and replicability of
-research findings and constrains the generalizability of psychological
-research (Yarkoni, 2022). Future research should sample a wider range of
-causes to address this pervasive but often ignored problem (Judd et al.,
-2012). This is particularly important when studying collective action as
-recent theorizing (Jost et al., 2017) and research (Osborne et al.,
-2019) highlighted the importance of differentiating between progressive
-and conservative causes. Second, our research was based on samples from
-two Western, Educated, Industrialized, Rich, and Democratic (WEIRD;
-Henrich et al., 2010) contexts---the United Kingdom and the United
+Despite its strengths, our research is qualified by several limitations.
+First, our research examined the hypothesized relationships for three
+political causes---protecting workers' rights, defunding the police,
+restricting abortion---and thus provided only limited evidence that our
+findings generalize beyond those causes. Relying on few stimuli to
+establish an effect threatens the validity and replicability of research
+findings and constrains the generalizability of psychological research
+(Yarkoni, 2022). Future research should sample a wider range of causes
+to address this pervasive but often ignored problem (Judd et al., 2012).
+This is particularly important when studying collective action as recent
+theorizing (Jost et al., 2017) and research (Osborne et al., 2019)
+highlighted the importance of differentiating between progressive and
+conservative causes. Second, our research was based on samples from two
+Western, Educated, Industrialized, Rich, and Democratic (WEIRD, Henrich
+et al., 2010) contexts---the United Kingdom and the United
 States---which limits the cross-cultural generalizability of our
@@
 
-Most broadly, our results showed that, for many controversial protest
+Growing evidence shows that, in many situations, observers apply a more
+lenient standards when judging moral transgressions by ingroup members
+(Abrams et al., 2013; Endevelt et al., 2021). And yet, our findings
+suggest that, unlike judgements in other domains, judgements about
+collective action are not subject to double standards based on group
+identities such as class, race, or gender---not even among highly
+identified group members. One reason might be that collective action,
+even in its controversial forms, is not judged as a moral transgression.
+Instead, observers might understand collective action primarily as a
+means to an end and judge how acceptable a form of collective action is
+in relation to its end. This aligns with our finding that how acceptable
+participants judged protest actions to be depended on the protesters'
+cause, not their group membership.
+
+Our findings showed that progressives and, to a lesser extent,
+conservatives consider the same controversial protest actions to be more
+acceptable when the protesters' cause aligns with their ideological
+position. So far, we explained this ideology-based double standard as
+resulting from conservatives being motivated to defend the system,
+progressives being motivated to challenge the system, and both being
+motivated to support collective action to those effects (Jost et al.,
+2017; Osborne et al., 2019). An alternative explanation is to understand
+those double standards as an expression of ingroup bias based on an
+opinion-based (McGarty et al., 2009) or partisan (Finkel et al., 2020)
+identity. This aligns with Verkuyten et al.'s (2022) finding that
+participants were more tolerant of the same transgressive protest
+actions when taken by their most-liked, rather than their least-liked,
+political group (e.g., climate activists). Another explanation is that
+their different moral concerns (Graham et al., 2009; Kivikangas et al.,
+2021) lead progressives and conservatives to see their support of
+certain causes as a fundamental matter of right or wrong (Skitka et al.,
+2021) and, by thus moralizing those causes, to accept more extreme means
+to achieving them. This aligns with Richardson \& Conway's (2022)
+finding that their different moral concerns explained why liberals rated
+protesting for liberal causes as more moral than conservatives and vice
+versa. Future research should seek to establish which of those plausible
+mechansims best explains partisan differences in judging collective
+action.
+
+Our findings raised the question why double standards in judging
+collective action were more pronounced among liberal than among
+conservative participants. One reason for this ideological asymmetry
+might be that collective action that violates prevailing norms or laws
+is seen as inherently threatening to the status quo, even if its aim is
+to bolster and defend the system. Political conservatism can be
+understood as having two psychologically related but distinguishable
+dimensions, resistance to change and justification of inequality (Jost
+et al., 2003). In this way, conservatives might be expected to condemn
+disruptive protest even if the protest's cause is to defend inequality.
+Future research should seek to establish how consistent this ideological
+asymmetry is across different political causes and cultural contexts.
+
+Most broadly, our findings showed that, for many controversial protest
 actions, there is no consensus on what is and is not an acceptable means
@@
 measuring where people draw the line between acceptable and unacceptable
-forms of collective action, we provided a paradigm that we hope will
+forms of collective action, we provide a paradigm that we hope will
 stimulate research on when and why people differ in making this
@@
 
-After all, this distinction can be consequential. If people believe that
-even extreme actions are acceptable means to advance a cause they
+After all, this distinction is often consequential. If people believe
+that even extreme actions are acceptable means to advance a cause they
 support, especially if they moralize their support for that cause
@@
 contrast to well-documented double standards in other domains, people do
-not apply a different standard when judging collective action by ingroup
-members or for a cause aligned with the ingroup's interests. Instead,
-our research demonstrated that people consider the same protest actions
-to be more acceptable when the cause of the protest aligns with their
-own ideological positions---in other words, that there are partisan
-double standards in judging collective action.
+not apply a different standard when judging collective action by members
+of the same class, race or gender or for a cause aligned with their own
+group's interests. Instead, our research demonstrated that people
+consider the same protest actions to be more acceptable when the cause
+of the protest aligns with their own ideological positions---in other
+words, that there are partisan double standards in judging collective
+action.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-finkel_political_2020}{}}%
+Finkel, E. J., Bail, C. A., Cikara, M., Ditto, P. H., Iyengar, S., Klar,
+S., Mason, L., McGrath, M. C., Nyhan, B., Rand, D. G., Skitka, L. J.,
+Tucker, J. A., Van Bavel, J. J., Wang, C. S., \& Druckman, J. N. (2020).
+Political sectarianism in {America}. \emph{Science}, \emph{370}(6516),
+533--536. \url{https://doi.org/10.1126/science.abe1715}
+
 \leavevmode\vadjust pre{\hypertarget{ref-foschi_double_2000}{}}%
@@
 \emph{19, 10}, 555. \url{https://doi.org/10.3390/e19100555}
+
+\leavevmode\vadjust pre{\hypertarget{ref-graham_liberals_2009}{}}%
+Graham, J., Haidt, J., \& Nosek, B. A. (2009). Liberals and
+conservatives rely on different sets of moral foundations. \emph{Journal
+of Personality and Social Psychology}, \emph{96}(5), 1029--1046.
+\url{https://doi.org/10.1037/a0015141}
+
+\leavevmode\vadjust pre{\hypertarget{ref-henrich_weirdest_2010}{}}%
+Henrich, J., Heine, S. J., \& Norenzayan, A. (2010). The weirdest people
+in the world? \emph{Behavioral and Brain Sciences}, \emph{33}(2-3),
+61--135. \url{https://doi.org/10.1017/S0140525X10000725}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kivikangas_moral_2021}{}}%
+Kivikangas, J. M., Fernández-Castilla, B., Järvelä, S., Ravaja, N., \&
+Lönnqvist, J.-E. (2021). Moral foundations and political orientation:
+{Systematic} review and meta-analysis. \emph{Psychological Bulletin},
+\emph{147}(1), 55--94. \url{https://doi.org/10.1037/bul0000308}
+
 \leavevmode\vadjust pre{\hypertarget{ref-lacina_nfl_2020}{}}%
@@
 \url{http://www.bethanylacina.com/docs/Lacina_NFL_paper.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcgarty_collective_2009}{}}%
+McGarty, C., Bliuc, A.-M., Thomas, E. F., \& Bongiorno, R. (2009).
+Collective action as the material expression of opinion-based group
+membership. \emph{Journal of Social Issues}, \emph{65}(4), 839--857.
+\url{https://doi.org/10.1111/j.1540-4560.2009.01627.x}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-richardson_standing_2022}{}}%
+Richardson, I., \& Conway, P. (2022). Standing up or giving up? {Moral}
+foundations mediate political differences in evaluations of {BLACK}
+{LIVES} {MATTER} and other protests. \emph{European Journal of Social
+Psychology}, \emph{52}(3), 553--569.
+\url{https://doi.org/10.1002/ejsp.2837}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ryan_reconsidering_2014}{}}%
@@
 \url{https://doi.org/10.1037/pspi0000333}
+
+\leavevmode\vadjust pre{\hypertarget{ref-skitka_psychology_2021}{}}%
+Skitka, L. J., Hanson, B. E., Morgan, G. S., \& Wisneski, D. C. (2021).
+The psychology of moral conviction. \emph{Annual Review of Psychology},
+\emph{72}(1), 347--366.
+\url{https://doi.org/10.1146/annurev-psych-063020-030612}
 

### 2022-08-26 Experiment 3/scripts/run_exploratory_analyses.R

@@ -118,7 +118,7 @@ rm(list = ls())
   # Extract posterior draws
   ide_post_ii <- ide_ps %>% spread_draws(alpha[ii], beta[ii]) %>% ungroup()
   ide_post_jj <- ide_ps %>% spread_draws(theta[jj]) %>% ungroup()
-  ide_post_kk <- ide_ps %>% spread_draws(delta[kk], b_ide_kk[kk]) %>% ungroup()
+  ide_post_kk <- ide_ps %>% spread_draws(delta[kk], b_ide_kk[kk], b_ide) %>% ungroup()
   
   # Calculate effect sizes
   ide_post <- dl %>% 
@@ -126,25 +126,43 @@ rm(list = ls())
     left_join(ide_post_kk, by = c("kk")) %>% 
     left_join(ide_post_jj, by = c("jj", ".draw", ".chain", ".iteration")) %>% 
     group_by(.draw) %>% 
-    summarize(
-      mean = mean(theta + delta + b_ide_kk * z_ide),
-      sd = sd(theta + delta + b_ide_kk * z_ide)
-    ) %>% 
-    crossing(
-      kk = 1:8, 
-      z_ide = -1:1
-    ) %>% 
-    left_join(ide_post_kk, by = c(".draw", "kk")) %>% 
     mutate(
-      z = ((delta + b_ide_kk * z_ide) - mean)/sd
+      b_ide = b_ide/sd(theta + delta + b_ide_kk*z_ide),
+      b_ide_kk = b_ide_kk/sd(theta + delta + b_ide_kk*z_ide),
     ) %>% 
-    right_join(
-      dl %>% 
-        select(kk, participant_ideology, participant_gender, protesters_cause) %>% 
-        distinct(),
-      by = "kk"
-    ) %>% 
-    select(.draw, kk, z_ide, z, participant_ideology, participant_gender, protesters_cause)
+    ungroup() %>%
+    distinct(.chain, .iteration, .draw, kk, b_ide, b_ide_kk) %>% 
+      right_join(
+        dl %>%
+          select(kk, participant_ideology, participant_gender, protesters_cause) %>%
+          distinct(),
+        by = "kk"
+      )
+    
+  # ide_post <- dl %>% 
+  #   distinct(jj, kk, z_ide) %>% 
+  #   left_join(ide_post_kk, by = c("kk")) %>% 
[…]

### 2022-09-13 Documents/manuscript.Rmd
preprint

@@
 Kneeling during the national anthem is not violent, disruptive, or illegal.
-And yet, only 29\% of White Americans (66\% of Black Americans) and 11\% of Republicans (59\% of Democrats) considered it appropriate for Black athletes to protest in this way [@yougov_nfl_2017].
-Indeed, Black athletes who had participated in the protests faced a backlash in public opinion [@lacina_nfl_2020].
-<!-- Likewise, 42\% of Americans thought that, in the 2020 Black Lives Matter protests, most protesters were trying to incite violence or destroy property, even though very few protests involved any violence or property destruction [@acled_year_2021].
--->
-This and other examples suggest that *who* the protesters are and *what* they are protesting influences how acceptable observers judge their actions to be.
+And yet, only 29\% of White Americans (compared to 66\% of Black Americans) and 11\% of Republicans (compared to 59\% of Democrats) considered it appropriate for Black athletes to protest in this way [@yougov_nfl_2017].
+Indeed, Black athletes who had participated in the protests faced a backlash in public opinion [@lacina_nfl_2020].<!-- Likewise, 42\% of Americans thought that, in the 2020 Black Lives Matter protests, most protesters were trying to incite violence or destroy property, even though very few protests involved any violence or property destruction [@acled_year_2021].
+--> This and other examples suggest that *who* the protesters are and *what* they are protesting influences how acceptable observers judge their actions to be.
 
@@
 
-<!-- In the next sections, we, first, introduce the consequential distinction between normative and nonnormative collective action and define double standards in judging what actions are acceptable and unacceptable means to effect social change.
-We, second, derive predictions from social psychological theories about identity- and ideology-based double standards that might affect judgements about collective action.
-We, third, introduce item response theory as a conceptual and statistical framework for study double standards in judging collective action.
-We, finally, provide an overview of the present research.
--->
-
 ## Double Standards in Judging Collective Action
@@
 Collective action can take many forms.
-Psychologists distinguish between *normative* collective action that seeks to achieve a political goal while conforming to the norms of the existing social system and *non-normative* collective action that violates those norms [@wright_responding_1990].
+Psychologists distinguish between *normative* collective action that seeks to achieve a political goal while conforming to the norms of the existing social system and *non-normative* collective action that violates those norms [@wright_responding_1990; @becker_dynamic_2015].
+Some instead refer to the two kinds of collective action as moderate and radical collective action [e.g., @jimenez-moya_by_2015].
+Broadly, this distinction captures the intuition that there are some forms of collective action that are generally considered acceptable and others that are not.
+
 Recent research has shown that this distinction matters for how people respond to collective action.
@@
 
-Research has, for the most part, relied on ad-hoc distinctions between what researchers themselves considered normative and non-normative collective action.
+These and other studies have, for the most part, relied on ad-hoc distinctions between what researchers themselves considered normative and non-normative collective action.
 In liberal democracies, researchers tend to consider actions such as signing petitions, voting in elections, or peaceful protest to be normative and actions such as blocking traffic, damaging property, or violent protest to be non-normative.
-But, as the examples in the first paragraph show, the distinction between normative and non-normative collective action might instead be in the eye of the beholder.
-
-In this research, we move this distinction from the researcher's intuition into the realm of scientific investigation and test for double standards in where people draw the line between acceptable and unacceptable forms of collective action.
+But, as reactions to Black athletes kneeling in protest suggest, the same action might be considered acceptable by some and unacceptable by others.
+That is, the distinction between normative and non-normative collective action might be in the eye of the beholder.
+
+In this research, we move this distinction from the researcher's intuition into the realm of scientific investigation and test for double standards in where people draw the line between acceptable (normative) and unacceptable (non-normative) forms of collective action.
 For that purpose, we define a double standard as judging the same action as more or less acceptable depending on who is performing the action and for what reason [for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.] In this section, we derive predictions from social psychological theories about how identity-based and ideology-based double standards could lead people to judge the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting.
@@
 
-One domain in which ingroup bias manifests is judgements about other people's actions.
-@hewstone_ultimate_1990 reviewed research showing that people make ingroup-serving causal attributions when judging actions by ingroup and outgroup members, for example, by attributing negative behavior by outgroup members to internal causes but attributing negative behavior by ingroup members to external causes.
+One domain in which ingroup bias manifests is judgments about other people's actions.
+@hewstone_ultimate_1990 reviewed research showing that people make ingroup-serving causal attributions when judging actions by ingroup and outgroup members, for example, by attributing negative behavior by outgroup members to internal causes but attributing negative behavior by ingroup members to external causes [see @pettigrew_ultimate_1979].
 @valdesolo_moral_2007 demonstrated that participants judged the same selfish action as less unfair when performed by themselves or an ingroup member than when performed by an outgroup member.
@@
 
-System Justification Theory [@jost_role_1994; for a recent review, see @jost_theory_2020] argues that people are motivated to defend, justify, and bolster the prevailing social, economic, and political system because doing so serves basic epistemic, existential, and relational needs.
+System Justification Theory [@jost_decade_2004; for a recent review, see @jost_theory_2020] argues that people are motivated to defend, justify, and bolster the prevailing social, economic, and political system because doing so serves basic epistemic, existential, and relational needs.
 Both advantaged and disadvantaged groups are thought to be motivated to justify the existing system, although the strength of this motivation and its expression vary across individuals and situations.
@@
 
-Just as system justification shapes support for, and opposition to, collective action, we propose that it results in ideology-based double standards in what actions people consider acceptable forms of collective action.
+Just as system justification affects support for, and opposition to, collective action, we propose that it results in ideology-based double standards in what actions people consider acceptable forms of collective action.
 First, we hypothesize that  people will judge the same protest actions as more acceptable if the protest supports, rather than challenges, the system (*Hypothesis 2a*).
@@
 
-Item response theory is a a conceptual and statistical framework for understanding how the characteristics of both items and respondents shape responses to a set of items [@demars_irt_2010]. 
-<!-- As such, it allows us to disentangle how acceptable protest actions are from how accepting respondents are of them and, in so doing, to detect double standards in judging collective action.  -->
-When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters:
+Item response theory is a conceptual and statistical framework for understanding how the characteristics of both items and respondents shape responses to a set of items [@demars_irt_2010].
+<!-- As such, it allows us to disentangle how acceptable protest actions are from how accepting respondents are of them and, in so doing, to detect double standards in judging collective action.
+--> When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two-parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters:
 $$ \Pr ( y_{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
@@
 First, item response theory provides a statistical framework for scale development.
-In Studies 1 and 2, we select the most discriminating and informative actions to build an instrument for capturing double standards in the experiments.
-Second, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters and what they are protesting.
+In Studies 1 and 2, we select the most discriminating and informative actions to build an instrument for capturing double standards in judging collective action.
+Second, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters are and what they are protesting.
 In Experiments 1--3, we operationalize double standards as the differences in $\theta_{(z)}$ between conditions which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
@@
 In Study 1, we ask participants to generate more and less extreme protest actions and compile a pool of protest actions for scale development.
-In Study 2, we ask participants to rate how acceptable they consider each of those protest action to be and use item response theory to select the most discriminating and informative protest actions for our instrument.
+In Study 2, we ask participants to rate how acceptable they consider each of those protest actions to be and use item response theory to select the most discriminating and informative protest actions for our instrument.
 In so doing, we build an instrument that enables both present and future research to capture double standards in judging collective action.
-<!-- In this way, our research situates the distinction between normative and non-normative collective action in the eye of the beholders and opens it to scientific investigation. -->
+<!-- In this way, our research situates the distinction between normative and non-normative collective action in the eye of the beholders and opens it to scientific investigation.
+-->
 
@@
 
-<!-- > **Hypothesis 1.** People will consider *more* controversial actions acceptable means of protest if (a) the protesters are ingroup members *or* if (b) the cause of the protest aligns with the ingroup's interests.
--->
-
-<!-- > **Hypothesis 2.** People will consider *more* controversial actions acceptable means of protest if (a) the protest's cause supports, rather than challenges, the system *or* if (b) the protest's cause aligns with their system-justifying beliefs.
--->
-
 # Scale Development
@@
 Assuming that collective action by lower-status group members is seen as more threatening to the unequal status quo than collective action by higher-status group members and that people are, to same extent, motivated to justify the system, we tested the preregistered hypothesis that, in line with *Hypothesis 2a*, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group than if they were from a lower-status group.
-In non-preregistered analyses, we further explored whether, in line with *Hypothesis 2b*, participants would judge the same protest actions as more or less acceptable depending on their own ideological positions.
+In non-preregistered analyses, we further explored whether participants would judge the same protest actions as more or less acceptable depending on their own ideological positions.
 In this way, Experiment 1 provided a first test of the hypothesized identity- and ideology-based double standards in judging collective action.
@@
 Participants in this lower-status group did not have a university degree and placed themselves on the bottom three ranks of the subjective socioeconomic status ladder.
-Another 233 participants considered their past, current job, or future jobs to be middle-class/professional jobs.
+Another 233 participants considered their past, current, or future jobs to be middle-class/professional jobs.
 Participants in this higher-status group had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socioeconomic status ladder.
@@
 
-We assessed reactions to the vignette by asking participants how outraged they would be if the government were to introduce this bill in real life, to what extent it would affect them personally, and to what extent the it would affect people like them (1 = *not at all*, 7 = *very much*).
+We assessed reactions to the vignette by asking participants how outraged they would be if the government were to introduce this bill in real life, to what extent it would affect them personally, and to what extent it would affect people like them (1 = *not at all*, 7 = *very much*).
 We also asked participants to what extent they identified with people with the kinds of jobs most affected by the proposed bill (1 = *not at all*, 7 = *very much*) and how often, if at all, they had participated in protest actions such as the ones we asked about (1 = *never*, 5 = *very often*).
@@
 We found that participants who reported a more right-wing political orientation ($\beta_{xy} = -0.25, [-0.37, -0.15]$) and who expressed more agreement with system-justifying beliefs ($\beta_{xy} = -0.20, [-0.30, -0.10]$) tended to find fewer collective actions to be acceptable.
-In contrast, we found that, after controlling for the other two variables, social dominance orientation was not associated with participants' judgements about how acceptable various collective actions are ($\beta_{xy} = -0.05, [-0.15, 0.05]$).
+In contrast, we found that, after controlling for the other two variables, social dominance orientation was not associated with participants' judgments about how acceptable various collective actions are ($\beta_{xy} = -0.05, [-0.15, 0.05]$).
 Overall, these findings suggest that people who are right-wing and endorse system-justifying beliefs tend to find various collective actions to be less acceptable than people who are left-wing and do not endorse system-justifying beliefs.^[We ran additional analyses to explore how these associations differed across experimental conditions.
@@
 Experiment 1 did not, however, provide a conclusive test of our hypotheses.
-First, by focusing on fictitious government bills, we might have chosen a scenario too far removed from current affairs to evoke strong ingroup bias in participants' judgements of protest actions.
+First, by focusing on fictitious government bills, we might have chosen a scenario too far removed from current affairs to evoke strong ingroup bias in participants' judgments of protest actions.
 That said, participants' reactions to the experimental manipulation suggested that it evoked outrage and was understood in ingroup--outgroup terms.
@@
 That is, we tested whether Black and White Americans applied different standards when judging collective action taken by either Black or White protesters to protest either for or against police divestment as a possible solution to racist police violence.
-We conducted this study in January 2021, a time when, after a year of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement.
+We conducted this study in January 2021, a time when, after months of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement.
 We focused on defunding the police as this position remained controversial among both Black and White Americans even as they broadly agreed on the need for police reform.^[In December 2020, only 34% of Black Americans and 18% of White Americans supported defunding the police while majorities of Black and White Americans supported police reforms such as eliminating qualified immunity (50% and 57%) or banning chokeholds [79% and 61%, @yougov_police_2020].]
@@
 Black participants strongly identified with their ingroup ($M = 6.46, \textit{SD} = 1.09$) and, across experimental conditions, identified more with Black protesters ($M = 4.66, \textit{SD} = 1.99$) than with White protesters ($M = 4.03, \textit{SD} = 2.16$; $d = 0.30$).
-In contrast, White participants identified less strongly with their ingroup ($M = 4.85, \textit{SD} = 1.53$; $d = 1.21$) and did not identify more with White protesters ($M = 3.58, \textit{SD} = 2.01$) than with Black protesters ($M = 3.81, \textit{SD} = 1.77$; $d = -0.12$).
+In contrast, White participants identified less strongly with their ingroup than Black participants ($M = 4.85, \textit{SD} = 1.53$; $d = 1.21$) and did not identify more with White protesters ($M = 3.58, \textit{SD} = 2.01$) than with Black protesters ($M = 3.81, \textit{SD} = 1.77$; $d = -0.12$).
 Both White ($d = 0.67$) and Black ($d = 0.53$) participants tended to identify more with protesters protesting *for* defunding the police.
@@
 Model 2 (M2) extended Model 1 by estimating participants' responses as a function of their *z*-standardized endorsement of system-justifying beliefs.
-As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgements about system-defending protest actions and one estimating their effect on judgements about system-challenging protest actions, and one varying effect estimating its variance across conditions.
+As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgments about system-defending protest actions and one estimating their effect on judgments about system-challenging protest actions, and one varying effect estimating its variance across conditions.
 Supporting *Hypothesis 2b*, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`).
@@
 We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast" & x_ide == 1L])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast" & x_ide == 1L])`).
-Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgements about collective action by ingroup and outgroup members.
+Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgments about collective action by ingroup and outgroup members.
 
@@
 In Experiment 3, we investigated potential double standards in judging collective action for or against restricting abortion in the United States.
-That is, we tested whether liberal and conservative women and men applied different standards when judging collective action opposing or supporting further restricting or banning abortion.
+That is, we tested whether liberal and conservative women and men applied different standards when judging collective action opposing or supporting further restricting abortion.
 We conducted this study in June 2022 when many expected the Supreme Court to overturn its *Roe v.
@@
 
-By varying the participants' political orientation and the protesters' cause, we tested the preregistered hypothesis that, in line with *symmetrical* ideology-based double standards, conservatives would judge the same protest actions to be more acceptable if the protesters supported restricting abortion and that liberals would judge the same protest actions to be more acceptable if the protesters opposed restricting abortion.
+By comparing participants of different political orientations and by varying the protesters' cause, we tested the preregistered hypothesis that, in line with *symmetrical* ideology-based double standards, conservatives would judge the same protest actions to be more acceptable if the protesters supported restricting abortion and that liberals would judge the same protest actions to be more acceptable if the protesters opposed restricting abortion.
 We tested the alternative hypothesis that, in line with *asymmetrical* ideology-based double standards, conservatives would judge all protest actions to be equally unacceptable while liberals would judge the same protest actions to be more acceptable if the protesters opposed restricting abortion.
-By varying the participants' gender, we further tested the preregistered hypothesis that, in line with identity-based double standards, female participants would judge protest actions against restricting abortion to be more acceptable, on average, than male participants or protest actions for restricting abortion.
+By comparing male and female participants, we further tested the preregistered hypothesis that, in line with identity-based double standards, women would judge protest actions against restricting abortion to be more acceptable, on average, than men or protest actions for restricting abortion.
 
@@
 First, we stopped collecting data before reaching the preregistered sample size.
-We did so as the Supreme Court decision on reproductive rights, announced on June 24, 2022, had rendered the experimental manipulation invalid.
+We did so as the Supreme Court decision on reproductive rights, announced on June 24, 2022, rendered the experimental manipulation invalid.
 Second, we do not report results for two further hypotheses---testing moral conviction as a potential mechanism underlying ideology-based double standards---as those findings are beyond the scope of this manuscript and will be reported in a future publication.
@@
 As noted above, we had to stop collecting data before reaching the preregistered sample size of 800 eligible participants.
-This resulted in a final sample of 733 participants ($\textit{Mdn} = 37$ years, age range: 18--93 years) of whom 181 were conservative men, 190 were conservative women, 185 were liberal men, and 177 were liberal women.
+This resulted in a final sample of 733 participants ($\textit{Mdn} = 37$ years, age range: 18--93 years; 610 White, 55 Asian, 31 mixed, 20 Black, 16 other) of whom 181 were conservative men, 190 were conservative women, 185 were liberal men, and 177 were liberal women.
 
@@
 
-After answering where they would you place themselves along the political spectrum, in line with their response in the Prolific prescreening questionnaire, participants read the following paragraph:
+After answering where they would place themselves along the political spectrum, in line with their response in the Prolific prescreening questionnaire, participants read the following paragraph:
 
@@
 
-\noindent Participants rated which of several protest actions they consider acceptable means to protest for or against restricting abortion.
+\noindent As in the previous experiments, participants rated which of several protest actions they consider acceptable means to protest for or against restricting abortion.
 On the next pages, participants completed the remaining measures.
@@
 
-We included additional measures to describe the sample, to report as descriptive statistics, to use in non-preregistered analyses or future analyses not reported in this paper.
-
 We measured support for restricting abortion with one item: "Do you support or oppose restricting (or banning) abortion?" (1 = *strongly oppose*, 5 = *strongly support*).
-We also included a  three-item measure of moral conviction [@ryan_reconsidering_2014] that assessed to what extent participants' feelings about restricting or banning abortion were, for example, connected to their core moral beliefs or convictions (1 = *not at all*, 5 = *very much*).
-
-We measured gender-related system-justifying beliefs with eight items [adapted from @jost_exposure_2005], for example, "in general, relations between men and women are fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .93$).
-A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]$.
-
-We measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
-We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
 
@@
 
+We included additional measures, not used in the analyses reported in this paper, including measures of moral conviction [@ryan_reconsidering_2014], gender-related system-justifying beliefs [@jost_exposure_2005], and the updated moral foundations questionnaire [@atari_morality_2022, for details, see SOM].
+Within the latter questionnaire, we embedded three further attention checks (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
+
 ## Results
@@
 
-As expected, participants who identified as conservative reported being far more conservative ($M = 6.05, SD = 0.79$) than participants who identified as liberal ($M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43$).
+As intended, participants who identified as conservative reported being far more conservative ($M = 6.05, SD = 0.79$) than participants who identified as liberal ($M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43$).
 Likewise, conservative participants tended to endorse system-justifying beliefs ($M = 5.10, SD = 0.95$) while liberal participants tended to reject system-justifying beliefs ($M = 2.86, SD = 1.03; d = 2.26$).
@@
 We found some evidence for identity-based double standards: Female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
-This pattern was, however, overshadowed by stronger ideology-based double standards that were consistent across subsamples: Conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and both liberal women (`r with(m1, d_text[contrast == "nh6_contrast"])`) and men (`r with(m1, d_text[contrast == "nh5_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
+This pattern was, however, overshadowed by stronger ideology-based double standards that were consistent across subsamples: Conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and both liberal women (`r with(m1, d_text[contrast == "nh6_contrast"])`) and liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
 In this way, Experiment 3 provided more consistent evidence for ideology-based than for identity-based double standards in judging collective action.
@@
 
-  # # Define contrasts
-  # ide <- ide %>% 
-  #   mutate(
-  #     h1_contrast = case_when(
-  #       participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
-  #       TRUE ~ -1L,
-  #     ),
-  #     nh1_contrast = case_when(
-  #       participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
-  #       participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
-  #       TRUE ~ NA_integer_
-  #     ),
-  #     nh2_contrast = case_when(
-  #       participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
-  #       participant_gender == "Man" & protesters_cause == "For" ~ -1L,
-  #       TRUE ~ NA_integer_
-  #     )
-  #   )
-  # 
-  # # Calculate contrasts
-  # ide <- ide %>% 
-  #   pivot_longer(
-  #     ends_with("_contrast"),
-  #     names_to = "contrast",
-  #     values_to = "direction",
-  #     values_drop_na = TRUE
-  #   ) %>% 
-  #   group_by(.draw, z_ide, contrast, direction) %>% 
-  #   summarize(z = mean(z)) %>% 
-  #   summarize(d = sum(z * direction)) %>% 
-  #   group_by(z_ide, contrast) %>% 
-  #   median_qi(d)
-  # 
-  # # Make text
-  # ide <- ide %>% 
-  #   mutate(
-  #     across(
-  #       c(d, .lower, .upper), 
-  #       f_num, digits = 2, retain.leading.zero = T
-  #     ),
-  #     d_text = glue(
-  #       "$d = {d}, [{.lower}, {.upper}]$"
-  #     )
-  #   )
-
-```
-
-We explored whether group identification moderated participants' judgements.
+```
+
+We explored whether group identification moderated participants' judgments.
 We found, however, that gender identification neither affected how women judged protesters opposing (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "For"])`) abortion restrictions nor how men judged protesters opposing (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "For"])`) abortion restrictions.
 Our non-preregistered analyses thus suggested that gender identification did not moderate identity-based double standards in judging collective action concerning reproductive rights.
-
-```{r, include = FALSE}
-
-  # Import results
-  sjb <- read_rds("../Experiment 3/results/sjb_post.rds")
-
-  # Hypothesis 2
-  sjb <- sjb %>% 
-    distinct(.draw, b_sjb_for, b_sjb_against) %>% 
-    pivot_longer(
-      c(b_sjb_for, b_sjb_against),
-      names_to = "cause",
-      names_prefix = "b_sjb_",
-      values_to = "b"
-    ) %>%
-    group_by(cause) %>% 
-    median_qi(b) %>% 
-    mutate(
-      across(
-        c(b, .lower, .upper), 
-        f_num, digits = 2, retain.leading.zero = T
-      ),
-      b_text = glue(
-        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
-        .open = "(", .close = ")"
-      )
-    )
-
-```
-
-<!-- We also explored whether gender-related system justification predicted how acceptable participants judged protest actions for and against restricting abortion to be.
-To that end, we added two fixed effects to the model used in the preregistered analyses that estimated participants' responses as a function of their *z*-standardized system-justifying beliefs.
-As in Experiment 2, we found evidence for ideological asymmetry in this relationship: Participants who endorsed system-justifying beliefs considered the same protest action less acceptable when judging protests *against* restricting abortion (`r with(sjb, b_text[cause == "against"])`) but did not differ from participants who rejected system-justifying beliefs when judging protests *for* restricting abortion (`r with(sjb, b_text[cause == "for"])`).
-As in Experiment 2, our non-preregistered analyses thus suggested that system justification affects judgements about protests for a progressive cause (opposing abortion restrictions) but not for a conservative cause (supporting abortion restrictions).
--->
 
@@
 As both conservative and liberal participants judged protests for a cause aligned with their ideological orientation to be more acceptable and as this double standard was stronger among liberal participants, Experiment 3 provided evidence for both symmetry and asymmetry in ideology-based double standards in judging collective action.
+
 In contrast to Experiments 1--2, Experiment 3 provided some evidence for identity-based double standards, although any gender differences in judging protests for or against restricting abortion were overshadowed by more consistent ideology-based double standards.
+Our finding that conservative women considered collective action in support of restricting abortion to be more acceptable while liberal women considered collective action in opposition to restricting abortion to be more acceptable aligned with Mikolajczak et al.'s [-@mikolajczak_women_2022] argument that, by itself, gender identity is too broad to explain support for collective action.
+Instead, we need to consider the content of those identities---for example, identification with feminism or with traditional women---to understand support for, and opposition to, progressive and reactionary collective action.
 
@@
 <!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.
--->
-In the remainder of this General Discussion, we consider some limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.
+--> In the remainder of this General Discussion, we consider some limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.
 
@@
 
-Our research applied item response theory to judgements about collective action to investigate whether where people draw the line between acceptable and unacceptable means of protest depends on who the protesters are and what they are protesting.
+Our research applied item response theory to judgments about collective action to investigate whether where people draw the line between acceptable and unacceptable means of protest depends on who the protesters are and what they are protesting.
 In so doing, we moved the distinction between normative and non-normative collective action from the realm of the researcher's intuition into the realm of scientific investigation, allowing us to test for double standards in judging collective action.
@@
 Growing evidence shows that, in many situations, observers apply a more lenient standards when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
-And yet, our findings suggest that, unlike judgements in other domains, judgements about collective action are not subject to double standards based on group identities such as class, race, or gender---not even among highly identified group members.
-One reason might be that collective action, even in its controversial forms, is not judged as a moral transgression.
-Instead, observers might understand collective action primarily as a means to an end and judge how acceptable a form of collective action is in relation to its end.
+And yet, our findings suggest that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
+Why did we not find any evidence for identity-based double standards?
+
+One reason for this discrepancy might be that participants did not identify strongly enough with the social categories examined in our studies.
+This explanation, however, is not plausible because, first, we examined social categories (class, race, gender) that are central to most people's identities and because, second, we did not find evidence for ingroup bias even among highly identified group members.
+Another reason for this discrepancy might be that participants did not perceive the protesters' causes as relevant to the social identities examined in our studies.
+This explanation, however, conflicts with participants' reactions to the manipulation which, at least in Experiment 1--2, showed that participants understood the vignette in ingroup--outgroup terms.
+
+A more plausible explanation might be that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013].
+Instead, observers might understand collective action primarily as a means to an end and judge how acceptable a form of collective action is in relation to its end ("the end justifies the mean").
 This aligns with our finding that how acceptable participants judged protest actions to be depended on the protesters' cause, not their group membership.
@@
 So far, we explained this ideology-based double standard as resulting from conservatives being motivated to defend the system, progressives being motivated to challenge the system, and both being motivated to support collective action to those effects [@jost_missing_2017; @osborne_protesting_2019].
-An alternative explanation is to understand those double standards as an expression of ingroup bias based on an opinion-based [@mcgarty_collective_2009] or partisan [@finkel_political_2020] identity.
+Other explanations, however, are plausible.
+
+First, we might instead understand those double standards as an expression of ingroup bias based on an opinion-based [@mcgarty_collective_2009] or partisan [@finkel_political_2020] identity.
+Opinion-based identities are often a stronger predictor of collective action than identities based on other social categories [@bliuc_opinion-based_2007].
 This aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of the same transgressive protest actions when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
-Another explanation is that their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
+
+Second, we might argue that their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
 This aligns with Richardson & Conway's [-@richardson_standing_2022] finding that their different moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives and vice versa.
-Future research should seek to establish which of those plausible mechansims best explains partisan differences in judging collective action.
+Future research should seek to disentangle which of those plausible mechanisms best explains partisan differences in judging collective action.
 
@@
 
-<!-- Why do we find ideology-based double standards?
-We argued for system justification as the relevant distinction.
-One alternative explanation is opinion-based groups, see Verkuhyten et al.
-Another alternative explanation is moral concerns, see Richardson et al.
-Ultimately, ideology-based double standards are probably multiply determined.
--->
-
-<!-- 1.
-There are well-documented identity-based double standards in other domains.
-And yet, we did not find consistent evidence that judgements about collective action are subject to double standards based on social categories such as class, race, and gender---not even among highly identified group members.
-Why is that?
-(1) Collective action is not a moral transgression.
-(2) Opinion-based groups is one way to look at ideology.
--->
-
-<!-- 2.
-Our findings provide strong evidence for ideology-based double standards such that ...
-This aligns with other recent research that provides converging evidence.
-This all suggest that the right cause can justify the means of protest.
--->
-
-<!-- 3.
-Why asymmetry?
-System Justification != Social Change == Collective Action.
-However, that we do find effects for a conservative cause ...
-One explanation might lie in the different moral values by ...
-moral conviction ...
--->
-
-<!-- Our findings raised the question why we did not find double standards in judging collective action, even though there are well-documents double standards in other domains. -->
-
-<!-- 1. What does it mean that we did not find identity-based double standards? -->
-<!-- We do not find ingroup bias, at least not for broad social categories such as race and class, in contrast to evidence and predictions from social identity theory.
-Protest actions are not the same as moral transgressions, perhaps.
-We might have reached a different conclusion when adopting the framework of opinion-based group memberships and understanding political ideology in partisan identity terms.
--->
-<!-- 2. What does it mean that we did not find symmetrical ideology-based double standards? -->
-<!-- It could just be that conservatives don't like protest.
-But consider Jan.
-6 --- maybe it's about the cause of the protest?
-We know about the distinct moral values of conservatives and liberals (see MFT, see MFT and political orientation) and how they shape attitudes toward policy issues (see MFT and culture war issues).
-This aligns with Verkuyten and Richardson's finding of morality and tolerance judgements.
--->
-
-<!-- In two preregistered experiments, we showed that, unlike judgements in other domains, judgements what are acceptable means of protest are not subject to double standards based on group identities (such as class and race) but instead depend on whether the protesters' cause aligns with one's ideological orientation.
--->
-
-<!-- Our research sought to understand the often divided response to protest movements.
-Reactions to Black athletes kneeling to protest racist police violence seem to suggest that Americans are divided not only about the ends of protests, but also about the means of protest.
--->
-
-<!-- Contradicting research showing that people sometimes apply a more lenient standard when judging moral transgressions by ingroup members [e.g., @abrams_double_2013; @valdesolo_moral_2007], ...
-that who the protesters are really does not matter all that much.
--->
-
-<!-- Our research was motivated by examples showing that different groups judge the same protest action more acceptable depending on who's doing the action.
-That is, in contrast to previous research, not the case.
--->
-
-<!-- Examples strongly suggest that who's protesting and why shapes where people draw the line ...
-However, public opinion polling cannot tell us whether the underlying psychological mechanism is one of identity or ideology.
-In the present research, we developed a systematic approach to ...
-and, across two experiments, test for ...
-double standards.
-Contrary to prior theorizing and research, we found no evidence for ingroup bias...
-Instead, we found ...
--->
-
-<!-- I should discuss how our findings could explain why people are so ready to perceive BLM protests as non-normative, in contrast to the evidence, because they applied a stricter standard when judging protests for a progressive cause.
--->
-
-<!-- 1. Summary of Findings -->
-<!-- 3. What does it mean that we did not find identity-based double standards? -->
-<!-- We do not find ingroup bias, at least not for broad social categories such as race and class, in contrast to evidence and predictions from social identity theory.
-Protest actions are not the same as moral transgressions, perhaps.
-We might have reached a different conclusion when adopting the framework of opinion-based group memberships and understanding political ideology in partisan identity terms.
--->
-<!-- 4. What does it mean that we did not  find symmetrical ideology-based double standards? -->
-<!-- It could just be that conservatives don't like protest.
-But consider Jan.
-6 --- maybe it's about the cause of the protest?
-We know about the distinct moral values of conservatives and liberals (see MFT, see MFT and political orientation) and how they shape attitudes toward policy issues (see MFT and culture war issues).
-This aligns with Verkuyten and Richardson's finding of morality and tolerance judgements.
--->
-<!-- 2. Is the distinction between normative and non-normative collective action in the eye of the beholder? -->
-<!-- Relate to "protect framework" (e.g., Leach et al.).
-Question asssumption that there is a universal standard for acceptable protest actions.
-Relate to legislative issues such as protest bans/NFL's decision to ban certain forms of political expression.
--->
-<!-- 5. What are the limitations of the present research? -->
-<!-- 6. Conclusion -->
-
-<!-- - Summary: No evidence for identity-based double standards, but some evidence for ideology-based double standards -->
-<!-- - Reject the alternative hypothesis that there is consensus about what is an acceptable means of protest. -->
-<!-- - Relate our findings to @verkuyten_intolerance_2022  -->
-<!-- - Two plausible interpretations: (1) conservatives oppose controversial protest (no matter the end) *or* (2) the end justifies the mean -->
-<!-- - Re (1) it's plausible but what about recent events? (see public polling from grant) -->
-<!-- - Re (2) why is there no asymmetry? -->
-<!-- - Outline the-end-justifies-the-mean idea, highlighting that just supporting a cause is not enough, it needs moralization -->
-<!-- - Discuss why high-status group are more accepting of controversial protest actions -->
-<!-- - Discuss limitations (two cultures, two issues, liberal samples; system-defending vs.
-system-supporting actions; what about other conservative causes?; ingroup bias might need zero-sum situation to succeed) -->
-<!-- - Invite researchers to use our paradigm -->
-
 Most broadly, our findings showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
-Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
+Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
 In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.

### 2022-09-13 Documents/manuscript.tex
preprint

@@
-\documentclass[twocolumn, 11pt, letterpaper]{article}
+\documentclass[12pt, letterpaper]{article}
 
@@
 \usepackage[utf8]{inputenc}
-\usepackage{MinionPro} 
-\input{glyphtounicode}
-\pdfgentounicode=1 
+\usepackage{amsmath}
 \usepackage{microtype}
@@
 \usepackage[letterpaper, margin = 1in]{geometry}
-\setlength{\columnsep}{18pt}
 \setcounter{secnumdepth}{0}
+\setlength{\parindent}{0.5in}
 \usepackage[compact]{titlesec}
-\titleformat*{\section}{\centering\normalfont\Large\bfseries}
+\titleformat*{\section}{\centering\normalfont\normalsize\bfseries}
 \titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
 \titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
-\usepackage{authblk}
-\renewcommand\Affilfont{\small}
-\newcommand{\refsection}{\section{References}}
+\newcommand{\refsection}{\newpage \section{References}}
+\usepackage{indentfirst}
+\usepackage[hang,flushmargin]{footmisc}
 \providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Header
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{14.5pt}
+\renewcommand{\headrulewidth}{0pt}
+\fancyhead[L]{DOUBLE STANDARDS}
+\fancyhead[R]{\thepage}
 
@@
 
+% Line spacing
+\usepackage[doublespacing]{setspace}
+
 % Frontmatter
-\title{ Double Standards in Judging Collective Action\thanks{} }
-\author[1]{Nils K. Reimer}
-\affil[1]{University of California, Santa Barbara, United States}
-\author[2]{Marija Brankovi\'{c}}
-\affil[2]{Singidunum University, Serbia}
-\author[3]{Iniobong Essien}
-\affil[3]{Leuphana University of L\"{u}neburg, Germany}
-\author[4]{Jin X. Goh}
-\affil[4]{Colby College, United States}
-\author[5]{S\'{e}bastien Goudeau}
-\affil[5]{University of Poitiers, France}
-\author[6]{N\'{o}ra A. Lantos}
-\affil[6]{E\"{o}tv\"{o}s Lor\'{o}nd University, Hungary}
-\author[7]{Jenny Veldman}
-\affil[7]{KU Leuven, Belgium}
+\title{ Double Standards in Judging Collective Action }
 
@@
 
-\twocolumn[
-  \begin{@twocolumnfalse}
-    \maketitle
-    \begin{abstract}
-      \noindent Collective action is a powerful force driving social
-change but often sparks contention about what actions are acceptable
-means to effect social change. In five studies (total \emph{N} = 2,979),
-we investigated double standards in judging collective action---that is,
+\section{Abstract}
+
+\noindent Collective action is a powerful force driving social change
+but often sparks contention about what actions are acceptable means to
+effect social change. In five studies (total \emph{N} = 2,979), we
+investigated double standards in judging collective action---that is,
 whether observers will judge the same protest actions as more acceptable
@@
 prominent social movements.
-    \end{abstract}
-  \end{@twocolumnfalse}
-  % \renewcommand{\abstractname}{Statement of Relevance}
-  % \begin{@twocolumnfalse}
-  %   \maketitle
-  %   \begin{abstract}
-  %   \noindent 
-  %   \end{abstract}
-  % \end{@twocolumnfalse}
-]
-
-{
-  \renewcommand{\thefootnote}%
-    {\fnsymbol{footnote}}
-  \footnotetext[1]{M.B.--J.V. are listed in alphabetical order. This research was supported by a Seedcorn Research Grant from the European Association of Social Psychology.}
-}
-
-\noindent Protest movements often spark contention about what actions
-are acceptable means of protest. Reactions to Black American athletes
+
+\textit{Keywords:} protest, collective action, double standards, social
+identity theory, system justification theory, Black Lives Matter
+
+\newpage
+
+\section{Double Standards in Judging Collective Action}
+
+Protest movements often spark contention about what actions are
+acceptable means of protest. Reactions to Black American athletes
 kneeling during the national anthem to protest racist police violence
 are a case in point. Kneeling during the national anthem is not violent,
-disruptive, or illegal. And yet, only 29\% of White Americans (66\% of
-Black Americans) and 11\% of Republicans (59\% of Democrats) considered
-it appropriate for Black athletes to protest in this way (YouGov, 2017).
-Indeed, Black athletes who had participated in the protests faced a
-backlash in public opinion (Lacina, 2020). This and other examples
-suggest that \emph{who} the protesters are and \emph{what} they are
-protesting influences how acceptable observers judge their actions to
-be.
+disruptive, or illegal. And yet, only 29\% of White Americans (compared
+to 66\% of Black Americans) and 11\% of Republicans (compared to 59\% of
+Democrats) considered it appropriate for Black athletes to protest in
+this way (YouGov, 2017). Indeed, Black athletes who had participated in
+the protests faced a backlash in public opinion (Lacina, 2020). This and
+other examples suggest that \emph{who} the protesters are and
+\emph{what} they are protesting influences how acceptable observers
+judge their actions to be.
 
@@
 goal while conforming to the norms of the existing social system and
-\emph{non-normative} collective action that violates those norms (Wright
-et al., 1990). Recent research has shown that this distinction matters
-for how people respond to collective action. Shuman et al. (2021) showed
-that non-normative, non-violent collective action is more effective than
+\emph{non-normative} collective action that violates those norms (Becker
+\& Tausch, 2015; Wright et al., 1990). Some instead refer to the two
+kinds of collective action as moderate and radical collective action
+(e.g., Jiménez-Moya et al., 2015). Broadly, this distinction captures
+the intuition that there are some forms of collective action that are
+generally considered acceptable and others that are not.
+
+Recent research has shown that this distinction matters for how people
+respond to collective action. Shuman et al. (2021) showed that
+non-normative, non-violent collective action is more effective than
 either normative or violent collective action at gaining concessions
@@
 
-Research has, for the most part, relied on ad-hoc distinctions between
-what researchers themselves considered normative and non-normative
-collective action. In liberal democracies, researchers tend to consider
-actions such as signing petitions, voting in elections, or peaceful
-protest to be normative and actions such as blocking traffic, damaging
-property, or violent protest to be non-normative. But, as the examples
-in the first paragraph show, the distinction between normative and
-non-normative collective action might instead be in the eye of the
-beholder.
+These and other studies have, for the most part, relied on ad-hoc
+distinctions between what researchers themselves considered normative
+and non-normative collective action. In liberal democracies, researchers
+tend to consider actions such as signing petitions, voting in elections,
+or peaceful protest to be normative and actions such as blocking
+traffic, damaging property, or violent protest to be non-normative. But,
+as reactions to Black athletes kneeling in protest suggest, the same
+action might be considered acceptable by some and unacceptable by
+others. That is, the distinction between normative and non-normative
+collective action might be in the eye of the beholder.
 
@@
 intuition into the realm of scientific investigation and test for double
-standards in where people draw the line between acceptable and
-unacceptable forms of collective action. For that purpose, we define a
-double standard as judging the same action as more or less acceptable
-depending on who is performing the action and for what reason (for an
-analogous definition, see Foschi, 2000).\footnote{Feinberg et al. (2020)
-  provided incidental evidence for double standards in judging
-  collective action as, in a manipulation check, Black participants
-  rated Black Lives Matter protests as less extreme than White
-  participants and liberal participants rated protest actions for
+standards in where people draw the line between acceptable (normative)
+and unacceptable (non-normative) forms of collective action. For that
+purpose, we define a double standard as judging the same action as more
+or less acceptable depending on who is performing the action and for
+what reason (for an analogous definition, see Foschi, 2000).\footnote{Feinberg
+  et al. (2020) provided incidental evidence for double standards in
+  judging collective action as, in a manipulation check, Black
+  participants rated Black Lives Matter protests as less extreme than
+  White participants and liberal participants rated protest actions for
   progressive causes as less extreme than conservative participants.} In
@@
 
-One domain in which ingroup bias manifests is judgements about other
+One domain in which ingroup bias manifests is judgments about other
 people's actions. Hewstone (1990) reviewed research showing that people
@@
 outgroup members to internal causes but attributing negative behavior by
-ingroup members to external causes. Valdesolo \& DeSteno (2007)
-demonstrated that participants judged the same selfish action as less
-unfair when performed by themselves or an ingroup member than when
-performed by an outgroup member. Other studies (Abrams et al., 2013;
-Endevelt et al., 2021; for exceptions, see Mendoza et al., 2014; Pinto
-et al., 2010) provided further evidence for ingroup bias in judging
-moral transgressions by ingroup and outgroup members.
+ingroup members to external causes (see Pettigrew, 1979). Valdesolo \&
+DeSteno (2007) demonstrated that participants judged the same selfish
+action as less unfair when performed by themselves or an ingroup member
+than when performed by an outgroup member. Other studies (Abrams et al.,
+2013; Endevelt et al., 2021; for exceptions, see Mendoza et al., 2014;
+Pinto et al., 2010) provided further evidence for ingroup bias in
+judging moral transgressions by ingroup and outgroup members.
 
@@
 
-System Justification Theory (Jost \& Banaji, 1994; for a recent review,
-see Jost, 2020) argues that people are motivated to defend, justify, and
+System Justification Theory (Jost et al., 2004; for a recent review, see
+Jost, 2020) argues that people are motivated to defend, justify, and
 bolster the prevailing social, economic, and political system because
@@
 
-Just as system justification shapes support for, and opposition to,
+Just as system justification affects support for, and opposition to,
 collective action, we propose that it results in ideology-based double
@@
 
-Item response theory is a a conceptual and statistical framework for
+Item response theory is a conceptual and statistical framework for
 understanding how the characteristics of both items and respondents
@@
 whether each of a set of actions is an acceptable means of protest (1 =
-\emph{yes}, 0 = \emph{no}), a two parameter logistic item response
+\emph{yes}, 0 = \emph{no}), a two-parameter logistic item response
 theory model estimates responses as a function of several latent
@@
 In Studies 1 and 2, we select the most discriminating and informative
-actions to build an instrument for capturing double standards in the
-experiments. Second, item response theory provides a formal and
+actions to build an instrument for capturing double standards in judging
+collective action. Second, item response theory provides a formal and
 statistical definition of double standards in judging collective action
 as respondents being more accepting of the same protest actions
-depending on who the protesters and what they are protesting. In
+depending on who the protesters are and what they are protesting. In
 Experiments 1--3, we operationalize double standards as the differences
@@
 scale development. In Study 2, we ask participants to rate how
-acceptable they consider each of those protest action to be and use item
-response theory to select the most discriminating and informative
+acceptable they consider each of those protest actions to be and use
+item response theory to select the most discriminating and informative
 protest actions for our instrument. In so doing, we build an instrument
@@
 lower-status group. In non-preregistered analyses, we further explored
-whether, in line with \emph{Hypothesis 2b}, participants would judge the
-same protest actions as more or less acceptable depending on their own
-ideological positions. In this way, Experiment 1 provided a first test
-of the hypothesized identity- and ideology-based double standards in
-judging collective action.
+whether participants would judge the same protest actions as more or
+less acceptable depending on their own ideological positions. In this
+way, Experiment 1 provided a first test of the hypothesized identity-
+and ideology-based double standards in judging collective action.
 
@@
 socioeconomic status ladder. Another 233 participants considered their
-past, current job, or future jobs to be middle-class/professional jobs.
+past, current, or future jobs to be middle-class/professional jobs.
 Participants in this higher-status group had at least an undergraduate
@@
 real life, to what extent it would affect them personally, and to what
-extent the it would affect people like them (1 = \emph{not at all}, 7 =
+extent it would affect people like them (1 = \emph{not at all}, 7 =
 \emph{very much}). We also asked participants to what extent they
@@
 for the other two variables, social dominance orientation was not
-associated with participants' judgements about how acceptable various
+associated with participants' judgments about how acceptable various
 collective actions are (\(\beta_{xy} = -0.05, [-0.15, 0.05]\)). Overall,
@@
 have chosen a scenario too far removed from current affairs to evoke
-strong ingroup bias in participants' judgements of protest actions. That
+strong ingroup bias in participants' judgments of protest actions. That
 said, participants' reactions to the experimental manipulation suggested
@@
 as a possible solution to racist police violence. We conducted this
-study in January 2021, a time when, after a year of unprecedented
+study in January 2021, a time when, after months of unprecedented
 protests for racial justice, most Americans could be expected to be
@@
 \(d = 0.30\)). In contrast, White participants identified less strongly
-with their ingroup (\(M = 4.85, \textit{SD} = 1.53\); \(d = 1.21\)) and
-did not identify more with White protesters
-(\(M = 3.58, \textit{SD} = 2.01\)) than with Black protesters
-(\(M = 3.81, \textit{SD} = 1.77\); \(d = -0.12\)). Both White
-(\(d = 0.67\)) and Black (\(d = 0.53\)) participants tended to identify
-more with protesters protesting \emph{for} defunding the police. On
-average, participants tended to describe their political orientation as
-moderately liberal (\(M = 2.92, \textit{SD} = 1.73\)) but, as expected,
-were divided in their support for defunding the police
+with their ingroup than Black participants
+(\(M = 4.85, \textit{SD} = 1.53\); \(d = 1.21\)) and did not identify
+more with White protesters (\(M = 3.58, \textit{SD} = 2.01\)) than with
+Black protesters (\(M = 3.81, \textit{SD} = 1.77\); \(d = -0.12\)). Both
+White (\(d = 0.67\)) and Black (\(d = 0.53\)) participants tended to
+identify more with protesters protesting \emph{for} defunding the
+police. On average, participants tended to describe their political
+orientation as moderately liberal (\(M = 2.92, \textit{SD} = 1.73\))
+but, as expected, were divided in their support for defunding the police
 (\(M = 3.37, \textit{SD} = 1.42\)). That is, 56\% supported defunding
@@
 effects, one estimating the effect of system-justifying beliefs on
-judgements about system-defending protest actions and one estimating
-their effect on judgements about system-challenging protest actions, and
+judgments about system-defending protest actions and one estimating
+their effect on judgments about system-challenging protest actions, and
 one varying effect estimating its variance across conditions. Supporting
@@
 analyses thus suggested that group identification did not moderate group
-differences in judgements about collective action by ingroup and
-outgroup members.
+differences in judgments about collective action by ingroup and outgroup
+members.
 
@@
 men applied different standards when judging collective action opposing
-or supporting further restricting or banning abortion. We conducted this
-study in June 2022 when many expected the Supreme Court to overturn its
-\emph{Roe v. Wade} decision that hitherto had guaranteed access to legal
-abortions in the first 20--24 weeks of pregnancy in the United States.
-We focused on restricting abortion as it is a conservative issue, as it
-primarily affects women's reproductive rights, and as the anticipated
-Supreme Court decision would prompt several states to restrict or ban
+or supporting further restricting abortion. We conducted this study in
+June 2022 when many expected the Supreme Court to overturn its \emph{Roe
+v. Wade} decision that hitherto had guaranteed access to legal abortions
+in the first 20--24 weeks of pregnancy in the United States. We focused
+on restricting abortion as it is a conservative issue, as it primarily
+affects women's reproductive rights, and as the anticipated Supreme
+Court decision would prompt several states to restrict or ban abortion.
+
+By comparing participants of different political orientations and by
+varying the protesters' cause, we tested the preregistered hypothesis
+that, in line with \emph{symmetrical} ideology-based double standards,
+conservatives would judge the same protest actions to be more acceptable
+if the protesters supported restricting abortion and that liberals would
+judge the same protest actions to be more acceptable if the protesters
+opposed restricting abortion. We tested the alternative hypothesis that,
+in line with \emph{asymmetrical} ideology-based double standards,
+conservatives would judge all protest actions to be equally unacceptable
+while liberals would judge the same protest actions to be more
+acceptable if the protesters opposed restricting abortion. By comparing
+male and female participants, we further tested the preregistered
+hypothesis that, in line with identity-based double standards, women
+would judge protest actions against restricting abortion to be more
+acceptable, on average, than men or protest actions for restricting
 abortion.
-
-By varying the participants' political orientation and the protesters'
-cause, we tested the preregistered hypothesis that, in line with
-\emph{symmetrical} ideology-based double standards, conservatives would
-judge the same protest actions to be more acceptable if the protesters
-supported restricting abortion and that liberals would judge the same
-protest actions to be more acceptable if the protesters opposed
-restricting abortion. We tested the alternative hypothesis that, in line
-with \emph{asymmetrical} ideology-based double standards, conservatives
-would judge all protest actions to be equally unacceptable while
-liberals would judge the same protest actions to be more acceptable if
-the protesters opposed restricting abortion. By varying the
-participants' gender, we further tested the preregistered hypothesis
-that, in line with identity-based double standards, female participants
-would judge protest actions against restricting abortion to be more
-acceptable, on average, than male participants or protest actions for
-restricting abortion.
 
@@
 as the Supreme Court decision on reproductive rights, announced on June
-24, 2022, had rendered the experimental manipulation invalid. Second, we
-do not report results for two further hypotheses---testing moral
-conviction as a potential mechanism underlying ideology-based double
-standards---as those findings are beyond the scope of this manuscript
-and will be reported in a future publication. We made all materials,
-data, and analysis scripts available online
+24, 2022, rendered the experimental manipulation invalid. Second, we do
+not report results for two further hypotheses---testing moral conviction
+as a potential mechanism underlying ideology-based double standards---as
+those findings are beyond the scope of this manuscript and will be
+reported in a future publication. We made all materials, data, and
+analysis scripts available online
 (\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}).
@@
 This resulted in a final sample of 733 participants
-(\(\textit{Mdn} = 37\) years, age range: 18--93 years) of whom 181 were
-conservative men, 190 were conservative women, 185 were liberal men, and
-177 were liberal women.
+(\(\textit{Mdn} = 37\) years, age range: 18--93 years; 610 White, 55
+Asian, 31 mixed, 20 Black, 16 other) of whom 181 were conservative men,
+190 were conservative women, 185 were liberal men, and 177 were liberal
+women.
 
@@
 
-After answering where they would you place themselves along the
-political spectrum, in line with their response in the Prolific
-prescreening questionnaire, participants read the following paragraph:
+After answering where they would place themselves along the political
+spectrum, in line with their response in the Prolific prescreening
+questionnaire, participants read the following paragraph:
 
@@
 
-\noindent Participants rated which of several protest actions they
-consider acceptable means to protest for or against restricting
-abortion. On the next pages, participants completed the remaining
-measures. On the final page, participants responded to two attention
-checks that assessed whether they paid attention to the experimental
-manipulation: ``In this study, you read about some people's view on a
-policy change. What was the policy change?'' and ``In this study, you
-then read about some protesters. Were the protesters for or against
-restricting abortion?''.
+\noindent As in the previous experiments, participants rated which of
+several protest actions they consider acceptable means to protest for or
+against restricting abortion. On the next pages, participants completed
+the remaining measures. On the final page, participants responded to two
+attention checks that assessed whether they paid attention to the
+experimental manipulation: ``In this study, you read about some people's
+view on a policy change. What was the policy change?'' and ``In this
+study, you then read about some protesters. Were the protesters for or
+against restricting abortion?''.
 
@@
 
-We included additional measures to describe the sample, to report as
-descriptive statistics, to use in non-preregistered analyses or future
-analyses not reported in this paper.
-
 We measured support for restricting abortion with one item: ``Do you
 support or oppose restricting (or banning) abortion?'' (1 =
-\emph{strongly oppose}, 5 = \emph{strongly support}). We also included a
-three-item measure of moral conviction (Ryan, 2014) that assessed to
-what extent participants' feelings about restricting or banning abortion
-were, for example, connected to their core moral beliefs or convictions
-(1 = \emph{not at all}, 5 = \emph{very much}).
-
-We measured gender-related system-justifying beliefs with eight items
-(adapted from Jost \& Kay, 2005), for example, ``in general, relations
-between men and women are fair'' (1 = \emph{strongly disagree}, 7 =
-\emph{strongly agree}; McDonald's \(\omega = .93\)). A confirmatory
-factor analysis model in which all items loaded onto a single factor
-showed acceptable fit,
-\(\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]\).
-
-We measured moral concerns with the 36-item moral foundations
-questionnaire (MFQ-2, Atari et al., 2022) which assesses to what extent
-participants endorse concerns about care, equality, proportionality,
-loyalty, authority, and purity (e.g., ``I believe chastity is an
-important virtue.''; 1 = \emph{does not describe me at all}, 5 =
-\emph{describes me extremely well}). We embedded three further attention
-checks within the questionnaire (e.g., ``To show that you are paying
-attention and giving your best effort, please select `moderately
-describes me'.'').
+\emph{strongly oppose}, 5 = \emph{strongly support}).
 
@@
 
+We included additional measures, not used in the analyses reported in
+this paper, including measures of moral conviction (Ryan, 2014),
+gender-related system-justifying beliefs (Jost \& Kay, 2005), and the
+updated moral foundations questionnaire (Atari et al., 2022, for
+details, see SOM). Within the latter questionnaire, we embedded three
+further attention checks (e.g., ``To show that you are paying attention
+and giving your best effort, please select `moderately describes
+me'.'').
+
 \hypertarget{results-2}{%
@@
 
-As expected, participants who identified as conservative reported being
+As intended, participants who identified as conservative reported being
 far more conservative (\(M = 6.05, SD = 0.79\)) than participants who
@@
 \emph{supported} restricting abortion and both liberal women
-(\(d = 1.04, [0.78, 1.32]\)) and men (\(d = 0.76, [0.49, 1.03]\))
-considered the same protest actions to be more acceptable when
-protesters \emph{opposed} restricting abortion. In this way, Experiment
-3 provided more consistent evidence for ideology-based than for
-identity-based double standards in judging collective action.
+(\(d = 1.04, [0.78, 1.32]\)) and liberal men
+(\(d = 0.76, [0.49, 1.03]\)) considered the same protest actions to be
+more acceptable when protesters \emph{opposed} restricting abortion. In
+this way, Experiment 3 provided more consistent evidence for
+ideology-based than for identity-based double standards in judging
+collective action.
 
@@
 We explored whether group identification moderated participants'
-judgements. We found, however, that gender identification neither
+judgments. We found, however, that gender identification neither
 affected how women judged protesters opposing
@@
 3 provided evidence for both symmetry and asymmetry in ideology-based
-double standards in judging collective action. In contrast to
-Experiments 1--2, Experiment 3 provided some evidence for identity-based
-double standards, although any gender differences in judging protests
-for or against restricting abortion were overshadowed by more consistent
-ideology-based double standards.
+double standards in judging collective action.
+
+In contrast to Experiments 1--2, Experiment 3 provided some evidence for
+identity-based double standards, although any gender differences in
+judging protests for or against restricting abortion were overshadowed
+by more consistent ideology-based double standards. Our finding that
+conservative women considered collective action in support of
+restricting abortion to be more acceptable while liberal women
+considered collective action in opposition to restricting abortion to be
+more acceptable aligned with Mikolajczak et al.'s (2022) argument that,
+by itself, gender identity is too broad to explain support for
+collective action. Instead, we need to consider the content of those
+identities---for example, identification with feminism or with
+traditional women---to understand support for, and opposition to,
+progressive and reactionary collective action.
 
@@
 
-Our research applied item response theory to judgements about collective
+Our research applied item response theory to judgments about collective
 action to investigate whether where people draw the line between
@@
 (Abrams et al., 2013; Endevelt et al., 2021). And yet, our findings
-suggest that, unlike judgements in other domains, judgements about
+suggest that, unlike judgments in other domains, judgments about
 collective action are not subject to double standards based on group
-identities such as class, race, or gender---not even among highly
-identified group members. One reason might be that collective action,
-even in its controversial forms, is not judged as a moral transgression.
-Instead, observers might understand collective action primarily as a
-means to an end and judge how acceptable a form of collective action is
-in relation to its end. This aligns with our finding that how acceptable
-participants judged protest actions to be depended on the protesters'
-cause, not their group membership.
+identities such as class, race, or gender. Why did we not find any
+evidence for identity-based double standards?
+
+One reason for this discrepancy might be that participants did not
+identify strongly enough with the social categories examined in our
+studies. This explanation, however, is not plausible because, first, we
+examined social categories (class, race, gender) that are central to
+most people's identities and because, second, we did not find evidence
+for ingroup bias even among highly identified group members. Another
+reason for this discrepancy might be that participants did not perceive
+the protesters' causes as relevant to the social identities examined in
+our studies. This explanation, however, conflicts with participants'
+reactions to the manipulation which, at least in Experiment 1--2, showed
+that participants understood the vignette in ingroup--outgroup terms.
+
+A more plausible explanation might be that collective action, even in
+its controversial forms, is not judged as a moral transgression (as in
+previous research on identity-based double standards, e.g., Abrams et
+al., 2013). Instead, observers might understand collective action
+primarily as a means to an end and judge how acceptable a form of
+collective action is in relation to its end (``the end justifies the
+mean''). This aligns with our finding that how acceptable participants
+judged protest actions to be depended on the protesters' cause, not
+their group membership.
 
@@
 motivated to support collective action to those effects (Jost et al.,
-2017; Osborne et al., 2019). An alternative explanation is to understand
-those double standards as an expression of ingroup bias based on an
-opinion-based (McGarty et al., 2009) or partisan (Finkel et al., 2020)
-identity. This aligns with Verkuyten et al.'s (2022) finding that
-participants were more tolerant of the same transgressive protest
-actions when taken by their most-liked, rather than their least-liked,
-political group (e.g., climate activists). Another explanation is that
-their different moral concerns (Graham et al., 2009; Kivikangas et al.,
-2021) lead progressives and conservatives to see their support of
-certain causes as a fundamental matter of right or wrong (Skitka et al.,
-2021) and, by thus moralizing those causes, to accept more extreme means
-to achieving them. This aligns with Richardson \& Conway's (2022)
-finding that their different moral concerns explained why liberals rated
-protesting for liberal causes as more moral than conservatives and vice
-versa. Future research should seek to establish which of those plausible
-mechansims best explains partisan differences in judging collective
-action.
+2017; Osborne et al., 2019). Other explanations, however, are plausible.
+
+First, we might instead understand those double standards as an
+expression of ingroup bias based on an opinion-based (McGarty et al.,
+2009) or partisan (Finkel et al., 2020) identity. Opinion-based
+identities are often a stronger predictor of collective action than
+identities based on other social categories (Bliuc et al., 2007). This
+aligns with Verkuyten et al.'s (2022) finding that participants were
+more tolerant of the same transgressive protest actions when taken by
+their most-liked, rather than their least-liked, political group (e.g.,
+climate activists).
+
+Second, we might argue that their different moral concerns (Graham et
+al., 2009; Kivikangas et al., 2021) lead progressives and conservatives
+to see their support of certain causes as a fundamental matter of right
+or wrong (Skitka et al., 2021) and, by thus moralizing those causes, to
+accept more extreme means to achieving them. This aligns with Richardson
+\& Conway's (2022) finding that their different moral concerns explained
+why liberals rated protesting for liberal causes as more moral than
+conservatives and vice versa. Future research should seek to disentangle
+which of those plausible mechanisms best explains partisan differences
+in judging collective action.
 
@@
 actions, there is no consensus on what is and is not an acceptable means
-to advance a cause. Instead, this judgement seems to depend on what the
+to advance a cause. Instead, this judgment seems to depend on what the
 cause of a protest is and how it aligns with one's beliefs and values.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-becker_dynamic_2015}{}}%
+Becker, J. C., \& Tausch, N. (2015). A dynamic model of engagement in
+normative and non-normative collective action: {Psychological}
+antecedents, consequences, and barriers. \emph{European Review of Social
+Psychology}, \emph{26}(1), 43--92.
+\url{https://doi.org/10.1080/10463283.2015.1094265}
+
 \leavevmode\vadjust pre{\hypertarget{ref-becker_committed_2011}{}}%
@@
 \emph{37}(8), 1104--1116. \url{https://doi.org/10.1177/0146167211407076}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bliuc_opinion-based_2007}{}}%
+Bliuc, A.-M., McGarty, C., Reynolds, K., \& Muntele, D. (2007).
+Opinion-based group membership as a predictor of commitment to political
+action. \emph{European Journal of Social Psychology}, \emph{37}(1),
+19--32. \url{https://doi.org/10.1002/ejsp.334}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-jimenez-moya_by_2015}{}}%
+Jiménez-Moya, G., Spears, R., Rodríguez-Bailón, R., \& Lemus, S. de.
+(2015). By any means necessary? {When} and why low group identification
+paradoxically predicts radical collective action: Predicting radical
+collective action. \emph{Journal of Social Issues}, \emph{71}(3),
+517--535. \url{https://doi.org/10.1111/josi.12126}
+
 \leavevmode\vadjust pre{\hypertarget{ref-jost_theory_2020}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-jost_role_1994}{}}%
-Jost, J. T., \& Banaji, M. R. (1994). The role of stereotyping in
-system-justification and the production of false consciousness.
-\emph{British Journal of Social Psychology}, \emph{33}(1), 1--27.
-\url{https://doi.org/10.1111/j.2044-8309.1994.tb01008.x}
+\leavevmode\vadjust pre{\hypertarget{ref-jost_decade_2004}{}}%
+Jost, J. T., Banaji, M. R., \& Nosek, B. A. (2004). A decade of system
+justification theory: Accumulated evidence of conscious and unconscious
+bolstering of the status quo. \emph{Political Psychology}, \emph{25}(6),
+881--919. \url{https://doi.org/10.1111/j.1467-9221.2004.00402.x}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mikolajczak_women_2022}{}}%
+Mikolajczak, G., Becker, J. C., \& Iyer, A. (2022). Women who challenge
+or defend the status quo: {Ingroup} identities as predictors of
+progressive and reactionary collective action. \emph{European Journal of
+Social Psychology}. \url{https://doi.org/10.1002/ejsp.2842}
+
 \leavevmode\vadjust pre{\hypertarget{ref-mooijman_moralization_2018}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pettigrew_ultimate_1979}{}}%
+Pettigrew, T. F. (1979). The ultimate attribution error: Extending
+{Allport}'s cognitive analysis of prejudice. \emph{Personality and
+Social Psychology Bulletin}, \emph{5}(4), 461--476.
+\url{https://doi.org/10.1177/014616727900500407}
+
 \leavevmode\vadjust pre{\hypertarget{ref-pinto_membership_2010}{}}%
@@
 anti-protest bills. \emph{Pew Charitable Trusts}.
-\url{https://www.pewtrusts.org/en/research-and-analysis/blogs/stateline/2021/02/04/republicans-respond-to-black-lives-matter-with-anti-protest-bills}
+\url{https://pew.org/3rhdGYw}
 

### 2022-09-13 Documents/supplemental-materials.Rmd
preprint

@@
 As preregistered, we excluded all participants who did not respond with an accurate description of the group they had read about.
+
+# Experiment 3
+
+## Method
+
+### Measures
+
+We included additional measures not used in the analyses reported in this paper.
+
+We included a three-item measure of moral conviction [@ryan_reconsidering_2014] that assessed to what extent participants' feelings about restricting or banning abortion were, for example, connected to their core moral beliefs or convictions (1 = *not at all*, 5 = *very much*).
+
+We measured gender-related system-justifying beliefs with eight items [adapted from @jost_exposure_2005], for example, "in general, relations between men and women are fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .93$).
+A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]$.
+
+We measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
+We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
 

### 2022-09-13 Experiment 3/scripts/prepare_data.R

@@ -70,7 +70,15 @@ rm(list = ls())
         q64_1 == "white" ~ NA_character_,
         q64_1 == "Women" ~ "woman"
       ),
-      age = as.integer(q64_2)
+      age = as.integer(q64_2),
+      race = recode_factor(
+        q65,
+        "1" = "White",
+        "2" = "Black",
+        "3" = "Asian",
+        "4" = "Mixed",
+        "5" = "Other"
+      )
     )
   
   # Transform predictor and outcome variables
@@ -280,7 +288,7 @@ rm(list = ls())
     select(
       id,
       participant_ideology, participant_gender, protesters_cause,
-      age,
+      age, race,
       sup, mor_1:mor_3,
       starts_with("mfq_"),
       starts_with("sjb_"),

### 2022-09-19 Documents/manuscript.Rmd
update preprint

@@
     keep_tex: yes
-    template: preprint-template.tex
+    template: apa-template.tex
 
@@
 @shuman_disrupting_2021 showed that non-normative, non-violent collective action is more effective than either normative or violent collective action at gaining concessions from those most resistant to social change.
-@feinberg_activists_2020 demonstrated, however, that observers are less supportive of social movements that use extreme protest actions.
-Similarly, @teixeira_is_2020 found that advantaged-group members perceive non-normative collective action by disadvantaged-group members as more damaging to their ingroup's social image and are thus less supportive of such action.
+@feinberg_activists_2020 demonstrated, however, that observers are less supportive of social movements that use extreme protest actions.^[In contrast, White Americans exposed to pictures of destructive Black Lives Matter protests were not, on average, any less supportive of systemic change than those exposed to pictures of peaceful protests [@teixeira_white_2022].] Similarly, @teixeira_is_2020 found that advantaged-group members perceive non-normative collective action by disadvantaged-group members as more damaging to their ingroup's social image and are thus less supportive of such action.
 @verkuyten_intolerance_2022 found that, in general, the public tends to be intolerant of non-normative collective action.
@@
 Ideology-based double standards could, for example, explain why far more Democrats than Republicans considered it appropriate to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].
-
-<!-- That is, we hypothesize that, at least to the extent that they hold system-justifying beliefs, people will judge the same protest actions as more acceptable if the protest supports, rather than challenges, the system.
--->
 
@@
 Future research should seek to establish how consistent this ideological asymmetry is across different political causes and cultural contexts.
+
+Our findings provided a relevant comparison to findings from a recent experiment in which White Americans viewed pictures of either peaceful or destructive Black Lives Matter protests and rated how legitimate (i.e., fair, reasonable, legitimate, and justified) they considered the pictured protests to be [@teixeira_white_2022].
+In this experiment, more conservative participants rated destructive and, to a lesser extent, peaceful protests to be less legitimate than more liberal participants.
+Likewise, participants who identified more strongly with their ethnic ingroup rated destructive and, to a lesser extent, peaceful protests to be less legitimate than participants who identified less strongly with their ingroup.
+However, participants' group identification was no longer associated with their protest ratings after controlling for their political orientation.^[These findings are based on our reanalysis, reported in the SOM, of Teixeira et al.'s [-@teixeira_white_2022] experiment.
+We thank the authors for making their data available online.] In this way, results from the experiment aligned with our conclusion that political ideologies are more important than social identities for shaping how people judge controversial forms of collective action.
 

### 2022-09-19 Documents/manuscript.tex
update preprint

@@
-\documentclass[12pt, letterpaper]{article}
+\documentclass[twocolumn, 11pt, letterpaper]{article}
 
@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath}
+\usepackage{MinionPro} 
+\input{glyphtounicode}
+\pdfgentounicode=1 
 \usepackage{microtype}
@@
 \usepackage[letterpaper, margin = 1in]{geometry}
+\setlength{\columnsep}{18pt}
 \setcounter{secnumdepth}{0}
-\setlength{\parindent}{0.5in}
 \usepackage[compact]{titlesec}
-\titleformat*{\section}{\centering\normalfont\normalsize\bfseries}
+\titleformat*{\section}{\centering\normalfont\Large\bfseries}
 \titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
 \titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
-\newcommand{\refsection}{\newpage \section{References}}
-\usepackage{indentfirst}
-\usepackage[hang,flushmargin]{footmisc}
+\usepackage{authblk}
+\renewcommand\Affilfont{\small}
+\newcommand{\refsection}{\section{References}}
 \providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% Header
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-\setlength{\headheight}{14.5pt}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{DOUBLE STANDARDS}
-\fancyhead[R]{\thepage}
 
@@
 
-% Line spacing
-\usepackage[doublespacing]{setspace}
-
 % Frontmatter
-\title{ Double Standards in Judging Collective Action }
+\title{ Double Standards in Judging Collective Action\thanks{} }
+\author[1]{Nils K. Reimer}
+\affil[1]{University of California, Santa Barbara, United States}
+\author[2]{Marija Brankovi\'{c}}
+\affil[2]{Singidunum University and University of Belgrade, Serbia}
+\author[3]{Iniobong Essien}
+\affil[3]{Leuphana University of L\"{u}neburg, Germany}
+\author[4]{Jin X. Goh}
+\affil[4]{Colby College, United States}
+\author[5]{S\'{e}bastien Goudeau}
+\affil[5]{University of Poitiers, France}
+\author[6]{N\'{o}ra A. Lantos}
+\affil[6]{E\"{o}tv\"{o}s Lor\'{o}nd University, Hungary}
+\author[7]{Jenny Veldman}
+\affil[7]{NYU Abu Dhabi, United Arab Emirates}
 
@@
 
-\section{Abstract}
-
-\noindent Collective action is a powerful force driving social change
-but often sparks contention about what actions are acceptable means to
-effect social change. In five studies (total \emph{N} = 2,979), we
-investigated double standards in judging collective action---that is,
+\twocolumn[
+  \begin{@twocolumnfalse}
+    \maketitle
+    \begin{abstract}
+      \noindent Collective action is a powerful force driving social
+change but often sparks contention about what actions are acceptable
+means to effect social change. In five studies (total \emph{N} = 2,979),
+we investigated double standards in judging collective action---that is,
 whether observers will judge the same protest actions as more acceptable
@@
 prominent social movements.
-
-\textit{Keywords:} protest, collective action, double standards, social
-identity theory, system justification theory, Black Lives Matter
-
-\newpage
-
-\section{Double Standards in Judging Collective Action}
-
-Protest movements often spark contention about what actions are
-acceptable means of protest. Reactions to Black American athletes
+    \end{abstract}
+  \end{@twocolumnfalse}
+  % \renewcommand{\abstractname}{Statement of Relevance}
+  % \begin{@twocolumnfalse}
+  %   \maketitle
+  %   \begin{abstract}
+  %   \noindent 
+  %   \end{abstract}
+  % \end{@twocolumnfalse}
+]
+
+{
+  \renewcommand{\thefootnote}%
+    {\fnsymbol{footnote}}
+  \footnotetext[1]{M.B.--J.V. are listed in alphabetical order. This research was supported by a Seedcorn Research Grant from the European Association of Social Psychology.}
+}
+
+\noindent Protest movements often spark contention about what actions
+are acceptable means of protest. Reactions to Black American athletes
 kneeling during the national anthem to protest racist police violence
@@
 demonstrated, however, that observers are less supportive of social
-movements that use extreme protest actions. Similarly, Teixeira et al.
-(2020) found that advantaged-group members perceive non-normative
-collective action by disadvantaged-group members as more damaging to
-their ingroup's social image and are thus less supportive of such
-action. Verkuyten et al. (2022) found that, in general, the public tends
-to be intolerant of non-normative collective action. This means that
-activists face the dilemma that non-normative collective action might be
-most effective at gaining concessions but that it also risks reducing
-popular support for a cause. Together, these studies show that the
-distinction between normative and non-normative collective action is
-psychologically and societally consequential.
+movements that use extreme protest actions.\footnote{In contrast, White
+  Americans exposed to pictures of destructive Black Lives Matter
+  protests were not, on average, any less supportive of systemic change
+  than those exposed to pictures of peaceful protests (Teixeira et al.,
+  2022).} Similarly, Teixeira et al. (2020) found that advantaged-group
+members perceive non-normative collective action by disadvantaged-group
+members as more damaging to their ingroup's social image and are thus
+less supportive of such action. Verkuyten et al. (2022) found that, in
+general, the public tends to be intolerant of non-normative collective
+action. This means that activists face the dilemma that non-normative
+collective action might be most effective at gaining concessions but
+that it also risks reducing popular support for a cause. Together, these
+studies show that the distinction between normative and non-normative
+collective action is psychologically and societally consequential.
 
@@
 
+Our findings provided a relevant comparison to findings from a recent
+experiment in which White Americans viewed pictures of either peaceful
+or destructive Black Lives Matter protests and rated how legitimate
+(i.e., fair, reasonable, legitimate, and justified) they considered the
+pictured protests to be (Teixeira et al., 2022). In this experiment,
+more conservative participants rated destructive and, to a lesser
+extent, peaceful protests to be less legitimate than more liberal
+participants. Likewise, participants who identified more strongly with
+their ethnic ingroup rated destructive and, to a lesser extent, peaceful
+protests to be less legitimate than participants who identified less
+strongly with their ingroup. However, participants' group identification
+was no longer associated with their protest ratings after controlling
+for their political orientation.\footnote{These findings are based on
+  our reanalysis, reported in the SOM, of Teixeira et al.'s (2022)
+  experiment. We thank the authors for making their data available
+  online.} In this way, results from the experiment aligned with our
+conclusion that political ideologies are more important than social
+identities for shaping how people judge controversial forms of
+collective action.
+
 Most broadly, our findings showed that, for many controversial protest
@@
 intergroup relations} (pp. 33--48). Brooks/Cole.
+
+\leavevmode\vadjust pre{\hypertarget{ref-teixeira_white_2022}{}}%
+Teixeira, C. P., Leach, C. W., \& Spears, R. (2022). White {Americans}'
+belief in systemic racial injustice and in-group identification affect
+reactions to ({peaceful} vs. {destructive}) {``{Black Lives Matter}''}
+protest. \emph{Psychology of Violence}, \emph{12}(4), 280--292.
+\url{https://doi.org/10.1037/vio0000425}
 

### 2022-09-19 Documents/supplemental-materials.Rmd
update preprint

@@
 We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
+
+# Reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment
+
+At the height of the 2020 Black Lives Matter protests, @teixeira_white_2022 exposed White Americans ($N = 399$) to photos of either peaceful or destructive Black Lives Matter protests and asked them to rate how legitimate (i.e., fair, reasonable, legitimate, and justified) they thought the pictured protests were.
+As such, their findings make for an interesting comparison to our findings.
+@teixeira_white_2022, however, reported results from a linear model estimating the three-way interaction between the experimental manipulation, belief in systemic racial injustice, and group identification while controlling for political orientation.
+This analysis did not provide a direct comparison to our findings.
+
+To compare theirs to our findings, we instead estimated a series of simpler linear regression models, using the brms R package [@burkner_brms_2017; @burkner_advanced_2018] and weakly informative Student-*t* $(3, 0, 2.5)$ prior distributions, with participants' ratings of how legitimate the pictured protests were as the outcome variable.^[We thank the authors for making their data available online ([https://osf.io/sud2y/](https://osf.io/sud2y/)).] First, we found that, as expected, participants considered peaceful protests far more legitimate than destructive protests ($\text{Cohen's}~d = 1.14, [0.98, 1.30]$).
+Second, we found that more conservative participants rated destructive protests ($\beta_{xy} = -0.59, [-0.68, -0.49]$) and, to a lesser extent, peaceful protests ($\beta_{xy} = -0.20, [-0.30, -0.10]$) to be less legitimate than more liberal participants.
+Third, we found that participants who identified more strongly with their (White American) ethnic identity rated destructive protests ($\beta_{xy} = -0.35, [-0.45, -0.24]$) and, to a lesser extent, peaceful protests ($\beta_{xy} = -0.10, [-0.21, 0.01]$) to be less legitimate than participants who identified less strongly with their (White American) ethnic identity.
+Finally, we found that participants' group identification was no longer associated with how legitimate they rated peaceful ($\beta_{xy} = -0.02, [-0.13, 0.05]$) and destructive ($\beta_{xy} = -0.06, [-0.17, 0.05]$) protests to be after controlling for their political orientation.
+We refer to findings from our reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment in the main text.
 

### 2022-09-19 Documents/supplemental-materials.tex
update preprint

@@
 
+\hypertarget{experiment-3}{%
+\section{Experiment 3}\label{experiment-3}}
+
+\hypertarget{method-1}{%
+\subsection{Method}\label{method-1}}
+
+\hypertarget{measures}{%
+\subsubsection{Measures}\label{measures}}
+
+We included additional measures not used in the analyses reported in
+this paper.
+
+We included a three-item measure of moral conviction (Ryan, 2014) that
+assessed to what extent participants' feelings about restricting or
+banning abortion were, for example, connected to their core moral
+beliefs or convictions (1 = \emph{not at all}, 5 = \emph{very much}).
+
+We measured gender-related system-justifying beliefs with eight items
+(adapted from Jost \& Kay, 2005), for example, ``in general, relations
+between men and women are fair'' (1 = \emph{strongly disagree}, 7 =
+\emph{strongly agree}; McDonald's \(\omega = .93\)). A confirmatory
+factor analysis model in which all items loaded onto a single factor
+showed acceptable fit,
+\(\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]\).
+
+We measured moral concerns with the 36-item moral foundations
+questionnaire (MFQ-2, Atari et al., 2022) which assesses to what extent
+participants endorse concerns about care, equality, proportionality,
+loyalty, authority, and purity (e.g., ``I believe chastity is an
+important virtue.''; 1 = \emph{does not describe me at all}, 5 =
+\emph{describes me extremely well}). We embedded three further attention
+checks within the questionnaire (e.g., ``To show that you are paying
+attention and giving your best effort, please select `moderately
+describes me'.'').
+
+\hypertarget{reanalysis-of-teixeira-et-al.s--teixeira_white_2022-experiment}{%
+\section{Reanalysis of Teixeira et al.'s (2022)
+experiment}\label{reanalysis-of-teixeira-et-al.s--teixeira_white_2022-experiment}}
+
+At the height of the 2020 Black Lives Matter protests, Teixeira et al.
+(2022) exposed White Americans (\(N = 399\)) to photos of either
+peaceful or destructive Black Lives Matter protests and asked them to
+rate how legitimate (i.e., fair, reasonable, legitimate, and justified)
+they thought the pictured protests were. As such, their findings make
+for an interesting comparison to our findings. Teixeira et al. (2022),
+however, reported results from a linear model estimating the three-way
+interaction between the experimental manipulation, belief in systemic
+racial injustice, and group identification while controlling for
+political orientation. This analysis did not provide a direct comparison
+to our findings.
+
+To compare theirs to our findings, we instead estimated a series of
+simpler linear regression models, using the brms R package (Bürkner,
+2017, 2018) and weakly informative Student-\emph{t} \((3, 0, 2.5)\)
+prior distributions, with participants' ratings of how legitimate the
+pictured protests were as the outcome variable.\footnote{We thank the
+  authors for making their data available online
+  (\url{https://osf.io/sud2y/}).} First, we found that, as expected,
+participants considered peaceful protests far more legitimate than
+destructive protests (\(\text{Cohen's}~d = 1.14, [0.98, 1.30]\)).
+Second, we found that more conservative participants rated destructive
+protests (\(\beta_{xy} = -0.59, [-0.68, -0.49]\)) and, to a lesser
+extent, peaceful protests (\(\beta_{xy} = -0.20, [-0.30, -0.10]\)) to be
+less legitimate than more liberal participants. Third, we found that
+participants who identified more strongly with their (White American)
+ethnic identity rated destructive protests
+(\(\beta_{xy} = -0.35, [-0.45, -0.24]\)) and, to a lesser extent,
+peaceful protests (\(\beta_{xy} = -0.10, [-0.21, 0.01]\)) to be less
+legitimate than participants who identified less strongly with their
+(White American) ethnic identity. Finally, we found that participants'
+group identification was no longer associated with how legitimate they
+rated peaceful (\(\beta_{xy} = -0.02, [-0.13, 0.05]\)) and destructive
+(\(\beta_{xy} = -0.06, [-0.17, 0.05]\)) protests to be after controlling
+for their political orientation. We refer to findings from our
+reanalysis of Teixeira et al.'s (2022) experiment in the main text.
+
 \newpage
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2022}{}}%
+Atari, M., Haidt, J., Graham, J., Koleva, S., Stevens, S. T., \&
+Dehghani, M. (2022). Morality beyond the {WEIRD}: How the nomological
+network of morality varies across cultures. \emph{PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/q6c9r}
+
 \leavevmode\vadjust pre{\hypertarget{ref-betancourt_ordinal_2019}{}}%
@@
 \url{https://betanalpha.github.io/assets/case_studies/ordinal_regression.html}
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner_brms_2017}{}}%
+Bürkner, P.-C. (2017). {brms}: An {R} package for bayesian multilevel
+models using {Stan}. \emph{Journal of Statistical Software},
+\emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner_advanced_2018}{}}%
+Bürkner, P.-C. (2018). Advanced bayesian multilevel modeling with the
+{R} package brms. \emph{The R Journal}, \emph{10}(1), 395--411.
+\url{https://doi.org/10.32614/RJ-2018-017}
 
@@
 \url{https://doi.org/10.18637/jss.v100.i05}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jost_exposure_2005}{}}%
+Jost, J. T., \& Kay, A. C. (2005). Exposure to benevolent sexism and
+complementary gender stereotypes: Consequences for specific and diffuse
+forms of system justification. \emph{Journal of Personality and Social
+Psychology}, \emph{88}(3), 498--509.
+\url{https://doi.org/10.1037/0022-3514.88.3.498}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ryan_reconsidering_2014}{}}%
+Ryan, T. J. (2014). Reconsidering moral issues in politics. \emph{The
+Journal of Politics}, \emph{76}(2), 380--397.
+\url{https://doi.org/10.1017/S0022381613001357}
 
@@
 Sargent.
+
+\leavevmode\vadjust pre{\hypertarget{ref-teixeira_white_2022}{}}%
+Teixeira, C. P., Leach, C. W., \& Spears, R. (2022). White {Americans}'
+belief in systemic racial injustice and in-group identification affect
+reactions to ({peaceful} vs. {destructive}) {``{Black Lives Matter}''}
+protest. \emph{Psychology of Violence}, \emph{12}(4), 280--292.
+\url{https://doi.org/10.1037/vio0000425}
 

### 2022-09-20 Documents/manuscript.Rmd
submit manuscript

@@
 header: "DOUBLE STANDARDS"
-keywords: protest, collective action, double standards, social identity theory, system justification theory, Black Lives Matter
+keywords: collective action, double standards, social identity theory, system justification theory, Black Lives Matter
 abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
@@
 
+### Transparency and Openness
+
 We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations ([https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b](https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b)).
@@
 That is, we tested whether Black and White Americans applied different standards when judging collective action taken by either Black or White protesters to protest either for or against police divestment as a possible solution to racist police violence.
-We conducted this study in January 2021, a time when, after months of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement.
+We conducted this study in January 2021, a time when, after months of unprecedented protests for racial justice, most Americans could be expected to be aware of, and have formed an opinion on, the Black Lives Matter movement [@leach_social_2017; @leach_understanding_2022].
 We focused on defunding the police as this position remained controversial among both Black and White Americans even as they broadly agreed on the need for police reform.^[In December 2020, only 34% of Black Americans and 18% of White Americans supported defunding the police while majorities of Black and White Americans supported police reforms such as eliminating qualified immunity (50% and 57%) or banning chokeholds [79% and 61%, @yougov_police_2020].]
@@
 ## Method
+
+### Transparency and Openness
 

### 2022-09-20 Documents/manuscript.tex
submit manuscript

@@
 
+\hypertarget{transparency-and-openness}{%
+\subsubsection{Transparency and
+Openness}\label{transparency-and-openness}}
+
 We preregistered the sample size as well as all hypotheses,
@@
 protests for racial justice, most Americans could be expected to be
-aware of, and have formed an opinion on, the Black Lives Matter
-movement. We focused on defunding the police as this position remained
-controversial among both Black and White Americans even as they broadly
-agreed on the need for police reform.\footnote{In December 2020, only
-  34\% of Black Americans and 18\% of White Americans supported
-  defunding the police while majorities of Black and White Americans
-  supported police reforms such as eliminating qualified immunity (50\%
-  and 57\%) or banning chokeholds (79\% and 61\%, YouGov, 2020).}
+aware of, and have formed an opinion on, the Black Lives Matter movement
+(Leach \& Allen, 2017; Leach \& Teixeira, 2022). We focused on defunding
+the police as this position remained controversial among both Black and
+White Americans even as they broadly agreed on the need for police
+reform.\footnote{In December 2020, only 34\% of Black Americans and 18\%
+  of White Americans supported defunding the police while majorities of
+  Black and White Americans supported police reforms such as eliminating
+  qualified immunity (50\% and 57\%) or banning chokeholds (79\% and
+  61\%, YouGov, 2020).}
 
@@
 \subsection{Method}\label{method-1}}
+
+\hypertarget{transparency-and-openness-1}{%
+\subsubsection{Transparency and
+Openness}\label{transparency-and-openness-1}}
 
@@
 
+\hypertarget{transparency-and-openness-2}{%
+\subsubsection{Transparency and
+Openness}\label{transparency-and-openness-2}}
+
 We preregistered the sample size, hypotheses, inclusion/exclusion
@@
 \url{http://www.bethanylacina.com/docs/Lacina_NFL_paper.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-leach_social_2017}{}}%
+Leach, C. W., \& Allen, A. M. (2017). The social psychology of the
+{Black Lives Matter} meme and movement. \emph{Current Directions in
+Psychological Science}, \emph{26}(6), 543--547.
+\url{https://doi.org/10.1177/0963721417719319}
+
+\leavevmode\vadjust pre{\hypertarget{ref-leach_understanding_2022}{}}%
+Leach, C. W., \& Teixeira, C. P. (2022). Understanding sentiment toward
+{``{Black Lives Matter}.''} \emph{Social Issues and Policy Review},
+\emph{16}(1), 3--32. \url{https://doi.org/10.1111/sipr.12084}
 

### 2023-04-19 Documents/manuscript.Rmd
switch to Quarto

@@
 
+### Are Partisan Double Standards Based in Ideology or Identity?
+
+So far, we ...
+
+<!-- Throughout this paper, we will use "identity-based" as a shorthand for double standards based on social categories such as race, gender, and class and "ideology-based" as a shorthand for double standards based on system-justification and political ideology.
+-->
+
 ## Using Item Response Theory to Detect Double Standards
@@
 --> When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two-parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters:
-$$ \Pr ( y_{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
-where $\Pr ( y_{ij} = 1 )$ is the probability that respondent $j$ considers action $i$ an acceptable means of protest, $\theta_j$ estimates how *accepting* respondent $j$ is of various protest actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.
+$$ \Pr ( y_\textit{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
+where $\Pr ( y_\textit{ij} = 1 )$ is the probability that respondent $j$ considers action $i$ an acceptable means of protest, $\theta_j$ estimates how *accepting* respondent $j$ is of various protest actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.
 

### 2023-04-19 Documents/manuscript.tex
switch to Quarto

@@
 % Font
-\usepackage[utf8]{inputenc}
-\usepackage{MinionPro} 
-\input{glyphtounicode}
-\pdfgentounicode=1 
+\usepackage[MnSymbol]{mathspec}
+\setmainfont{Minion Pro}
+\setmathsfont(Digits,Latin,Greek)
+[Numbers={Lining,Proportional}]{Minion Pro}
 \usepackage{microtype}
@@
 \renewcommand\Affilfont{\small}
+
+% References/Footnotes
 \newcommand{\refsection}{\section{References}}
-\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Header
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{14.5pt}
+\renewcommand{\headrulewidth}{0pt}
+\fancyhead[L]{ DOUBLE STANDARDS }
+\fancyhead[R]{\thepage}
 
@@
 \usepackage[colorlinks = true, linkcolor = black, urlcolor = black, citecolor = black]{hyperref}
+\usepackage{xurl}
 
@@
 
+% Lists
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 % References
@@
 
-% Frontmatter
+% Front matter
 \title{ Double Standards in Judging Collective Action\thanks{} }
@@
 when the protesters' cause aligns with their own ideological
-positions---but do not show consistent ingroup bias when judging
-collective action by ingroup and outgroup members.
+positions---but do not show consistent ingroup bias (in terms of class,
+race, and gender) when judging collective action by ingroup and outgroup
+members.
 
@@
 2012; van Zomeren, 2016)---is a powerful force driving social change.
-For example, Black Lives Matter protests lastingly shifted public
-discourse about racial inequity toward antiracist ideas (Dunivin et al.,
-2022).
+For example, Black Lives Matter protests shifted public discourse about
+racial inequity toward antiracist ideas (Dunivin et al., 2022).
 
@@
 and unacceptable (non-normative) forms of collective action. For that
-purpose, we define a double standard as judging the same action as more
-or less acceptable depending on who is performing the action and for
-what reason (for an analogous definition, see Foschi, 2000).\footnote{Feinberg
-  et al. (2020) provided incidental evidence for double standards in
-  judging collective action as, in a manipulation check, Black
-  participants rated Black Lives Matter protests as less extreme than
-  White participants and liberal participants rated protest actions for
-  progressive causes as less extreme than conservative participants.} In
-this section, we derive predictions from social psychological theories
-about how identity-based and ideology-based double standards could lead
-people to judge the same protest actions as more or less acceptable
-depending on who the protesters are and what they are protesting.
+purpose, we define a double standard as judging the same protest actions
+as more or less acceptable depending on who the protesters are and what
+they are protesting. (for an analogous definition, see Foschi,
+2000).\footnote{Feinberg et al. (2020) provided incidental evidence for
+  double standards in judging collective action as, in a manipulation
+  check, Black participants rated Black Lives Matter protests as less
+  extreme than White participants and liberal participants rated protest
+  actions for progressive causes as less extreme than conservative
+  participants.}
 
@@
 
-In the same vain, we propose that ingroup bias results in identity-based
+In the same vein, we propose that ingroup bias results in identity-based
 double standards in judging collective action by ingroup and outgroup
@@
 differences in what actions liberals and conservatives consider
-acceptable means of protest for progressive and conservative causes.
-Ideology-based double standards could, for example, explain why far more
-Democrats than Republicans considered it appropriate to kneel during the
-national anthem to protest racist police violence (YouGov, 2017).
+acceptable means of protest for progressive and conservative
+causes---for example, why far more Democrats than Republicans considered
+it appropriate to kneel during the national anthem to protest racist
+police violence (YouGov, 2017).
+
+\hypertarget{are-partisan-double-standards-based-in-ideology-or-identity}{%
+\subsubsection{Are Partisan Double Standards Based in Ideology or
+Identity?}\label{are-partisan-double-standards-based-in-ideology-or-identity}}
+
+\begin{itemize}
+\tightlist
+\item
+  Are Partisan Double Standards Based in Ideology or Identity?
+\item
+  Evidence for partisanship as a elective affinity (Jost, 2021)
+\item
+  Evidence for partisanship as a social identity
+\item
+  We will use ``identity-based'' as a shorthand for double standards
+  based on assigned categories such as class, race, and gender and
+  ``ideology-based'' as a shorthand for double standards based on
+  system-justification and political orientation.
+\end{itemize}
 
@@
 (unobserved) parameters:
-\[ \Pr ( y_{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) \]
-where \(\Pr ( y_{ij} = 1 )\) is the probability that respondent \(j\)
-considers action \(i\) an acceptable means of protest, \(\theta_j\)
-estimates how \emph{accepting} respondent \(j\) is of various protest
-actions, \(\beta_i\) estimates how \emph{acceptable} action \(i\) is,
-and \(\alpha_i\) estimates how \emph{discriminating} action \(i\) is.
+\[ \Pr ( y_\textit{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) \]
+where \(\Pr ( y_\textit{ij} = 1 )\) is the probability that respondent
+\(j\) considers action \(i\) an acceptable means of protest,
+\(\theta_j\) estimates how \emph{accepting} respondent \(j\) is of
+various protest actions, \(\beta_i\) estimates how \emph{acceptable}
+action \(i\) is, and \(\alpha_i\) estimates how \emph{discriminating}
+action \(i\) is.
 
@@
 as the ones we had asked about (1 = \emph{never}, 5 = \emph{very often})
-and whether they has participated in protests for reforming, defunding
+and whether they had participated in protests for reforming, defunding
 or abolishing the police; against reforming, defunding or abolishing the
@@
 (\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}).
+Our sample size was determined by budget and time constraints.
 
@@
 collective action are not subject to double standards based on group
-identities such as class, race, or gender. Why did we not find any
-evidence for identity-based double standards?
+identities such as class, race, or gender.
 
@@
 asymmetry is across different political causes and cultural contexts.
-
-Our findings provided a relevant comparison to findings from a recent
-experiment in which White Americans viewed pictures of either peaceful
-or destructive Black Lives Matter protests and rated how legitimate
-(i.e., fair, reasonable, legitimate, and justified) they considered the
-pictured protests to be (Teixeira et al., 2022). In this experiment,
-more conservative participants rated destructive and, to a lesser
-extent, peaceful protests to be less legitimate than more liberal
-participants. Likewise, participants who identified more strongly with
-their ethnic ingroup rated destructive and, to a lesser extent, peaceful
-protests to be less legitimate than participants who identified less
-strongly with their ingroup. However, participants' group identification
-was no longer associated with their protest ratings after controlling
-for their political orientation.\footnote{These findings are based on
-  our reanalysis, reported in the SOM, of Teixeira et al.'s (2022)
-  experiment. We thank the authors for making their data available
-  online.} In this way, results from the experiment aligned with our
-conclusion that political ideologies are more important than social
-identities for shaping how people judge controversial forms of
-collective action.
 
@@
 University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-jost_left_2021}{}}%
+Jost, J. T. (2021). \emph{Left and right: The psychological significance
+of a political distinction}. Oxford University Press.
 

### 2023-04-19 Documents/supplemental-materials.tex
switch to Quarto

@@
 \usepackage{xltabular}
+
+% Lists
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 
+\setcounter{table}{3}
+
+\begin{table}[t!]
+\caption{Correlations between variables in Experiment 1}
+\centering
+\begin{tabular}{rlrrrr}
+\toprule
+\# & Variable              &    1 &    2 &    3 &    4 \\ \midrule
+ 1 & Proportion            &      & -.31 & -.16 & -.23 \\
+ 2 & Political Orientation & -.31 &      &  .41 &  .41 \\
+ 3 & Social Dominance      & -.16 &  .41 &      &  .29 \\
+ 4 & System Justification  & -.23 &  .41 &  .29 &      \\ \bottomrule
+\addlinespace
+\end{tabular}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Political Orientation (1 = \textit{left}, 7 = \textit{right}); Social Dominance Orientation (\textit{factor score}); Systen-Justifying Beliefs (\textit{factor score})}
+\end{table}
+
+\setcounter{table}{4}
+
+\begin{table}[t!]
+\caption{Correlations between variables in Experiment 2}
+\centering
+\begin{tabular}{rlrrrrr}
+\toprule
+\# & Variable              &    1 &    2 &    3 &    4 &    5 \\ \midrule
+ 1 & Proportion            &      & -.21 &  .29 & -.23 & -.07 \\
+ 2 & System Justification  & -.21 &      & -.45 &  .47 &  .04 \\
+ 3 & Policy Support        &  .29 & -.45 &      & -.49 &  .00 \\
+ 4 & Political Orientation & -.23 &  .47 & -.49 &      &  .03 \\
+ 5 & Identification        & -.07 &  .04 &  .00 &  .03 &      \\ \bottomrule
+\addlinespace
+\end{tabular}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Systen-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Racial Identification (1 = \textit{not at all}, 7 = \textit{very much})}
+\end{table}
+
+\setcounter{table}{5}
+
+\begin{table}[t!]
+\caption{Correlations between variables in Experiment 3}
+\centering
+\begin{tabular}{rlrrrrr}
+\toprule
+\# & Variable              &    1 &    2 &    3 &    4 &    5 \\ \midrule
+ 1 & Proportion            &      & -.27 & -.22 & -.32 & -.11 \\
+ 2 & System Justification  & -.27 &      &  .61 &  .78 &  .16 \\
+ 3 & Policy Support        & -.22 &  .61 &      &  .74 &  .16 \\
+ 4 & Political Orientation & -.32 &  .78 &  .74 &      &  .20 \\
+ 5 & Identification        & -.11 &  .16 &  .16 &  .20 &      \\ \bottomrule
+\addlinespace
+\end{tabular}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Systen-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Gender Identification (1 = \textit{not at all}, 7 = \textit{very much})}
+\end{table}
+
 \end{document}

### 2023-04-19 Experiment 2/figures/figure-5.R

@@ -238,9 +238,9 @@ rm(list = ls())
     geom_hline(yintercept = c(4.5), size = 3*0.455, colour = "white") +
     scale_x_continuous(minor_breaks = NULL) +
     scale_y_discrete(expand = c(0, 0)) +
-    scale_shape_manual(values = c("+1SD" = 19, "-1SD" = 21)) +
-    scale_colour_manual(values = c("+1SD" = "#648FFF", "-1SD" = "#DC267F")) +
-    scale_fill_manual(values = c("+1SD" = "#648FFF", "-1SD" = "#DC267F")) +
+    scale_shape_manual(values = c("+1SD" = 21, "-1SD" = 19)) +
+    scale_colour_manual(values = c("+1SD" = "#DC267F", "-1SD" = "#648FFF")) +
+    scale_fill_manual(values = c("+1SD" = "#DC267F", "-1SD" = "#648FFF")) +
     facet_grid(. ~ predictor) +
     coord_cartesian(ylim = c(0.5, 8.5)) +
     theme_grey(base_size = 10) +

### 2023-04-19 Experiment 3/figures/figure-6.R

@@ -84,9 +84,9 @@ rm(list = ls())
     geom_hline(yintercept = c(4.5), size = 3*0.455, colour = "white") +
     scale_x_continuous(minor_breaks = NULL) +
     scale_y_reverse(expand = c(0, 0)) +
-    scale_shape_manual(values = c("Conservative" = 19, "Liberal" = 21)) +
-    scale_colour_manual(values = c("Conservative" = "#648FFF", "Liberal" = "#DC267F")) +
-    scale_fill_manual(values = c("Conservative" = "#648FFF", "Liberal" = "#DC267F")) +
+    scale_shape_manual(values = c("Conservative" = 21, "Liberal" = 19)) +
+    scale_colour_manual(values = c("Conservative" = "#DC267F", "Liberal" = "#648FFF")) +
+    scale_fill_manual(values = c("Conservative" = "#DC267F", "Liberal" = "#648FFF")) +
     coord_cartesian(xlim = c(-0.65, 1.15)) +
     facet_grid(. ~ "") +
     theme_grey(base_size = 10) +

### 2023-04-19 Documents/manuscript.tex
switch to Quarto

@@
-\documentclass[twocolumn, 11pt, letterpaper]{article}
+\documentclass[12pt, letterpaper]{article}
 
 % Font
-\usepackage[MnSymbol]{mathspec}
-\setmainfont{Minion Pro}
-\setmathsfont(Digits,Latin,Greek)
-[Numbers={Lining,Proportional}]{Minion Pro}
+\usepackage{amsmath, amssymb}
+\usepackage{lmodern}
 \usepackage{microtype}
@@
 \usepackage[letterpaper, margin = 1in]{geometry}
-\setlength{\columnsep}{18pt}
 \setcounter{secnumdepth}{0}
+\setlength{\parindent}{0.5in}
 \usepackage[compact]{titlesec}
-\titleformat*{\section}{\centering\normalfont\Large\bfseries}
+\titleformat*{\section}{\centering\normalfont\normalsize\bfseries}
 \titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
 \titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
-\usepackage{authblk}
-\renewcommand\Affilfont{\small}
+\usepackage{indentfirst}
 
 % References/Footnotes
-\newcommand{\refsection}{\section{References}}
+\newcommand{\refsection}{\newpage \section{References}}
+\usepackage[hang, flushmargin]{footmisc}
 
@@
 
-% Front matter
-\title{ Double Standards in Judging Collective Action\thanks{} }
-\author[1]{Nils K. Reimer}
-\affil[1]{University of California, Santa Barbara, United States}
-\author[2]{Marija Brankovi\'{c}}
-\affil[2]{Singidunum University and University of Belgrade, Serbia}
-\author[3]{Iniobong Essien}
-\affil[3]{Leuphana University of L\"{u}neburg, Germany}
-\author[4]{Jin X. Goh}
-\affil[4]{Colby College, United States}
-\author[5]{S\'{e}bastien Goudeau}
-\affil[5]{University of Poitiers, France}
-\author[6]{N\'{o}ra A. Lantos}
-\affil[6]{E\"{o}tv\"{o}s Lor\'{o}nd University, Hungary}
-\author[7]{Jenny Veldman}
-\affil[7]{NYU Abu Dhabi, United Arab Emirates}
+% Line spacing
+\usepackage[doublespacing]{setspace}
 
@@
 prominent social movements.
-    \end{abstract}
-  \end{@twocolumnfalse}
-  % \renewcommand{\abstractname}{Statement of Relevance}
-  % \begin{@twocolumnfalse}
-  %   \maketitle
-  %   \begin{abstract}
-  %   \noindent 
-  %   \end{abstract}
-  % \end{@twocolumnfalse}
-]
-
-{
-  \renewcommand{\thefootnote}%
-    {\fnsymbol{footnote}}
-  \footnotetext[1]{M.B.--J.V. are listed in alphabetical order. This research was supported by a Seedcorn Research Grant from the European Association of Social Psychology.}
-}
-
-\noindent Protest movements often spark contention about what actions
-are acceptable means of protest. Reactions to Black American athletes
+
+\textit{Keywords:} collective action, double standards, social identity
+theory, system justification theory, Black Lives Matter
+
+\newpage
+
+\section{ Double Standards in Judging Collective Action }
+
+Protest movements often spark contention about what actions are
+acceptable means of protest. Reactions to Black American athletes
 kneeling during the national anthem to protest racist police violence

### 2023-04-24 Documents/manuscript.qmd
switch to Quarto

--- 2023-04-19 Documents/manuscript.Rmd
@@
 header: "DOUBLE STANDARDS"
-keywords: collective action, double standards, social identity theory, system justification theory, Black Lives Matter
 abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
@@
 Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
+keywords: collective action, double standards, social identity theory, system justification theory, Black Lives Matter
 bibliography: references.bib
 csl: apa.csl
-output:
-  pdf_document:
-    latex_engine: pdflatex
-    keep_tex: yes
+execute:
+  echo: false
+format:
+  pdf:
+    keep-tex: true
+    pdf-engine: xelatex
     template: apa-template.tex
-
 ---
 
-```{r setup, include = FALSE}
+```{r}
+#| include: false
 
@@
 library(tidyverse); library(tidybayes); library(numform); library(glue)
-
-  # knitr options
-  knitr::opts_chunk$set(echo = FALSE)
 
@@
 In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
-We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show consistent ingroup bias when judging collective action by ingroup and outgroup members.
+We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show consistent ingroup bias (in terms of class, race, and gender) when judging collective action by ingroup and outgroup members.
 
@@
 Collective action---that is, any action taken by group members to advance a shared political goal [for similar definitions, see @becker_virtual_2012; @van_zomeren_building_2016]---is a powerful force driving social change.
-For example, Black Lives Matter protests lastingly shifted public discourse about racial inequity toward antiracist ideas [@dunivin_black_2022].
+For example, Black Lives Matter protests shifted public discourse about racial inequity toward antiracist ideas [@dunivin_black_2022].
 
@@
 In this research, we move this distinction from the researcher's intuition into the realm of scientific investigation and test for double standards in where people draw the line between acceptable (normative) and unacceptable (non-normative) forms of collective action.
-For that purpose, we define a double standard as judging the same action as more or less acceptable depending on who is performing the action and for what reason [for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.] In this section, we derive predictions from social psychological theories about how identity-based and ideology-based double standards could lead people to judge the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting.
+For that purpose, we define a double standard as judging the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting.
+[for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.]
 
@@
 
-In the same vain, we propose that ingroup bias results in identity-based double standards in judging collective action by ingroup and outgroup members.
+In the same vein, we propose that ingroup bias results in identity-based double standards in judging collective action by ingroup and outgroup members.
 That is, we hypothesize that people will judge the same protest actions as more acceptable if the protesters are ingroup members (*Hypothesis 1a*) or if the cause of the protest aligns with the ingroup's interests (*Hypothesis 1b*).
@@
 
-As system-justifying beliefs are closely related to political conservatism [@jost_political_2003], this hypothesis could explain partisan differences in what actions liberals and conservatives consider acceptable means of protest for progressive and conservative causes.
-Ideology-based double standards could, for example, explain why far more Democrats than Republicans considered it appropriate to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].
+As system-justifying beliefs are closely related to political conservatism [@jost_political_2003], this hypothesis could explain partisan differences in what actions liberals and conservatives consider acceptable means of protest for progressive and conservative causes---for example, why far more Democrats than Republicans considered it appropriate to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].
 
@@
 
-So far, we ...
-
-<!-- Throughout this paper, we will use "identity-based" as a shorthand for double standards based on social categories such as race, gender, and class and "ideology-based" as a shorthand for double standards based on system-justification and political ideology.
--->
+<!-- Write it in your own words and structure it (with references) later. -->
+<!-- Political orientation can be understood as ideology and identity. -->
+
+- Are Partisan Double Standards Based in Ideology or Identity?
+- Evidence for partisanship as a elective affinity [@jost_left_2021]
+- Evidence for partisanship as a social identity
+- We will use "identity-based" as a shorthand for double standards based on assigned categories such as class, race, and gender and "ideology-based" as a shorthand for double standards based on system-justification and political orientation.
+
+<!-- I could move the last paragraph of the previous section to here. -->
 
@@
 In addition to demographic questions, we asked participants how outraged they were about recent incidents of police violence against Black Americans, to what extent they identifed with the protesters described in the study, and to what extent they identifed with their racial ingroup (1 = *not at all*, 7 = *very much*).
-We also asked how often, if at all, participants had participated in protest actions such as the ones we had asked about (1 = *never*, 5 = *very often*) and whether they has participated in protests for reforming, defunding or abolishing the police; against reforming, defunding or abolishing the police; or in neither.
+We also asked how often, if at all, participants had participated in protest actions such as the ones we had asked about (1 = *never*, 5 = *very often*) and whether they had participated in protests for reforming, defunding or abolishing the police; against reforming, defunding or abolishing the police; or in neither.
 We measured political orientation with a one-item semantic differential scale: "People often describe their political orientation as liberal or conservative.
@@
 We made all materials, data, and analysis scripts available online ([https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)).
+Our sample size was determined by budget and time constraints.
 
@@
 And yet, our findings suggest that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
-Why did we not find any evidence for identity-based double standards?
+<!-- Why did we not find any evidence for identity-based double standards? -->
 
@@
 
-Our findings provided a relevant comparison to findings from a recent experiment in which White Americans viewed pictures of either peaceful or destructive Black Lives Matter protests and rated how legitimate (i.e., fair, reasonable, legitimate, and justified) they considered the pictured protests to be [@teixeira_white_2022].
+<!-- Our findings provided a relevant comparison to findings from a recent experiment in which White Americans viewed pictures of either peaceful or destructive Black Lives Matter protests and rated how legitimate (i.e., fair, reasonable, legitimate, and justified) they considered the pictured protests to be [@teixeira_white_2022].
 In this experiment, more conservative participants rated destructive and, to a lesser extent, peaceful protests to be less legitimate than more liberal participants.
@@
 We thank the authors for making their data available online.] In this way, results from the experiment aligned with our conclusion that political ideologies are more important than social identities for shaping how people judge controversial forms of collective action.
+-->
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 

### 2023-04-24 Documents/supplemental-materials.qmd
switch to Quarto

--- 2022-09-19 Documents/supplemental-materials.Rmd
@@
 bibliography: references.bib
-output:
-  pdf_document:
-    latex_engine: pdflatex
-    keep_tex: yes
+csl: apa.csl
+execute:
+  echo: false
+format:
+  pdf:
+    keep-tex: true
+    pdf-engine: xelatex
     template: som-template.tex
-csl: apa.csl
 ---
@@
 ```
+
+```{=latex}
+\setcounter{table}{3}
+
+\begin{table}[t!]
+\caption{Correlations between variables in Experiment 1}
+\centering
+\begin{tabular}{rlrrrr}
+\toprule
+\# & Variable              &    1 &    2 &    3 &    4 \\ \midrule
+ 1 & Proportion            &      & -.31 & -.16 & -.23 \\
+ 2 & Political Orientation & -.31 &      &  .41 &  .41 \\
+ 3 & Social Dominance      & -.16 &  .41 &      &  .29 \\
+ 4 & System Justification  & -.23 &  .41 &  .29 &      \\ \bottomrule
+\addlinespace
+\end{tabular}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Political Orientation (1 = \textit{left}, 7 = \textit{right}); Social Dominance Orientation (\textit{factor score}); Systen-Justifying Beliefs (\textit{factor score})}
+\end{table}
+```
+
+```{=latex}
+\setcounter{table}{4}
+
+\begin{table}[t!]
+\caption{Correlations between variables in Experiment 2}
+\centering
+\begin{tabular}{rlrrrrr}
+\toprule
+\# & Variable              &    1 &    2 &    3 &    4 &    5 \\ \midrule
+ 1 & Proportion            &      & -.21 &  .29 & -.23 & -.07 \\
+ 2 & System Justification  & -.21 &      & -.45 &  .47 &  .04 \\
+ 3 & Policy Support        &  .29 & -.45 &      & -.49 &  .00 \\
+ 4 & Political Orientation & -.23 &  .47 & -.49 &      &  .03 \\
+ 5 & Identification        & -.07 &  .04 &  .00 &  .03 &      \\ \bottomrule
+\addlinespace
+\end{tabular}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Systen-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Racial Identification (1 = \textit{not at all}, 7 = \textit{very much})}
+\end{table}
+```
+
+```{=latex}
+\setcounter{table}{5}
+
+\begin{table}[t!]
+\caption{Correlations between variables in Experiment 3}
+\centering
+\begin{tabular}{rlrrrrr}
+\toprule
+\# & Variable              &    1 &    2 &    3 &    4 &    5 \\ \midrule
+ 1 & Proportion            &      & -.27 & -.22 & -.32 & -.11 \\
+ 2 & System Justification  & -.27 &      &  .61 &  .78 &  .16 \\
+ 3 & Policy Support        & -.22 &  .61 &      &  .74 &  .16 \\
+ 4 & Political Orientation & -.32 &  .78 &  .74 &      &  .20 \\
+ 5 & Identification        & -.11 &  .16 &  .16 &  .20 &      \\ \bottomrule
+\addlinespace
+\end{tabular}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Systen-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Gender Identification (1 = \textit{not at all}, 7 = \textit{very much})}
+\end{table}
+```

### 2023-04-26 Documents/manuscript.qmd
add paragraph on partisanship as a social identity

@@
 In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
-We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no consistent ingroup bias when judging collective action by ingroup and outgroup members.
+We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no consistent ingroup bias in terms of class, race, and gender.
 Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
@@
 
-<!-- Write it in your own words and structure it (with references) later. -->
-<!-- Political orientation can be understood as ideology and identity. -->
-
-- Are Partisan Double Standards Based in Ideology or Identity?
-- Evidence for partisanship as a elective affinity [@jost_left_2021]
-- Evidence for partisanship as a social identity
-- We will use "identity-based" as a shorthand for double standards based on assigned categories such as class, race, and gender and "ideology-based" as a shorthand for double standards based on system-justification and political orientation.
-
-<!-- I could move the last paragraph of the previous section to here. -->
+Partisan differences can result from two social psychological processes.
+On the one hand, political partisanship reflects ideological differences in resistance to social change and acceptance of inequality which, in turn, are motivated by epistemic, existential, and relational needs [@jost_political_2003; @jost_left_2021].
+As motivated cognition, ideological orientations shape how we think and feel about issues, actions, and events.
+On the other hand, political partisanship functions as a social identity [@finkel_political_2020; @osborne_psychology_2022] with evidence suggesting that partisans' policy preferences depends more on the stated position of their party than policy content [@cohen_party_2003].
+Like other identities, political identities result in group-serving biases in how we evaluate ingroup and outgroup members.
+In practice, the two processes are often difficult to disentangle.
+
+For now, we use "identity-based" to refer to double standards based on assigned social categories such as class, race, and gender and "ideology-based" to refer to double standards based on ideological alignment.
+We compare the evidence for identity and ideology processes in the General Discussion.
 
@@
 And yet, our findings suggest that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
-<!-- Why did we not find any evidence for identity-based double standards? -->
 
@@
 This explanation, however, conflicts with participants' reactions to the manipulation which, at least in Experiment 1--2, showed that participants understood the vignette in ingroup--outgroup terms.
+
+<!-- Instead, I should offer three explanations for the observed partisan biases: (1) ideology (Jost, 2020, 2021), (2) identity, and (3) moralizaton.
+I could close by suggesting experiments to tease them apart.
+Assymmetry is evidence against identity-based processes as we would expect ingroup bias in both directions.
+-->
 

### 2023-04-26 Documents/manuscript.tex
add paragraph on partisanship as a social identity

@@
 the protesters' cause aligned with their own ideological position---but
-showed no consistent ingroup bias when judging collective action by
-ingroup and outgroup members. Our findings have theoretical and
-practical implications for understanding the often divided response to
-prominent social movements.
+showed no consistent ingroup bias in terms of class, race, and gender.
+Our findings have theoretical and practical implications for
+understanding the often divided response to prominent social movements.
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Are Partisan Double Standards Based in Ideology or Identity?
-\item
-  Evidence for partisanship as a elective affinity (Jost, 2021)
-\item
-  Evidence for partisanship as a social identity
-\item
-  We will use ``identity-based'' as a shorthand for double standards
-  based on assigned categories such as class, race, and gender and
-  ``ideology-based'' as a shorthand for double standards based on
-  system-justification and political orientation.
-\end{itemize}
+Partisan differences can result from two social psychological processes.
+On the one hand, political partisanship reflects ideological differences
+in resistance to social change and acceptance of inequality which, in
+turn, are motivated by epistemic, existential, and relational needs
+(Jost et al., 2003; Jost, 2021). As motivated cognition, ideological
+orientations shape how we think and feel about issues, actions, and
+events. On the other hand, political partisanship functions as a social
+identity (Finkel et al., 2020; Harris et al., 2022) with evidence
+suggesting that partisans' policy preferences depends more on the stated
+position of their party than policy content (Cohen, 2003). Like other
+identities, political identities result in group-serving biases in how
+we evaluate ingroup and outgroup members. In practice, the two processes
+are often difficult to disentangle.
+
+For now, we use ``identity-based'' to refer to double standards based on
+assigned social categories such as class, race, and gender and
+``ideology-based'' to refer to double standards based on ideological
+alignment. We compare the evidence for identity and ideology processes
+in the General Discussion.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cohen_party_2003}{}}%
+Cohen, G. L. (2003). Party over policy: The dominating impact of group
+influence on political beliefs. \emph{Journal of Personality and Social
+Psychology}, \emph{85}(5), 808--822.
+\url{https://doi.org/10.1037/0022-3514.85.5.808}
+
 \leavevmode\vadjust pre{\hypertarget{ref-demars_irt_2010}{}}%
@@
 \url{https://doi.org/10.1037/a0015141}
+
+\leavevmode\vadjust pre{\hypertarget{ref-osborne_psychology_2022}{}}%
+Harris, E., Pärnamets, P., Sternisko, A., Robertson, C., \& Van Bavel,
+J. J. (2022). The psychology and neuroscience of partisanship. In D.
+Osborne \& C. G. Sibley (Eds.), \emph{The {Cambridge} handbook of
+political psychology} (pp. 50--67). Cambridge University Press.
+\url{https://doi.org/10.1017/9781108779104.005}
 

### 2023-05-02 Documents/manuscript.qmd
first draft

@@
 As motivated cognition, ideological orientations shape how we think and feel about issues, actions, and events.
-On the other hand, political partisanship functions as a social identity [@finkel_political_2020; @osborne_psychology_2022] with evidence suggesting that partisans' policy preferences depends more on the stated position of their party than policy content [@cohen_party_2003].
+On the other hand, political partisanship functions as a social identity [@finkel_political_2020; @osborne_psychology_2022] with evidence suggesting that partisans' policy preferences depend more on the stated position of their party than policy content [@cohen_party_2003].
 Like other identities, political identities result in group-serving biases in how we evaluate ingroup and outgroup members.
@@
 ## Using Item Response Theory to Detect Double Standards
+
+Item response theory is a conceptual and statistical framework for understanding how the characteristics of both items and respondents shape responses to a set of items [@demars_irt_2010].
+<!-- As such, it allows us to disentangle how acceptable protest actions are from how accepting respondents are of them and, in so doing, to detect double standards in judging collective action.
+--> When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two-parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters:
+$$ \Pr ( y_\textit{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
+where $\Pr ( y_\textit{ij} = 1 )$ is the probability that respondent $j$ considers action $i$ an acceptable means of protest, $\theta_j$ estimates how *accepting* respondent $j$ is of various protest actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.
+
+Figure \ref{fig:f1} illustrates those relationships for three hypothetical protest actions.
+$\theta_{(z)}$ is the *z*-standardized tendency for a participant to be more accepting of various actions so that $\beta_i$ determines the probability that the average respondent ($\theta_{(z)} = 0$) considers an action an acceptable means of protest.
+Action 2 ($\beta_2 = 1.28$) is more acceptable than Actions 1 and 2 ($\{ \beta_2, \beta_3 \} = 0$) as the average respondent is more likely to consider the former ($\Pr = .78$) than the latter ($\Pr = .50$) acceptable means of protest.
+Action 3 ($\alpha_3 = 3.60$) is more discriminating than Actions 1 and 2 ($\{ \alpha_1, \alpha_2 \} = 1$) as the slope for the relationship between how accepting a respondent is and how likely they are to consider an action acceptable is steeper for the former action.
+This means that knowing whether someone considers Action 3 acceptable provides more information about them than knowing whether they consider Actions 1 or 2 acceptable.
 
@@
 ```
-
-Item response theory is a conceptual and statistical framework for understanding how the characteristics of both items and respondents shape responses to a set of items [@demars_irt_2010].
-<!-- As such, it allows us to disentangle how acceptable protest actions are from how accepting respondents are of them and, in so doing, to detect double standards in judging collective action.
---> When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two-parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters:
-$$ \Pr ( y_\textit{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
-where $\Pr ( y_\textit{ij} = 1 )$ is the probability that respondent $j$ considers action $i$ an acceptable means of protest, $\theta_j$ estimates how *accepting* respondent $j$ is of various protest actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.
-
-Figure \ref{fig:f1} illustrates those relationships for three hypothetical protest actions.
-$\theta_{(z)}$ is the *z*-standardized tendency for a participant to be more accepting of various actions so that $\beta_i$ determines the probability that the average respondent ($\theta_{(z)} = 0$) considers an action an acceptable means of protest.
-Action 2 ($\beta_2 = 1.28$) is more acceptable than Actions 1 and 2 ($\{ \beta_2, \beta_3 \} = 0$) as the average respondent is more likely to consider the former ($\Pr = .78$) than the latter ($\Pr = .50$) acceptable means of protest.
-Action 3 ($\alpha_3 = 3.60$) is more discriminating than Actions 1 and 2 ($\{ \alpha_1, \alpha_2 \} = 1$) as the slope for the relationship between how accepting a respondent is and how likely they are to consider an action acceptable is steeper for the former action.
-This means that knowing whether someone considers Action 3 acceptable provides more information about them than knowing whether they consider Actions 1 or 2 acceptable.
 
@@
 <!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.
---> In the remainder of this General Discussion, we consider some limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.
+--> In the remainder of this General Discussion, we consider limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.
 
@@
 Growing evidence shows that, in many situations, observers apply a more lenient standards when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
-And yet, our findings suggest that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
-
-One reason for this discrepancy might be that participants did not identify strongly enough with the social categories examined in our studies.
-This explanation, however, is not plausible because, first, we examined social categories (class, race, gender) that are central to most people's identities and because, second, we did not find evidence for ingroup bias even among highly identified group members.
-Another reason for this discrepancy might be that participants did not perceive the protesters' causes as relevant to the social identities examined in our studies.
-This explanation, however, conflicts with participants' reactions to the manipulation which, at least in Experiment 1--2, showed that participants understood the vignette in ingroup--outgroup terms.
-
-<!-- Instead, I should offer three explanations for the observed partisan biases: (1) ideology (Jost, 2020, 2021), (2) identity, and (3) moralizaton.
-I could close by suggesting experiments to tease them apart.
-Assymmetry is evidence against identity-based processes as we would expect ingroup bias in both directions.
--->
-
-A more plausible explanation might be that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013].
-Instead, observers might understand collective action primarily as a means to an end and judge how acceptable a form of collective action is in relation to its end ("the end justifies the mean").
-This aligns with our finding that how acceptable participants judged protest actions to be depended on the protesters' cause, not their group membership.
-
-Our findings showed that progressives and, to a lesser extent, conservatives consider the same controversial protest actions to be more acceptable when the protesters' cause aligns with their ideological position.
-So far, we explained this ideology-based double standard as resulting from conservatives being motivated to defend the system, progressives being motivated to challenge the system, and both being motivated to support collective action to those effects [@jost_missing_2017; @osborne_protesting_2019].
-Other explanations, however, are plausible.
-
-First, we might instead understand those double standards as an expression of ingroup bias based on an opinion-based [@mcgarty_collective_2009] or partisan [@finkel_political_2020] identity.
-Opinion-based identities are often a stronger predictor of collective action than identities based on other social categories [@bliuc_opinion-based_2007].
-This aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of the same transgressive protest actions when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
-
-Second, we might argue that their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
-This aligns with Richardson & Conway's [-@richardson_standing_2022] finding that their different moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives and vice versa.
-Future research should seek to disentangle which of those plausible mechanisms best explains partisan differences in judging collective action.
-
-Our findings raised the question why double standards in judging collective action were more pronounced among liberal than among conservative participants.
-One reason for this ideological asymmetry might be that collective action that violates prevailing norms or laws is seen as inherently threatening to the status quo, even if its aim is to bolster and defend the system.
-Political conservatism can be understood as having two psychologically related but distinguishable dimensions, resistance to change and justification of inequality [@jost_political_2003].
-In this way, conservatives might be expected to condemn disruptive protest even if the protest's cause is to defend inequality.
-Future research should seek to establish how consistent this ideological asymmetry is across different political causes and cultural contexts.
-
-<!-- Our findings provided a relevant comparison to findings from a recent experiment in which White Americans viewed pictures of either peaceful or destructive Black Lives Matter protests and rated how legitimate (i.e., fair, reasonable, legitimate, and justified) they considered the pictured protests to be [@teixeira_white_2022].
-In this experiment, more conservative participants rated destructive and, to a lesser extent, peaceful protests to be less legitimate than more liberal participants.
-Likewise, participants who identified more strongly with their ethnic ingroup rated destructive and, to a lesser extent, peaceful protests to be less legitimate than participants who identified less strongly with their ingroup.
-However, participants' group identification was no longer associated with their protest ratings after controlling for their political orientation.^[These findings are based on our reanalysis, reported in the SOM, of Teixeira et al.'s [-@teixeira_white_2022] experiment.
-We thank the authors for making their data available online.] In this way, results from the experiment aligned with our conclusion that political ideologies are more important than social identities for shaping how people judge controversial forms of collective action.
--->
-
-Most broadly, our findings showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause.
-Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
-In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.
+And yet, our findings suggested that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
+We did not find evidence for those double standards even though we examined social categories that are central to most people's identities, even when we considered only highly identified group members, and even though participants' reactions to the manipulation showed that, at least in Experiments 1--2, they understood the protesters' causes in ingroup--outgroup terms.
+
+Instead, our findings showed that progressives and, to a lesser extent, conservatives judge the same controversial protest actions to be more acceptable when the protesters' cause aligns with their own ideological position.
+We consider social, political, and moral psychological explanations for the observed ideology-based double standards.
+
+First, we might understand ideology-based double standards as an expression of ingroup bias based on opinion-based [@mcgarty_collective_2009] or partisan [@osborne_psychology_2022; @finkel_political_2020] identities.
+This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
+An identity-based account, however, implies that both conservatives and progressives would show ingroup bias and, therefore, cannot explain the observed asymmetry in the two groups' double standards.
+
+Second, we might explain ideology-based double standards, as we have done so far, as resulting from conservatives being motivated to defend the system, progressives being motivated to challenge the system, and both being motivated to support collective action to those effects [@jost_missing_2017; @osborne_protesting_2019].
+An ideology-based account could also  explain the asymmetry in how progressives and conservatives judged collective action for various causes.
+Political conservatism can be understood as uniting two motives, to resist social change and to maintain social inequality [@jost_political_2003; @jost_left_2021].
+If conservatives perceive disruptive protest actions as inherently threatening to the social order, they will be motivated to condemn such actions even when the protesters' cause supports the unequal status quo (e.g., protesting *against* defunding the police).
+
+Lastly, we might consider that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013]---but is instead understood as a means to an end and judged in relation to its end ("the end justifies the mean").
+Their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] might lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
+This explanation aligns with Richardson & Conway's [-@richardson_standing_2022] finding that moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives (and vice versa).
+A values-based account could also explain why conservatives who, more so than liberals, endorse moral concerns related to loyalty and authority  tend to reject disruptive protest actions even when the protesters' cause otherwise aligns with their moral concerns.
+
+Future research should, across different causes, contexts, and cultures, seek to disentangle which of the proposed social, political, and moral psychological processes best explains partisan differences in judging collective action.
+
+Most broadly, our findings showed that, for many controversial protest actions, there is no consensus on whether they are acceptable means to advance a cause.
+<!--Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
+-->This confirms our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.
 By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provide a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
 
-After all, this distinction is often consequential.
+After all, this distinction is consequential.
 If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
@@
 If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
-In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].
-For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.
+In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].<!-- For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.-->
 
@@
 Our research showed that, in contrast to well-documented double standards in other domains, people do not apply a different standard when judging collective action by members of the same class, race or gender or for a cause aligned with their own group's interests.
-Instead, our research demonstrated that people consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological positions---in other words, that there are partisan double standards in judging collective action.
+Instead, our research demonstrated that people consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological position<!-- ---in other words, that there are partisan double standards in judging collective action -->.
 

### 2023-05-02 Documents/manuscript.tex
first draft

@@
 identity (Finkel et al., 2020; Harris et al., 2022) with evidence
-suggesting that partisans' policy preferences depends more on the stated
+suggesting that partisans' policy preferences depend more on the stated
 position of their party than policy content (Cohen, 2003). Like other
@@
 Standards}\label{using-item-response-theory-to-detect-double-standards}}
-
-\begin{figure}
-\centering
-\caption{Item response curves for three hypothetical protest actions}
-\includegraphics[scale=1]{../Scale Development/figures/figure-1}
-\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
-\label{fig:f1}
-\end{figure}
 
@@
 Actions 1 or 2 acceptable.
+
+\begin{figure}
+\centering
+\caption{Item response curves for three hypothetical protest actions}
+\includegraphics[scale=1]{../Scale Development/figures/figure-1}
+\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
+\label{fig:f1}
+\end{figure}
 
@@
 restricting abortion---aligned with their own ideological position. In
-the remainder of this General Discussion, we consider some limiting
+the remainder of this General Discussion, we consider limiting
 conditions on our findings and discuss theoretical and practical
@@
 (Abrams et al., 2013; Endevelt et al., 2021). And yet, our findings
-suggest that, unlike judgments in other domains, judgments about
+suggested that, unlike judgments in other domains, judgments about
 collective action are not subject to double standards based on group
-identities such as class, race, or gender.
-
-One reason for this discrepancy might be that participants did not
-identify strongly enough with the social categories examined in our
-studies. This explanation, however, is not plausible because, first, we
-examined social categories (class, race, gender) that are central to
-most people's identities and because, second, we did not find evidence
-for ingroup bias even among highly identified group members. Another
-reason for this discrepancy might be that participants did not perceive
-the protesters' causes as relevant to the social identities examined in
-our studies. This explanation, however, conflicts with participants'
-reactions to the manipulation which, at least in Experiment 1--2, showed
-that participants understood the vignette in ingroup--outgroup terms.
-
-A more plausible explanation might be that collective action, even in
-its controversial forms, is not judged as a moral transgression (as in
+identities such as class, race, or gender. We did not find evidence for
+those double standards even though we examined social categories that
+are central to most people's identities, even when we considered only
+highly identified group members, and even though participants' reactions
+to the manipulation showed that, at least in Experiments 1--2, they
+understood the protesters' causes in ingroup--outgroup terms.
+
+Instead, our findings showed that progressives and, to a lesser extent,
+conservatives judge the same controversial protest actions to be more
+acceptable when the protesters' cause aligns with their own ideological
+position. We consider social, political, and moral psychological
+explanations for the observed ideology-based double standards.
+
+First, we might understand ideology-based double standards as an
+expression of ingroup bias based on opinion-based (McGarty et al., 2009)
+or partisan (Finkel et al., 2020; Harris et al., 2022) identities. This
+explanation aligns with Verkuyten et al.'s (2022) finding that
+participants were more tolerant of transgressive protest actions when
+taken by their most-liked, rather than their least-liked, political
+group (e.g., climate activists). An identity-based account, however,
+implies that both conservatives and progressives would show ingroup bias
+and, therefore, cannot explain the observed asymmetry in the two groups'
+double standards.
+
+Second, we might explain ideology-based double standards, as we have
+done so far, as resulting from conservatives being motivated to defend
+the system, progressives being motivated to challenge the system, and
+both being motivated to support collective action to those effects (Jost
+et al., 2017; Osborne et al., 2019). An ideology-based account could
+also explain the asymmetry in how progressives and conservatives judged
+collective action for various causes. Political conservatism can be
+understood as uniting two motives, to resist social change and to
+maintain social inequality (Jost et al., 2003; Jost, 2021). If
+conservatives perceive disruptive protest actions as inherently
+threatening to the social order, they will be motivated to condemn such
+actions even when the protesters' cause supports the unequal status quo
+(e.g., protesting \emph{against} defunding the police).
+
+Lastly, we might consider that collective action, even in its
+controversial forms, is not judged as a moral transgression (as in
 previous research on identity-based double standards, e.g., Abrams et
-al., 2013). Instead, observers might understand collective action
-primarily as a means to an end and judge how acceptable a form of
-collective action is in relation to its end (``the end justifies the
-mean''). This aligns with our finding that how acceptable participants
-judged protest actions to be depended on the protesters' cause, not
-their group membership.
-
-Our findings showed that progressives and, to a lesser extent,
-conservatives consider the same controversial protest actions to be more
-acceptable when the protesters' cause aligns with their ideological
-position. So far, we explained this ideology-based double standard as
-resulting from conservatives being motivated to defend the system,
-progressives being motivated to challenge the system, and both being
-motivated to support collective action to those effects (Jost et al.,
-2017; Osborne et al., 2019). Other explanations, however, are plausible.
-
-First, we might instead understand those double standards as an
-expression of ingroup bias based on an opinion-based (McGarty et al.,
-2009) or partisan (Finkel et al., 2020) identity. Opinion-based
-identities are often a stronger predictor of collective action than
-identities based on other social categories (Bliuc et al., 2007). This
-aligns with Verkuyten et al.'s (2022) finding that participants were
-more tolerant of the same transgressive protest actions when taken by
-their most-liked, rather than their least-liked, political group (e.g.,
-climate activists).
-
-Second, we might argue that their different moral concerns (Graham et
-al., 2009; Kivikangas et al., 2021) lead progressives and conservatives
-to see their support of certain causes as a fundamental matter of right
-or wrong (Skitka et al., 2021) and, by thus moralizing those causes, to
-accept more extreme means to achieving them. This aligns with Richardson
-\& Conway's (2022) finding that their different moral concerns explained
-why liberals rated protesting for liberal causes as more moral than
-conservatives and vice versa. Future research should seek to disentangle
-which of those plausible mechanisms best explains partisan differences
-in judging collective action.
-
-Our findings raised the question why double standards in judging
-collective action were more pronounced among liberal than among
-conservative participants. One reason for this ideological asymmetry
-might be that collective action that violates prevailing norms or laws
-is seen as inherently threatening to the status quo, even if its aim is
-to bolster and defend the system. Political conservatism can be
-understood as having two psychologically related but distinguishable
-dimensions, resistance to change and justification of inequality (Jost
-et al., 2003). In this way, conservatives might be expected to condemn
-disruptive protest even if the protest's cause is to defend inequality.
-Future research should seek to establish how consistent this ideological
-asymmetry is across different political causes and cultural contexts.
+al., 2013)---but is instead understood as a means to an end and judged
+in relation to its end (``the end justifies the mean''). Their different
+moral concerns (Graham et al., 2009; Kivikangas et al., 2021) might lead
+progressives and conservatives to see their support of certain causes as
+a fundamental matter of right or wrong (Skitka et al., 2021) and, by
+thus moralizing those causes, to accept more extreme means to achieving
+them. This explanation aligns with Richardson \& Conway's (2022) finding
+that moral concerns explained why liberals rated protesting for liberal
+causes as more moral than conservatives (and vice versa). A values-based
+account could also explain why conservatives who, more so than liberals,
+endorse moral concerns related to loyalty and authority tend to reject
+disruptive protest actions even when the protesters' cause otherwise
+aligns with their moral concerns.
+
+Future research should, across different causes, contexts, and cultures,
+seek to disentangle which of the proposed social, political, and moral
+psychological processes best explains partisan differences in judging
+collective action.
 
 Most broadly, our findings showed that, for many controversial protest
-actions, there is no consensus on what is and is not an acceptable means
-to advance a cause. Instead, this judgment seems to depend on what the
-cause of a protest is and how it aligns with one's beliefs and values.
-In this way, our results confirm our contention that the distinction
+actions, there is no consensus on whether they are acceptable means to
+advance a cause. This confirms our contention that the distinction
 between normative and non-normative forms of collective action should be
@@
 
-After all, this distinction is often consequential. If people believe
-that even extreme actions are acceptable means to advance a cause they
+After all, this distinction is consequential. If people believe that
+even extreme actions are acceptable means to advance a cause they
 support, especially if they moralize their support for that cause
@@
 highways, Quinton, 2021) that might be most effective at challenging
-injustice (Shuman et al., 2022, 2021). For those reasons, we encourage
-researchers to use our paradigm to further investigate when and why
-people apply different standards when judging collective action for
-different causes.
+injustice (Shuman et al., 2022, 2021).
 
@@
 consider the same protest actions to be more acceptable when the cause
-of the protest aligns with their own ideological positions---in other
-words, that there are partisan double standards in judging collective
-action.
+of the protest aligns with their own ideological position.
 
@@
 \emph{37}(8), 1104--1116. \url{https://doi.org/10.1177/0146167211407076}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bliuc_opinion-based_2007}{}}%
-Bliuc, A.-M., McGarty, C., Reynolds, K., \& Muntele, D. (2007).
-Opinion-based group membership as a predictor of commitment to political
-action. \emph{European Journal of Social Psychology}, \emph{37}(1),
-19--32. \url{https://doi.org/10.1002/ejsp.334}
 

### 2023-05-31 Documents/manuscript.qmd
submit

@@
 keywords: collective action, double standards, social identity theory, system justification theory, Black Lives Matter
+statement: |
+  This research shows that what people consider an acceptable *means* of protest depends on what the *end* of a protest is.
+That is, people tend to judge the *same* controversial protest actions to be more acceptable when the protesters' cause aligned with their own ideological positions.
+This finding helps explain the often divided response to recent protest movements.
+word-count: "11,546 words"
 bibliography: references.bib
@@
 In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action.
-In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
+In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights, for or against defunding the police, and for or against restricting abortion.
 We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show consistent ingroup bias (in terms of class, race, and gender) when judging collective action by ingroup and outgroup members.
@@
 In this research, we move this distinction from the researcher's intuition into the realm of scientific investigation and test for double standards in where people draw the line between acceptable (normative) and unacceptable (non-normative) forms of collective action.
-For that purpose, we define a double standard as judging the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting.
-[for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.]
+For that purpose, we define a double standard as judging the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting [for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.]
 
@@
 On the one hand, political partisanship reflects ideological differences in resistance to social change and acceptance of inequality which, in turn, are motivated by epistemic, existential, and relational needs [@jost_political_2003; @jost_left_2021].
-As motivated cognition, ideological orientations shape how we think and feel about issues, actions, and events.
+As motivated cognition, ideological differences shape how we think and feel about issues, actions, and events.
 On the other hand, political partisanship functions as a social identity [@finkel_political_2020; @osborne_psychology_2022] with evidence suggesting that partisans' policy preferences depend more on the stated position of their party than policy content [@cohen_party_2003].
@@
 The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more acceptable if the protesters' are ingroup members or the protesters' cause aligns with the ingroup's interests (*Hypothesis 1a*, *1b*) or if the protesters' cause aligns with their own system-justifying or system-challenging motives (*Hypotheses 2a*, *2b*).
+
 In Experiment 1, we use the instrument developed in Studies 1 and 2 to test for double standards in how acceptable participants consider collective action for workers' rights in the United Kingdom.
@@
 
-Model 1 (M1) estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause.
+Model 1 estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause.
 Contradicting *Hypothesis 1a*, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`).
@@
 
-Model 2 (M2) extended Model 1 by estimating participants' responses as a function of their *z*-standardized endorsement of system-justifying beliefs.
+Model 2 extended Model 1 by estimating participants' responses as a function of their *z*-standardized endorsement of system-justifying beliefs.
 As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgments about system-defending protest actions and one estimating their effect on judgments about system-challenging protest actions, and one varying effect estimating its variance across conditions.
@@
 
-Model 3 (M3) extended Model 1 by estimating participants' responses as a function of their self-reported support for the cause of the protest.
+Model 3 extended Model 1 by estimating participants' responses as a function of their self-reported support for the cause of the protest.
 We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police.
@@
 If they selected the wrong answer, they were instructed to reread the text and select the right answer.
-Participants then rated to what extent they support or oppose restricting abortion and to what extent their feelings about this issue are a moral conviction.
+Participants then rated to what extent they supported or opposed restricting abortion and to what extent their feelings about this issue were a moral conviction.
 
@@
 On the next pages, participants completed the remaining measures.
-On the final page, participants responded to two attention checks that assessed whether they paid attention to the experimental manipulation: "In this study, you read about some people's view on a policy change.
+On the final page, participants responded to two attention checks that assessed whether they had paid attention to the experimental manipulation: "In this study, you read about some people's view on a policy change.
 What was the policy change?" and "In this study, you then read about some protesters.
@@
 
-Growing evidence shows that, in many situations, observers apply a more lenient standards when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
+Growing evidence shows that, in many situations, observers apply a more lenient standard when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
 And yet, our findings suggested that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
@@
 This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
-An identity-based account, however, implies that both conservatives and progressives would show ingroup bias and, therefore, cannot explain the observed asymmetry in the two groups' double standards.
+An identity-based account, however, implies that both conservatives and progressives would show ingroup bias and, therefore, cannot explain the observed asymmetry in the two groups' judgements.
 
@@
 
-Lastly, we might consider that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013]---but is instead understood as a means to an end and judged in relation to its end ("the end justifies the mean").
+Lastly, we might consider that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013]---but is instead understood as a means to an end and judged in relation to its end ("the end justifies the means").
 Their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] might lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
-This explanation aligns with Richardson & Conway's [-@richardson_standing_2022] finding that moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives (and vice versa).
+This explanation aligns with Richardson and Conway's [-@richardson_standing_2022] finding that moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives (and vice versa).
 A values-based account could also explain why conservatives who, more so than liberals, endorse moral concerns related to loyalty and authority  tend to reject disruptive protest actions even when the protesters' cause otherwise aligns with their moral concerns.

### 2023-05-31 Documents/manuscript.tex
submit

@@
 
+\vfill
+
+\begin{center}
+\noindent 11,546 words
+\end{center}
+
+\vfill
+
+\section{Public Significance Statement}
+
+\noindent This research shows that what people consider an acceptable
+\emph{means} of protest depends on what the \emph{end} of a protest is.
+That is, people tend to judge the \emph{same} controversial protest
+actions to be more acceptable when the protesters' cause aligned with
+their own ideological positions. This finding helps explain the often
+divided response to recent protest movements.
+
 \newpage
@@
 to test for double standards in judging protest actions for workers'
-rights in the United Kingdom and for or against defunding the police and
-for or against restricting abortion in the United States. We find
-evidence for ideology-based, but not for identity-based, double
-standards: People judge the same protest actions to be more acceptable
-when the protesters' cause aligns with their own ideological
-positions---but do not show consistent ingroup bias (in terms of class,
-race, and gender) when judging collective action by ingroup and outgroup
-members.
+rights, for or against defunding the police, and for or against
+restricting abortion. We find evidence for ideology-based, but not for
+identity-based, double standards: People judge the same protest actions
+to be more acceptable when the protesters' cause aligns with their own
+ideological positions---but do not show consistent ingroup bias (in
+terms of class, race, and gender) when judging collective action by
+ingroup and outgroup members.
 
@@
 as more or less acceptable depending on who the protesters are and what
-they are protesting. (for an analogous definition, see Foschi,
+they are protesting (for an analogous definition, see Foschi,
 2000).\footnote{Feinberg et al. (2020) provided incidental evidence for
@@
 (Jost et al., 2003; Jost, 2021). As motivated cognition, ideological
-orientations shape how we think and feel about issues, actions, and
+differences shape how we think and feel about issues, actions, and
 events. On the other hand, political partisanship functions as a social
@@
 with their own system-justifying or system-challenging motives
-(\emph{Hypotheses 2a}, \emph{2b}). In Experiment 1, we use the
-instrument developed in Studies 1 and 2 to test for double standards in
-how acceptable participants consider collective action for workers'
-rights in the United Kingdom. By varying both the participants' and the
-protesters' group memberships (working/middle class), Experiment 1
-compares the relative evidence for identity-based (\emph{Hypothesis 1a},
-\emph{1b}) and ideology-based double standards (\emph{Hypothesis 2a}).
-Extending Experiment 1, Experiment 2 considers both system-challenging
-and system-defending collective action in the context of the 2020 Black
-Lives Matter protests in the United States. By varying the protesters'
-cause (for/against defunding the police) as well as the participants'
-and the protesters' group memberships (Black/White), Experiment 2
-compares the relative evidence for identity-based (\emph{Hypothesis 1a},
-\emph{1b}) and ideology-based double standards (\emph{Hypothesis 2a},
-\emph{2b}). Building on Experiments 1--2, Experiment 3 examines how
-liberal and conservative men and women judge collective action for or
-against restricting abortion in anticipation of the Supreme Court
-overturning its \emph{Roe v. Wade} decision on reproductive rights in
-the United States. By varying the protesters' cause as well as the
-participants' ideology and gender, Experiment 3 compares the relative
-evidence for identity-based (\emph{Hypothesis 1b}) and ideology-based
+(\emph{Hypotheses 2a}, \emph{2b}).
+
+In Experiment 1, we use the instrument developed in Studies 1 and 2 to
+test for double standards in how acceptable participants consider
+collective action for workers' rights in the United Kingdom. By varying
+both the participants' and the protesters' group memberships
+(working/middle class), Experiment 1 compares the relative evidence for
+identity-based (\emph{Hypothesis 1a}, \emph{1b}) and ideology-based
+double standards (\emph{Hypothesis 2a}). Extending Experiment 1,
+Experiment 2 considers both system-challenging and system-defending
+collective action in the context of the 2020 Black Lives Matter protests
+in the United States. By varying the protesters' cause (for/against
+defunding the police) as well as the participants' and the protesters'
+group memberships (Black/White), Experiment 2 compares the relative
+evidence for identity-based (\emph{Hypothesis 1a}, \emph{1b}) and
+ideology-based double standards (\emph{Hypothesis 2a}, \emph{2b}).
+Building on Experiments 1--2, Experiment 3 examines how liberal and
+conservative men and women judge collective action for or against
+restricting abortion in anticipation of the Supreme Court overturning
+its \emph{Roe v. Wade} decision on reproductive rights in the United
+States. By varying the protesters' cause as well as the participants'
+ideology and gender, Experiment 3 compares the relative evidence for
+identity-based (\emph{Hypothesis 1b}) and ideology-based
 (\emph{Hypothesis 2b}) double standards in judging collective action
@@
 
-Model 1 (M1) estimated varying intercepts for the eight conditions to
-test whether, in line with Hypotheses 1a, 1b, or 2a, participants'
-responses depended on their own group membership, the protesters' group
+Model 1 estimated varying intercepts for the eight conditions to test
+whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses
+depended on their own group membership, the protesters' group
 membership, or the protesters' cause. Contradicting \emph{Hypothesis
@@
 
-Model 2 (M2) extended Model 1 by estimating participants' responses as a
+Model 2 extended Model 1 by estimating participants' responses as a
 function of their \emph{z}-standardized endorsement of system-justifying
@@
 
-Model 3 (M3) extended Model 1 by estimating participants' responses as a
+Model 3 extended Model 1 by estimating participants' responses as a
 function of their self-reported support for the cause of the protest. We
@@
 acceptable than the same actions performed by the relevant outgroup
-(\(d = 0.01, [-0.00, 0.02]\)) and did not consider protest actions for a
+(\(d = 0.05, [-0.01, 0.13]\)) and did not consider protest actions for a
 cause that was aligned with their ingroup's interests to be more
 acceptable than protest actions for a cause not aligned with their
-ingroup's interests (\(d = 0.00, [-0.01, 0.01]\)). Our non-preregistered
+ingroup's interests (\(d = 0.01, [-0.06, 0.08]\)). Our non-preregistered
 analyses thus suggested that group identification did not moderate group
@@
 the text and select the right answer. Participants then rated to what
-extent they support or oppose restricting abortion and to what extent
-their feelings about this issue are a moral conviction.
+extent they supported or opposed restricting abortion and to what extent
+their feelings about this issue were a moral conviction.
 
@@
 the remaining measures. On the final page, participants responded to two
-attention checks that assessed whether they paid attention to the
+attention checks that assessed whether they had paid attention to the
 experimental manipulation: ``In this study, you read about some people's
@@
 Growing evidence shows that, in many situations, observers apply a more
-lenient standards when judging moral transgressions by ingroup members
+lenient standard when judging moral transgressions by ingroup members
 (Abrams et al., 2013; Endevelt et al., 2021). And yet, our findings
@@
 and, therefore, cannot explain the observed asymmetry in the two groups'
-double standards.
+judgements.
 
@@
 al., 2013)---but is instead understood as a means to an end and judged
-in relation to its end (``the end justifies the mean''). Their different
-moral concerns (Graham et al., 2009; Kivikangas et al., 2021) might lead
-progressives and conservatives to see their support of certain causes as
-a fundamental matter of right or wrong (Skitka et al., 2021) and, by
-thus moralizing those causes, to accept more extreme means to achieving
-them. This explanation aligns with Richardson \& Conway's (2022) finding
-that moral concerns explained why liberals rated protesting for liberal
-causes as more moral than conservatives (and vice versa). A values-based
-account could also explain why conservatives who, more so than liberals,
-endorse moral concerns related to loyalty and authority tend to reject
-disruptive protest actions even when the protesters' cause otherwise
-aligns with their moral concerns.
+in relation to its end (``the end justifies the means''). Their
+different moral concerns (Graham et al., 2009; Kivikangas et al., 2021)
+might lead progressives and conservatives to see their support of
+certain causes as a fundamental matter of right or wrong (Skitka et al.,
+2021) and, by thus moralizing those causes, to accept more extreme means
+to achieving them. This explanation aligns with Richardson and Conway's
+(2022) finding that moral concerns explained why liberals rated
+protesting for liberal causes as more moral than conservatives (and vice
+versa). A values-based account could also explain why conservatives who,
+more so than liberals, endorse moral concerns related to loyalty and
+authority tend to reject disruptive protest actions even when the
+protesters' cause otherwise aligns with their moral concerns.
 

### 2023-05-31 Documents/supplemental-materials.qmd
submit

@@
 
-# Reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment
+<!-- # Reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment -->
 
-At the height of the 2020 Black Lives Matter protests, @teixeira_white_2022 exposed White Americans ($N = 399$) to photos of either peaceful or destructive Black Lives Matter protests and asked them to rate how legitimate (i.e., fair, reasonable, legitimate, and justified) they thought the pictured protests were.
+<!-- At the height of the 2020 Black Lives Matter protests, @teixeira_white_2022 exposed White Americans ($N = 399$) to photos of either peaceful or destructive Black Lives Matter protests and asked them to rate how legitimate (i.e., fair, reasonable, legitimate, and justified) they thought the pictured protests were.
 As such, their findings make for an interesting comparison to our findings.
@@
 This analysis did not provide a direct comparison to our findings.
+-->
 
-To compare theirs to our findings, we instead estimated a series of simpler linear regression models, using the brms R package [@burkner_brms_2017; @burkner_advanced_2018] and weakly informative Student-*t* $(3, 0, 2.5)$ prior distributions, with participants' ratings of how legitimate the pictured protests were as the outcome variable.^[We thank the authors for making their data available online ([https://osf.io/sud2y/](https://osf.io/sud2y/)).] First, we found that, as expected, participants considered peaceful protests far more legitimate than destructive protests ($\text{Cohen's}~d = 1.14, [0.98, 1.30]$).
+<!-- To compare theirs to our findings, we instead estimated a series of simpler linear regression models, using the brms R package [@burkner_brms_2017; @burkner_advanced_2018] and weakly informative Student-*t* $(3, 0, 2.5)$ prior distributions, with participants' ratings of how legitimate the pictured protests were as the outcome variable.^[We thank the authors for making their data available online ([https://osf.io/sud2y/](https://osf.io/sud2y/)).] First, we found that, as expected, participants considered peaceful protests far more legitimate than destructive protests ($\text{Cohen's}~d = 1.14, [0.98, 1.30]$).
 Second, we found that more conservative participants rated destructive protests ($\beta_{xy} = -0.59, [-0.68, -0.49]$) and, to a lesser extent, peaceful protests ($\beta_{xy} = -0.20, [-0.30, -0.10]$) to be less legitimate than more liberal participants.
@@
 We refer to findings from our reanalysis of Teixeira et al.'s [-@teixeira_white_2022] experiment in the main text.
+-->
 

### 2023-05-31 Documents/supplemental-materials.tex
submit

@@
 
-\hypertarget{reanalysis-of-teixeira-et-al.s--teixeira_white_2022-experiment}{%
-\section{Reanalysis of Teixeira et al.'s (2022)
-experiment}\label{reanalysis-of-teixeira-et-al.s--teixeira_white_2022-experiment}}
-
-At the height of the 2020 Black Lives Matter protests, Teixeira et al.
-(2022) exposed White Americans (\(N = 399\)) to photos of either
-peaceful or destructive Black Lives Matter protests and asked them to
-rate how legitimate (i.e., fair, reasonable, legitimate, and justified)
-they thought the pictured protests were. As such, their findings make
-for an interesting comparison to our findings. Teixeira et al. (2022),
-however, reported results from a linear model estimating the three-way
-interaction between the experimental manipulation, belief in systemic
-racial injustice, and group identification while controlling for
-political orientation. This analysis did not provide a direct comparison
-to our findings.
-
-To compare theirs to our findings, we instead estimated a series of
-simpler linear regression models, using the brms R package (Bürkner,
-2017, 2018) and weakly informative Student-\emph{t} \((3, 0, 2.5)\)
-prior distributions, with participants' ratings of how legitimate the
-pictured protests were as the outcome variable.\footnote{We thank the
-  authors for making their data available online
-  (\url{https://osf.io/sud2y/}).} First, we found that, as expected,
-participants considered peaceful protests far more legitimate than
-destructive protests (\(\text{Cohen's}~d = 1.14, [0.98, 1.30]\)).
-Second, we found that more conservative participants rated destructive
-protests (\(\beta_{xy} = -0.59, [-0.68, -0.49]\)) and, to a lesser
-extent, peaceful protests (\(\beta_{xy} = -0.20, [-0.30, -0.10]\)) to be
-less legitimate than more liberal participants. Third, we found that
-participants who identified more strongly with their (White American)
-ethnic identity rated destructive protests
-(\(\beta_{xy} = -0.35, [-0.45, -0.24]\)) and, to a lesser extent,
-peaceful protests (\(\beta_{xy} = -0.10, [-0.21, 0.01]\)) to be less
-legitimate than participants who identified less strongly with their
-(White American) ethnic identity. Finally, we found that participants'
-group identification was no longer associated with how legitimate they
-rated peaceful (\(\beta_{xy} = -0.02, [-0.13, 0.05]\)) and destructive
-(\(\beta_{xy} = -0.06, [-0.17, 0.05]\)) protests to be after controlling
-for their political orientation. We refer to findings from our
-reanalysis of Teixeira et al.'s (2022) experiment in the main text.
-
 \newpage
@@
 \url{https://betanalpha.github.io/assets/case_studies/ordinal_regression.html}
-
-\leavevmode\vadjust pre{\hypertarget{ref-burkner_brms_2017}{}}%
-Bürkner, P.-C. (2017). {brms}: An {R} package for bayesian multilevel
-models using {Stan}. \emph{Journal of Statistical Software},
-\emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
-
-\leavevmode\vadjust pre{\hypertarget{ref-burkner_advanced_2018}{}}%
-Bürkner, P.-C. (2018). Advanced bayesian multilevel modeling with the
-{R} package brms. \emph{The R Journal}, \emph{10}(1), 395--411.
-\url{https://doi.org/10.32614/RJ-2018-017}
 
@@
 Sargent.
-
-\leavevmode\vadjust pre{\hypertarget{ref-teixeira_white_2022}{}}%
-Teixeira, C. P., Leach, C. W., \& Spears, R. (2022). White {Americans}'
-belief in systemic racial injustice and in-group identification affect
-reactions to ({peaceful} vs. {destructive}) {``{Black Lives Matter}''}
-protest. \emph{Psychology of Violence}, \emph{12}(4), 280--292.
-\url{https://doi.org/10.1037/vio0000425}
 

### 2023-05-31 Experiment 1/figures/figure-3.R

@@ -233,13 +233,11 @@ rm(list = ls())
   # Export as .png
   ggsave(
     "Experiment 1/figures/figure-3.png",
-    width = 6.5, height = 6.5/3*2, units = "in", dpi = 600,
-    type = "cairo"
+    width = 6.5, height = 6.5/3*2, units = "in", dpi = 600
   )
   
   # Export as .pdf
   ggsave(
     "Experiment 1/figures/figure-3.pdf",
-    width = 6.5, height = 6.5/3*2, units = "in", dpi = 600,
-    device = cairo_pdf
+    width = 6.5, height = 6.5/3*2, units = "in", dpi = 600
   )  

### 2023-05-31 Experiment 2/figures/figure-4.R

@@ -2,7 +2,6 @@ rm(list = ls())
 
 # Notes -------------------------------------------------------------------
 
-  # CHECK FIGURE M1
 
 # Library -----------------------------------------------------------------
 
@@ -428,14 +427,12 @@ rm(list = ls())
   # Export as .png
   ggsave(
     "Experiment 2/figures/figure-4.png",
-    width = 6.5, height = 6.5/4*3, units = "in", dpi = 600,
-    type = "cairo"
+    width = 6.5, height = 6.5/4*3, units = "in", dpi = 600
   )
   
   # Export as .pdf
   ggsave(
     "Experiment 2/figures/figure-4.pdf",
-    width = 6.5, height = 6.5/4*3, units = "in",
-    device = cairo_pdf
+    width = 6.5, height = 6.5/4*3, units = "in"
   )  
   

### 2023-05-31 Experiment 2/figures/figure-5.R

@@ -278,14 +278,12 @@ rm(list = ls())
   # Export as .png
   ggsave(
     "Experiment 2/figures/figure-5.png",
-    width = 6.5, height = 6.5/5*2, units = "in", dpi = 600,
-    type = "cairo"
+    width = 6.5, height = 6.5/5*2, units = "in", dpi = 600
   )
   
   # Export as .pdf
   ggsave(
     "Experiment 2/figures/figure-5.pdf",
-    width = 6.5, height = 6.5/5*2, units = "in",
-    device = cairo_pdf
+    width = 6.5, height = 6.5/5*2, units = "in"
   )  
   

### 2023-05-31 Experiment 3/figures/figure-6.R

@@ -118,14 +118,12 @@ rm(list = ls())
   # Export as .png
   ggsave(
     "Experiment 3/figures/figure-6.png",
-    width = 6.5/4*3, height = 6.5/5*2, units = "in", dpi = 600,
-    type = "cairo"
+    width = 6.5/4*3, height = 6.5/5*2, units = "in", dpi = 600
   )
   
   # Export as .pdf
   ggsave(
     "Experiment 3/figures/figure-6.pdf",
-    width = 6.5/4*3, height = 6.5/5*2, units = "in",
-    device = cairo_pdf
+    width = 6.5/4*3, height = 6.5/5*2, units = "in"
   )    
   

### 2023-05-31 Scale Development/figures/figure-1.R

@@ -129,13 +129,12 @@ rm(list = ls())
   ggsave(
     "Scale Development/figures/figure-1.png",
     width = (6.5 - 18/72.27)/2, height = (6.5 - 18/72.27)/2, units = "in", 
-    dpi = 600, type = "cairo"
+    dpi = 600
   )
   
   # Export as .pdf
   ggsave(
     "Scale Development/figures/figure-1.pdf",
-    width = (6.5 - 18/72.27)/2, height = (6.5 - 18/72.27)/2, units = "in",
-    device = cairo_pdf
+    width = (6.5 - 18/72.27)/2, height = (6.5 - 18/72.27)/2, units = "in"
   )
   

### 2023-05-31 Scale Development/figures/figure-2.R

@@ -89,13 +89,12 @@ rm(list = ls())
   ggsave(
     "Scale Development/figures/figure-2.png",
     width = 6.5, height = 6.5/7*3, units = "in", 
-    dpi = 600, type = "cairo"
+    dpi = 600
   )
   
   # Export as .pdf
   ggsave(
     "Scale Development/figures/figure-2.pdf",
-    width = 6.5, height = 6.5/7*3, units = "in", 
-    device = cairo_pdf
+    width = 6.5, height = 6.5/7*3, units = "in"
   )
   

### 2024-04-16 Documents/manuscript.qmd
resubmit

@@
 header: "DOUBLE STANDARDS"
-abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
-In five studies (total *N* = 2,979), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*).
-In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action.
-In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States.
-We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no consistent ingroup bias in terms of class, race, and gender.
-Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
-keywords: collective action, double standards, social identity theory, system justification theory, Black Lives Matter
+abstract: |
+    Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
+Five studies (total $N = 2,979$) investigated double standards in judging collective action---that is, whether observers judge the same protest actions to be more acceptable depending on who the protesters are and what they are protesting.
+Two studies used item response theory to develop an instrument of 25 controversial protest actions to measure where people draw the line between acceptable and unacceptable forms of collective action.
+Three preregistered experiments found no consistent evidence for ingroup bias in terms of social class when judging protests for workers' rights (Experiment 1), in terms of race when judging protests for and against defunding the police (Experiment 2), and in terms of gender when judging protests for and against restricting abortion (Experiment 3).
+Instead, we found that participants who identified as left-wing/liberal (Experiments 1--3) and rejected system-justifying beliefs (Experiments 1--2) considered the same protest actions more acceptable when a cause aligned with their ideological orientation (for workers' rights, for defunding the police, against restricting abortion) than when it did not (against defunding the police, for restricting abortion).
+Conservative participants considered the same actions somewhat more acceptable when protesters supported, rather than opposed, restricting abortion (Experiment 3) but considered all protest actions, for and against defunding the police, equally unacceptable (Experiment 2).
+Our findings have theoretical and practical implications for understanding the often-divided response to social movements.
+keywords: collective action, double standards, social identity theory, system justification theory, partisanship, Black Lives Matter, abortion
 statement: |
   This research shows that what people consider an acceptable *means* of protest depends on what the *end* of a protest is.
-That is, people tend to judge the *same* controversial protest actions to be more acceptable when the protesters' cause aligned with their own ideological positions.
-This finding helps explain the often divided response to recent protest movements.
-word-count: "11,546 words"
+That is, observers sometimes judge the *same* controversial protest action (e.g., blocking a road) to be more acceptable when the protesters' cause (e.g., defunding the police) aligns with their own ideological position.
+Conservatives, who support social order and oppose social change, tend to consider all protest actions less acceptable than progressives.
+Progressives, in turn, tend to consider disruptive but non-violent actions acceptable means to achieve social change and challenge social injustice.
+This finding helps explain the often-divided response to recent protest movements.
+word-count: "11,654 words"
 bibliography: references.bib
@@
 format:
+  docx:
+    reference-doc: word-template.docx
   pdf:
@@
   # Load packages
-library(tidyverse); library(tidybayes); library(numform); library(glue)
+  library(tidyverse)
+  library(tidybayes)
+  library(numform)
+  library(glue)
 
@@
 Kneeling during the national anthem is not violent, disruptive, or illegal.
-And yet, only 29\% of White Americans (compared to 66\% of Black Americans) and 11\% of Republicans (compared to 59\% of Democrats) considered it appropriate for Black athletes to protest in this way [@yougov_nfl_2017].
-Indeed, Black athletes who had participated in the protests faced a backlash in public opinion [@lacina_nfl_2020].<!-- Likewise, 42\% of Americans thought that, in the 2020 Black Lives Matter protests, most protesters were trying to incite violence or destroy property, even though very few protests involved any violence or property destruction [@acled_year_2021].
---> This and other examples suggest that *who* the protesters are and *what* they are protesting influences how acceptable observers judge their actions to be.
-
-In this paper, we examine double standards in judging collective action---that is, whether observers judge the same protest actions as more or less acceptable depending on their own and the protesters' group memberships (*identity-based double standards*) or on the protesters' cause and how it aligns with their own ideological positions (*ideology-based double standards*).
-In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action.
+And yet, only 29\% of White Americans (compared to 66\% of Black Americans) and 11\% of Republicans (compared to 59\% of Democrats) considered it appropriate for Black athletes to protest in this way [@yougov_nfl_2017].<!-- Indeed, Black athletes who had participated in the protests faced a backlash in public opinion [@lacina_nfl_2020].--> <!--Likewise, 42\% of Americans thought that, in the 2020 Black Lives Matter protests, most protesters were trying to incite violence or destroy property, even though very few protests involved any violence or property destruction [@acled_year_2021].
+-->This and other examples suggest that *who* the protesters are and *what* they are protesting influences how acceptable their actions are judged to be.
+
+In this paper, we examine double standards in judging collective action---that is, whether observers judge the *same* protest actions as more or less acceptable depending on their own and the protesters' group memberships *or* on the protesters' cause and how it aligns with their own ideological positions.<!-- In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action.
 In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights, for or against defunding the police, and for or against restricting abortion.
-We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show consistent ingroup bias (in terms of class, race, and gender) when judging collective action by ingroup and outgroup members.
+We do not find consistent evidence for ingroup bias in terms of class, race, and gender when judging collective action by ingroup and outgroup members.
+Instead, we find that liberals and, less so, conservatives judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions.
+-->
 
@@
 
-Recent research has shown that this distinction matters for how people respond to collective action.
+Recent research has shown that this distinction matters for how people respond to collective action [for a review, see @shuman_when_2023].
 @shuman_disrupting_2021 showed that non-normative, non-violent collective action is more effective than either normative or violent collective action at gaining concessions from those most resistant to social change.
 @feinberg_activists_2020 demonstrated, however, that observers are less supportive of social movements that use extreme protest actions.^[In contrast, White Americans exposed to pictures of destructive Black Lives Matter protests were not, on average, any less supportive of systemic change than those exposed to pictures of peaceful protests [@teixeira_white_2022].] Similarly, @teixeira_is_2020 found that advantaged-group members perceive non-normative collective action by disadvantaged-group members as more damaging to their ingroup's social image and are thus less supportive of such action.
-@verkuyten_intolerance_2022 found that, in general, the public tends to be intolerant of non-normative collective action.
 This means that activists face the dilemma that non-normative collective action might be most effective at gaining concessions but that it also risks reducing popular support for a cause.
@@
 
-These and other studies have, for the most part, relied on ad-hoc distinctions between what researchers themselves considered normative and non-normative collective action.
+Past research has, for the most part, relied on ad-hoc distinctions between what researchers themselves considered normative and non-normative collective action.
 In liberal democracies, researchers tend to consider actions such as signing petitions, voting in elections, or peaceful protest to be normative and actions such as blocking traffic, damaging property, or violent protest to be non-normative.
-But, as reactions to Black athletes kneeling in protest suggest, the same action might be considered acceptable by some and unacceptable by others.
-That is, the distinction between normative and non-normative collective action might be in the eye of the beholder.
-
-In this research, we move this distinction from the researcher's intuition into the realm of scientific investigation and test for double standards in where people draw the line between acceptable (normative) and unacceptable (non-normative) forms of collective action.
-For that purpose, we define a double standard as judging the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting [for an analogous definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.]
-
-### Identity-Based Double Standards
+But, as reactions to Black athletes kneeling in protest suggest, the same action might be considered acceptable by some and unacceptable by others.<!-- That is, the distinction between normative and non-normative collective action might be in the eye of the beholder.
+--> In this research, we <!--move this distinction from the researcher's intuition into the realm of scientific investigation and -->test for double standards in where people draw the line between what is an acceptable (normative) and an unacceptable (non-normative) form of collective action.^[In this research, we use *acceptable* ["agreed or approved of by most people in a society", @acceptable] as a plain-language synonym for *normative* ("conforming to the norms of the existing social system").] For that purpose, we define a double standard as judging the same protest actions as more or less acceptable depending on who the protesters are and what they are protesting [for a similar definition, see @foschi_double_2000].^[@feinberg_activists_2020 provided incidental evidence for double standards in judging collective action as, in a manipulation check, Black participants rated Black Lives Matter protests as less extreme than White participants and liberal participants rated protest actions for progressive causes as less extreme than conservative participants.]
+
+### Group-Based Double Standards
+
+Far more Black ($66\%$) than White ($29\%$) Americans considered it appropriate for Black athletes to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].
+This suggests that what protest actions people consider acceptable might depend on their own and the protesters' group memberships.
 
@@
 One domain in which ingroup bias manifests is judgments about other people's actions.
-@hewstone_ultimate_1990 reviewed research showing that people make ingroup-serving causal attributions when judging actions by ingroup and outgroup members, for example, by attributing negative behavior by outgroup members to internal causes but attributing negative behavior by ingroup members to external causes [see @pettigrew_ultimate_1979].
+@hewstone_ultimate_1990 reviewed research showing that people make ingroup-serving causal attributions when judging actions by ingroup and outgroup members.
+For example, people tend to attribute negative behavior by outgroup members to internal causes but attribute negative behavior by ingroup members to external causes [see @pettigrew_ultimate_1979].
 @valdesolo_moral_2007 demonstrated that participants judged the same selfish action as less unfair when performed by themselves or an ingroup member than when performed by an outgroup member.
@@
 
-In the same vein, we propose that ingroup bias results in identity-based double standards in judging collective action by ingroup and outgroup members.
-That is, we hypothesize that people will judge the same protest actions as more acceptable if the protesters are ingroup members (*Hypothesis 1a*) or if the cause of the protest aligns with the ingroup's interests (*Hypothesis 1b*).
-This hypothesis could explain why, for example, far more Black than White Americans considered it appropriate for Black athletes to kneel down to protest police violence against Black Americans [@yougov_nfl_2017].
-Notably, this hypothesis applies equally to historically disadvantaged groups mobilizing against social injustice and to historically advantaged groups defending their group's position.
+In the same vein, we propose that ingroup bias results in double standards in judging collective action by ingroup and outgroup members.
+That is, we hypothesize that people will judge the same protest actions as more acceptable if the protesters are ingroup members and if the cause of the protest aligns with the ingroup's interests.
+Notably, this hypothesis applies equally to historically disadvantaged groups mobilizing against social injustice and to historically advantaged groups defending their group's position.<!-- In this research, we investigate group-based double standards in terms of class, race, and gender.
+-->
 
@@
 
-System Justification Theory [@jost_decade_2004; for a recent review, see @jost_theory_2020] argues that people are motivated to defend, justify, and bolster the prevailing social, economic, and political system because doing so serves basic epistemic, existential, and relational needs.
+Far more Democrats ($59\%$) than Republicans ($11\%$) considered it appropriate to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].
+This suggests that what protest actions people consider acceptable might depend on the protesters' cause and how it aligns with their own ideological positions.
+
+One explanation for ideology-based double standards is that they result from the same intergroup processes as other group-based double standards.
+In this view, political partisanship functions as a social identity [@finkel_political_2020; @osborne_psychology_2022] and results from universal cognitive mechanisms for forming, detecting, and maintaining alliances [@pinsof_strange_2023].
+Like other identities, political identities result in group-serving biases in how we evaluate ingroup and outgroup members.<!-- Likewise, political partisanship might result from universal cognitive mechanisms for forming, detecting, and maintaining alliances [@pinsof_strange_2023] which, in turn, cause biases that serve to support political allies and oppose political rivals.--> This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions (e.g., sending hate mail) when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
+
+Another explanation for ideology-based double standards is that they are rooted in ideological differences in how conservatives and progressives relate to social change and social inequality which, in turn, are motivated by epistemic, existential, and relational needs [@jost_political_2003; @jost_left_2021].
+As motivated social cognition, ideological differences shape how we think and feel about political issues, actions, and events.
+In this view, ideology-based double standards depend on whether the cause of a protest aligns with someone's ideological positions.
+
+System Justification Theory [@jost_decade_2004; for a recent review, see @jost_theory_2020], in particular, argues that people are motivated to defend, justify, and bolster the prevailing social, economic, and political system because doing so serves basic epistemic, existential, and relational needs.
 Both advantaged and disadvantaged groups are thought to be motivated to justify the existing system, although the strength of this motivation and its expression vary across individuals and situations.
+Conservatives, far more so than progressives, endorse system-justifying beliefs.
 
@@
 Just as system justification affects support for, and opposition to, collective action, we propose that it results in ideology-based double standards in what actions people consider acceptable forms of collective action.
-First, we hypothesize that  people will judge the same protest actions as more acceptable if the protest supports, rather than challenges, the system (*Hypothesis 2a*).
-This hypothesis reflects the assumption that the motivation to justify the system is, to some extent, universal.
-Second, we hypothesize that people will judge the same protest actions as more acceptable if they endorse system-justifying beliefs and the protest supports the system or if they reject system-justifying beliefs and the protest challenges the system (*Hypothesis 2b*).
-Notably, this hypothesis implies a symmetry in how people with different ideological orientations judge system-supporting and system-challenging collective action.
-
-As system-justifying beliefs are closely related to political conservatism [@jost_political_2003], this hypothesis could explain partisan differences in what actions liberals and conservatives consider acceptable means of protest for progressive and conservative causes---for example, why far more Democrats than Republicans considered it appropriate to kneel during the national anthem to protest racist police violence [@yougov_nfl_2017].
-
-### Are Partisan Double Standards Based in Ideology or Identity?
-
-Partisan differences can result from two social psychological processes.
-On the one hand, political partisanship reflects ideological differences in resistance to social change and acceptance of inequality which, in turn, are motivated by epistemic, existential, and relational needs [@jost_political_2003; @jost_left_2021].
-As motivated cognition, ideological differences shape how we think and feel about issues, actions, and events.
+That is, we hypothesize that people will judge the same protest actions as more acceptable if they endorse system-justifying beliefs and the protest supports the system or if they reject system-justifying beliefs and the protest challenges the system.
+<!-- Notably, this hypothesis implies a symmetry in how people with different ideological orientations judge system-supporting and system-challenging collective action.
+-->
+
+<!-- What both explanations have in common is [ingroup bias].
+[This perspective would explain ideology-based double standards as a group-based double standard.]  [Ingroup bias, in this account, is symmetric.] -->
+
+<!-- [In this view, the content of those coalitions is arbitrary and a historical accident.] [Cohen, 2003] -->
+
+<!-- A related perspective ... What both perspectives have in common, is that the content of ... and its effects are symmetric. -->
+
+<!-- When we understand political partisanship as an identity or alliance, we might understand ideology-based double standards as an expression of ingroup bias based on partisan identities and alliances.
+-->
+
+<!-- Partisan differences can result from two social psychological processes.
 On the other hand, political partisanship functions as a social identity [@finkel_political_2020; @osborne_psychology_2022] with evidence suggesting that partisans' policy preferences depend more on the stated position of their party than policy content [@cohen_party_2003].
 Like other identities, political identities result in group-serving biases in how we evaluate ingroup and outgroup members.
-In practice, the two processes are often difficult to disentangle.
-
-For now, we use "identity-based" to refer to double standards based on assigned social categories such as class, race, and gender and "ideology-based" to refer to double standards based on ideological alignment.
-We compare the evidence for identity and ideology processes in the General Discussion.
+In this view, ...
+-->
+
+<!-- In practice, the two processes are often difficult to disentangle. -->
+
+<!-- One explanation is that political partisanship reflects ideological differences in resistance to social change and acceptance of inequality which, in turn, are motivated by epistemic, existential, and relational needs (Jost et al., 2003; Jost, 2021).
+In particular, SJT ...
+-->
+
+<!-- On the one hand, political partisanship reflects ideological differences in resistance to social change and acceptance of inequality which, in turn, are motivated by epistemic, existential, and relational needs [@jost_political_2003; @jost_left_2021].
+As motivated cognition, ideological differences shape how we think and feel about issues, actions, and events.
+-->
+<!-- This perspective, however, assumes that ideologies are essentially arbitrary in content and symmetric.
+OR While this perspective, on its own, might explain double standards, it is incomplete.
+At the same time, there are real ...
+OR By assuming that political ideologies are arbitrary and symmetric, this perspective ...
+OR Individual Differences and Needs OR not SUFFICIENT TO EXPLAIN -->
+
+In principle, the two explanations for ideology-based double standards---in terms of intergroup processes and ideological differences---could be disentangled by separating political identities and causes.
+An explanation in terms of intergroup processes posits that political identities, not causes, shape partisan behavior.
+Indeed, research suggests that partisans' policy preferences depend more on the stated position of their party than policy content [@cohen_party_2003].
+An explanation in terms of ideological differences instead posits that there is an "elective affinity" between political causes and the psychological needs and motives of their supporters [@jost_left_2021].
+In other words, the content of political ideologies is not arbitrary or a historical accident, as implied by an explanation in terms of intergroup processes [see @pinsof_strange_2023].
+In practice, however, the two explanations are difficult to disentangle in the context of real-world protest movements in which political identities and causes can seldom be separated.
+
+For now, we use "group-based" to refer to double standards based on ascribed categories such as class, race, and gender and "ideology-based" to refer to double standards based on ideological alignment.
+We compare the evidence for different explanations of those double standards in the General Discussion.
 
@@
 \caption{Item response curves for three hypothetical protest actions}
-\includegraphics[scale=1]{../Scale Development/figures/figure-1}
+\includegraphics[scale=1]{../Scale Development/figures/figure-1.png}
 \caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
@@
 In this research, item response theory serves two purposes.
-First, item response theory provides a statistical framework for scale development.
+First, it provides a statistical framework for scale development.
 In Studies 1 and 2, we select the most discriminating and informative actions to build an instrument for capturing double standards in judging collective action.
-Second, item response theory provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters are and what they are protesting.
-In Experiments 1--3, we operationalize double standards as the differences in $\theta_{(z)}$ between conditions which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
+Second, it provides a formal and statistical definition of double standards in judging collective action as respondents being more accepting of the same protest actions depending on who the protesters are and what they are protesting.
+In Experiments 1--3, we operationalize double standards as differences in $\theta_{(z)}$ which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
 
@@
 
-The first purpose of our research is to develop an instrument to measure where people draw the line between acceptable and unacceptable forms of collective action.
+Five studies (total $N = 2,979$) investigate double standards in judging collective actions---that is, whether observers judge the same controversial protest actions to be more acceptable depending on who the protesters are and what they are protesting.
+
+Our first purpose is to develop an instrument to measure where people draw the line between normative (acceptable) and non-normative (unacceptable) forms of collective action.
 In Study 1, we ask participants to generate more and less extreme protest actions and compile a pool of protest actions for scale development.
@@
 
-The second purpose of our research is to test for double standards in judging collective action---that is, whether people judge the same protest actions as more acceptable if the protesters' are ingroup members or the protesters' cause aligns with the ingroup's interests (*Hypothesis 1a*, *1b*) or if the protesters' cause aligns with their own system-justifying or system-challenging motives (*Hypotheses 2a*, *2b*).
-
-In Experiment 1, we use the instrument developed in Studies 1 and 2 to test for double standards in how acceptable participants consider collective action for workers' rights in the United Kingdom.
+Our second purpose is to investigate potential double standards in judging collective action.
+Three preregistered experiments used the newly developed instrument to compare how acceptable participants controversial protest actions in different conditions.
+
+By varying the participants' and the protesters' group memberships, we test for ingroup bias in terms of social class when judging protests for workers' rights (Experiment 1), in terms of race when judging protests for and against defunding the police (Experiment 2), and in terms of gender when judging protests for and against restricting abortion (Experiment 3).
+In this way, we test the hypothesis that people will judge the same protest actions as more acceptable if the protesters are ingroup members or the cause of the protest aligns with the ingroup's interests.
+
+By varying the protesters' causes, we test whether participants who identify as left-wing/liberal (Experiments 1--3) and reject system-justifying beliefs (Experiments 1--2) judge the same protest actions to be more acceptable when the protesters' cause is progressive and system challenging (for workers' rights, for defunding the police, against restricting abortion).
+Likewise, we test whether participants who identify as right-wing/conservative and endorse system-justifying beliefs judge the same protest actions to be more acceptable when the protesters' cause is conservative and system defending (against defunding the police, against restricting abortion).
+In this way, we test the hypothesis that people will judge the same protest actions as more acceptable if the cause of the protest aligns with their own ideological positions.
+
+<!-- In Experiment 1, we use the instrument developed in Studies 1 and 2 to test for double standards in how acceptable participants consider collective action for workers' rights in the United Kingdom.
 By varying both the participants' and the protesters' group memberships (working/middle class), Experiment 1 compares the relative evidence for identity-based (*Hypothesis 1a*, *1b*) and ideology-based double standards (*Hypothesis 2a*).
@@
 Together, the preregistered experiments provide a comprehensive test of the hypothesized double standards in judging collective action.
+-->
 
@@
 In two studies, we developed an instrument of 25 protest actions to capture double standards in judging collective action.
+
+In Study 1, we compiled protest actions from participants' responses and other sources.
+We recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US.
+To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students.
+Participants first read an accessible definition of what a social group is and that collective action is any action group members take to promote the interests of their social group.
+
+> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong.
+Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club, has a religion, or belongs to any number of other such groups.
+Individuals often act in ways to promote the interests of the social groups to which they belong.
+
+\noindent Participants were then asked to name between five and ten actions that fit that definition.
+Participants were encouraged to think of actions that varied in how acceptable they were in their opinion.
+We recoded responses into a smaller set of unique actions, which we supplemented with protest actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973].
+This process resulted in 72 actions that varied in how acceptable we would expect them to be.
+For details, see Supplemental Online Material (SOM).
+
+In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action.
+We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US.
+To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students.
+We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.
 
@@
 
-## Study 1
-
-In Study 1, we compiled protest actions from participants' responses and other sources.
-We recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US.
-To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students.
-
-Participants first read an accessible definition of what a social group is and that collective action is any action group members take to promote the interests of their social group.
-Participants were then asked to name between five and ten actions that fit that definition.
-Participants were encouraged to think of actions that varied in how acceptable they were in their opinion.
-We recoded responses into a smaller set of unique actions, which we supplemented with protest actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973].
-This process resulted in 72 actions that varied in how acceptable we would expect them to be.
-For details, see Supplemental Online Material (SOM).
-
-## Study 2
-
-In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action.
-We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US.
-To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students.
-We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.
-
 Participants again read an accessible definition of collective action.
@@
 
-We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997]---an item response theory model for ordinal response variables---for participants' ratings of how often an action would be an acceptable form of collective action.^[Ratings of how often a given action would be an acceptable form of collective action were strongly and negatively correlated with how disruptive, violent, and extreme participants considered the action to be and how negative they felt about it (Table S3).] Based on the results, we selected the 25 most discriminating, informative, and relevant protest actions to form an instrument for measuring double standards in judging collective action in Experiments 1 and 2.
+We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997]---an item response theory model for ordinal response variables---for participants' ratings of how often an action would be an acceptable form of collective action.^[Ratings of how often a given action would be an acceptable form of collective action were strongly and negatively correlated with how disruptive, violent, and extreme participants considered the action to be and how negative they felt about it (Table S3).] Based on the results, we selected the 25 most discriminating, informative, and relevant protest actions to form an instrument for measuring double standards in judging collective action in Experiments 1--3.
 We selected controversial, and thus diagnostic, actions (e.g., disrupting traffic) over actions that almost all respondents considered acceptable (e.g., signing petitions) or unacceptable (e.g., political violence; Figure \ref{fig:f2}).
@@
 In Experiment 1, we used the newly developed instrument to test for double standards in judging collective action for workers' rights in the United Kingdom.
-That is, we tested whether people with either working-class or professional jobs applied different standards when judging collective action taken by people with either working-class or middle-class jobs to protest against a fictitious government bill threatening their rights.
-
-By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that, in line with *Hypothesis 1a* and *1b*, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights than if they were outgroup members protesting for the outgroup's rights.
-By comparing reactions to working-class and middle-class protesters, we tested how observers judge collective action by lower-status (i.e.
-working-class) and higher-status (middle-class) group members.
-Assuming that collective action by lower-status group members is seen as more threatening to the unequal status quo than collective action by higher-status group members and that people are, to same extent, motivated to justify the system, we tested the preregistered hypothesis that, in line with *Hypothesis 2a*, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group than if they were from a lower-status group.
-In non-preregistered analyses, we further explored whether participants would judge the same protest actions as more or less acceptable depending on their own ideological positions.
-In this way, Experiment 1 provided a first test of the hypothesized identity- and ideology-based double standards in judging collective action.
+That is, we tested whether people with either working-class or professional jobs applied different standards when judging collective action taken by people with either working-class or middle-class jobs to protest against a fictitious government bill threatening their group's rights.
+
+By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights than if they were outgroup members protesting for the outgroup's rights.
+By comparing reactions to working-class and middle-class protesters, we further tested how observers judge collective action by lower-status (i.e.
+working-class) and higher-status (middle-class) group members.^[For Experiments 1 and 2, we had preregistered the additional hypotheses that all participants would  judge  the  same actions  as  more  acceptable when the protesters were from the higher-status group rather than the lower-status group (Experiment 1) and when the goal of the protest was to support rather than challenge the system (Experiment 2).
+These hypotheses reflected the assumption that the motivation to justify the system was, to some extent, universal.
+With hindsight, we no longer think that this assumption is tenable and, therefore, omit the hypotheses based on it from the revised manuscript.
+We still report the same preregistered analyses and their results.] In non-preregistered analyses, we tested whether participants would judge the same protest actions as more or less acceptable depending on their own ideological positions in terms of political orientation, system-justifying beliefs, and social dominance orientation.^[In Experiment 1, we explored a broader range of operationalizations of political ideology, including social dominance orientation, than in the other experiments.] In this way, Experiment 1 provided a first test of the hypothesized class- and ideology-based double standards in judging collective action.
 
@@
 Participants in this higher-status group had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socioeconomic status ladder.
+On average, participants tended to describe their political orientation as somewhat left-wing ($\textit{M} = 3.54$, $\textit{SD} = 1.31$) with $36\%$ and $20\%$ describing their political orientation as, respectively, at least somewhat left-wing and right-wing.
 
@@
 We used a screening survey to recruit participants who satisfied our preregistered inclusion criteria for the lower-status and higher-status groups.
-For the lower-status group, we recruited 475 participants who did not have a university degree and placed themselves on the bottom three ranks of a subjective socioeconomic status ladder.
-For the higher-status group, we recruited 400 participants who had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socioeconomic status ladder.
+For the lower-status group, we recruited participants who did not have a university degree and placed themselves on the bottom three ranks of a subjective socioeconomic status ladder.
+For the higher-status group, we recruited participants who had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socioeconomic status ladder.
 
@@
 
-We recruited 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group.
+We recruited 500 participants from the remaining participants, 250 from the lower-status group and 250 from the higher-status group.
 Participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (lower-status protesters) or people in professional jobs (higher-status protesters).
@@
 Overall, we found that participants' responses depended on both the protesters' and the participants' group status---but not in the directions predicted by our hypotheses.
-Contradicting our first hypothesis, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = 0.00, [-0.20, 0.20]$).
+Contradicting our hypothesis, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = 0.00, [-0.20, 0.20]$).
 Instead, we found that participants from higher-status backgrounds considered protest actions performed by both lower-status ($d = 0.34, [0.07, 0.59$]) and higher-status ($d = 0.33, [0.03, 0.63]$) protesters to be, on average, more acceptable than participants from lower-status backgrounds.
-Contradicting our second hypothesis, participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.40, -0.00]$).
+All participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.40, -0.00]$).
 
@@
 
-Figure \ref{fig:f3} (B) the $z$-standardized propensity to consider more controversial protest actions acceptable as a function of the three ideological orientation variables.
+Figure \ref{fig:f3} shows (B) the $z$-standardized propensity to consider more controversial protest actions acceptable as a function of the three ideological orientation variables.
 We found that participants who reported a more right-wing political orientation ($\beta_{xy} = -0.25, [-0.37, -0.15]$) and who expressed more agreement with system-justifying beliefs ($\beta_{xy} = -0.20, [-0.30, -0.10]$) tended to find fewer collective actions to be acceptable.
@@
 
-Experiment 1 provided a first test of the hypothesized identity- and ideology-based double standards in judging collective action.
-That is, we tested whether, in line with *Hypothesis 1a* and *1b*, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights or whether, in line with *Hypothesis 2a*, participants would judge the same protest actions as more acceptable if the protesters were from a higher-status group.
+Experiment 1 provided a first test of the hypothesized group-based double standards in judging collective action.
+That is, we tested whether, as hypothesized, participants would judge the same protest actions as more acceptable if the protesters were ingroup members protesting for the ingroup's rights.
 We found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the hypothesized directions.
-Instead, we found that both lower-status and higher-status participants tended to find the same protest actions more acceptable when lower-status protesters protested a bill threatening their rights.
-In non-preregistered analyses, we further found that people who were right-wing and endorsed system-justifying beliefs tended to find all protest actions to be less acceptable.
+Both lower-status and higher-status participants tended to find the same protest actions more acceptable when lower-status protesters protested a bill threatening their rights.
+In non-preregistered analyses, we found evidence for ideology-based double standards as people who were right-wing and endorsed system-justifying beliefs tended to find all protest actions to be less acceptable.
 
@@
 That said, participants' reactions to the experimental manipulation suggested that it evoked outrage and was understood in ingroup--outgroup terms.
-Second, by focusing on collective action to defend workers' rights, we arguably studied reactions to collective action for a progressive cause.
-To test for ideology-based double standards, we relied on the assumption that collective action by lower-status group members would be seen as more challenging to the prevailing system than collective action by higher-status group members.
-This assumption, however, was contradicted by our observation that participants who endorsed system-justifying beliefs tended to find actions by lower-status and higher-status protesters to be less acceptable.
+Second, by focusing on collective action to defend workers' rights, we studied reactions to collective action for a progressive cause.
+To test for ideology-based double standards, we needed to vary the protesters' cause to compare reactions to both progressive and conservative causes.
 Experiment 2 addressed those limitations by focusing on a scenario we expected to evoke stronger responses from participants and by examining reactions to both system-challenging and system-defending collective action.
@@
 
-By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the protesters were ingroup members (*Hypothesis 1a*).
-By assuming that defunding the police aligns more closely with the interests of Black than White Americans, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the cause of the protest aligned with their ingroup's interests (*Hypothesis 1b*).
-By varying whether the protesters protested for or against defunding the police, we tested the preregistered hypotheses that participants would judge the same protest actions as more acceptable if the protest supported, rather than challenged, the system (*Hypothesis 2a*) and if they endorsed system-justifying beliefs and the protest supported the system or if they rejected system-justifying beliefs and the protest challenged the system (*Hypothesis 2b*).
-In this way, Experiment 2 provided a complete test of the hypothesized identity- and ideology-based double standards in judging collective action.
-In addition, we tested the simpler, alternative hypothesis that participants, in general, would judge the same protest actions as more acceptable if they supported the protesters' cause (*Hypothesis 3*).
+By varying the participants' and the protesters' group memberships, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the protesters were ingroup members.
+By assuming that defunding the police aligns more closely with the interests of Black than White Americans, we tested the preregistered hypothesis that participants would judge the same protest actions as more acceptable if the cause of the protest aligned with their ingroup's interests.
+By varying whether the protesters protested for or against defunding the police, we tested the preregistered hypotheses that participants would judge the same protest actions as more acceptable if they endorsed system-justifying beliefs and the protest supported the system or if they rejected system-justifying beliefs and the protest challenged the system.
+In this way, Experiment 2 provided a complete test of the hypothesized race- and ideology-based double standards in judging collective action.
+In addition, we tested the simpler, alternative hypothesis that participants, in general, would judge the same protest actions as more acceptable if they supported the protesters' cause.
 
@@
 This resulted in a final sample of 1,600 participants ($\textit{Mdn} = 31$ years, age range: 18--84 years; 864 women, 708 men, 27 sex/gender diverse) of whom 800 identified as Black and 800 identified as White.
+On average, participants tended to describe their political orientation as moderately liberal ($\textit{M} = 2.92$, $\textit{SD} = 1.73$) with $63\%$ and $16\%$ describing their political orientation as, respectively, at least somewhat liberal and conservative.
 
@@
 
-Model 1 estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause.
-Contradicting *Hypothesis 1a*, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`).
-Contradicting *Hypothesis 1b*, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`).
-Contradicting *Hypothesis 2a*, participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`).
-Instead, we found that, on average, White participants considered all protest actions to be more acceptable than Black participants (`r with(m1, d_text[contrast == "nh1_contrast"])`) and both Black (`r with(m1, d_text[contrast == "nh3_contrast"])`) and White (`r with(m1, d_text[contrast == "nh2_contrast"])`) participants considered the same actions to be more acceptable when protests were for, rather than against, defunding the police.
+Model 1 estimated varying intercepts for the eight conditions to test whether, as hypothesized, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause.
+Contradicting the hypothesized race-based double standards, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`).
+Likewise, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`).
+<!--Participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`).
+-->Instead, we found that, on average, White participants considered all protest actions to be more acceptable than Black participants (`r with(m1, d_text[contrast == "nh1_contrast"])`) and both Black (`r with(m1, d_text[contrast == "nh3_contrast"])`) and White (`r with(m1, d_text[contrast == "nh2_contrast"])`) participants considered the same actions to be more acceptable when protests were for, rather than against, defunding the police.
 As in Experiment 1, we thus found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the directions predicted by our hypotheses.
@@
 As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgments about system-defending protest actions and one estimating their effect on judgments about system-challenging protest actions, and one varying effect estimating its variance across conditions.
-Supporting *Hypothesis 2b*, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`).
-We did not, however, find evidence for the ideological symmetry implied by *Hypothesis 2b* since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).
+Supporting the hypothesized ideology-based double standards, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`).
+We did not, however, find evidence for the hypothesized ideological symmetry since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).
 
@@
 Participants who endorsed system-justifying beliefs tended to consider system-challenging and system-defending protest actions to be equally unacceptable.
-In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with *Hypothesis 2b*, they considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.
+In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: They considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.
 
@@
 We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police.
-As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Contradicting *Hypothesis 3*, participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
+As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
 Ergo, our results did not support the alternative hypothesis that ideology-based double standards can be reduced to support for, or opposition to, the cause of a protest.
@@
 
-Experiment 2 provided a complete test of the hypothesized identity- and ideology-based double standards in judging collective action.
+Experiment 2 provided a complete test of the hypothesized race- and ideology-based double standards in judging collective action.
 As in Experiment 1, we found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the hypothesized directions.
-We found that, contrary to *Hypothesis 1a*, participants considered protest actions taken by ingroup and outgroup members to be equally acceptable and that, contrary to *Hypothesis 1b*, participants did not consider the same protest actions to be more acceptable when the protesters' cause aligned with their ingroup's interests.
-Like Experiment 1, Experiment 2 thus found no evidence for identity-based double standards---even though it focused on an issue of direct and current relevance to the participants.
-
-Expanding Experiment 1, Experiment 2 considered reactions to both system-challenging and system-defending collective action and, in so doing, provided a stronger test of the hypothesized ideology-based double standards.
-Contrary to *Hypothesis 2a* and the underlying assumption of a universal system-justifying motive, we found that participants considered the same protest actions to be *more* acceptable when judging system-challenging collective action (for defunding the police) than when judging system-defending collective action.
-This finding might reflect our liberal-leaning sample.
-
-In line with *Hypothesis 2b*, we found that participants who rejected system-justifying beliefs considered the same protest action more acceptable when judging system-challenging collective action (for defunding the police).
-We did not, however, find evidence for ideological symmetry in this relationship as participants who endorsed system-justifying beliefs considered system-challenging and system-defending collective action to be equally unacceptable.
-In addition, we did not find evidence for the alternative, simpler *Hypothesis 3* that that participants would judge the same protest actions to be acceptable when they supported the protesters' cause.
+We found that, contrary to the hypothesized race-based double standards, participants considered protest actions taken by ingroup and outgroup members to be equally acceptable and participants did not consider the same protest actions to be more acceptable when the protesters' cause aligned with their ingroup's interests.
+Like Experiment 1, Experiment 2 thus found no evidence for group-based double standards---even though it focused on an issue of direct and current relevance to the participants.
+
+Expanding Experiment 1, Experiment 2 considered reactions to both system-challenging and system-defending collective action and, in so doing, provided a stronger test of the hypothesized ideology-based double standards.<!--We found that, on average, participants considered the same protest actions to be *more* acceptable when judging system-challenging collective action (for defunding the police) than when judging system-defending collective action.
+This finding might reflect our liberal-leaning sample.--> As hypothesized, we found that participants who *rejected* system-justifying beliefs considered the same protest action more acceptable when judging system-challenging collective action (for defunding the police).
+We did not, however, find evidence for ideological symmetry in this relationship as participants who *endorsed* system-justifying beliefs considered system-challenging and system-defending collective action to be equally unacceptable.
+In addition, we did not find evidence for the alternative, simpler hypothesis that that participants would judge the same protest actions to be acceptable when they supported the protesters' cause.
 
 One reason why we did not find evidence for ideological symmetry might be that participants judged protests for or against a progressive cause and that *opposing* a progressive cause is not equivalent to *supporting* a conservative cause.
-Another reason might be that our sample skewed liberal and lacked committed conservatives who might accept more extreme means to oppose a progressive cause.
+Another reason might be that, as in Experiment 1, our sample skewed liberal and lacked committed conservatives who might accept more extreme means to oppose a progressive cause.
 Experiment 3 addressed those limitations by testing for ideology-based double standards in judging collective action for or against a conservative cause in a balanced sample of conservatives and progressives.
@@
 We tested the alternative hypothesis that, in line with *asymmetrical* ideology-based double standards, conservatives would judge all protest actions to be equally unacceptable while liberals would judge the same protest actions to be more acceptable if the protesters opposed restricting abortion.
-By comparing male and female participants, we further tested the preregistered hypothesis that, in line with identity-based double standards, women would judge protest actions against restricting abortion to be more acceptable, on average, than men or protest actions for restricting abortion.
+By comparing male and female participants, we further tested the preregistered hypothesis that, in line with gender-based double standards, women would judge protest actions against restricting abortion to be more acceptable, on average, than men or protest actions for restricting abortion.
+
+To achieve an ideologically balanced sample, we moved from an operationalization of political ideology in terms of system-justifying beliefs to an operationalization in terms of ideological self-placement.
+We did so for three reasons.
+First, there was the practical reason that Prolific and similar platforms allow prescreening participants in terms of ideological self-placement but not system-justifying beliefs.
+Second, there was the *a priori* empirical reason that, when comparing the evidence for ideology-based double standards across the two operationalizations of political ideology, the patterns of results were nearly identical in terms of direction and magnitude for system-justifying beliefs and ideological self-placement (see, for example, Figure \ref{fig:f5}).
+Third, there was the *a posteriori* empirical reason that, as reported in the Results section, dividing the sample into self-identified conservatives and liberals achieved a nearly complete separation of the sample in terms of system-justifying beliefs.
+For those reasons, selecting participants based on ideological self-placement was a feasible and effective quasi-experimental strategy to achieve an ideologically balanced sample.
 
@@
 
-As intended, participants who identified as conservative reported being far more conservative ($M = 6.05, SD = 0.79$) than participants who identified as liberal ($M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43$).
-Likewise, conservative participants tended to endorse system-justifying beliefs ($M = 5.10, SD = 0.95$) while liberal participants tended to reject system-justifying beliefs ($M = 2.86, SD = 1.03; d = 2.26$).
+As expected, conservative participants tended to endorse system-justifying beliefs ($M = 5.10, SD = 0.95$) while liberal participants tended to reject system-justifying beliefs ($M = 2.86, SD = 1.03; d = 2.26$).
+Indeed, in $93.5\%$ of all possible pairings of conservative and liberal participants, the conservative participant reported stronger system-justifying beliefs.
+This shows that ideological self-placement was an effective quasi-experimental approach to varying system justification.
+
 As expected, conservative and liberal participants differed in their reactions to the manipulation as the former tended to support restricting abortion ($M = 3.77, SD = 1.48$) while the latter tended to oppose restricting abortion ($M = 1.28, SD = 0.85; d = 2.06$).
@@
 
+  # Gender differences
+  m1_gender <- read_rds("../Experiment 3/results/m1_post.rds") |> 
+    group_by(.draw, participant_gender) |> 
+    summarize(z = mean(z)) |> 
+    summarize(d = z[1] - z[2]) |> 
+    median_qi(d) |> 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
 ```
@@
 
-We found some evidence for identity-based double standards: Female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
+We found some evidence for gender-based double standards: Female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
 This pattern was, however, overshadowed by stronger ideology-based double standards that were consistent across subsamples: Conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and both liberal women (`r with(m1, d_text[contrast == "nh6_contrast"])`) and liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
-In this way, Experiment 3 provided more consistent evidence for ideology-based than for identity-based double standards in judging collective action.
+In this way, Experiment 3 provided more consistent evidence for ideology-based than for gender-based double standards in judging collective action.
+Across conditions, men tended to consider more protest actions acceptable than women (`r m1_gender$d_text`).
 
@@
 We found, however, that gender identification neither affected how women judged protesters opposing (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "For"])`) abortion restrictions nor how men judged protesters opposing (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "For"])`) abortion restrictions.
-Our non-preregistered analyses thus suggested that gender identification did not moderate identity-based double standards in judging collective action concerning reproductive rights.
+Our non-preregistered analyses thus suggested that gender identification did not moderate gender-based double standards in judging collective action concerning reproductive rights.
 
@@
 Experiment 3 tested for double standards in judging collective action for and against restricting abortion in a sample balanced by gender and political orientation.
-As in Experiments 1--2, participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their ideological orientation.
+As hypothesized, participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their ideological orientation.
 While participants in Experiment 2 judged protests opposing or supporting a progressive cause (defunding police), participants in Experiment 3 considered protests opposing or supporting a conservative cause (restricting abortion).
@@
 
-In contrast to Experiments 1--2, Experiment 3 provided some evidence for identity-based double standards, although any gender differences in judging protests for or against restricting abortion were overshadowed by more consistent ideology-based double standards.
+In contrast to Experiments 1--2, Experiment 3 provided some evidence for group-based double standards distinct from ideology-based double standards, although any gender differences in judging protests for or against restricting abortion were overshadowed by more consistent ideology-based double standards.
 Our finding that conservative women considered collective action in support of restricting abortion to be more acceptable while liberal women considered collective action in opposition to restricting abortion to be more acceptable aligned with Mikolajczak et al.'s [-@mikolajczak_women_2022] argument that, by itself, gender identity is too broad to explain support for collective action.
@@
 Our findings, however, showed that it is not so simple.
-Contradicting our hypothesis of identity-based double standards, participants in three preregistered experiments did not show consistent ingroup bias---in terms of class, race, and gender---when judging controversial protest actions taken by ingroup and outgroup members to advance their group's interests.
-Instead, supporting our hypothesis of ideology-based double standards, participants judged the same controversial protest actions to be more acceptable when the protesters' cause---protecting workers' rights, defunding the police, restricting abortion---aligned with their own ideological position.
-<!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.
---> In the remainder of this General Discussion, we consider limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.
+
+Contradicting our hypothesis of group-based double standards, participants in three preregistered experiments did not show consistent ingroup bias in terms of social class when judging protests for workers' rights (Experiment 1), in terms of race when judging protests for and against defunding the police (Experiment 2), and in terms of gender when judging protests for and against restricting abortion (Experiment 3).
+Instead, we found that members of advantaged groups---middle-class (<!--$d = 0.33$, -->Experiment 1), White (<!--$d = 0.09$, -->Experiment 2), and male (<!--$d = 0.16$,-->Experiment 3) participants---tended to consider all protest actions more acceptable than members of relatively disadvantaged groups.
+That is, our findings showed that participants' judgements of protest actions depended on their group memberships---but not in the directions implied by the hypothesized group-based double standards.
+
+<!-- All participants tended to consider protest actions for a progressive cause---for the rights of working-class people ($d = 0.20$, Experiment 1), for defunding the police ($d = 0.29$, Experiment 2), and against restricting abortion ($d = 0.33$, Experiment 3) participants---more acceptable than protest actions for a conservative cause.
+-->
+
+Supporting our hypothesis of ideology-based double standards, progressive participants who rejected system-justifying beliefs judged the same controversial protest actions to be more acceptable when the protesters' cause---for workers' rights (<!--$\beta_{xy} =0.20$, -->Experiment 1), for defunding the police (<!--$\beta_{xy} = 0.38$, -->Experiment 2), against restricting abortion (<!--$d = 0.90$, -->Experiment 3)---aligned with their own ideological position.<!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.--> While conservative participants who endorsed system-justifying beliefs considered the same actions somewhat more acceptable when protesters supported, rather than opposed, restricting abortion (Experiment 3), they were more consistent than liberal participants and considered all protest actions, for and against defunding the police (Experiment 2), less acceptable than progressive participants.
+In this way, our findings provided evidence for asymmetrical ideology-based double standards.
+
+In the remainder of this General Discussion, we discuss limiting conditions on our findings as well as theoretical and practical implications for understanding the often divided response to collective action.
 
@@
 
-Despite its strengths, our research is qualified by several limitations.
+Despite its strengths, our research has several limitations that constrain the generalizability of our findings.
 First, our research examined the hypothesized relationships for three political causes---protecting workers' rights, defunding the police, restricting abortion---and thus provided only limited evidence that our findings generalize beyond those causes.
@@
 
+### Theoretical Implications
+
 Growing evidence shows that, in many situations, observers apply a more lenient standard when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
-And yet, our findings suggested that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on group identities such as class, race, or gender.
+And yet, our findings suggested that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on ascribed group memberships such as class, race, or gender.
 We did not find evidence for those double standards even though we examined social categories that are central to most people's identities, even when we considered only highly identified group members, and even though participants' reactions to the manipulation showed that, at least in Experiments 1--2, they understood the protesters' causes in ingroup--outgroup terms.
@@
 
-First, we might understand ideology-based double standards as an expression of ingroup bias based on opinion-based [@mcgarty_collective_2009] or partisan [@osborne_psychology_2022; @finkel_political_2020] identities.
-This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions when taken by their most-liked, rather than their least-liked, political group (e.g., climate activists).
+First, we might understand ideology-based double standards as an expression of ingroup bias based on partisan identities [@osborne_psychology_2022; @finkel_political_2020].
+This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions when taken by their most-liked, rather than their least-liked, political group.
 An identity-based account, however, implies that both conservatives and progressives would show ingroup bias and, therefore, cannot explain the observed asymmetry in the two groups' judgements.
@@
 
-Most broadly, our findings showed that, for many controversial protest actions, there is no consensus on whether they are acceptable means to advance a cause.
-<!--Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.
--->This confirms our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder.
+### Practical Implications
+
+Most broadly, our findings showed that, for many protest actions (e.g., blocking roads), there is no consensus on whether they are acceptable means to advance a cause.<!-- Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.--> This confirms our contention that the distinction between normative and non-normative collective action should be situated in the eye of the beholder.
 By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provide a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
@@
 After all, this distinction is consequential.
+Research on how people respond to collective action showed that disruptive yet non-violent collective action might be most effective at gaining concessions from those resistant to social change [@shuman_disrupting_2021; @shuman_when_2023].
+If, as our research suggests, people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
+This could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice.
+In this way, our research complements, and adds to, the growing program of research on how people respond to different kinds of collective action.
+
 If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
-In this way, our findings could explain why 45% of Republicans supported the storming of the U.S.
+For example, 45% of Republicans supported the storming of the U.S.
 Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
-If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
-In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022].<!-- For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.-->
+Recent research, most of it by political scientists, investigated support for partisan violence [@kalmoe_radical_2022], albeit with mixed results [@westwood_current_2022].
+While our research focused on controversial but non-violent actions, we encourage researchers to adapt our paradigm to investigate when and why people apply different standards when judging violent and non-violent collective action for different causes.
 
@@
 Our research showed that, in contrast to well-documented double standards in other domains, people do not apply a different standard when judging collective action by members of the same class, race or gender or for a cause aligned with their own group's interests.
-Instead, our research demonstrated that people consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological position<!-- ---in other words, that there are partisan double standards in judging collective action -->.
+Instead, our research demonstrated that progressives and, less so, conservatives consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological position<!-- ---in other words, that there are partisan double standards in judging collective action -->.
 

### 2024-04-16 Documents/manuscript.tex
resubmit

@@
 but often sparks contention about what actions are acceptable means to
-effect social change. In five studies (total \emph{N} = 2,979), we
-investigated double standards in judging collective action---that is,
-whether observers will judge the same protest actions as more acceptable
-if the protesters are ingroup members and their cause aligns with the
-ingroup's interests (\emph{identity-based double standards}) or if the
-protesters' cause aligns with the observers' ideological positions
-(\emph{ideology-based double standards}). In two studies, we developed
-an instrument of 25 controversial protest actions, based in item
-response theory, to measure where people draw the line between
-acceptable and unacceptable forms of collective action. In three
-preregistered experiments, we used this instrument to test for double
-standards in judging collective action for workers' rights in the United
-Kingdom and for or against defunding the police and for or against
-restricting abortion in the United States. We found evidence for
-ideology-based, but not for identity-based, double standards:
-Participants judged the same protest actions to be more acceptable when
-the protesters' cause aligned with their own ideological position---but
-showed no consistent ingroup bias in terms of class, race, and gender.
-Our findings have theoretical and practical implications for
-understanding the often divided response to prominent social movements.
+effect social change. Five studies (total \(N = 2,979\)) investigated
+double standards in judging collective action---that is, whether
+observers judge the same protest actions to be more acceptable depending
+on who the protesters are and what they are protesting. Two studies used
+item response theory to develop an instrument of 25 controversial
+protest actions to measure where people draw the line between acceptable
+and unacceptable forms of collective action. Three preregistered
+experiments found no consistent evidence for ingroup bias in terms of
+social class when judging protests for workers' rights (Experiment 1),
+in terms of race when judging protests for and against defunding the
+police (Experiment 2), and in terms of gender when judging protests for
+and against restricting abortion (Experiment 3). Instead, we found that
+participants who identified as left-wing/liberal (Experiments 1--3) and
+rejected system-justifying beliefs (Experiments 1--2) considered the
+same protest actions more acceptable when a cause aligned with their
+ideological orientation (for workers' rights, for defunding the police,
+against restricting abortion) than when it did not (against defunding
+the police, for restricting abortion). Conservative participants
+considered the same actions somewhat more acceptable when protesters
+supported, rather than opposed, restricting abortion (Experiment 3) but
+considered all protest actions, for and against defunding the police,
+equally unacceptable (Experiment 2). Our findings have theoretical and
+practical implications for understanding the often-divided response to
+social movements.
 
 \textit{Keywords:} collective action, double standards, social identity
-theory, system justification theory, Black Lives Matter
+theory, system justification theory, partisanship, Black Lives Matter,
+abortion
 
@@
 \begin{center}
-\noindent 11,546 words
+\noindent 11,654 words
 \end{center}
@@
 \emph{means} of protest depends on what the \emph{end} of a protest is.
-That is, people tend to judge the \emph{same} controversial protest
-actions to be more acceptable when the protesters' cause aligned with
-their own ideological positions. This finding helps explain the often
-divided response to recent protest movements.
+That is, observers sometimes judge the \emph{same} controversial protest
+action (e.g., blocking a road) to be more acceptable when the
+protesters' cause (e.g., defunding the police) aligns with their own
+ideological position. Conservatives, who support social order and oppose
+social change, tend to consider all protest actions less acceptable than
+progressives. Progressives, in turn, tend to consider disruptive but
+non-violent actions acceptable means to achieve social change and
+challenge social injustice. This finding helps explain the often-divided
+response to recent protest movements.
 
@@
 Democrats) considered it appropriate for Black athletes to protest in
-this way (YouGov, 2017). Indeed, Black athletes who had participated in
-the protests faced a backlash in public opinion (Lacina, 2020). This and
-other examples suggest that \emph{who} the protesters are and
-\emph{what} they are protesting influences how acceptable observers
-judge their actions to be.
+this way (YouGov, 2017). This and other examples suggest that \emph{who}
+the protesters are and \emph{what} they are protesting influences how
+acceptable their actions are judged to be.
 
 In this paper, we examine double standards in judging collective
-action---that is, whether observers judge the same protest actions as
-more or less acceptable depending on their own and the protesters' group
-memberships (\emph{identity-based double standards}) or on the
-protesters' cause and how it aligns with their own ideological positions
-(\emph{ideology-based double standards}). In two studies, we first
-develop an instrument of 25 controversial protest actions, based in item
-response theory, to capture double standards in judging collective
-action. In three preregistered experiments, we then use this instrument
-to test for double standards in judging protest actions for workers'
-rights, for or against defunding the police, and for or against
-restricting abortion. We find evidence for ideology-based, but not for
-identity-based, double standards: People judge the same protest actions
-to be more acceptable when the protesters' cause aligns with their own
-ideological positions---but do not show consistent ingroup bias (in
-terms of class, race, and gender) when judging collective action by
-ingroup and outgroup members.
+action---that is, whether observers judge the \emph{same} protest
+actions as more or less acceptable depending on their own and the
+protesters' group memberships \emph{or} on the protesters' cause and how
+it aligns with their own ideological positions.
 
@@
 Recent research has shown that this distinction matters for how people
-respond to collective action. Shuman et al. (2021) showed that
-non-normative, non-violent collective action is more effective than
-either normative or violent collective action at gaining concessions
-from those most resistant to social change. Feinberg et al. (2020)
-demonstrated, however, that observers are less supportive of social
-movements that use extreme protest actions.\footnote{In contrast, White
-  Americans exposed to pictures of destructive Black Lives Matter
-  protests were not, on average, any less supportive of systemic change
-  than those exposed to pictures of peaceful protests (Teixeira et al.,
-  2022).} Similarly, Teixeira et al. (2020) found that advantaged-group
-members perceive non-normative collective action by disadvantaged-group
-members as more damaging to their ingroup's social image and are thus
-less supportive of such action. Verkuyten et al. (2022) found that, in
-general, the public tends to be intolerant of non-normative collective
+respond to collective action (for a review, see Shuman et al., 2023).
+Shuman et al. (2021) showed that non-normative, non-violent collective
+action is more effective than either normative or violent collective
+action at gaining concessions from those most resistant to social
+change. Feinberg et al. (2020) demonstrated, however, that observers are
+less supportive of social movements that use extreme protest
+actions.\footnote{In contrast, White Americans exposed to pictures of
+  destructive Black Lives Matter protests were not, on average, any less
+  supportive of systemic change than those exposed to pictures of
+  peaceful protests (Teixeira et al., 2022).} Similarly, Teixeira et al.
+(2020) found that advantaged-group members perceive non-normative
+collective action by disadvantaged-group members as more damaging to
+their ingroup's social image and are thus less supportive of such
 action. This means that activists face the dilemma that non-normative
@@
 
-These and other studies have, for the most part, relied on ad-hoc
-distinctions between what researchers themselves considered normative
-and non-normative collective action. In liberal democracies, researchers
+Past research has, for the most part, relied on ad-hoc distinctions
+between what researchers themselves considered normative and
+non-normative collective action. In liberal democracies, researchers
 tend to consider actions such as signing petitions, voting in elections,
@@
 action might be considered acceptable by some and unacceptable by
-others. That is, the distinction between normative and non-normative
-collective action might be in the eye of the beholder.
-
-In this research, we move this distinction from the researcher's
-intuition into the realm of scientific investigation and test for double
-standards in where people draw the line between acceptable (normative)
-and unacceptable (non-normative) forms of collective action. For that
-purpose, we define a double standard as judging the same protest actions
-as more or less acceptable depending on who the protesters are and what
-they are protesting (for an analogous definition, see Foschi,
-2000).\footnote{Feinberg et al. (2020) provided incidental evidence for
-  double standards in judging collective action as, in a manipulation
-  check, Black participants rated Black Lives Matter protests as less
-  extreme than White participants and liberal participants rated protest
-  actions for progressive causes as less extreme than conservative
-  participants.}
-
-\hypertarget{identity-based-double-standards}{%
-\subsubsection{Identity-Based Double
-Standards}\label{identity-based-double-standards}}
+others. In this research, we test for double standards in where people
+draw the line between what is an acceptable (normative) and an
+unacceptable (non-normative) form of collective action.\footnote{In this
+  research, we use \emph{acceptable} (``agreed or approved of by most
+  people in a society,'' Oxford Advanced Learner's Dictionary, n.d.) as
+  a plain-language synonym for \emph{normative} (``conforming to the
+  norms of the existing social system'').} For that purpose, we define a
+double standard as judging the same protest actions as more or less
+acceptable depending on who the protesters are and what they are
+protesting (for a similar definition, see Foschi, 2000).\footnote{Feinberg
+  et al. (2020) provided incidental evidence for double standards in
+  judging collective action as, in a manipulation check, Black
+  participants rated Black Lives Matter protests as less extreme than
+  White participants and liberal participants rated protest actions for
+  progressive causes as less extreme than conservative participants.}
+
+\hypertarget{group-based-double-standards}{%
+\subsubsection{Group-Based Double
+Standards}\label{group-based-double-standards}}
+
+Far more Black (\(66\%\)) than White (\(29\%\)) Americans considered it
+appropriate for Black athletes to kneel during the national anthem to
+protest racist police violence (YouGov, 2017). This suggests that what
+protest actions people consider acceptable might depend on their own and
+the protesters' group memberships.
 
@@
 make ingroup-serving causal attributions when judging actions by ingroup
-and outgroup members, for example, by attributing negative behavior by
-outgroup members to internal causes but attributing negative behavior by
-ingroup members to external causes (see Pettigrew, 1979). Valdesolo \&
-DeSteno (2007) demonstrated that participants judged the same selfish
-action as less unfair when performed by themselves or an ingroup member
-than when performed by an outgroup member. Other studies (Abrams et al.,
-2013; Endevelt et al., 2021; for exceptions, see Mendoza et al., 2014;
-Pinto et al., 2010) provided further evidence for ingroup bias in
-judging moral transgressions by ingroup and outgroup members.
-
-In the same vein, we propose that ingroup bias results in identity-based
-double standards in judging collective action by ingroup and outgroup
-members. That is, we hypothesize that people will judge the same protest
-actions as more acceptable if the protesters are ingroup members
-(\emph{Hypothesis 1a}) or if the cause of the protest aligns with the
-ingroup's interests (\emph{Hypothesis 1b}). This hypothesis could
-explain why, for example, far more Black than White Americans considered
-it appropriate for Black athletes to kneel down to protest police
-violence against Black Americans (YouGov, 2017). Notably, this
+and outgroup members. For example, people tend to attribute negative
+behavior by outgroup members to internal causes but attribute negative
+behavior by ingroup members to external causes (see Pettigrew, 1979).
+Valdesolo \& DeSteno (2007) demonstrated that participants judged the
+same selfish action as less unfair when performed by themselves or an
+ingroup member than when performed by an outgroup member. Other studies
+(Abrams et al., 2013; Endevelt et al., 2021; for exceptions, see Mendoza
+et al., 2014; Pinto et al., 2010) provided further evidence for ingroup
+bias in judging moral transgressions by ingroup and outgroup members.
+
+In the same vein, we propose that ingroup bias results in double
+standards in judging collective action by ingroup and outgroup members.
+That is, we hypothesize that people will judge the same protest actions
+as more acceptable if the protesters are ingroup members and if the
+cause of the protest aligns with the ingroup's interests. Notably, this
 hypothesis applies equally to historically disadvantaged groups
@@
 
+Far more Democrats (\(59\%\)) than Republicans (\(11\%\)) considered it
+appropriate to kneel during the national anthem to protest racist police
+violence (YouGov, 2017). This suggests that what protest actions people
+consider acceptable might depend on the protesters' cause and how it
+aligns with their own ideological positions.
+
+One explanation for ideology-based double standards is that they result
+from the same intergroup processes as other group-based double
+standards. In this view, political partisanship functions as a social
+identity (Finkel et al., 2020; Harris et al., 2022) and results from
+universal cognitive mechanisms for forming, detecting, and maintaining
+alliances (Pinsof et al., 2023). Like other identities, political
+identities result in group-serving biases in how we evaluate ingroup and
+outgroup members. This explanation aligns with Verkuyten et al.'s (2022)
+finding that participants were more tolerant of transgressive protest
+actions (e.g., sending hate mail) when taken by their most-liked, rather
+than their least-liked, political group (e.g., climate activists).
+
+Another explanation for ideology-based double standards is that they are
+rooted in ideological differences in how conservatives and progressives
+relate to social change and social inequality which, in turn, are
+motivated by epistemic, existential, and relational needs (Jost et al.,
+2003; Jost, 2021). As motivated social cognition, ideological
+differences shape how we think and feel about political issues, actions,
+and events. In this view, ideology-based double standards depend on
+whether the cause of a protest aligns with someone's ideological
+positions.
+
 System Justification Theory (Jost et al., 2004; for a recent review, see
-Jost, 2020) argues that people are motivated to defend, justify, and
-bolster the prevailing social, economic, and political system because
-doing so serves basic epistemic, existential, and relational needs. Both
-advantaged and disadvantaged groups are thought to be motivated to
-justify the existing system, although the strength of this motivation
-and its expression vary across individuals and situations.
+Jost, 2020), in particular, argues that people are motivated to defend,
+justify, and bolster the prevailing social, economic, and political
+system because doing so serves basic epistemic, existential, and
+relational needs. Both advantaged and disadvantaged groups are thought
+to be motivated to justify the existing system, although the strength of
+this motivation and its expression vary across individuals and
+situations. Conservatives, far more so than progressives, endorse
+system-justifying beliefs.
 
@@
 standards in what actions people consider acceptable forms of collective
-action. First, we hypothesize that people will judge the same protest
-actions as more acceptable if the protest supports, rather than
-challenges, the system (\emph{Hypothesis 2a}). This hypothesis reflects
-the assumption that the motivation to justify the system is, to some
-extent, universal. Second, we hypothesize that people will judge the
-same protest actions as more acceptable if they endorse
-system-justifying beliefs and the protest supports the system or if they
-reject system-justifying beliefs and the protest challenges the system
-(\emph{Hypothesis 2b}). Notably, this hypothesis implies a symmetry in
-how people with different ideological orientations judge
-system-supporting and system-challenging collective action.
-
-As system-justifying beliefs are closely related to political
-conservatism (Jost et al., 2003), this hypothesis could explain partisan
-differences in what actions liberals and conservatives consider
-acceptable means of protest for progressive and conservative
-causes---for example, why far more Democrats than Republicans considered
-it appropriate to kneel during the national anthem to protest racist
-police violence (YouGov, 2017).
-
-\hypertarget{are-partisan-double-standards-based-in-ideology-or-identity}{%
-\subsubsection{Are Partisan Double Standards Based in Ideology or
-Identity?}\label{are-partisan-double-standards-based-in-ideology-or-identity}}
-
-Partisan differences can result from two social psychological processes.
-On the one hand, political partisanship reflects ideological differences
-in resistance to social change and acceptance of inequality which, in
-turn, are motivated by epistemic, existential, and relational needs
-(Jost et al., 2003; Jost, 2021). As motivated cognition, ideological
-differences shape how we think and feel about issues, actions, and
-events. On the other hand, political partisanship functions as a social
-identity (Finkel et al., 2020; Harris et al., 2022) with evidence
-suggesting that partisans' policy preferences depend more on the stated
-position of their party than policy content (Cohen, 2003). Like other
-identities, political identities result in group-serving biases in how
-we evaluate ingroup and outgroup members. In practice, the two processes
-are often difficult to disentangle.
-
-For now, we use ``identity-based'' to refer to double standards based on
-assigned social categories such as class, race, and gender and
+action. That is, we hypothesize that people will judge the same protest
+actions as more acceptable if they endorse system-justifying beliefs and
+the protest supports the system or if they reject system-justifying
+beliefs and the protest challenges the system.
+
+In principle, the two explanations for ideology-based double
+standards---in terms of intergroup processes and ideological
+differences---could be disentangled by separating political identities
+and causes. An explanation in terms of intergroup processes posits that
+political identities, not causes, shape partisan behavior. Indeed,
+research suggests that partisans' policy preferences depend more on the
+stated position of their party than policy content (Cohen, 2003). An
+explanation in terms of ideological differences instead posits that
+there is an ``elective affinity'' between political causes and the
+psychological needs and motives of their supporters (Jost, 2021). In
+other words, the content of political ideologies is not arbitrary or a
+historical accident, as implied by an explanation in terms of intergroup
+processes (see Pinsof et al., 2023). In practice, however, the two
+explanations are difficult to disentangle in the context of real-world
+protest movements in which political identities and causes can seldom be
+separated.
+
+For now, we use ``group-based'' to refer to double standards based on
+ascribed categories such as class, race, and gender and
 ``ideology-based'' to refer to double standards based on ideological
-alignment. We compare the evidence for identity and ideology processes
-in the General Discussion.
+alignment. We compare the evidence for different explanations of those
+double standards in the General Discussion.
 
@@
 
-In this research, item response theory serves two purposes. First, item
-response theory provides a statistical framework for scale development.
-In Studies 1 and 2, we select the most discriminating and informative
-actions to build an instrument for capturing double standards in judging
-collective action. Second, item response theory provides a formal and
-statistical definition of double standards in judging collective action
-as respondents being more accepting of the same protest actions
-depending on who the protesters are and what they are protesting. In
-Experiments 1--3, we operationalize double standards as the differences
-in \(\theta_{(z)}\) between conditions which, as \(\theta_{(z)}\) is
-\(z\)-standardized, correspond to Cohen's \(d\) effect sizes.
+In this research, item response theory serves two purposes. First, it
+provides a statistical framework for scale development. In Studies 1 and
+2, we select the most discriminating and informative actions to build an
+instrument for capturing double standards in judging collective action.
+Second, it provides a formal and statistical definition of double
+standards in judging collective action as respondents being more
+accepting of the same protest actions depending on who the protesters
+are and what they are protesting. In Experiments 1--3, we operationalize
+double standards as differences in \(\theta_{(z)}\) which, as
+\(\theta_{(z)}\) is \(z\)-standardized, correspond to Cohen's \(d\)
+effect sizes.
 
@@
 
-The first purpose of our research is to develop an instrument to measure
-where people draw the line between acceptable and unacceptable forms of
-collective action. In Study 1, we ask participants to generate more and
-less extreme protest actions and compile a pool of protest actions for
-scale development. In Study 2, we ask participants to rate how
-acceptable they consider each of those protest actions to be and use
-item response theory to select the most discriminating and informative
-protest actions for our instrument. In so doing, we build an instrument
-that enables both present and future research to capture double
-standards in judging collective action.
-
-The second purpose of our research is to test for double standards in
-judging collective action---that is, whether people judge the same
-protest actions as more acceptable if the protesters' are ingroup
-members or the protesters' cause aligns with the ingroup's interests
-(\emph{Hypothesis 1a}, \emph{1b}) or if the protesters' cause aligns
-with their own system-justifying or system-challenging motives
-(\emph{Hypotheses 2a}, \emph{2b}).
-
-In Experiment 1, we use the instrument developed in Studies 1 and 2 to
-test for double standards in how acceptable participants consider
-collective action for workers' rights in the United Kingdom. By varying
-both the participants' and the protesters' group memberships
-(working/middle class), Experiment 1 compares the relative evidence for
-identity-based (\emph{Hypothesis 1a}, \emph{1b}) and ideology-based
-double standards (\emph{Hypothesis 2a}). Extending Experiment 1,
-Experiment 2 considers both system-challenging and system-defending
-collective action in the context of the 2020 Black Lives Matter protests
-in the United States. By varying the protesters' cause (for/against
-defunding the police) as well as the participants' and the protesters'
-group memberships (Black/White), Experiment 2 compares the relative
-evidence for identity-based (\emph{Hypothesis 1a}, \emph{1b}) and
-ideology-based double standards (\emph{Hypothesis 2a}, \emph{2b}).
-Building on Experiments 1--2, Experiment 3 examines how liberal and
-conservative men and women judge collective action for or against
-restricting abortion in anticipation of the Supreme Court overturning
-its \emph{Roe v. Wade} decision on reproductive rights in the United
-States. By varying the protesters' cause as well as the participants'
-ideology and gender, Experiment 3 compares the relative evidence for
-identity-based (\emph{Hypothesis 1b}) and ideology-based
-(\emph{Hypothesis 2b}) double standards in judging collective action
-concerning a conservative cause in a sample balanced by gender and
-ideology. Together, the preregistered experiments provide a
-comprehensive test of the hypothesized double standards in judging
-collective action.
+Five studies (total \(N = 2,979\)) investigate double standards in
+judging collective actions---that is, whether observers judge the same
+controversial protest actions to be more acceptable depending on who the
+protesters are and what they are protesting.
+
+Our first purpose is to develop an instrument to measure where people
+draw the line between normative (acceptable) and non-normative
+(unacceptable) forms of collective action. In Study 1, we ask
+participants to generate more and less extreme protest actions and
+compile a pool of protest actions for scale development. In Study 2, we
+ask participants to rate how acceptable they consider each of those
+protest actions to be and use item response theory to select the most
+discriminating and informative protest actions for our instrument. In so
+doing, we build an instrument that enables both present and future
+research to capture double standards in judging collective action.
+
+Our second purpose is to investigate potential double standards in
+judging collective action. Three preregistered experiments used the
+newly developed instrument to compare how acceptable participants
+controversial protest actions in different conditions.
+
+By varying the participants' and the protesters' group memberships, we
+test for ingroup bias in terms of social class when judging protests for
+workers' rights (Experiment 1), in terms of race when judging protests
+for and against defunding the police (Experiment 2), and in terms of
+gender when judging protests for and against restricting abortion
+(Experiment 3). In this way, we test the hypothesis that people will
+judge the same protest actions as more acceptable if the protesters are
+ingroup members or the cause of the protest aligns with the ingroup's
+interests.
+
+By varying the protesters' causes, we test whether participants who
+identify as left-wing/liberal (Experiments 1--3) and reject
+system-justifying beliefs (Experiments 1--2) judge the same protest
+actions to be more acceptable when the protesters' cause is progressive
+and system challenging (for workers' rights, for defunding the police,
+against restricting abortion). Likewise, we test whether participants
+who identify as right-wing/conservative and endorse system-justifying
+beliefs judge the same protest actions to be more acceptable when the
+protesters' cause is conservative and system defending (against
+defunding the police, against restricting abortion). In this way, we
+test the hypothesis that people will judge the same protest actions as
+more acceptable if the cause of the protest aligns with their own
+ideological positions.
 
@@
 capture double standards in judging collective action.
-
-\begin{figure*}[!t]
-\centering
-\caption{Examples of protest actions rated in Study 2}
-\includegraphics[scale=1]{../Scale Development/figures/figure-2}
-\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause.
-Darker bars mark actions that were included in the final scale.}
-\label{fig:f2}
-\end{figure*}
-
-\hypertarget{study-1}{%
-\subsection{Study 1}\label{study-1}}
 
@@
 without a university degree, 15 non-students with a university degree,
-and 15 current university students.
-
-Participants first read an accessible definition of what a social group
-is and that collective action is any action group members take to
-promote the interests of their social group. Participants were then
-asked to name between five and ten actions that fit that definition.
-Participants were encouraged to think of actions that varied in how
-acceptable they were in their opinion. We recoded responses into a
-smaller set of unique actions, which we supplemented with protest
-actions from the psychological and political science literature (e.g.,
-Sharp, 1973). This process resulted in 72 actions that varied in how
-acceptable we would expect them to be. For details, see Supplemental
-Online Material (SOM).
-
-\hypertarget{study-2}{%
-\subsection{Study 2}\label{study-2}}
+and 15 current university students. Participants first read an
+accessible definition of what a social group is and that collective
+action is any action group members take to promote the interests of
+their social group.
+
+\begin{quote}
+Society is not only made up of individuals, but consists of `social
+groups' to which these individuals belong. Each person lives in a place,
+has a job (or not) at an organisation, is a fan of a specific sports
+club, has a religion, or belongs to any number of other such groups.
+Individuals often act in ways to promote the interests of the social
+groups to which they belong.
+\end{quote}
+
+\noindent Participants were then asked to name between five and ten
+actions that fit that definition. Participants were encouraged to think
+of actions that varied in how acceptable they were in their opinion. We
+recoded responses into a smaller set of unique actions, which we
+supplemented with protest actions from the psychological and political
+science literature (e.g., Sharp, 1973). This process resulted in 72
+actions that varied in how acceptable we would expect them to be. For
+details, see Supplemental Online Material (SOM).
 
@@
 
+\begin{figure*}[!t]
+\centering
+\caption{Examples of protest actions rated in Study 2}
+\includegraphics[scale=1]{../Scale Development/figures/figure-2}
+\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause.
+Darker bars mark actions that were included in the final scale.}
+\label{fig:f2}
+\end{figure*}
+
 Participants again read an accessible definition of collective action.
@@
 an instrument for measuring double standards in judging collective
-action in Experiments 1 and 2. We selected controversial, and thus
+action in Experiments 1--3. We selected controversial, and thus
 diagnostic, actions (e.g., disrupting traffic) over actions that almost
@@
 middle-class jobs to protest against a fictitious government bill
-threatening their rights.
+threatening their group's rights.
 
 By varying the participants' and the protesters' group memberships, we
-tested the preregistered hypothesis that, in line with \emph{Hypothesis
-1a} and \emph{1b}, participants would judge the same protest actions as
-more acceptable if the protesters were ingroup members protesting for
-the ingroup's rights than if they were outgroup members protesting for
-the outgroup's rights. By comparing reactions to working-class and
-middle-class protesters, we tested how observers judge collective action
-by lower-status (i.e.~working-class) and higher-status (middle-class)
-group members. Assuming that collective action by lower-status group
-members is seen as more threatening to the unequal status quo than
-collective action by higher-status group members and that people are, to
-same extent, motivated to justify the system, we tested the
-preregistered hypothesis that, in line with \emph{Hypothesis 2a},
-participants would judge the same protest actions as more acceptable if
-the protesters were from a higher-status group than if they were from a
-lower-status group. In non-preregistered analyses, we further explored
-whether participants would judge the same protest actions as more or
-less acceptable depending on their own ideological positions. In this
-way, Experiment 1 provided a first test of the hypothesized identity-
-and ideology-based double standards in judging collective action.
+tested the preregistered hypothesis that participants would judge the
+same protest actions as more acceptable if the protesters were ingroup
+members protesting for the ingroup's rights than if they were outgroup
+members protesting for the outgroup's rights. By comparing reactions to
+working-class and middle-class protesters, we further tested how
+observers judge collective action by lower-status (i.e.~working-class)
+and higher-status (middle-class) group members.\footnote{For Experiments
+  1 and 2, we had preregistered the additional hypotheses that all
+  participants would judge the same actions as more acceptable when the
+  protesters were from the higher-status group rather than the
+  lower-status group (Experiment 1) and when the goal of the protest was
+  to support rather than challenge the system (Experiment 2). These
+  hypotheses reflected the assumption that the motivation to justify the
+  system was, to some extent, universal. With hindsight, we no longer
+  think that this assumption is tenable and, therefore, omit the
+  hypotheses based on it from the revised manuscript. We still report
+  the same preregistered analyses and their results.} In
+non-preregistered analyses, we tested whether participants would judge
+the same protest actions as more or less acceptable depending on their
+own ideological positions in terms of political orientation,
+system-justifying beliefs, and social dominance orientation.\footnote{In
+  Experiment 1, we explored a broader range of operationalizations of
+  political ideology, including social dominance orientation, than in
+  the other experiments.} In this way, Experiment 1 provided a first
+test of the hypothesized class- and ideology-based double standards in
+judging collective action.
 
@@
 degree and placed themselves on the top four ranks of the subjective
-socioeconomic status ladder.
+socioeconomic status ladder. On average, participants tended to describe
+their political orientation as somewhat left-wing
+(\(\textit{M} = 3.54\), \(\textit{SD} = 1.31\)) with \(36\%\) and
+\(20\%\) describing their political orientation as, respectively, at
+least somewhat left-wing and right-wing.
 
@@
 preregistered inclusion criteria for the lower-status and higher-status
-groups. For the lower-status group, we recruited 475 participants who
-did not have a university degree and placed themselves on the bottom
-three ranks of a subjective socioeconomic status ladder. For the
-higher-status group, we recruited 400 participants who had at least an
-undergraduate degree and placed themselves on the top four ranks of the
-subjective socioeconomic status ladder.
+groups. For the lower-status group, we recruited participants who did
+not have a university degree and placed themselves on the bottom three
+ranks of a subjective socioeconomic status ladder. For the higher-status
+group, we recruited participants who had at least an undergraduate
+degree and placed themselves on the top four ranks of the subjective
+socioeconomic status ladder.
 
@@
 
-We recruited 500 participants from the remaining 687 participants, 250
-from the lower-status group and 250 from the higher-status group.
+We recruited 500 participants from the remaining participants, 250 from
+the lower-status group and 250 from the higher-status group.
 Participants were randomly assigned to read a vignette about a
@@
 participants' group status---but not in the directions predicted by our
-hypotheses. Contradicting our first hypothesis, participants did not
-consider protest actions performed by their ingroup to be more
-acceptable, on average, than the same actions performed by the relevant
-outgroup (Cohen's \(d = 0.00, [-0.20, 0.20]\)). Instead, we found that
+hypotheses. Contradicting our hypothesis, participants did not consider
+protest actions performed by their ingroup to be more acceptable, on
+average, than the same actions performed by the relevant outgroup
+(Cohen's \(d = 0.00, [-0.20, 0.20]\)). Instead, we found that
 participants from higher-status backgrounds considered protest actions
@@
 higher-status (\(d = 0.33, [0.03, 0.63]\)) protesters to be, on average,
-more acceptable than participants from lower-status backgrounds.
-Contradicting our second hypothesis, participants considered protest
-actions performed by higher-status protesters to be less acceptable, on
-average, than the same actions performed by lower-status protesters
-(\(d = -0.20, [-0.40, -0.00]\)).
+more acceptable than participants from lower-status backgrounds. All
+participants considered protest actions performed by higher-status
+protesters to be less acceptable, on average, than the same actions
+performed by lower-status protesters (\(d = -0.20, [-0.40, -0.00]\)).
 
@@
 
-Figure \ref{fig:f3} (B) the \(z\)-standardized propensity to consider
-more controversial protest actions acceptable as a function of the three
-ideological orientation variables. We found that participants who
-reported a more right-wing political orientation
+Figure \ref{fig:f3} shows (B) the \(z\)-standardized propensity to
+consider more controversial protest actions acceptable as a function of
+the three ideological orientation variables. We found that participants
+who reported a more right-wing political orientation
 (\(\beta_{xy} = -0.25, [-0.37, -0.15]\)) and who expressed more
@@
 
-Experiment 1 provided a first test of the hypothesized identity- and
-ideology-based double standards in judging collective action. That is,
-we tested whether, in line with \emph{Hypothesis 1a} and \emph{1b},
-participants would judge the same protest actions as more acceptable if
-the protesters were ingroup members protesting for the ingroup's rights
-or whether, in line with \emph{Hypothesis 2a}, participants would judge
-the same protest actions as more acceptable if the protesters were from
-a higher-status group. We found that participants' responses depended on
-both the participants' and the protesters' group memberships---but not
-in the hypothesized directions. Instead, we found that both lower-status
+Experiment 1 provided a first test of the hypothesized group-based
+double standards in judging collective action. That is, we tested
+whether, as hypothesized, participants would judge the same protest
+actions as more acceptable if the protesters were ingroup members
+protesting for the ingroup's rights. We found that participants'
+responses depended on both the participants' and the protesters' group
+memberships---but not in the hypothesized directions. Both lower-status
 and higher-status participants tended to find the same protest actions
 more acceptable when lower-status protesters protested a bill
-threatening their rights. In non-preregistered analyses, we further
-found that people who were right-wing and endorsed system-justifying
-beliefs tended to find all protest actions to be less acceptable.
+threatening their rights. In non-preregistered analyses, we found
+evidence for ideology-based double standards as people who were
+right-wing and endorsed system-justifying beliefs tended to find all
+protest actions to be less acceptable.
 
@@
 Second, by focusing on collective action to defend workers' rights, we
-arguably studied reactions to collective action for a progressive cause.
-To test for ideology-based double standards, we relied on the assumption
-that collective action by lower-status group members would be seen as
-more challenging to the prevailing system than collective action by
-higher-status group members. This assumption, however, was contradicted
-by our observation that participants who endorsed system-justifying
-beliefs tended to find actions by lower-status and higher-status
-protesters to be less acceptable. Experiment 2 addressed those
-limitations by focusing on a scenario we expected to evoke stronger
-responses from participants and by examining reactions to both
-system-challenging and system-defending collective action.
+studied reactions to collective action for a progressive cause. To test
+for ideology-based double standards, we needed to vary the protesters'
+cause to compare reactions to both progressive and conservative causes.
+Experiment 2 addressed those limitations by focusing on a scenario we
+expected to evoke stronger responses from participants and by examining
+reactions to both system-challenging and system-defending collective
+action.
 
@@
 same protest actions as more acceptable if the protesters were ingroup
-members (\emph{Hypothesis 1a}). By assuming that defunding the police
-aligns more closely with the interests of Black than White Americans, we
-tested the preregistered hypothesis that participants would judge the
-same protest actions as more acceptable if the cause of the protest
-aligned with their ingroup's interests (\emph{Hypothesis 1b}). By
-varying whether the protesters protested for or against defunding the
-police, we tested the preregistered hypotheses that participants would
-judge the same protest actions as more acceptable if the protest
-supported, rather than challenged, the system (\emph{Hypothesis 2a}) and
-if they endorsed system-justifying beliefs and the protest supported the
+members. By assuming that defunding the police aligns more closely with
+the interests of Black than White Americans, we tested the preregistered
+hypothesis that participants would judge the same protest actions as
+more acceptable if the cause of the protest aligned with their ingroup's
+interests. By varying whether the protesters protested for or against
+defunding the police, we tested the preregistered hypotheses that
+participants would judge the same protest actions as more acceptable if
+they endorsed system-justifying beliefs and the protest supported the
 system or if they rejected system-justifying beliefs and the protest
-challenged the system (\emph{Hypothesis 2b}). In this way, Experiment 2
-provided a complete test of the hypothesized identity- and
-ideology-based double standards in judging collective action. In
-addition, we tested the simpler, alternative hypothesis that
-participants, in general, would judge the same protest actions as more
-acceptable if they supported the protesters' cause (\emph{Hypothesis
-3}).
+challenged the system. In this way, Experiment 2 provided a complete
+test of the hypothesized race- and ideology-based double standards in
+judging collective action. In addition, we tested the simpler,
+alternative hypothesis that participants, in general, would judge the
+same protest actions as more acceptable if they supported the
+protesters' cause.
 
@@
 women, 708 men, 27 sex/gender diverse) of whom 800 identified as Black
-and 800 identified as White.
+and 800 identified as White. On average, participants tended to describe
+their political orientation as moderately liberal
+(\(\textit{M} = 2.92\), \(\textit{SD} = 1.73\)) with \(63\%\) and
+\(16\%\) describing their political orientation as, respectively, at
+least somewhat liberal and conservative.
 
@@
 Model 1 estimated varying intercepts for the eight conditions to test
-whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses
-depended on their own group membership, the protesters' group
-membership, or the protesters' cause. Contradicting \emph{Hypothesis
-1a}, participants did not consider protest actions performed by their
-ingroup to be more acceptable, on average, than the same actions
-performed by the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)).
-Contradicting \emph{Hypothesis 1b}, participants did not consider
-protest actions for a cause that was nominally aligned with their
-ingroup's interests to be more acceptable, on average, than protest
-actions for a cause not aligned with their ingroup's interests
-(\(d = -0.01, [-0.06, 0.04]\)). Contradicting \emph{Hypothesis 2a},
-participants did not consider protest actions for a system-defending
-cause to be more acceptable, on average, than protest actions for a
-system-challenging cause (\(d = -0.14, [-0.20, -0.09]\)). Instead, we
-found that, on average, White participants considered all protest
-actions to be more acceptable than Black participants
-(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.13, [0.06, 0.20]\))
-and White (\(d = 0.16, [0.08, 0.23]\)) participants considered the same
-actions to be more acceptable when protests were for, rather than
-against, defunding the police. As in Experiment 1, we thus found that
+whether, as hypothesized, participants' responses depended on their own
+group membership, the protesters' group membership, or the protesters'
+cause. Contradicting the hypothesized race-based double standards,
+participants did not consider protest actions performed by their ingroup
+to be more acceptable, on average, than the same actions performed by
+the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)). Likewise,
+participants did not consider protest actions for a cause that was
+nominally aligned with their ingroup's interests to be more acceptable,
+on average, than protest actions for a cause not aligned with their
+ingroup's interests (\(d = -0.01, [-0.06, 0.04]\)). Instead, we found
+that, on average, White participants considered all protest actions to
+be more acceptable than Black participants (\(d = 0.09, [0.04, 0.14]\))
+and both Black (\(d = 0.13, [0.06, 0.20]\)) and White
+(\(d = 0.16, [0.08, 0.23]\)) participants considered the same actions to
+be more acceptable when protests were for, rather than against,
+defunding the police. As in Experiment 1, we thus found that
 participants' responses depended on both the participants' and the
@@
 one varying effect estimating its variance across conditions. Supporting
-\emph{Hypothesis 2b}, participants who \emph{rejected} system-justifying
-beliefs were \emph{more} likely to consider system-challenging protest
-actions (for defunding the police) to be acceptable means of protest
-(\(\beta_{xy} = 0.38, [0.16, 0.58]\)). We did not, however, find
-evidence for the ideological symmetry implied by \emph{Hypothesis 2b}
-since participants who \emph{endorsed} system-justifying beliefs were
-\emph{not} more likely to consider system-defending protest actions
-(against defunding the police) to be acceptable means of protest
+the hypothesized ideology-based double standards, participants who
+\emph{rejected} system-justifying beliefs were \emph{more} likely to
+consider system-challenging protest actions (for defunding the police)
+to be acceptable means of protest (\(\beta_{xy} = 0.38, [0.16, 0.58]\)).
+We did not, however, find evidence for the hypothesized ideological
+symmetry since participants who \emph{endorsed} system-justifying
+beliefs were \emph{not} more likely to consider system-defending protest
+actions (against defunding the police) to be acceptable means of protest
 (\(\beta_{xy} = -0.18, [-0.40, 0.02]\)).
@@
 contrast, participants who rejected system-justifying beliefs evinced
-ideology-based double standards: In line with \emph{Hypothesis 2b}, they
-considered the same protest actions to be more acceptable when the
-protesters challenged the system or, to a lesser extent, when the
-protesters were from the disadvantaged group.
+ideology-based double standards: They considered the same protest
+actions to be more acceptable when the protesters challenged the system
+or, to a lesser extent, when the protesters were from the disadvantaged
+group.
 
@@
   proportions of the overall change that was expected to occur between
-  each of the four pairs of predictor categories.} Contradicting
-\emph{Hypothesis 3}, participants who supported the protesters' cause
-did not, on average, consider the same protest actions to be more
-acceptable than participants who opposed the protesters' cause
-(\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our results did not support
-the alternative hypothesis that ideology-based double standards can be
-reduced to support for, or opposition to, the cause of a protest.
+  each of the four pairs of predictor categories.} Participants who
+supported the protesters' cause did not, on average, consider the same
+protest actions to be more acceptable than participants who opposed the
+protesters' cause (\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our
+results did not support the alternative hypothesis that ideology-based
+double standards can be reduced to support for, or opposition to, the
+cause of a protest.
 
@@
 
-Experiment 2 provided a complete test of the hypothesized identity- and
+Experiment 2 provided a complete test of the hypothesized race- and
 ideology-based double standards in judging collective action. As in
@@
 participants' and the protesters' group memberships---but not in the
-hypothesized directions. We found that, contrary to \emph{Hypothesis
-1a}, participants considered protest actions taken by ingroup and
-outgroup members to be equally acceptable and that, contrary to
-\emph{Hypothesis 1b}, participants did not consider the same protest
-actions to be more acceptable when the protesters' cause aligned with
-their ingroup's interests. Like Experiment 1, Experiment 2 thus found no
-evidence for identity-based double standards---even though it focused on
-an issue of direct and current relevance to the participants.
+hypothesized directions. We found that, contrary to the hypothesized
+race-based double standards, participants considered protest actions
+taken by ingroup and outgroup members to be equally acceptable and
+participants did not consider the same protest actions to be more
+acceptable when the protesters' cause aligned with their ingroup's
+interests. Like Experiment 1, Experiment 2 thus found no evidence for
+group-based double standards---even though it focused on an issue of
+direct and current relevance to the participants.
 
@@
 doing, provided a stronger test of the hypothesized ideology-based
-double standards. Contrary to \emph{Hypothesis 2a} and the underlying
-assumption of a universal system-justifying motive, we found that
-participants considered the same protest actions to be \emph{more}
-acceptable when judging system-challenging collective action (for
-defunding the police) than when judging system-defending collective
-action. This finding might reflect our liberal-leaning sample.
-
-In line with \emph{Hypothesis 2b}, we found that participants who
-rejected system-justifying beliefs considered the same protest action
-more acceptable when judging system-challenging collective action (for
-defunding the police). We did not, however, find evidence for
-ideological symmetry in this relationship as participants who endorsed
-system-justifying beliefs considered system-challenging and
-system-defending collective action to be equally unacceptable. In
+double standards. As hypothesized, we found that participants who
+\emph{rejected} system-justifying beliefs considered the same protest
+action more acceptable when judging system-challenging collective action
+(for defunding the police). We did not, however, find evidence for
+ideological symmetry in this relationship as participants who
+\emph{endorsed} system-justifying beliefs considered system-challenging
+and system-defending collective action to be equally unacceptable. In
 addition, we did not find evidence for the alternative, simpler
-\emph{Hypothesis 3} that that participants would judge the same protest
-actions to be acceptable when they supported the protesters' cause.
+hypothesis that that participants would judge the same protest actions
+to be acceptable when they supported the protesters' cause.
 
@@
 and that \emph{opposing} a progressive cause is not equivalent to
-\emph{supporting} a conservative cause. Another reason might be that our
-sample skewed liberal and lacked committed conservatives who might
-accept more extreme means to oppose a progressive cause. Experiment 3
-addressed those limitations by testing for ideology-based double
-standards in judging collective action for or against a conservative
-cause in a balanced sample of conservatives and progressives.
+\emph{supporting} a conservative cause. Another reason might be that, as
+in Experiment 1, our sample skewed liberal and lacked committed
+conservatives who might accept more extreme means to oppose a
+progressive cause. Experiment 3 addressed those limitations by testing
+for ideology-based double standards in judging collective action for or
+against a conservative cause in a balanced sample of conservatives and
+progressives.
 
@@
 male and female participants, we further tested the preregistered
-hypothesis that, in line with identity-based double standards, women
-would judge protest actions against restricting abortion to be more
+hypothesis that, in line with gender-based double standards, women would
+judge protest actions against restricting abortion to be more
 acceptable, on average, than men or protest actions for restricting
 abortion.
+
+To achieve an ideologically balanced sample, we moved from an
+operationalization of political ideology in terms of system-justifying
+beliefs to an operationalization in terms of ideological self-placement.
+We did so for three reasons. First, there was the practical reason that
+Prolific and similar platforms allow prescreening participants in terms
+of ideological self-placement but not system-justifying beliefs. Second,
+there was the \emph{a priori} empirical reason that, when comparing the
+evidence for ideology-based double standards across the two
+operationalizations of political ideology, the patterns of results were
+nearly identical in terms of direction and magnitude for
+system-justifying beliefs and ideological self-placement (see, for
+example, Figure \ref{fig:f5}). Third, there was the \emph{a posteriori}
+empirical reason that, as reported in the Results section, dividing the
+sample into self-identified conservatives and liberals achieved a nearly
+complete separation of the sample in terms of system-justifying beliefs.
+For those reasons, selecting participants based on ideological
+self-placement was a feasible and effective quasi-experimental strategy
+to achieve an ideologically balanced sample.
 
@@
 
-As intended, participants who identified as conservative reported being
-far more conservative (\(M = 6.05, SD = 0.79\)) than participants who
-identified as liberal
-(\(M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43\)). Likewise,
-conservative participants tended to endorse system-justifying beliefs
-(\(M = 5.10, SD = 0.95\)) while liberal participants tended to reject
-system-justifying beliefs (\(M = 2.86, SD = 1.03; d = 2.26\)). As
-expected, conservative and liberal participants differed in their
+As expected, conservative participants tended to endorse
+system-justifying beliefs (\(M = 5.10, SD = 0.95\)) while liberal
+participants tended to reject system-justifying beliefs
+(\(M = 2.86, SD = 1.03; d = 2.26\)). Indeed, in \(93.5\%\) of all
+possible pairings of conservative and liberal participants, the
+conservative participant reported stronger system-justifying beliefs.
+This shows that ideological self-placement was an effective
+quasi-experimental approach to varying system justification.
+
+As expected, conservative and liberal participants differed in their
 reactions to the manipulation as the former tended to support
@@
 
-We found some evidence for identity-based double standards: Female
+We found some evidence for gender-based double standards: Female
 participants considered protest actions \emph{against} restricting
@@
 this way, Experiment 3 provided more consistent evidence for
-ideology-based than for identity-based double standards in judging
-collective action.
+ideology-based than for gender-based double standards in judging
+collective action. Across conditions, men tended to consider more
+protest actions acceptable than women (\(d = 0.16, [0.02, 0.30]\)).
 
@@
 restrictions. Our non-preregistered analyses thus suggested that gender
-identification did not moderate identity-based double standards in
-judging collective action concerning reproductive rights.
+identification did not moderate gender-based double standards in judging
+collective action concerning reproductive rights.
 
@@
 for and against restricting abortion in a sample balanced by gender and
-political orientation. As in Experiments 1--2, participants judged the
-same protest actions to be more acceptable when the protesters' cause
-aligned with their ideological orientation. While participants in
-Experiment 2 judged protests opposing or supporting a progressive cause
-(defunding police), participants in Experiment 3 considered protests
-opposing or supporting a conservative cause (restricting abortion). As
-both conservative and liberal participants judged protests for a cause
+political orientation. As hypothesized, participants judged the same
+protest actions to be more acceptable when the protesters' cause aligned
+with their ideological orientation. While participants in Experiment 2
+judged protests opposing or supporting a progressive cause (defunding
+police), participants in Experiment 3 considered protests opposing or
+supporting a conservative cause (restricting abortion). As both
+conservative and liberal participants judged protests for a cause
 aligned with their ideological orientation to be more acceptable and as
@@
 In contrast to Experiments 1--2, Experiment 3 provided some evidence for
-identity-based double standards, although any gender differences in
-judging protests for or against restricting abortion were overshadowed
-by more consistent ideology-based double standards. Our finding that
-conservative women considered collective action in support of
-restricting abortion to be more acceptable while liberal women
-considered collective action in opposition to restricting abortion to be
-more acceptable aligned with Mikolajczak et al.'s (2022) argument that,
-by itself, gender identity is too broad to explain support for
-collective action. Instead, we need to consider the content of those
-identities---for example, identification with feminism or with
-traditional women---to understand support for, and opposition to,
-progressive and reactionary collective action.
+group-based double standards distinct from ideology-based double
+standards, although any gender differences in judging protests for or
+against restricting abortion were overshadowed by more consistent
+ideology-based double standards. Our finding that conservative women
+considered collective action in support of restricting abortion to be
+more acceptable while liberal women considered collective action in
+opposition to restricting abortion to be more acceptable aligned with
+Mikolajczak et al.'s (2022) argument that, by itself, gender identity is
+too broad to explain support for collective action. Instead, we need to
+consider the content of those identities---for example, identification
+with feminism or with traditional women---to understand support for, and
+opposition to, progressive and reactionary collective action.
 
@@
 judging collective action. Our findings, however, showed that it is not
-so simple. Contradicting our hypothesis of identity-based double
-standards, participants in three preregistered experiments did not show
-consistent ingroup bias---in terms of class, race, and gender---when
-judging controversial protest actions taken by ingroup and outgroup
-members to advance their group's interests. Instead, supporting our
-hypothesis of ideology-based double standards, participants judged the
-same controversial protest actions to be more acceptable when the
-protesters' cause---protecting workers' rights, defunding the police,
-restricting abortion---aligned with their own ideological position. In
-the remainder of this General Discussion, we consider limiting
-conditions on our findings and discuss theoretical and practical
+so simple.
+
+Contradicting our hypothesis of group-based double standards,
+participants in three preregistered experiments did not show consistent
+ingroup bias in terms of social class when judging protests for workers'
+rights (Experiment 1), in terms of race when judging protests for and
+against defunding the police (Experiment 2), and in terms of gender when
+judging protests for and against restricting abortion (Experiment 3).
+Instead, we found that members of advantaged groups---middle-class
+(Experiment 1), White (Experiment 2), and male (Experiment 3)
+participants---tended to consider all protest actions more acceptable
+than members of relatively disadvantaged groups. That is, our findings
+showed that participants' judgements of protest actions depended on
+their group memberships---but not in the directions implied by the
+hypothesized group-based double standards.
+
+Supporting our hypothesis of ideology-based double standards,
+progressive participants who rejected system-justifying beliefs judged
+the same controversial protest actions to be more acceptable when the
+protesters' cause---for workers' rights (Experiment 1), for defunding
+the police (Experiment 2), against restricting abortion (Experiment
+3)---aligned with their own ideological position. While conservative
+participants who endorsed system-justifying beliefs considered the same
+actions somewhat more acceptable when protesters supported, rather than
+opposed, restricting abortion (Experiment 3), they were more consistent
+than liberal participants and considered all protest actions, for and
+against defunding the police (Experiment 2), less acceptable than
+progressive participants. In this way, our findings provided evidence
+for asymmetrical ideology-based double standards.
+
+In the remainder of this General Discussion, we discuss limiting
+conditions on our findings as well as theoretical and practical
 implications for understanding the often divided response to collective
@@
 
-Despite its strengths, our research is qualified by several limitations.
-First, our research examined the hypothesized relationships for three
-political causes---protecting workers' rights, defunding the police,
-restricting abortion---and thus provided only limited evidence that our
-findings generalize beyond those causes. Relying on few stimuli to
-establish an effect threatens the validity and replicability of research
-findings and constrains the generalizability of psychological research
-(Yarkoni, 2022). Future research should sample a wider range of causes
-to address this pervasive but often ignored problem (Judd et al., 2012).
-This is particularly important when studying collective action as recent
+Despite its strengths, our research has several limitations that
+constrain the generalizability of our findings. First, our research
+examined the hypothesized relationships for three political
+causes---protecting workers' rights, defunding the police, restricting
+abortion---and thus provided only limited evidence that our findings
+generalize beyond those causes. Relying on few stimuli to establish an
+effect threatens the validity and replicability of research findings and
+constrains the generalizability of psychological research (Yarkoni,
+2022). Future research should sample a wider range of causes to address
+this pervasive but often ignored problem (Judd et al., 2012). This is
+particularly important when studying collective action as recent
 theorizing (Jost et al., 2017) and research (Osborne et al., 2019)
@@
 
+\hypertarget{theoretical-implications}{%
+\subsubsection{Theoretical
+Implications}\label{theoretical-implications}}
+
 Growing evidence shows that, in many situations, observers apply a more
@@
 suggested that, unlike judgments in other domains, judgments about
-collective action are not subject to double standards based on group
-identities such as class, race, or gender. We did not find evidence for
-those double standards even though we examined social categories that
-are central to most people's identities, even when we considered only
-highly identified group members, and even though participants' reactions
-to the manipulation showed that, at least in Experiments 1--2, they
-understood the protesters' causes in ingroup--outgroup terms.
+collective action are not subject to double standards based on ascribed
+group memberships such as class, race, or gender. We did not find
+evidence for those double standards even though we examined social
+categories that are central to most people's identities, even when we
+considered only highly identified group members, and even though
+participants' reactions to the manipulation showed that, at least in
+Experiments 1--2, they understood the protesters' causes in
+ingroup--outgroup terms.
 
@@
 First, we might understand ideology-based double standards as an
-expression of ingroup bias based on opinion-based (McGarty et al., 2009)
-or partisan (Finkel et al., 2020; Harris et al., 2022) identities. This
-explanation aligns with Verkuyten et al.'s (2022) finding that
-participants were more tolerant of transgressive protest actions when
-taken by their most-liked, rather than their least-liked, political
-group (e.g., climate activists). An identity-based account, however,
-implies that both conservatives and progressives would show ingroup bias
-and, therefore, cannot explain the observed asymmetry in the two groups'
-judgements.
+expression of ingroup bias based on partisan identities (Finkel et al.,
+2020; Harris et al., 2022). This explanation aligns with Verkuyten et
+al.'s (2022) finding that participants were more tolerant of
+transgressive protest actions when taken by their most-liked, rather
+than their least-liked, political group. An identity-based account,
+however, implies that both conservatives and progressives would show
+ingroup bias and, therefore, cannot explain the observed asymmetry in
+the two groups' judgements.
 
@@
 
-Most broadly, our findings showed that, for many controversial protest
-actions, there is no consensus on whether they are acceptable means to
-advance a cause. This confirms our contention that the distinction
-between normative and non-normative forms of collective action should be
-situated in the eye of the beholder. By developing an instrument for
+\hypertarget{practical-implications}{%
+\subsubsection{Practical Implications}\label{practical-implications}}
+
+Most broadly, our findings showed that, for many protest actions (e.g.,
+blocking roads), there is no consensus on whether they are acceptable
+means to advance a cause. This confirms our contention that the
+distinction between normative and non-normative collective action should
+be situated in the eye of the beholder. By developing an instrument for
 measuring where people draw the line between acceptable and unacceptable
@@
 
-After all, this distinction is consequential. If people believe that
-even extreme actions are acceptable means to advance a cause they
-support, especially if they moralize their support for that cause
-(Mooijman et al., 2018), they might support violent extremism. In this
-way, our findings could explain why 45\% of Republicans supported the
-storming of the U.S. Capitol on January 6, 2021 (YouGov, 2021) even
-though it was illegal, destructive, and violent. If people believe that
-even non-violent actions are unacceptable means to advance a cause they
-oppose, they might support stifling dissent. In this way, our findings
-could explain why Republicans responded to Black Lives Matter protests
-by seeking to further criminalize disruptive protest (e.g., blocking
-highways, Quinton, 2021) that might be most effective at challenging
-injustice (Shuman et al., 2022, 2021).
+After all, this distinction is consequential. Research on how people
+respond to collective action showed that disruptive yet non-violent
+collective action might be most effective at gaining concessions from
+those resistant to social change (Shuman et al., 2021, 2023). If, as our
+research suggests, people believe that even non-violent actions are
+unacceptable means to advance a cause they oppose, they might support
+stifling dissent. This could explain why Republicans responded to Black
+Lives Matter protests by seeking to further criminalize disruptive
+protest (e.g., blocking highways, Quinton, 2021) that might be most
+effective at challenging injustice. In this way, our research
+complements, and adds to, the growing program of research on how people
+respond to different kinds of collective action.
+
+If people believe that even extreme actions are acceptable means to
+advance a cause they support, especially if they moralize their support
+for that cause (Mooijman et al., 2018), they might support violent
+extremism. For example, 45\% of Republicans supported the storming of
+the U.S. Capitol on January 6, 2021 (YouGov, 2021) even though it was
+illegal, destructive, and violent. Recent research, most of it by
+political scientists, investigated support for partisan violence (Kalmoe
+\& Mason, 2022), albeit with mixed results (Westwood et al., 2022).
+While our research focused on controversial but non-violent actions, we
+encourage researchers to adapt our paradigm to investigate when and why
+people apply different standards when judging violent and non-violent
+collective action for different causes.
 
@@
 of the same class, race or gender or for a cause aligned with their own
-group's interests. Instead, our research demonstrated that people
-consider the same protest actions to be more acceptable when the cause
-of the protest aligns with their own ideological position.
+group's interests. Instead, our research demonstrated that progressives
+and, less so, conservatives consider the same protest actions to be more
+acceptable when the cause of the protest aligns with their own
+ideological position.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kalmoe_radical_2022}{}}%
+Kalmoe, N. P., \& Mason, L. (2022). \emph{Radical {American}
+partisanship: Mapping violent hostility, its causes, and the
+consequences for democracy}. The University of Chicago Press.
+
 \leavevmode\vadjust pre{\hypertarget{ref-kay_complementary_2003}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lacina_nfl_2020}{}}%
-Lacina, B. (2020). \emph{National anthem protests and whites' views of
-black NFL players} {[}Unpublished manuscript{]}. Department of Political
-Science, University of Rochester.
-\url{http://www.bethanylacina.com/docs/Lacina_NFL_paper.pdf}
-
 \leavevmode\vadjust pre{\hypertarget{ref-leach_social_2017}{}}%
@@
 \emph{16}(1), 3--32. \url{https://doi.org/10.1111/sipr.12084}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcgarty_collective_2009}{}}%
-McGarty, C., Bliuc, A.-M., Thomas, E. F., \& Bongiorno, R. (2009).
-Collective action as the material expression of opinion-based group
-membership. \emph{Journal of Social Issues}, \emph{65}(4), 839--857.
-\url{https://doi.org/10.1111/j.1540-4560.2009.01627.x}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-acceptable}{}}%
+Oxford Advanced Learner's Dictionary. (n.d.). \emph{Acceptable}.
+\url{https://www.oxfordlearnersdictionaries.com/us/definition/english/acceptable}
+
 \leavevmode\vadjust pre{\hypertarget{ref-pettigrew_ultimate_1979}{}}%
@@
 \url{https://doi.org/10.1177/014616727900500407}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pinsof_strange_2023}{}}%
+Pinsof, D., Sears, D. O., \& Haselton, M. G. (2023). Strange bedfellows:
+The alliance theory of political belief systems. \emph{Psychological
+Inquiry}, \emph{34}(3), 139--160.
+\url{https://doi.org/10.1080/1047840X.2023.2274433}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-shuman_protest_2022}{}}%
-Shuman, E., Hasan-Aslih, S., Zomeren, M. van, Saguy, T., \& Halperin, E.
-(2022). Protest movements involving limited violence can sometimes be
-effective: Evidence from the 2020 {Black Lives Matter} protests.
-\emph{Proceedings of the National Academy of Sciences}, \emph{119}(14),
-e2118990119. \url{https://doi.org/10.1073/pnas.2118990119}
+\leavevmode\vadjust pre{\hypertarget{ref-shuman_when_2023}{}}%
+Shuman, E., Goldenberg, A., Saguy, T., Halperin, E., \& Van Zomeren, M.
+(2023). When are social protests effective? \emph{Trends in Cognitive
+Sciences}, S1364661323002619.
+\url{https://doi.org/10.1016/j.tics.2023.10.003}
 
@@
 Bulletin}. \url{https://doi.org/10.1177/01461672221099709}
+
+\leavevmode\vadjust pre{\hypertarget{ref-westwood_current_2022}{}}%
+Westwood, S. J., Grimmer, J., Tyler, M., \& Nall, C. (2022). Current
+research overstates {American} support for political violence.
+\emph{Proceedings of the National Academy of Sciences}, \emph{119}(12),
+e2116870119. \url{https://doi.org/10.1073/pnas.2116870119}
 

### 2024-04-16 Documents/supplemental-materials.qmd
resubmit

@@
   # Load packages
-  library(tidyverse); library(tidybayes); library(numform)
+  library(tidyverse)
+  library(tidybayes)
+  library(numform)
+  library(glue)
 
@@
 For each item $i$, the model estimated four acceptability thresholds ($\beta_{ik}$), separating the five response options, and one discrimination parameter ($\alpha_i$) indicating how well the item differentiates between participants with different propensities to consider collective actions acceptable.
-We estimated the model in *CmdStanR* using similar prior distribution as in Experiment.
+We estimated the model in *CmdStanR* using similar prior distribution as in Experiment 1.
 We used an induced Dirichlet prior [@betancourt_ordinal_2019] for the item-specific difficulty thresholds.
 
-Table S1 shows each actions' item information and discrimination as well as proportions for the observed responses.
+Table S1 shows each protest action's information and discrimination parameters as well as the proportions of participants choosing each answer option.
 Table S2 shows participants' averaged ratings of how acceptable, disruptive, violent, extreme, and negative they considered each action to be.
-Table S3 shows that all five ratings were strongy correlated across actions and participants.
-We considered both estimates from the graded response model and their relevance to the social contexts to select actions for Experiments 1 and 2.
+Table S3 shows that all five ratings were strongly correlated across actions and participants.
+
+Figure \ref{fig:s1} shows the scale information curves for scales composed of, respectively, the 1--72 most informative and discriminating protest actions.
+We selected the 25 most informative actions to construct the final scale to be used in Experiments 1--3.
+While adding more actions to the scale would have improved the information provided by the instrument, the added benefit of each additional action included in the scale would have been small (see Figure \ref{fig:s1}).
+We limited the number of protest actions to 25 so as to not overwhelm participants and to prevent satisficing.
+Tables A1, B1, and C1 list the protest actions used in Experiments 1--3.
+
+```{=latex}
+\begin{figure*}[!t]
+\centering
+\caption{Scale information curves for scales composed of the 1--72 most informative and discriminating protest actions}
+\includegraphics[scale=1]{../Scale Development/figures/figure-s1}
+\caption*{\textit{Note.} The information curve for the final scale of 25 actions is highlighted in red.
+As in Figure 1, $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
+\label{fig:s1}
+\end{figure*}
+```
 
@@
 -->
+
+# Reanalyses Without Exclusions
+
+We replicated the analyses for Experiments 1-3 using the complete data without excluding participants based on the (preregistered) exclusion criteria.
+
+## Experiment 1
+
+```{=latex}
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered (\textbf{A}) and non-preregistered (\textbf{B}) analyses for Experiment 1 without exclusions}
+\includegraphics[scale=1]{../Experiment 1/figures/figure-s2}
+\caption*{\textit{Note.} HS = Higher Status, LS = Lower Status.
+$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
+(\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition.
+Bars enclose the 95\% most plausible estimates.
+(\textbf{B}) Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
+\label{fig:s2}
+\end{figure*}
+```
+
+Figure \ref{fig:s2} (A) shows estimates for each combination of the protesters' and the participants' group membership.
+Overall, we found that participants' responses depended on both the protesters' and the participants' group status---but not in the directions predicted by our hypotheses.
+Contradicting our first hypothesis, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (Cohen's $d = -0.01, [-0.20, 0.16]$).
+Instead, we found that participants from higher-status backgrounds considered protest actions performed by both lower-status ($d = 0.40, [0.15, 0.66$]) and higher-status ($d = 0.37, [0.10, 0.63]$) protesters to be, on average, more acceptable than participants from lower-status backgrounds.
+Contradicting our second hypothesis, participants considered protest actions performed by higher-status protesters to be less acceptable, on average, than the same actions performed by lower-status protesters ($d = -0.20, [-0.37, -0.02]$).
+
+Figure \ref{fig:s2} shows (B) the $z$-standardized propensity to consider more controversial protest actions acceptable as a function of the three ideological orientation variables.
+We found that participants who reported a more right-wing political orientation ($\beta_{xy} = -0.24, [-0.34, -0.15]$) and who expressed more agreement with system-justifying beliefs ($\beta_{xy} = -0.21, [-0.31, -0.12]$) tended to find fewer collective actions to be acceptable.
+In contrast, we found that, after controlling for the other two variables, social dominance orientation was not associated with participants' judgments about how acceptable various collective actions are ($\beta_{xy} = -0.06, [-0.16, 0.03]$).
+Overall, these findings suggest that people who are right-wing and endorse system-justifying beliefs tend to find various collective actions to be less acceptable than people who are left-wing and do not endorse system-justifying beliefs.
+
+## Experiment 2
+
+### Preregistered Analyses
+
+```{=latex}
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered analyses for Experiment 2}
+\includegraphics[scale=1]{../Experiment 2/figures/figure-s3}
+\caption*{\textit{Note.} Against = Protesters oppose defunding the police.
+For = Protesters support defunding the police.
+$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
+Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
+\label{fig:s3}
+\end{figure*}
+```
+
+```{=latex}
+\begin{figure*}[!t]
+\centering
+\caption{Predictions from the preregistered (system justification) and non-preregistered (political orientation) analyses for Experiment 2 without exclusions}
+\includegraphics[scale=1]{../Experiment 2/figures/figure-s4}
+\caption*{\textit{Note.} Against = Protesters oppose defunding the police.
+For = Protesters support defunding the police.
+$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
+\label{fig:s4}
+\end{figure*}
+```
+
+Figure \ref{fig:s3} shows results from three preregistered models we estimated to test our hypotheses.
+
+```{r, include = FALSE}
+
+  # Import results
+  m1 <- read_rds("../Experiment 2/results/m1_post_without_exclusions.rds")
+
+  # Define contrasts
+  m1 <- m1 %>% 
+    mutate(
+      h1a_contrast = case_when(
+        Participant == Protesters ~ +1/8,
+        Participant != Protesters ~ -1/8,
+      ),
+      h1b_contrast = case_when(
+        Participant == "Black" & Cause == "For" ~ +1/8,
+        Participant == "White" & Cause == "Against" ~ +1/8,
+        TRUE ~ -1/8,
+      ),
+      h2a_contrast = case_when(
+        Cause == "For" ~ -1/8,
+        Cause == "Against" ~ +1/8
+      ),
+      nh1_contrast = case_when(
+        Participant == "Black" ~ -1/8,
+        Participant == "White" ~ +1/8
+      ),
+      nh2_contrast = case_when(
+        Participant == "White" & Cause == "Against" ~ -1/4,
+        Participant == "White" & Cause == "For" ~ +1/4,
+        Participant == "Black" ~ 0
+      ),
+      nh3_contrast = case_when(
+        Participant == "Black" & Cause == "Against" ~ -1/4,
+        Participant == "Black" & Cause == "For" ~ +1/4,
+        Participant == "White" ~ 0
+      )
+    )
+
+  # Calculate contrasts
+  m1 <- m1 %>% 
+    group_by(.draw) %>% 
+    summarize(
+      across(
+        ends_with("_contrast"),
+        ~sum(z * .)
+      )
+    ) %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "d"
+    ) %>% 
+    group_by(contrast) %>% 
+    median_qi(d)
+
+  # Make text
+  m1 <- m1 %>% 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
+```
+
+Model 1 estimated varying intercepts for the eight conditions to test whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses depended on their own group membership, the protesters' group membership, or the protesters' cause.
+Contradicting *Hypothesis 1a*, participants did not consider protest actions performed by their ingroup to be more acceptable, on average, than the same actions performed by the relevant outgroup (`r with(m1, d_text[contrast == "h1a_contrast"])`).
+Contradicting *Hypothesis 1b*, participants did not consider protest actions for a cause that was nominally aligned with their ingroup's interests to be more acceptable, on average, than protest actions for a cause not aligned with their ingroup's interests (`r with(m1, d_text[contrast == "h1b_contrast"])`).
+Contradicting *Hypothesis 2a*, participants did not consider protest actions for a system-defending cause to be more acceptable, on average, than protest actions for a system-challenging cause (`r with(m1, d_text[contrast == "h2a_contrast"])`).
+Instead, we found that, on average, White participants considered all protest actions to be more acceptable than Black participants (`r with(m1, d_text[contrast == "nh1_contrast"])`) and both Black (`r with(m1, d_text[contrast == "nh3_contrast"])`) and White (`r with(m1, d_text[contrast == "nh2_contrast"])`) participants considered the same actions to be more acceptable when protests were for, rather than against, defunding the police.
+As in Experiment 1, we thus found that participants' responses depended on both the participants' and the protesters' group memberships---but not in the directions predicted by our hypotheses.
+
+```{r, include = FALSE}
+
+  # Import results
+  m2 <- read_rds("../Experiment 2/results/m2_post_without_exclusions.rds")
+
+  # Hypothesis 2
+  m2_h2 <- m2 %>% 
+    distinct(.draw, b_sjb_for, b_sjb_against) %>% 
+    pivot_longer(
+      c(b_sjb_for, b_sjb_against),
+      names_to = "Cause",
+      names_prefix = "b_sjb_",
+      values_to = "b"
+    ) %>% 
+    group_by(Cause) %>% 
+    mutate(
+      b = case_when(
+        Cause == "for" ~ -b,
+        Cause == "against" ~ b
+      )
+    ) %>% 
+    median_qi(b) %>% 
+    mutate(
+      across(
+        c(b, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      b_text = glue(
+        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
+        .open = "(", .close = ")"
+      )
+    )
+
+```
+
+Model 2 extended Model 1 by estimating participants' responses as a function of their *z*-standardized endorsement of system-justifying beliefs.
+As preregistered, we modeled this relationship with two fixed effects, one estimating the effect of system-justifying beliefs on judgments about system-defending protest actions and one estimating their effect on judgments about system-challenging protest actions, and one varying effect estimating its variance across conditions.
+Supporting *Hypothesis 2b*, participants who *rejected* system-justifying beliefs were *more* likely to consider system-challenging protest actions (for defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "for"])`).
+We did not, however, find evidence for the ideological symmetry implied by *Hypothesis 2b* since participants who *endorsed* system-justifying beliefs were *not* more likely to consider system-defending protest actions (against defunding the police) to be acceptable means of protest (`r with(m2_h2, b_text[Cause == "against"])`).
+
+Figure \ref{fig:s4} shows the estimated pattern of condition-wise differences underlying those fixed effects.
+Participants who endorsed system-justifying beliefs tended to consider system-challenging and system-defending protest actions to be equally unacceptable.
+In contrast, participants who rejected system-justifying beliefs evinced ideology-based double standards: In line with *Hypothesis 2b*, they considered the same protest actions to be more acceptable when the protesters challenged the system or, to a lesser extent, when the protesters were from the disadvantaged group.
+
+```{r, include = FALSE}
+
+  # Import results
+  m3 <- read_rds("../Experiment 2/results/m3_post_without_exclusions.rds")
+
+  # Hypothesis 2
+  m3_h3 <- m3 %>% 
+    distinct(.draw, b_sup) %>%
+    rename(b = b_sup) %>% 
+    median_qi(b) %>% 
+    mutate(
+      across(
+        c(b, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      b_text = glue(
+        "$\\beta_{y} = (b), [(.lower), (.upper)]$",
+        .open = "(", .close = ")"
+      )
+    )
+
+```
+
+Model 3 extended Model 1 by estimating participants' responses as a function of their self-reported support for the cause of the protest.
+We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police.
+As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Contradicting *Hypothesis 3*, participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
+Ergo, our results did not support the alternative hypothesis that ideology-based double standards can be reduced to support for, or opposition to, the cause of a protest.
+
+### Non-Preregistered Analyses
+
+```{r, include = FALSE}
+
+  # Import results
+  m5 <- read_rds("../Experiment 2/results/m5_post_without_exclusions.rds")
+
+  # Define contrasts
+  m5 <- m5 %>% 
+    mutate(
+      h1a_contrast = case_when(
+        Participant == Protesters ~ +1/8,
+        Participant != Protesters ~ -1/8
+      ),
+      h1b_contrast = case_when(
+        Participant == "Black" & Cause == "For" ~ +1/8,
+        Participant == "White" & Cause == "Against" ~ +1/8,
+        TRUE ~ -1/8,
+      )
+    )
+
+  # Calculate contrasts
+  m5 <- m5 %>% 
+    group_by(.draw, x_ide) %>% 
+    summarize(
+      across(
+        ends_with("_contrast"),
+        ~sum(z * .)
+      )
+    ) %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "d"
+    ) %>% 
+    group_by(x_ide, contrast) %>% 
+    median_qi(d)
+
+  # Make text
+  m5 <- m5 %>% 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
+```
+
+We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses.
+To that end, we extended Model 1 by estimating participants' responses as a function of their *z*-standardized identification with their racial ingroup.
+As in Model 2, we modeled this relationship with a fixed and a varying effect.
+We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast" & x_ide == 1L])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast" & x_ide == 1L])`).
+Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgments about collective action by ingroup and outgroup members.
+
+```{r, include = FALSE}
+
+  # Import results
+  m4 <- read_rds("../Experiment 2/results/m4_post_without_exclusions.rds")
+
+  # Hypothesis 2
+  m4_exp <- m4 %>% 
+    distinct(.draw, b_pol_for, b_pol_against) %>% 
+    pivot_longer(
+      c(b_pol_for, b_pol_against),
+      names_to = "Cause",
+      names_prefix = "b_pol_",
+      values_to = "b"
+    ) %>% 
+    group_by(Cause) %>% 
+    mutate(
+      b = case_when(
+        Cause == "for" ~ -b,
+        Cause == "against" ~ b
+      )
+    ) %>% 
+    median_qi(b) %>% 
+    mutate(
+      across(
+        c(b, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      b_text = glue(
+        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
+        .open = "(", .close = ")"
+      )
+    )
+
+```
+
+We also explored whether we would find ideology-based double standards when operationalizing ideology as political orientation instead of as system justification.
+To that end, we reran Model 2 with political orientation as the *z*-standardized predictor variable.
+Our results mirrored the preregistered analyses: More liberal participants were more likely to consider protest actions for defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "for"])`) but more conservative participants were not more likely to consider protest actions against defunding the police to be acceptable (`r with(m4_exp, b_text[Cause == "against"])`).
+As Figure \ref{fig:s4} shows, conservative participants tended to consider all protest actions to be equally unacceptable while liberal participants considered the same protest actions to be more acceptable when protesters rallied around a progressive cause.
+Our non-preregistered analyses thus replicated the ideology-based double standards from the preregistered analyses with a different operationalization of ideology.
+
+## Experiment 3
+
+```{=latex}
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered analyses for Experiment 3 without exclusions}
+\includegraphics[scale=1]{../Experiment 3/figures/figure-s5}
+\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
+\label{fig:s5}
+\end{figure*}
+```
+
+### Preregistered Analyses
+
+```{r, include = FALSE}
+
+  # Import results
+  m1 <- read_rds("../Experiment 3/results/m1_post_without_exclusions.rds")
+
+  # Define contrasts
+  m1 <- m1 %>% 
+    mutate(
+      h1_contrast = case_when(
+        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L
+      ),
+      h2a_contrast = case_when(
+        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L
+      ),
+      h2b_contrast = case_when(
+        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ -1L
+      ),
+      nh1_contrast = case_when(
+        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Conservative" & protesters_cause == "Against" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh2_contrast = case_when(
+        participant_ideology == "Liberal" & protesters_cause == "For" ~ -1L,
+        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ NA_integer_
+      ),
+      nh3_contrast = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "Against" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh4_contrast = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For" ~ +1L,
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "Against" ~ -1L,
+        TRUE ~ NA_integer_
+      ),
+      nh5_contrast = case_when(
+        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "For" ~ -1L,
+        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ NA_integer_
+      ),
+      nh6_contrast = case_when(
+        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
+        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
+        TRUE ~ NA_integer_
+      )
+    )
+
+  # Calculate contrasts
+  m1 <- m1 %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "direction",
+      values_drop_na = TRUE
+    ) %>% 
+    group_by(.draw, contrast, direction) %>% 
+    summarize(z = mean(z)) %>% 
+    summarize(d = sum(z * direction)) %>% 
+    pivot_wider(
+      names_from = contrast,
+      values_from = d
+    ) %>% 
+    mutate(nh7_contrast = nh2_contrast - nh1_contrast) %>% 
+    pivot_longer(
+      ends_with("_contrast"),
+      names_to = "contrast",
+      values_to = "d"
+    ) %>% 
+    group_by(contrast) %>% 
+    median_qi(d)
+
+  # Make text
+  m1 <- m1 %>% 
+    mutate(
+      across(
+        c(d, .lower, .upper), 
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      d_text = glue(
+        "$d = {d}, [{.lower}, {.upper}]$"
+      )
+    )
+
+```
+
+Figure \ref{fig:s5} shows the results of our preregistered analyses without exclusions.
+
+As in Experiments 1 and 2, we found evidence for ideology-based double standards: Conservative participants considered the same protest actions to be more acceptable, on average, when protesters *supported* restricting abortion (`r with(m1, d_text[contrast == "nh1_contrast"])`) while liberal participants considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion (`r with(m1, d_text[contrast == "nh2_contrast"])`).
+This double standard was more pronounced among liberal participants (`r with(m1, d_text[contrast == "nh7_contrast"])`).
+In this way, Experiment 3 provided evidence for both ideological symmetry and asymmetry in judging collective action for and against restricting abortion.
+
+We found some evidence for identity-based double standards: Female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`).
+This pattern was, however, overshadowed by stronger ideology-based double standards that were consistent across subsamples: Conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and both liberal women (`r with(m1, d_text[contrast == "nh6_contrast"])`) and liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion.
+In this way, Experiment 3 provided more consistent evidence for ideology-based than for identity-based double standards in judging collective action.
+
+### Non-Preregistered Analyses
+
+```{r, include = FALSE}
+
+  # Import results
+  ide <- read_rds("../Experiment 3/results/ide_post_without_exclusions.rds")
+
+  # Make text
+  ide <- ide %>%
+    group_by(.draw, participant_gender, protesters_cause) %>% 
+    summarize(
+      b = mean(b_ide_kk)
+    ) %>% 
+    group_by(participant_gender, protesters_cause) %>%
+    median_qi(b) %>% 
+    mutate(
+      across(
+        c(b, .lower, .upper),
+        f_num, digits = 2, retain.leading.zero = T
+      ),
+      b_text = glue(
+        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
+        .open = "(", .close = ")"
+      )
+    )
+
+```
+
+We explored whether group identification moderated participants' judgments.
+We found, however, that gender identification neither affected how women judged protesters opposing (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Woman" & protesters_cause == "For"])`) abortion restrictions nor how men judged protesters opposing (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "Against"])`) or supporting (`r with(ide, b_text[participant_gender == "Man" & protesters_cause == "For"])`) abortion restrictions.
+Our non-preregistered analyses thus suggested that gender identification did not moderate identity-based double standards in judging collective action concerning reproductive rights.
 
@@
 \end{tabular}
-\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Political Orientation (1 = \textit{left}, 7 = \textit{right}); Social Dominance Orientation (\textit{factor score}); Systen-Justifying Beliefs (\textit{factor score})}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Political Orientation (1 = \textit{left}, 7 = \textit{right}); Social Dominance Orientation (\textit{factor score}); System-Justifying Beliefs (\textit{factor score})}
 \end{table}
@@
 \end{tabular}
-\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Systen-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Racial Identification (1 = \textit{not at all}, 7 = \textit{very much})}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); System-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Racial Identification (1 = \textit{not at all}, 7 = \textit{very much})}
 \end{table}
@@
 \end{tabular}
-\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); Systen-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Gender Identification (1 = \textit{not at all}, 7 = \textit{very much})}
+\caption*{\textit{Note.} Proportion of actions considered acceptable (0--1); System-Justifying Beliefs (\textit{factor score}); Policy Support (1 = \textit{strongly oppose}, 5 = \textit{strongly support}); Political Orientation (1 = \textit{liberal}, 7 = \textit{conservative}); Gender Identification (1 = \textit{not at all}, 7 = \textit{very much})}
 \end{table}

### 2024-04-16 Documents/supplemental-materials.tex
resubmit

@@
 estimated the model in \emph{CmdStanR} using similar prior distribution
-as in Experiment. We used an induced Dirichlet prior (Betancourt, 2019)
-for the item-specific difficulty thresholds.
+as in Experiment 1. We used an induced Dirichlet prior (Betancourt,
+2019) for the item-specific difficulty thresholds.
 
-Table S1 shows each actions' item information and discrimination as well
-as proportions for the observed responses. Table S2 shows participants'
-averaged ratings of how acceptable, disruptive, violent, extreme, and
-negative they considered each action to be. Table S3 shows that all five
-ratings were strongy correlated across actions and participants. We
-considered both estimates from the graded response model and their
-relevance to the social contexts to select actions for Experiments 1 and
-2.
+Table S1 shows each protest action's information and discrimination
+parameters as well as the proportions of participants choosing each
+answer option. Table S2 shows participants' averaged ratings of how
+acceptable, disruptive, violent, extreme, and negative they considered
+each action to be. Table S3 shows that all five ratings were strongly
+correlated across actions and participants.
+
+Figure \ref{fig:s1} shows the scale information curves for scales
+composed of, respectively, the 1--72 most informative and discriminating
+protest actions. We selected the 25 most informative actions to
+construct the final scale to be used in Experiments 1--3. While adding
+more actions to the scale would have improved the information provided
+by the instrument, the added benefit of each additional action included
+in the scale would have been small (see Figure \ref{fig:s1}). We limited
+the number of protest actions to 25 so as to not overwhelm participants
+and to prevent satisficing. Tables A1, B1, and C1 list the protest
+actions used in Experiments 1--3.
+
+\begin{figure*}[!t]
+\centering
+\caption{Scale information curves for scales composed of the 1--72 most informative and discriminating protest actions}
+\includegraphics[scale=1]{../Scale Development/figures/figure-s1}
+\caption*{\textit{Note.} The information curve for the final scale of 25 actions is highlighted in red.
+As in Figure 1, $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
+\label{fig:s1}
+\end{figure*}
 
@@
 
+\hypertarget{reanalyses-without-exclusions}{%
+\section{Reanalyses Without
+Exclusions}\label{reanalyses-without-exclusions}}
+
+We replicated the analyses for Experiments 1-3 using the complete data
+without excluding participants based on the (preregistered) exclusion
+criteria.
+
+\hypertarget{experiment-1-1}{%
+\subsection{Experiment 1}\label{experiment-1-1}}
+
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered (\textbf{A}) and non-preregistered (\textbf{B}) analyses for Experiment 1 without exclusions}
+\includegraphics[scale=1]{../Experiment 1/figures/figure-s2}
+\caption*{\textit{Note.} HS = Higher Status, LS = Lower Status.
+$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
+(\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition.
+Bars enclose the 95\% most plausible estimates.
+(\textbf{B}) Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
+\label{fig:s2}
+\end{figure*}
+
+Figure \ref{fig:s2} (A) shows estimates for each combination of the
+protesters' and the participants' group membership. Overall, we found
+that participants' responses depended on both the protesters' and the
+participants' group status---but not in the directions predicted by our
+hypotheses. Contradicting our first hypothesis, participants did not
+consider protest actions performed by their ingroup to be more
+acceptable, on average, than the same actions performed by the relevant
+outgroup (Cohen's \(d = -0.01, [-0.20, 0.16]\)). Instead, we found that
+participants from higher-status backgrounds considered protest actions
+performed by both lower-status (\(d = 0.40, [0.15, 0.66\){]}) and
+higher-status (\(d = 0.37, [0.10, 0.63]\)) protesters to be, on average,
+more acceptable than participants from lower-status backgrounds.
+Contradicting our second hypothesis, participants considered protest
+actions performed by higher-status protesters to be less acceptable, on
+average, than the same actions performed by lower-status protesters
+(\(d = -0.20, [-0.37, -0.02]\)).
+
+Figure \ref{fig:s2} shows (B) the \(z\)-standardized propensity to
+consider more controversial protest actions acceptable as a function of
+the three ideological orientation variables. We found that participants
+who reported a more right-wing political orientation
+(\(\beta_{xy} = -0.24, [-0.34, -0.15]\)) and who expressed more
+agreement with system-justifying beliefs
+(\(\beta_{xy} = -0.21, [-0.31, -0.12]\)) tended to find fewer collective
+actions to be acceptable. In contrast, we found that, after controlling
+for the other two variables, social dominance orientation was not
+associated with participants' judgments about how acceptable various
+collective actions are (\(\beta_{xy} = -0.06, [-0.16, 0.03]\)). Overall,
+these findings suggest that people who are right-wing and endorse
+system-justifying beliefs tend to find various collective actions to be
+less acceptable than people who are left-wing and do not endorse
+system-justifying beliefs.
+
+\hypertarget{experiment-2}{%
+\subsection{Experiment 2}\label{experiment-2}}
+
+\hypertarget{preregistered-analyses}{%
+\subsubsection{Preregistered Analyses}\label{preregistered-analyses}}
+
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered analyses for Experiment 2}
+\includegraphics[scale=1]{../Experiment 2/figures/figure-s3}
+\caption*{\textit{Note.} Against = Protesters oppose defunding the police.
+For = Protesters support defunding the police.
+$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
+Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
+\label{fig:s3}
+\end{figure*}
+
+\begin{figure*}[!t]
+\centering
+\caption{Predictions from the preregistered (system justification) and non-preregistered (political orientation) analyses for Experiment 2 without exclusions}
+\includegraphics[scale=1]{../Experiment 2/figures/figure-s4}
+\caption*{\textit{Note.} Against = Protesters oppose defunding the police.
+For = Protesters support defunding the police.
+$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
+\label{fig:s4}
+\end{figure*}
+
+Figure \ref{fig:s3} shows results from three preregistered models we
+estimated to test our hypotheses.
+
+Model 1 estimated varying intercepts for the eight conditions to test
+whether, in line with Hypotheses 1a, 1b, or 2a, participants' responses
+depended on their own group membership, the protesters' group
+membership, or the protesters' cause. Contradicting \emph{Hypothesis
+1a}, participants did not consider protest actions performed by their
+ingroup to be more acceptable, on average, than the same actions
+performed by the relevant outgroup (\(d = 0.01, [-0.03, 0.06]\)).
+Contradicting \emph{Hypothesis 1b}, participants did not consider
+protest actions for a cause that was nominally aligned with their
+ingroup's interests to be more acceptable, on average, than protest
+actions for a cause not aligned with their ingroup's interests
+(\(d = -0.01, [-0.06, 0.03]\)). Contradicting \emph{Hypothesis 2a},
+participants did not consider protest actions for a system-defending
+cause to be more acceptable, on average, than protest actions for a
+system-challenging cause (\(d = -0.14, [-0.19, -0.08]\)). Instead, we
+found that, on average, White participants considered all protest
+actions to be more acceptable than Black participants
+(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.12, [0.05, 0.19]\))
+and White (\(d = 0.15, [0.08, 0.22]\)) participants considered the same
+actions to be more acceptable when protests were for, rather than
+against, defunding the police. As in Experiment 1, we thus found that
+participants' responses depended on both the participants' and the
+protesters' group memberships---but not in the directions predicted by
+our hypotheses.
+
+Model 2 extended Model 1 by estimating participants' responses as a
+function of their \emph{z}-standardized endorsement of system-justifying
+beliefs. As preregistered, we modeled this relationship with two fixed
+effects, one estimating the effect of system-justifying beliefs on
+judgments about system-defending protest actions and one estimating
+their effect on judgments about system-challenging protest actions, and
+one varying effect estimating its variance across conditions. Supporting
+\emph{Hypothesis 2b}, participants who \emph{rejected} system-justifying
+beliefs were \emph{more} likely to consider system-challenging protest
+actions (for defunding the police) to be acceptable means of protest
+(\(\beta_{xy} = 0.40, [0.20, 0.57]\)). We did not, however, find
+evidence for the ideological symmetry implied by \emph{Hypothesis 2b}
+since participants who \emph{endorsed} system-justifying beliefs were
+\emph{not} more likely to consider system-defending protest actions
+(against defunding the police) to be acceptable means of protest
+(\(\beta_{xy} = -0.17, [-0.37, -0.01]\)).
+
+Figure \ref{fig:s4} shows the estimated pattern of condition-wise
+differences underlying those fixed effects. Participants who endorsed
+system-justifying beliefs tended to consider system-challenging and
+system-defending protest actions to be equally unacceptable. In
+contrast, participants who rejected system-justifying beliefs evinced
+ideology-based double standards: In line with \emph{Hypothesis 2b}, they
+considered the same protest actions to be more acceptable when the
+protesters challenged the system or, to a lesser extent, when the
+protesters were from the disadvantaged group.
+
+Model 3 extended Model 1 by estimating participants' responses as a
+function of their self-reported support for the cause of the protest. We
+recoded participants' responses to create a predictor variable that
+encoded support for defunding the police when protesters supported
+defunding the police and opposition to defunding the police when
+protesters opposed defunding the police. As preregistered, we modeled
+this relationship as a monotonic effect (Bürkner \& Charpentier, 2020)
+that estimated the average change in the outcome variable across
+predictor categories as well as how much of this change occurred between
+each of the four pairs of adjacent predictor categories.\footnote{Our
+  model assigned a Dirichlet prior, \(\alpha = {1, 1, 1, 1}\), to the
+  proportions of the overall change that was expected to occur between
+  each of the four pairs of predictor categories.} Contradicting
+\emph{Hypothesis 3}, participants who supported the protesters' cause
+did not, on average, consider the same protest actions to be more
+acceptable than participants who opposed the protesters' cause
+(\(\beta_{y} = 0.14, [-0.10, 0.39]\)). Ergo, our results did not support
+the alternative hypothesis that ideology-based double standards can be
+reduced to support for, or opposition to, the cause of a protest.
+
+\hypertarget{non-preregistered-analyses}{%
+\subsubsection{Non-Preregistered
+Analyses}\label{non-preregistered-analyses}}
+
+We explored whether group identification moderated how the participants'
+and the protesters' group memberships affected participants' responses.
+To that end, we extended Model 1 by estimating participants' responses
+as a function of their \emph{z}-standardized identification with their
+racial ingroup. As in Model 2, we modeled this relationship with a fixed
+and a varying effect. We found, however, that even participants who
+strongly identified with their racial ingroup (+1\emph{SD}) did not
+consider protest actions performed by their ingroup to be more
+acceptable than the same actions performed by the relevant outgroup
+(\(d = 0.06, [-0.01, 0.13]\)) and did not consider protest actions for a
+cause that was aligned with their ingroup's interests to be more
+acceptable than protest actions for a cause not aligned with their
+ingroup's interests (\(d = -0.00, [-0.07, 0.07]\)). Our
+non-preregistered analyses thus suggested that group identification did
+not moderate group differences in judgments about collective action by
+ingroup and outgroup members.
+
+We also explored whether we would find ideology-based double standards
+when operationalizing ideology as political orientation instead of as
+system justification. To that end, we reran Model 2 with political
+orientation as the \emph{z}-standardized predictor variable. Our results
+mirrored the preregistered analyses: More liberal participants were more
+likely to consider protest actions for defunding the police to be
+acceptable (\(\beta_{xy} = 0.40, [0.20, 0.58]\)) but more conservative
+participants were not more likely to consider protest actions against
+defunding the police to be acceptable
+(\(\beta_{xy} = -0.16, [-0.35, 0.01]\)). As Figure \ref{fig:s4} shows,
+conservative participants tended to consider all protest actions to be
+equally unacceptable while liberal participants considered the same
+protest actions to be more acceptable when protesters rallied around a
+progressive cause. Our non-preregistered analyses thus replicated the
+ideology-based double standards from the preregistered analyses with a
+different operationalization of ideology.
+
+\hypertarget{experiment-3-1}{%
+\subsection{Experiment 3}\label{experiment-3-1}}
+
+\begin{figure*}[!t]
+\centering
+\caption{Results from the preregistered analyses for Experiment 3 without exclusions}
+\includegraphics[scale=1]{../Experiment 3/figures/figure-s5}
+\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
+\label{fig:s5}
+\end{figure*}
+
+\hypertarget{preregistered-analyses-1}{%
+\subsubsection{Preregistered Analyses}\label{preregistered-analyses-1}}
+
+Figure \ref{fig:s5} shows the results of our preregistered analyses
+without exclusions.
+
+As in Experiments 1 and 2, we found evidence for ideology-based double
+standards: Conservative participants considered the same protest actions
+to be more acceptable, on average, when protesters \emph{supported}
+restricting abortion (\(d = 0.21, [0.03, 0.39]\)) while liberal
+participants considered the same protest actions to be more acceptable
+when protesters \emph{opposed} restricting abortion
+(\(d = 0.87, [0.70, 1.06]\)). This double standard was more pronounced
+among liberal participants (\(d = 0.66, [0.42, 0.93]\)). In this way,
+Experiment 3 provided evidence for both ideological symmetry and
+asymmetry in judging collective action for and against restricting
+abortion.
+
+We found some evidence for identity-based double standards: Female
+participants considered protest actions \emph{against} restricting
+abortion to be more acceptable, on average, than male participants or
+protest actions \emph{for} restricting abortion
+(\(d = 0.14, [-0.01, 0.29]\)). This pattern was, however, overshadowed
+by stronger ideology-based double standards that were consistent across
+subsamples: Conservative women (\(d = 0.24, [-0.01, 0.48]\)) and, to a
+lesser extent, conservative men (\(d = 0.18, [-0.08, 0.44]\)) considered
+the same protest actions to be more acceptable when protesters
+\emph{supported} restricting abortion and both liberal women
+(\(d = 1.00, [0.73, 1.27]\)) and liberal men
+(\(d = 0.75, [0.51, 1.01]\)) considered the same protest actions to be
+more acceptable when protesters \emph{opposed} restricting abortion. In
+this way, Experiment 3 provided more consistent evidence for
+ideology-based than for identity-based double standards in judging
+collective action.
+
+\hypertarget{non-preregistered-analyses-1}{%
+\subsubsection{Non-Preregistered
+Analyses}\label{non-preregistered-analyses-1}}
+
+We explored whether group identification moderated participants'
+judgments. We found, however, that gender identification neither
+affected how women judged protesters opposing
+(\(\beta_{xy} = -0.06, [-0.15, 0.04]\)) or supporting
+(\(\beta_{xy} = -0.07, [-0.16, 0.02]\)) abortion restrictions nor how
+men judged protesters opposing (\(\beta_{xy} = -0.06, [-0.15, 0.03]\))
+or supporting (\(\beta_{xy} = -0.06, [-0.14, 0.03]\)) abortion
+restrictions. Our non-preregistered analyses thus suggested that gender
+identification did not moderate identity-based double standards in
+judging collective action concerning reproductive rights.
+
 \newpage
@@
 \url{https://doi.org/10.18637/jss.v100.i05}
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner_modelling_2020}{}}%
+Bürkner, P.-C., \& Charpentier, E. (2020). Modelling monotonic effects
+of ordinal predictors in {Bayesian} regression models. \emph{British
+Journal of Mathematical and Statistical Psychology}, \emph{73}(3),
+420--451. \url{https://doi.org/10.1111/bmsp.12195}
 

### 2024-04-16 Experiment 1/figures/figure-s2.R

@@ -1,0 +1,243 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("ggtext")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(ggtext)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import results (preregistered analyses)
+  d_prereg <- read_rds("Experiment 1/results/results_pregistered_analyses_without_exclusions.rds")
+  
+  # Import results (exploratory analyses)
+  d_explor <- read_rds("Experiment 1/results/results_exploratory_analyses_without_exclusions.rds")
+  
+
+# Figure 3 (top) ----------------------------------------------------------
+
+  # Make Figure 3a (z)
+  f_3a_z <- ggplot(d_prereg, aes(x = z, y = actor)) +
+    stat_pointinterval(
+      aes(shape = ingroup),
+      .width = 0.95,
+      fill = "white",
+      size = 0.455,
+      point_size = 0.455 * 3
+    ) +
+    scale_x_continuous(minor_breaks = NULL) +
+    scale_shape_manual(values = c("Ingroup" = 21, "Outgroup" = 19)) +
+    coord_cartesian(xlim = c(-0.5, 0.5)) +
+    facet_grid(observer ~ .) +
+    theme_grey(base_size = 10) +
+    labs(
+      tag = "A",
+      x = expression(theta[(italic(z))]),
+      y = "Protesters"
+    ) +
+    theme(
+      legend.position = "none",
+      # legend.position = c(0.95, 0.95),
+      # legend.justification = c(1, 1),
+      # legend.title = element_blank(),
+      # legend.spacing = unit(0, "pt"),
+      strip.background = element_blank(),
+      # strip.text = element_blank(),
+      strip.placement = "outside",
+      axis.line.x = element_line(size = rel(0.5)),
+      axis.title = element_text(size = rel(0.8)),
+      axis.text = element_text(colour = "black"),
[…]

### 2024-04-16 Experiment 1/scripts/run_exploratory_analyses_without_exclusions.R

@@ -1,0 +1,242 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 1/data/dl.rds")
+  dw <- read_rds("Experiment 1/data/dw.rds")
+  ds <- read_rds("Experiment 1/data/ds.rds")
+  
+  # Exclude participants
+  # dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  # dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Remove duplicates
+  dw <- dw |> group_by(id) |> filter(n() == 1) |> ungroup()
+  dl <- dl |> filter(id %in% dw$id)
+  ds <- ds |> filter(id %in% dw$id)
+  
+  # Code predictors (preregistered analyses)
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      x_1 = if_else(target == "PJ", 1L, 0L),
+      x_2 = if_else(target == sample, 1L, 0L),
+      x_3 = if_else(target == "PJ" & sample == "PJ", 1L, 0L),
+      y  = response
+    )
+  
+  # Code predictors (exploratory analyses)
+  dw <- dw %>% 
+    left_join(
+      ds %>% select(id, political_orientation = q10),
+      by = "id"
+    ) %>% 
+    rename_with(~str_replace(., "q11", "sdo")) %>% 
+    rename_with(~str_replace(., "q12", "sjb")) %>% 
[…]

### 2024-04-16 Experiment 1/scripts/run_factor_analyses.R

@@ -49,3 +49,4 @@
 # Remove ------------------------------------------------------------------
   detach(package:lavaan)
   rm("sdo_model", "sdo_fit", "sjb_model", "sjb_fit")
+  

### 2024-04-16 Experiment 1/scripts/run_preregistered_analyses_without_exclusions.R

@@ -1,0 +1,175 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+
+  # Load packages
+  library(tidyverse)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 1/data/dl.rds")
+  
+  # Exclude participants
+  # dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  
+  # Code predictors
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      x_1 = if_else(target == "PJ", 1L, 0L),
+      x_2 = if_else(target == sample, 1L, 0L),
+      x_3 = if_else(target == "PJ" & sample == "PJ", 1L, 0L),
+      y  = response
+    )
+  
+  # Compile data list
+  dlist <- with(dl, list(
+    N = length(ii),
+    I = max(ii),
+    J = max(jj),
+    K = 3L,
+    ii = ii,
+    jj = jj,
+    X = cbind(x_1, x_2, x_3),
+    y = y
+  ))
+
+
+# Estimate ----------------------------------------------------------------
+
+  # Compile model
+  model <- cmdstan_model("Experiment 1/models/s1_2pl_model.stan")
+
+  # Run model
+  fit <- model$sample(
+    data = dlist,
+    seed = 36939726,
+    chains = 8,
+    parallel_chains = parallel::detectCores(),
+    iter_sampling = 1000,
+    iter_warmup = 1000
+  )
+
+  # Save results as .rds
+  fit$save_object("Experiment 1/results/fit_pregistered_analyses_without_exclusions.rds")
+
+
[…]

### 2024-04-16 Experiment 2/figures/figure-s3.R

@@ -1,0 +1,438 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  # dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  # dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = case_when(
+        protest == "for" ~ q5,
+        protest == "against" ~ 6L - q5
+      )
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
[…]

### 2024-04-16 Experiment 2/figures/figure-s4.R

@@ -1,0 +1,289 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("tidybayes")
+  # install.packages("patchwork")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  # dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  # dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = (q5 - mean(q5))/sd(q5),
+      x_pol = (q19_1 - mean(q19_1))/sd(q19_1)
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
+        observers == "White" & protesters == "Black" & protest == "for"     ~ 5L, 
[…]

### 2024-04-16 Experiment 2/scripts/run_exploratory_analyses_without_exclusions.R

@@ -1,0 +1,253 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+  
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  # dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  # dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = (q5 - mean(q5))/sd(q5),
+      x_pol = (q19_1 - mean(q19_1))/sd(q19_1),
+      x_ide = (q16 - mean(q16, na.rm = T))/sd(q16, na.rm = T)
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
[…]

### 2024-04-16 Experiment 2/scripts/run_preregistered_analyses_without_exclusions.R

@@ -1,0 +1,270 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+  
+
+# Prepare -----------------------------------------------------------------
+  
+  # Import data
+  dl <- read_rds("Experiment 2/data/dl.rds")
+  dw <- read_rds("Experiment 2/data/dw.rds")
+  
+  # Exclude participants
+  # dl <- dl %>% filter(!exclude) %>% select(-exclude)
+  # dw <- dw %>% filter(!exclude) %>% select(-exclude)
+  
+  # Calculate factor scores (system-justifying beliefs)
+  source("Experiment 2/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    transmute(
+      id,
+      x_sjb = (sjb - mean(sjb))/sd(sjb),
+      x_sup = case_when(
+        protest == "for" ~ q5,
+        protest == "against" ~ 6L - q5
+      )
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        observers == "Black" & protesters == "Black" & protest == "for"     ~ 1L, 
+        observers == "Black" & protesters == "Black" & protest == "against" ~ 2L, 
+        observers == "Black" & protesters == "White" & protest == "for"     ~ 3L, 
+        observers == "Black" & protesters == "White" & protest == "against" ~ 4L, 
+        observers == "White" & protesters == "Black" & protest == "for"     ~ 5L, 
[…]

### 2024-04-16 Experiment 3/figures/figure-s5.R

@@ -1,0 +1,129 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+
+  # Load packages
+  library(tidyverse)
+  library(patchwork)
+  library(cmdstanr)
+  library(posterior)
+  library(tidybayes)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Load results
+  m1_post <- read_rds("Experiment 3/results/m1_post_without_exclusions.rds")
+  
+  # Prepare data for figure
+  d_fig <- m1_post %>% mutate(order = kk)
+  
+
+# Figure 6 ----------------------------------------------------------------
+
+  # Make table
+  f_6_t <- d_fig %>% 
+    distinct(order, participant_ideology, participant_gender, protesters_cause) %>% 
+    rename(Gender = participant_gender, Ideology = participant_ideology, Cause = protesters_cause) %>% 
+    pivot_longer(
+      c(Gender, Ideology, Cause),
+      names_to = "column",
+      values_to = "text"
+    ) %>% 
+    mutate(
+      column = factor(column, levels = c("Ideology", "Gender", "Cause"))
+    ) %>% 
+    ggplot(., aes(x = 0, y = order)) +
+    geom_text(
+      aes(label = text),
+      size = 10/.pt,
+      hjust = 0
+    ) +
+    geom_hline(yintercept = c(0.5, 8.5), size = 0.455) +
+    scale_y_reverse(expand = c(0, 0)) +
+    xlim(0, 1) +
+    facet_grid(. ~ column) +
+    theme_grey(base_size = 10) +
+    theme(
+      legend.position = "none",
+      strip.background = element_blank(),
+      strip.text.x = element_text(size = 10, colour = "black", face = "bold"),
+      strip.text.y = element_blank(),
+      axis.text.x = element_text(colour = NA),
+      axis.ticks.x = element_line(colour = NA),
+      axis.title.x = element_text(colour = NA),
+      axis.text.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      axis.title.y = element_blank(),
+      panel.background = element_blank(),
+      plot.margin = margin(2, 1, 2, 0)
+    )
+  
+  # Make plot
+  f_6_p <- d_fig %>% 
[…]

### 2024-04-16 Experiment 3/scripts/run_exploratory_analyses.R

@@ -9,7 +9,7 @@ rm(list = ls())
   # install.packages("tidyverse")
   # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
   # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
-  # install_cmdstan() 
+  # install_cmdstan()
   
   # Load packages
   library(tidyverse)
@@ -49,7 +49,7 @@ rm(list = ls())
     transmute(
       id,
       across(
-        c(ide, sjb),
+        c(ide, sjb, mfq_care, mfq_equa, mfq_prop, mfq_loya, mfq_auth, mfq_puri),
         ~(. - mean(., na.rm = T))/sd(., na.rm = T),
         .names = "z_{.col}"
       )
@@ -84,7 +84,13 @@ rm(list = ls())
     y = y,
     x_for = c(1L, 1L, 0L, 0L, 1L, 1L, 0L, 0L),
     x_sjb = z_sjb,
-    x_ide = z_ide
+    x_ide = z_ide,
+    x_mfq_care = z_mfq_care,
+    x_mfq_equa = z_mfq_equa,
+    x_mfq_prop = z_mfq_prop,
+    x_mfq_loya = z_mfq_loya,
+    x_mfq_auth = z_mfq_auth,
+    x_mfq_puri = z_mfq_puri
   ))
   
 
@@ -111,65 +117,65 @@ rm(list = ls())
 
   # Import results
   ide_fit <- read_rds("Experiment 3/results/ide_fit.rds")
-  
+
   # Extract draws
   ide_ps <- ide_fit$draws() %>% as_draws_df()
-  
+
   # Extract posterior draws
   ide_post_ii <- ide_ps %>% spread_draws(alpha[ii], beta[ii]) %>% ungroup()
   ide_post_jj <- ide_ps %>% spread_draws(theta[jj]) %>% ungroup()
   ide_post_kk <- ide_ps %>% spread_draws(delta[kk], b_ide_kk[kk], b_ide) %>% ungroup()
-  
+
   # Calculate effect sizes
-  ide_post <- dl %>% 
-    distinct(jj, kk, z_ide) %>% 
-    left_join(ide_post_kk, by = c("kk")) %>% 
-    left_join(ide_post_jj, by = c("jj", ".draw", ".chain", ".iteration")) %>% 
-    group_by(.draw) %>% 
+  ide_post <- dl %>%
+    distinct(jj, kk, z_ide) %>%
[…]

### 2024-04-16 Experiment 3/scripts/run_exploratory_analyses_without_exclusions.R

@@ -1,0 +1,217 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+  
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 3/data/dl.rds")
+  dw <- read_rds("Experiment 3/data/dw.rds")
+  
+  # Exclude participants
+  dl <- dl %>% 
+    filter(participant_gender %in% c("Man", "Woman"))
+  dw <- dw %>% 
+    filter(participant_gender %in% c("Man", "Woman"))
+  
+  # Calculate factor scores
+  source("Experiment 3/scripts/run_factor_analyses.R")
+  
+  # Code predictors
+  dl <- dw %>%
+    mutate(
+      ide = case_when(
+        participant_gender == "Woman" ~ ide_women,
+        participant_gender == "Man" ~ ide_men
+      )
+    ) %>% 
+    transmute(
+      id,
+      across(
+        c(ide, sjb),
+        ~(. - mean(., na.rm = T))/sd(., na.rm = T),
+        .names = "z_{.col}"
+      )
+    ) %>% 
+    left_join(dl, ., by = "id") %>% 
+    drop_na() %>% 
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Man"   & protesters_cause == "For"     ~ 1L, 
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For"     ~ 2L, 
[…]

### 2024-04-16 Experiment 3/scripts/run_factor_analyses.R

@@ -21,8 +21,36 @@ library(lavaan)
   dw <- predict(sjb_fit) %>%
     as.data.frame() %>%
     bind_cols(dw, .)
+  
+  
+
+# Moral Foundations Questionnaire (MFQ-2) ---------------------------------
+
+  # Load model
+  mfq_model <- '
+        mfq_care =~ mfq_care_1 + mfq_care_2 + mfq_care_3 + mfq_care_4 + mfq_care_5 + mfq_care_6 
+        mfq_equa =~ mfq_equa_1 + mfq_equa_2 + mfq_equa_3 + mfq_equa_4 + mfq_equa_5 + mfq_equa_6 
+        mfq_prop =~ mfq_prop_1 + mfq_prop_2 + mfq_prop_3 + mfq_prop_4 + mfq_prop_5 + mfq_prop_6 
+        mfq_loya =~ mfq_loya_1 + mfq_loya_2 + mfq_loya_3 + mfq_loya_4 + mfq_loya_5 + mfq_loya_6 
+        mfq_auth =~ mfq_auth_1 + mfq_auth_2 + mfq_auth_3 + mfq_auth_4 + mfq_auth_5 + mfq_auth_6 
+        mfq_puri =~ mfq_puri_1 + mfq_puri_2 + mfq_puri_3 + mfq_puri_4 + mfq_puri_5 + mfq_puri_6 
+      '
+  
+  # Run models
+  mfq_fit <- cfa(
+    mfq_model,
+    estimator = "mlr",
+    missing = "fiml",
+    std.lv = TRUE,
+    data = dw
+  )
+  
+  # Export factor scores
+  dw <- predict(mfq_fit) %>%
+    as.data.frame() %>%
+    bind_cols(dw, .)
 
 
 # Remove ------------------------------------------------------------------
 detach(package:lavaan)
-rm("sjb_model", "sjb_fit")
+rm("sjb_model", "sjb_fit", "mfq_model", "mfq_fit")

### 2024-04-16 Experiment 3/scripts/run_preregistered_analyses_without_exclusions.R

@@ -1,0 +1,132 @@
+rm(list = ls())
+
+# Notes -------------------------------------------------------------------
+
+
+# Library -----------------------------------------------------------------
+  
+  # Install packages
+  # install.packages("tidyverse")
+  # install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install.packages("posterior", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
+  # install_cmdstan() 
+  
+  # Load packages
+  library(tidyverse)
+  library(tidybayes)
+  library(cmdstanr)
+  library(posterior)
+  
+  # Functions
+  inv_logit <- function(a) exp(a) / ( 1 + exp(a) )
+
+
+# Prepare -----------------------------------------------------------------
+
+  # Import data
+  dl <- read_rds("Experiment 3/data/dl.rds")
+  
+  # Exclude participants
+  dl <- dl %>% 
+    filter(participant_gender %in% c("Man", "Woman"))
+  
+  # Code predictors
+  dl <- dl %>%
+    mutate(
+      ii = action,
+      jj = as.integer(factor(id)),
+      kk = case_when(
+        participant_ideology == "Conservative" & participant_gender == "Man"   & protesters_cause == "For"     ~ 1L, 
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For"     ~ 2L, 
+        participant_ideology == "Conservative" & participant_gender == "Man"   & protesters_cause == "Against" ~ 3L, 
+        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "Against" ~ 4L, 
+        participant_ideology == "Liberal"      & participant_gender == "Man"   & protesters_cause == "For"     ~ 5L, 
+        participant_ideology == "Liberal"      & participant_gender == "Woman" & protesters_cause == "For"     ~ 6L, 
+        participant_ideology == "Liberal"      & participant_gender == "Man"   & protesters_cause == "Against" ~ 7L, 
+        participant_ideology == "Liberal"      & participant_gender == "Woman" & protesters_cause == "Against" ~ 8L
+      ),
+      y  = response
+    )
+  
[…]

### 2024-04-16 Scale Development/figures/figure-1.R

@@ -30,8 +30,8 @@ rm(list = ls())
   
   # Make figure
   ggplot() +
-    geom_vline(xintercept = 0.0, colour = "grey92", size = 0.455) +
-    geom_hline(yintercept = 0.5, colour = "grey92", size = 0.455) +
+    geom_vline(xintercept = 0.0, colour = "grey92", linewidth = 0.455) +
+    geom_hline(yintercept = 0.5, colour = "grey92", linewidth = 0.455) +
     stat_function(
       aes(colour = "Action 1"),
       fun = irt, 
@@ -80,7 +80,7 @@ rm(list = ls())
     annotate(
       geom = "rich_text",
       x = qnorm(0.995), y = 0,
-      label = "<span style='color:#648FFF'>&alpha;<sub>1</sub> = 1.00, &beta;<sub>1</sub> = 0.00</span><br><span style='color:#DC267F'>&alpha;<sub>2</sub> = 1.00, &beta;<sub>2</sub> = 1.28</span><br><span style='color:#FFB000'>&alpha;<sub>3</sub> = 3.60, &beta;<sub>3</sub> = 0.00</span>",
+      label = "<span style='color:#648FFF'> &alpha;<sub>1</sub> = 1.00, &beta;<sub>1</sub> = 0.00</span><br><span style='color:#DC267F'>&alpha;<sub>2</sub> = 1.00, &beta;<sub>2</sub> = 1.28</span><br><span style='color:#FFB000'>&alpha;<sub>3</sub> = 3.60, &beta;<sub>3</sub> = 0.00</span>",
       vjust = 0, hjust = 1,
       label.colour = NA,
       label.padding = unit(0, "lines"),
@@ -129,7 +129,7 @@ rm(list = ls())
   ggsave(
     "Scale Development/figures/figure-1.png",
     width = (6.5 - 18/72.27)/2, height = (6.5 - 18/72.27)/2, units = "in", 
-    dpi = 600
+    dpi = 1200
   )
   
   # Export as .pdf

### 2024-04-16 Scale Development/scripts/run_analyses.R

@@ -86,9 +86,9 @@ rm(list = ls())
   
   # Save results as .rds
   fit$save_object("Scale Development/results/fit_scale_development.rds")
-  
 
-# Extract -----------------------------------------------------------------
+
+# Process -----------------------------------------------------------------
 
   # Load results
   fit <- read_rds("Scale Development/results/fit_scale_development.rds")
@@ -144,15 +144,14 @@ rm(list = ls())
   # Total Information Function
   I <- function(theta, ii = 1:72, standardized = FALSE) {
     if (standardized == TRUE) theta <- theta * m_theta$sd + m_theta$mean
-    sum(map_dbl(ii, ~I_i(theta, .)))
+    map_dbl(theta, \(theta) sum(map_dbl(ii, \(ii) I_i(theta, ii))))
   }
   
   # Vectorize
   p_ik <- Vectorize(p_ik)
   I_ik <- Vectorize(I_ik)
   I_i  <- Vectorize(I_i)
-  I    <- Vectorize(I)
-  
+
   # Summarize
   results <- post %>%
     mutate(
@@ -176,5 +175,68 @@ rm(list = ls())
 
 # Export ------------------------------------------------------------------
 
-  # Export as .rds
+  # Save as .rds
   write_rds(results, "Scale Development/results/results_scale_development.rds")
+
+  
+# Show --------------------------------------------------------------------
+
+  # Plot Total Information Curves
+  p <- ggplot()
+  for (n_i in c(1:24, 24:72)) {
+    p <- p + stat_function(
+      fun = I,
+      args = list(ii = pull(slice_max(results, n = n_i, I), ii), standardized = T),
+      color = grey.colors(72)[73 - n_i],
+      linewidth = 0.455
+    )
+  }
+  p <- p + stat_function(
+    fun = I,
+    args = list(ii = pull(slice_max(results, n = 25, I), ii), standardized = T),
+    color = "red",
+    linewidth = 0.455 * 2
+  ) 
+  p + scale_x_continuous(
+    limits = c(-5, 5),
+    breaks = qnorm(c(0.005, 0.1, 0.5, 0.9, 0.995)),
+    minor_breaks = NULL,
+    labels = scales::number(qnorm(c(0.005, 0.1, 0.5, 0.9, 0.995)), accuracy = 0.01),
+    sec.axis = dup_axis(
+      name = NULL,
+      breaks = qnorm(c(0.005, 0.1, 0.5, 0.9, 0.995)),
[…]

### 2024-11-05 Documents/manuscript.qmd
2nd revision

@@
     Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change.
-Five studies (total $N = 2,979$) investigated double standards in judging collective action---that is, whether observers judge the same protest actions to be more acceptable depending on who the protesters are and what they are protesting.
-Two studies used item response theory to develop an instrument of 25 controversial protest actions to measure where people draw the line between acceptable and unacceptable forms of collective action.
-Three preregistered experiments found no consistent evidence for ingroup bias in terms of social class when judging protests for workers' rights (Experiment 1), in terms of race when judging protests for and against defunding the police (Experiment 2), and in terms of gender when judging protests for and against restricting abortion (Experiment 3).
-Instead, we found that participants who identified as left-wing/liberal (Experiments 1--3) and rejected system-justifying beliefs (Experiments 1--2) considered the same protest actions more acceptable when a cause aligned with their ideological orientation (for workers' rights, for defunding the police, against restricting abortion) than when it did not (against defunding the police, for restricting abortion).
+We investigated double standards in judging collective action---that is, whether observers judge the same protest actions to be more acceptable depending on who the protesters are and what they are protesting.
+In two studies, we used item response theory to develop an instrument of 25 controversial protest actions to measure where people draw the line between acceptable and unacceptable forms of collective action.
+In three preregistered experiments ($N = 2776$), we found no consistent evidence for ingroup bias in terms of social class when judging protests for workers' rights (Experiment 1), in terms of race when judging protests for and against defunding the police (Experiment 2), and in terms of gender when judging protests for and against restricting abortion (Experiment 3).
+Instead, we found that progressive participants (Experiments 1--3) who rejected system-justifying beliefs (Experiments 1--2) considered the same protest actions more acceptable when a cause aligned with their ideological orientation (for workers' rights, for defunding the police, against restricting abortion) than when it did not (against defunding the police, for restricting abortion).
 Conservative participants considered the same actions somewhat more acceptable when protesters supported, rather than opposed, restricting abortion (Experiment 3) but considered all protest actions, for and against defunding the police, equally unacceptable (Experiment 2).
@@
 Progressives, in turn, tend to consider disruptive but non-violent actions acceptable means to achieve social change and challenge social injustice.
-This finding helps explain the often-divided response to recent protest movements.
-word-count: "11,654 words"
+These findings help explain the often-divided response to recent protest movements.
+word-count: "12,257 words"
 bibliography: references.bib
@@
 format:
-  docx:
-    reference-doc: word-template.docx
+  # docx:
+  #   reference-doc: word-template.docx
   pdf:
@@
 
-Another explanation for ideology-based double standards is that they are rooted in ideological differences in how conservatives and progressives relate to social change and social inequality which, in turn, are motivated by epistemic, existential, and relational needs [@jost_political_2003; @jost_left_2021].
+Another explanation for ideology-based double standards is that they are rooted in ideological differences in how conservatives and progressives relate to social change and social inequality which, in turn, are motivated by basic psychological needs [@jost_political_2003; @jost_left_2021].
 As motivated social cognition, ideological differences shape how we think and feel about political issues, actions, and events.
@@
 Item response theory is a conceptual and statistical framework for understanding how the characteristics of both items and respondents shape responses to a set of items [@demars_irt_2010].
-<!-- As such, it allows us to disentangle how acceptable protest actions are from how accepting respondents are of them and, in so doing, to detect double standards in judging collective action.
---> When respondents judge whether each of a set of actions is an acceptable means of protest (1 = *yes*, 0 = *no*), a two-parameter logistic item response theory model estimates responses as a function of several latent (unobserved) parameters:
+Applied to judging protest actions, this framework allows us to disentangle how *acceptable* different protest actions are from how *accepting* different respondents are of them.
+In line with the long-standing distinction between normative and non-normative collective action [@wright_responding_1990], it recognizes that some protest actions (e.g., signing petitions) are more acceptable than others (e.g., blocking roads).
+In line with the hypothesized double standards, it recognizes that some people consider more protest actions acceptable than others.
+By comparing how accepting respondents are of the *same* protest actions in *different* situations, we can use item response theory to detect double standards in judging collective action.
+
+In this research, respondents judge whether each of a set of more and less controversial protest actions is an acceptable means of protest (1 = *yes*, 0 = *no*).
+We use two-parameter item response theory models to estimate the probability $\Pr ( y_\textit{ij} = 1 )$ that respondent $j$ considers action $i$ an acceptable means of protest as a function of three latent model parameters:
 $$ \Pr ( y_\textit{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) $$
-where $\Pr ( y_\textit{ij} = 1 )$ is the probability that respondent $j$ considers action $i$ an acceptable means of protest, $\theta_j$ estimates how *accepting* respondent $j$ is of various protest actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.
-
-Figure \ref{fig:f1} illustrates those relationships for three hypothetical protest actions.
-$\theta_{(z)}$ is the *z*-standardized tendency for a participant to be more accepting of various actions so that $\beta_i$ determines the probability that the average respondent ($\theta_{(z)} = 0$) considers an action an acceptable means of protest.
-Action 2 ($\beta_2 = 1.28$) is more acceptable than Actions 1 and 2 ($\{ \beta_2, \beta_3 \} = 0$) as the average respondent is more likely to consider the former ($\Pr = .78$) than the latter ($\Pr = .50$) acceptable means of protest.
-Action 3 ($\alpha_3 = 3.60$) is more discriminating than Actions 1 and 2 ($\{ \alpha_1, \alpha_2 \} = 1$) as the slope for the relationship between how accepting a respondent is and how likely they are to consider an action acceptable is steeper for the former action.
-This means that knowing whether someone considers Action 3 acceptable provides more information about them than knowing whether they consider Actions 1 or 2 acceptable.
+where $\theta_j$ estimates how *accepting* respondent $j$ is of various actions, $\beta_i$ estimates how *acceptable* action $i$ is, and $\alpha_i$ estimates how *discriminating* action $i$ is.
 
@@
 \includegraphics[scale=1]{../Scale Development/figures/figure-1.png}
-\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
+\caption*{\textit{Note.} $\Pr ( y_\textit{ij} = 1 )$ is the probability that that respondent $j$ considers action $i$ an acceptable means of protest.
+$\theta_{(z)}$ is the $z$-standardized tendency for a participant to be more accepting of various actions.
+It follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).
+$\beta_i$ and $\alpha_i$ capture, respectively, how acceptable and discriminating an action is.}
 \label{fig:f1}
@@
 ```
+
+Figure \ref{fig:f1} illustrates those relationships for three hypothetical protest actions.
+$\theta_{(z)}$ is the *z*-standardized tendency for a participant to be more accepting of various actions.
+It follows a normal distribution so that $\theta_{(z)} = 1.28$ means that a respondent is more accepting than 90\% of respondents.
+$\beta_i$ determines the probability that the average respondent ($\theta_{(z)} = 0$) considers an action an acceptable means of protest.
+Action 2 ($\beta_2 = 1.28$) is more acceptable than Actions 1 and 3 ($\{ \beta_1, \beta_3 \} = 0$) as the average respondent is more likely to consider the former ($\Pr = .78$) than the latter ($\Pr = .50$) acceptable means of protest.
+$\alpha_i$ determines how much the probability that a respondent considers an action acceptable depends on how accepting that respondent is.
+Action 3 ($\alpha_3 = 3.60$) is more discriminating than Actions 1 and 2 ($\{ \alpha_1, \alpha_2 \} = 1$) as the slope for the relationship between how accepting a respondent is and how likely they are to consider an action acceptable is steeper for the former action.
+This means that knowing whether someone considers Action 3 acceptable provides more information about them than knowing whether they consider Actions 1 or 2 acceptable.
 
@@
 In Experiments 1--3, we operationalize double standards as differences in $\theta_{(z)}$ which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
+As $\theta_{(z)}$ determines the probability of respondents judging different protest actions acceptable, it can be converted to the predicted proportion $p$ and number $n$ of actions a respondent considers acceptable.
+Figure \ref{fig:f3} (A) illustrates these relationships.
+Figures \ref{fig:f3} (B), \ref{fig:f4}, \ref{fig:f5}, and \ref{fig:f6} show the predicted proportion of actions a participant would consider acceptable as an intuitive and easily interpretable effect size.
 
@@
 
-In Study 1, we compiled protest actions from participants' responses and other sources.
-We recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US.
-To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students.
-Participants first read an accessible definition of what a social group is and that collective action is any action group members take to promote the interests of their social group.
-
-> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong.
-Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club, has a religion, or belongs to any number of other such groups.
-Individuals often act in ways to promote the interests of the social groups to which they belong.
-
-\noindent Participants were then asked to name between five and ten actions that fit that definition.
-Participants were encouraged to think of actions that varied in how acceptable they were in their opinion.
-We recoded responses into a smaller set of unique actions, which we supplemented with protest actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973].
-This process resulted in 72 actions that varied in how acceptable we would expect them to be.
-For details, see Supplemental Online Material (SOM).
-
-In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action.
-We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US.
-To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students.
-We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.
-
 ```{=latex}
@@
 
+In Study 1, we compiled protest actions from participants' responses and other sources.
+We recruited 60 participants from the Prolific subject pool, all of whom were citizens of the UK or the US.
+To increase the socioeconomic diversity of our sample, we recruited 30 non-students without a university degree, 15 non-students with a university degree, and 15 current university students.
+Participants first read an accessible definition of what a social group is and that collective action is any action group members take to promote the interests of their social group.
+
+> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong.
+Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club, has a religion, or belongs to any number of other such groups.
+Individuals often act in ways to promote the interests of the social groups to which they belong.
+
+\noindent Participants were then asked to name between five and ten actions that fit that definition.
+Participants were encouraged to think of actions that varied in how acceptable they were in their opinion.
+We recoded responses into a smaller set of unique actions, which we supplemented with protest actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973].
+This process resulted in 72 actions that varied in how acceptable we would expect them to be.
+For details, see Supplemental Online Material A (SOM-A).
+
+In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action.
+We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US.
+To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students.
+We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.
+
 Participants again read an accessible definition of collective action.
@@
 We selected controversial, and thus diagnostic, actions (e.g., disrupting traffic) over actions that almost all respondents considered acceptable (e.g., signing petitions) or unacceptable (e.g., political violence; Figure \ref{fig:f2}).
-For detailed ratings and results, see SOM.
+For detailed ratings and results, see SOM-A and Tables S1--S3.
 
@@
 By comparing reactions to working-class and middle-class protesters, we further tested how observers judge collective action by lower-status (i.e.
-working-class) and higher-status (middle-class) group members.^[For Experiments 1 and 2, we had preregistered the additional hypotheses that all participants would  judge  the  same actions  as  more  acceptable when the protesters were from the higher-status group rather than the lower-status group (Experiment 1) and when the goal of the protest was to support rather than challenge the system (Experiment 2).
-These hypotheses reflected the assumption that the motivation to justify the system was, to some extent, universal.
-With hindsight, we no longer think that this assumption is tenable and, therefore, omit the hypotheses based on it from the revised manuscript.
-We still report the same preregistered analyses and their results.] In non-preregistered analyses, we tested whether participants would judge the same protest actions as more or less acceptable depending on their own ideological positions in terms of political orientation, system-justifying beliefs, and social dominance orientation.^[In Experiment 1, we explored a broader range of operationalizations of political ideology, including social dominance orientation, than in the other experiments.] In this way, Experiment 1 provided a first test of the hypothesized class- and ideology-based double standards in judging collective action.
+working-class) and higher-status (middle-class) group members.
+In non-preregistered analyses, we tested whether participants would judge the same protest actions as more or less acceptable depending on their own ideological positions in terms of political orientation, system-justifying beliefs, and social dominance orientation.^[In Experiment 1, we explored a broader range of operationalizations of political ideology, including social dominance orientation, than in the other experiments.] In this way, Experiment 1 provided a first test of the hypothesized class- and ideology-based double standards in judging collective action.
 
@@
 We preregistered the sample size as well as all hypotheses, inclusion/exclusion criteria, measures, and manipulations ([https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b](https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b)).
+For Experiments 1 and 2, we had preregistered the additional hypothesis that all participants would judge  the  same actions  to be more  acceptable when the protesters were from the higher-status group rather than the lower-status group (Experiment 1) and when the goal of the protest was to support rather than challenge the system (Experiment 2).
+This hypothesis reflected the assumption that the motivation to justify the system was, to some extent, universal.
+With hindsight, we no longer think that this assumption is tenable and, therefore, do not discuss the hypothesis based on it.
+We still follow the same preregistered analysis plan and report all results related to our original hypotheses.
+
 We made all materials, data, and analysis scripts available online ([https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)).
@@
 
-In the screening survey, participants read an accessible definition of what working-class and middle-class/professional jobs are (for details, see SOM).
+In the screening survey, participants read an accessible definition of what working-class and middle-class/professional jobs are (for details, see SOM-B).
 Participants then answered, among other questions, whether they considered their current job---or the jobs they had had in the past or expected to have in the future---to be a working-class job or a middle-class/professional job.
@@
 \caption*{\textit{Note.} HS = Higher Status, LS = Lower Status.
-$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
-(\textbf{A}) $p$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition.
+(\textbf{A}) $\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
+$\Pr$ and $n$ are, respectively, the predicted proportion and number of actions a participant would consider acceptable means of protest in each condition.
 Bars enclose the 95\% most plausible estimates.
@@
 Bayesian inference applies Bayes' theorem to update prior distributions in light of the observed data to produce posterior distributions.
-In contrast to $p$-values and confidence intervals, posterior distributions have a straight-forward interpretation as stating how plausible different values of the model parameters are given the observed data.
+Unlike $p$-values and confidence intervals, posterior distributions have a straight-forward interpretation as stating how plausible different values of the model parameters are given the observed data.
+We report point estimates, based on the median of posterior samples, and uncertainty intervals, based on the quantiles of posterior samples, that enclose the 95\% most plausible estimates.
 
@@
 We used partial pooling to estimate $\alpha_i$, $\beta_i$, and $\theta_j$.
-Our model assigned weakly informative prior distributions [@gelman_prior_2017] to all model parameters.^[Our model assigned $\text{Half-Cauchy} (0, 3)$ prior distributions to the standard deviations of $\alpha_i$, $\beta_i$, and $\theta_j$ and $\text{Normal} (0, 3)$ prior distributions to all other model parameters.] We report point estimates, based on the median of posterior samples, and uncertainty intervals, based on the quantiles of posterior samples, that enclose the 95% most plausible estimates.
+Our model assigned weakly informative prior distributions [@gelman_prior_2017] to all model parameters: $\text{Half-Cauchy} (0, 3)$ prior distributions to the standard deviations of $\alpha_i$, $\beta_i$, and $\theta_j$ and $\text{Normal} (0, 3)$ prior distributions to all other model parameters.
+
+We report differences between conditions in terms of differences in the estimated tendency for a participant to consider more protest actions acceptable ($\theta_{(z)}$) which, as $\theta_{(z)}$ is $z$-standardized, correspond to Cohen's $d$ effect sizes.
+Likewise, we report standardized regression coefficients ($\beta_{xy}$) that correspond to the estimated change in the $z$-standardized tendency to consider more protest actions acceptable for each $+1\textit{SD}$ change in the predictor variable.
+All figures show the predicted proportion of protest actions participants would consider acceptable in each condition based on the estimated models.
 
@@
 
-Figure \ref{fig:f3} shows (B) the $z$-standardized propensity to consider more controversial protest actions acceptable as a function of the three ideological orientation variables.
+Figure \ref{fig:f3} shows (B) how participants' tendency to consider more controversial protest actions acceptable varied as a function of the three ideological orientation variables.
 We found that participants who reported a more right-wing political orientation ($\beta_{xy} = -0.25, [-0.37, -0.15]$) and who expressed more agreement with system-justifying beliefs ($\beta_{xy} = -0.20, [-0.30, -0.10]$) tended to find fewer collective actions to be acceptable.
@@
 For = Protesters support defunding the police.
-$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.
+$\Pr$ is the predicted proportion of actions a participant would consider acceptable means of protest.
 Ribbons enclose, from the darkest to the lightest shade, the 50\%, 80\%, 95\%, and 99\% most plausible estimates.}
@@
 For = Protesters support defunding the police.
-$\theta_{(z)}$ is the \textit{z}-standardized tendency to consider more controversial actions acceptable means of protest.}
+$\Pr$ is the predicted proportion of actions a participant would consider acceptable means of protest in each condition if they were higher ($+1\text{SD}$) or lower ($-1\text{SD}$) than average in system justification (left) or political conservatism (right).}
 \label{fig:f5}
@@
 This resulted in partial pooling---condition-wise estimates were shrunk towards each other---and allowed us to compare conditions without multiple comparison problems [@gelman_why_2012].
-Our models assigned weakly informative prior distributions to all model parameters.^[Our model assigned $\text{LKJ} (2)$ prior distributions to the Cholesky-transformed correlation matrices for varying effects and $(\text{Half-})\text{Cauchy} (0, 5)$ prior distributions to all other model parameters.] Figure \ref{fig:f4} shows results from three preregistered models we estimated to test our hypotheses.
+Our models assigned weakly informative prior distributions: $\text{LKJ} (2)$ prior distributions to the Cholesky-transformed correlation matrices for varying effects and $(\text{Half-})\text{Cauchy} (0, 5)$ prior distributions to all other model parameters.
+Figure \ref{fig:f4} shows results from three preregistered models we estimated to test our hypotheses.
 
@@
 We recoded participants' responses to create a predictor variable that encoded support for defunding the police when protesters supported defunding the police and opposition to defunding the police when protesters opposed defunding the police.
-As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.^[Our model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.] Participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
+As preregistered, we modeled this relationship as a monotonic effect [@burkner_modelling_2020] that estimated the average change in the outcome variable across predictor categories as well as how much of this change occurred between each of the four pairs of adjacent predictor categories.
+This model assigned a Dirichlet prior, $\alpha = {1, 1, 1, 1}$, to the proportions of the overall change that was expected to occur between each of the four pairs of predictor categories.
+We found that participants who supported the protesters' cause did not, on average, consider the same protest actions to be more acceptable than participants who opposed the protesters' cause (`r with(m3_h3, b_text)`).
 Ergo, our results did not support the alternative hypothesis that ideology-based double standards can be reduced to support for, or opposition to, the cause of a protest.
@@
 
-We included additional measures, not used in the analyses reported in this paper, including measures of moral conviction [@ryan_reconsidering_2014], gender-related system-justifying beliefs [@jost_exposure_2005], and the updated moral foundations questionnaire [@atari_morality_2022, for details, see SOM].
+We included additional measures, not used in the analyses reported in this paper, including measures of moral conviction [@ryan_reconsidering_2014], gender-related system-justifying beliefs [@jost_exposure_2005], and the updated moral foundations questionnaire [@atari_morality_2023, for details, see SOM-C].
 Within the latter questionnaire, we embedded three further attention checks (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
@@
 \includegraphics[scale=1]{../Experiment 3/figures/figure-6}
-\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
+\caption*{\textit{Note.} Results show the predicted proportion ($\Pr$) of actions a participant would consider acceptable means of protest for or against restricting abortion as a function of the participants' ideology and gender.}
 \label{fig:f6}
@@
 Instead, we found that members of advantaged groups---middle-class (<!--$d = 0.33$, -->Experiment 1), White (<!--$d = 0.09$, -->Experiment 2), and male (<!--$d = 0.16$,-->Experiment 3) participants---tended to consider all protest actions more acceptable than members of relatively disadvantaged groups.
-That is, our findings showed that participants' judgements of protest actions depended on their group memberships---but not in the directions implied by the hypothesized group-based double standards.
+That is, our findings showed that participants' judgments of protest actions depended on their group memberships---but not in the directions implied by the hypothesized group-based double standards.
 
@@
 
-In the remainder of this General Discussion, we discuss limiting conditions on our findings as well as theoretical and practical implications for understanding the often divided response to collective action.
-
-## Limitations
+In the remainder of this General Discussion, we discuss theoretical and practical implications for understanding the often divided response to collective action as well as limitations that constraint the generality of our findings.
+
+## Implications
+
+### Theoretical Implications
+
+Growing evidence shows that, in many situations, observers apply a more lenient standard when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
+And yet, our findings suggested that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on ascribed group memberships such as class, race, or gender.
+We did not find evidence for those double standards even though we examined social categories that are central to most people's identities, even when we considered only highly identified group members, and even though participants' reactions to the manipulation showed that, at least in Experiments 1--2, they understood the protesters' causes in ingroup--outgroup terms.
+
+Instead, our findings showed that progressives and, to a lesser extent, conservatives judge the same controversial protest actions to be more acceptable when the protesters' cause aligns with their own ideological position.
+We consider social, political, and moral psychological explanations for the observed ideology-based double standards.
+
+First, we might understand ideology-based double standards as an expression of ingroup bias based on partisan identities [@osborne_psychology_2022; @finkel_political_2020].
+This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions when taken by their most-liked, rather than their least-liked, political group.
+Indeed, researchers argued that opinion-based identities are more proximate and important predictors of collective action than social categories such as gender and race [@bliuc_opinion-based_2007].
+Supporting this argument, meta-analyses found that politicized identification---that is, identifying with a social movement or activist group---has a stronger association with collective action than identification with other social categories [@van_zomeren_toward_2008; @agostini_toward_2021].
+An identity-based account, however, implies that both conservatives and progressives would show ingroup bias and cannot, on its own, explain the observed asymmetry in the two groups' judgments.
+
+Second, we might explain ideology-based double standards, as we have done so far, as resulting from conservatives being motivated to defend the system, progressives being motivated to challenge the system, and both being motivated to support collective action to those effects [@jost_missing_2017; @osborne_protesting_2019].
+An ideology-based account could also  explain the asymmetry in how progressives and conservatives judged collective action for various causes.
+Political conservatism can be understood as uniting two motives, to resist social change and to maintain social inequality [@jost_political_2003; @jost_left_2021].
+If conservatives perceive disruptive protest actions as inherently threatening to the social order, they will be motivated to condemn such actions even when the protesters' cause supports the unequal status quo (e.g., protesting *against* defunding the police).
+
+Lastly, we might consider that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013]---but is instead understood as a means to an end and judged in relation to its end ("the end justifies the means").
+Their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] might lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
+This explanation aligns with Richardson and Conway's [-@richardson_standing_2022] finding that moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives (and vice versa).
+A values-based account could also explain why conservatives who, more so than liberals, endorse moral concerns related to loyalty and authority  tend to reject disruptive protest actions even when the protesters' cause otherwise aligns with their moral concerns.
+
+Future research should, across different causes, contexts, and cultures, seek to disentangle which of the proposed social, political, and moral psychological processes best explains partisan differences in judging collective action.
+
+### Practical Implications
+
+Our findings showed that, for many protest actions (e.g., blocking roads), there is no consensus on whether they are acceptable means to advance a cause.<!-- Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.--> This confirms our contention that the distinction between normative and non-normative collective action should be situated in the eye of the beholder.
+By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provide a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
+
+After all, this distinction is consequential.
+Research on how people respond to collective action showed that disruptive yet non-violent collective action might be most effective at gaining concessions from those resistant to social change [@shuman_disrupting_2021; @shuman_when_2023].
+If, as our research suggests, people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
+This could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice.
+In this way, our research complements, and adds to, the growing program of research on how people respond to different kinds of collective action.
+
+If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
+For example, 45\% of Republicans supported the storming of the U.S.
+Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
+Recent research, most of it by political scientists, investigated support for partisan violence [@kalmoe_radical_2022], albeit with mixed results [@westwood_current_2022].
+While our research focused on controversial but non-violent actions, we encourage researchers to adapt our paradigm to investigate when and why people apply different standards when judging violent and non-violent collective action for different causes.
+
+## Constraints on Generality
 
@@
 
-Despite its strengths, our research has several limitations that constrain the generalizability of our findings.
+Despite its strengths, our research has several limitations that constrain the generality of our findings.
 First, our research examined the hypothesized relationships for three political causes---protecting workers' rights, defunding the police, restricting abortion---and thus provided only limited evidence that our findings generalize beyond those causes.
@@
 Second, our research was based on samples from two Western, Educated, Industrialized, Rich, and Democratic [WEIRD, @henrich_weirdest_2010] contexts---the United Kingdom and the United States---which limits the cross-cultural generalizability of our findings.
-
-## Implications
-
-### Theoretical Implications
-
-Growing evidence shows that, in many situations, observers apply a more lenient standard when judging moral transgressions by ingroup members [@abrams_double_2013; @endevelt_everyone_2021].
-And yet, our findings suggested that, unlike judgments in other domains, judgments about collective action are not subject to double standards based on ascribed group memberships such as class, race, or gender.
-We did not find evidence for those double standards even though we examined social categories that are central to most people's identities, even when we considered only highly identified group members, and even though participants' reactions to the manipulation showed that, at least in Experiments 1--2, they understood the protesters' causes in ingroup--outgroup terms.
-
-Instead, our findings showed that progressives and, to a lesser extent, conservatives judge the same controversial protest actions to be more acceptable when the protesters' cause aligns with their own ideological position.
-We consider social, political, and moral psychological explanations for the observed ideology-based double standards.
-
-First, we might understand ideology-based double standards as an expression of ingroup bias based on partisan identities [@osborne_psychology_2022; @finkel_political_2020].
-This explanation aligns with Verkuyten et al.'s [-@verkuyten_intolerance_2022] finding that participants were more tolerant of transgressive protest actions when taken by their most-liked, rather than their least-liked, political group.
-An identity-based account, however, implies that both conservatives and progressives would show ingroup bias and, therefore, cannot explain the observed asymmetry in the two groups' judgements.
-
-Second, we might explain ideology-based double standards, as we have done so far, as resulting from conservatives being motivated to defend the system, progressives being motivated to challenge the system, and both being motivated to support collective action to those effects [@jost_missing_2017; @osborne_protesting_2019].
-An ideology-based account could also  explain the asymmetry in how progressives and conservatives judged collective action for various causes.
-Political conservatism can be understood as uniting two motives, to resist social change and to maintain social inequality [@jost_political_2003; @jost_left_2021].
-If conservatives perceive disruptive protest actions as inherently threatening to the social order, they will be motivated to condemn such actions even when the protesters' cause supports the unequal status quo (e.g., protesting *against* defunding the police).
-
-Lastly, we might consider that collective action, even in its controversial forms, is not judged as a moral transgression [as in previous research on identity-based double standards, e.g., @abrams_double_2013]---but is instead understood as a means to an end and judged in relation to its end ("the end justifies the means").
-Their different moral concerns [@graham_liberals_2009; @kivikangas_moral_2021] might lead progressives and conservatives to see their support of certain causes as a fundamental matter of right or wrong [@skitka_psychology_2021] and, by thus moralizing those causes, to accept more extreme means to achieving them.
-This explanation aligns with Richardson and Conway's [-@richardson_standing_2022] finding that moral concerns explained why liberals rated protesting for liberal causes as more moral than conservatives (and vice versa).
-A values-based account could also explain why conservatives who, more so than liberals, endorse moral concerns related to loyalty and authority  tend to reject disruptive protest actions even when the protesters' cause otherwise aligns with their moral concerns.
-
-Future research should, across different causes, contexts, and cultures, seek to disentangle which of the proposed social, political, and moral psychological processes best explains partisan differences in judging collective action.
-
-### Practical Implications
-
-Most broadly, our findings showed that, for many protest actions (e.g., blocking roads), there is no consensus on whether they are acceptable means to advance a cause.<!-- Instead, this judgment seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values.--> This confirms our contention that the distinction between normative and non-normative collective action should be situated in the eye of the beholder.
-By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provide a paradigm that we hope will stimulate research on when and why people differ in making this distinction.
-
-After all, this distinction is consequential.
-Research on how people respond to collective action showed that disruptive yet non-violent collective action might be most effective at gaining concessions from those resistant to social change [@shuman_disrupting_2021; @shuman_when_2023].
-If, as our research suggests, people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent.
-This could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice.
-In this way, our research complements, and adds to, the growing program of research on how people respond to different kinds of collective action.
-
-If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism.
-For example, 45% of Republicans supported the storming of the U.S.
-Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent.
-Recent research, most of it by political scientists, investigated support for partisan violence [@kalmoe_radical_2022], albeit with mixed results [@westwood_current_2022].
-While our research focused on controversial but non-violent actions, we encourage researchers to adapt our paradigm to investigate when and why people apply different standards when judging violent and non-violent collective action for different causes.
+Both contexts are marked by severe political polarization which limits how well our findings generalize to less polarized contexts.
+Likewise, both contexts are liberal democracies which limits how well our findings generalize to more repressive contexts in which even peaceful protest might violate laws and social norms.
 
@@
 Our research showed that, in contrast to well-documented double standards in other domains, people do not apply a different standard when judging collective action by members of the same class, race or gender or for a cause aligned with their own group's interests.
-Instead, our research demonstrated that progressives and, less so, conservatives consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological position<!-- ---in other words, that there are partisan double standards in judging collective action -->.
+Instead, our research demonstrated that progressives and, less so, conservatives consider the same protest actions to be more acceptable when the cause of the protest aligns with their own ideological position.
+In other words, we found partisan double standards in judging collective action.
 

### 2024-11-05 Documents/manuscript.tex
2nd revision

@@
 but often sparks contention about what actions are acceptable means to
-effect social change. Five studies (total \(N = 2,979\)) investigated
-double standards in judging collective action---that is, whether
-observers judge the same protest actions to be more acceptable depending
-on who the protesters are and what they are protesting. Two studies used
-item response theory to develop an instrument of 25 controversial
-protest actions to measure where people draw the line between acceptable
-and unacceptable forms of collective action. Three preregistered
-experiments found no consistent evidence for ingroup bias in terms of
-social class when judging protests for workers' rights (Experiment 1),
-in terms of race when judging protests for and against defunding the
-police (Experiment 2), and in terms of gender when judging protests for
-and against restricting abortion (Experiment 3). Instead, we found that
-participants who identified as left-wing/liberal (Experiments 1--3) and
-rejected system-justifying beliefs (Experiments 1--2) considered the
-same protest actions more acceptable when a cause aligned with their
-ideological orientation (for workers' rights, for defunding the police,
-against restricting abortion) than when it did not (against defunding
-the police, for restricting abortion). Conservative participants
-considered the same actions somewhat more acceptable when protesters
-supported, rather than opposed, restricting abortion (Experiment 3) but
-considered all protest actions, for and against defunding the police,
-equally unacceptable (Experiment 2). Our findings have theoretical and
-practical implications for understanding the often-divided response to
-social movements.
+effect social change. We investigated double standards in judging
+collective action---that is, whether observers judge the same protest
+actions to be more acceptable depending on who the protesters are and
+what they are protesting. In two studies, we used item response theory
+to develop an instrument of 25 controversial protest actions to measure
+where people draw the line between acceptable and unacceptable forms of
+collective action. In three preregistered experiments (\(N = 2776\)), we
+found no consistent evidence for ingroup bias in terms of social class
+when judging protests for workers' rights (Experiment 1), in terms of
+race when judging protests for and against defunding the police
+(Experiment 2), and in terms of gender when judging protests for and
+against restricting abortion (Experiment 3). Instead, we found that
+progressive participants (Experiments 1--3) who rejected
+system-justifying beliefs (Experiments 1--2) considered the same protest
+actions more acceptable when a cause aligned with their ideological
+orientation (for workers' rights, for defunding the police, against
+restricting abortion) than when it did not (against defunding the
+police, for restricting abortion). Conservative participants considered
+the same actions somewhat more acceptable when protesters supported,
+rather than opposed, restricting abortion (Experiment 3) but considered
+all protest actions, for and against defunding the police, equally
+unacceptable (Experiment 2). Our findings have theoretical and practical
+implications for understanding the often-divided response to social
+movements.
 
@@
 \begin{center}
-\noindent 11,654 words
+\noindent 12,257 words
 \end{center}
@@
 non-violent actions acceptable means to achieve social change and
-challenge social injustice. This finding helps explain the often-divided
-response to recent protest movements.
+challenge social injustice. These findings help explain the
+often-divided response to recent protest movements.
 
@@
 relate to social change and social inequality which, in turn, are
-motivated by epistemic, existential, and relational needs (Jost et al.,
-2003; Jost, 2021). As motivated social cognition, ideological
-differences shape how we think and feel about political issues, actions,
-and events. In this view, ideology-based double standards depend on
-whether the cause of a protest aligns with someone's ideological
-positions.
+motivated by basic psychological needs (Jost et al., 2003; Jost, 2021).
+As motivated social cognition, ideological differences shape how we
+think and feel about political issues, actions, and events. In this
+view, ideology-based double standards depend on whether the cause of a
+protest aligns with someone's ideological positions.
 
@@
 understanding how the characteristics of both items and respondents
-shape responses to a set of items (DeMars, 2010). When respondents judge
-whether each of a set of actions is an acceptable means of protest (1 =
-\emph{yes}, 0 = \emph{no}), a two-parameter logistic item response
-theory model estimates responses as a function of several latent
-(unobserved) parameters:
+shape responses to a set of items (DeMars, 2010). Applied to judging
+protest actions, this framework allows us to disentangle how
+\emph{acceptable} different protest actions are from how
+\emph{accepting} different respondents are of them. In line with the
+long-standing distinction between normative and non-normative collective
+action (Wright et al., 1990), it recognizes that some protest actions
+(e.g., signing petitions) are more acceptable than others (e.g.,
+blocking roads). In line with the hypothesized double standards, it
+recognizes that some people consider more protest actions acceptable
+than others. By comparing how accepting respondents are of the
+\emph{same} protest actions in \emph{different} situations, we can use
+item response theory to detect double standards in judging collective
+action.
+
+In this research, respondents judge whether each of a set of more and
+less controversial protest actions is an acceptable means of protest (1
+= \emph{yes}, 0 = \emph{no}). We use two-parameter item response theory
+models to estimate the probability \(\Pr ( y_\textit{ij} = 1 )\) that
+respondent \(j\) considers action \(i\) an acceptable means of protest
+as a function of three latent model parameters:
 \[ \Pr ( y_\textit{ij} = 1 ) = \text{logit}^{-1} ( \alpha_i ( \theta_j + \beta_i ) ) \]
-where \(\Pr ( y_\textit{ij} = 1 )\) is the probability that respondent
-\(j\) considers action \(i\) an acceptable means of protest,
-\(\theta_j\) estimates how \emph{accepting} respondent \(j\) is of
-various protest actions, \(\beta_i\) estimates how \emph{acceptable}
-action \(i\) is, and \(\alpha_i\) estimates how \emph{discriminating}
-action \(i\) is.
-
-Figure \ref{fig:f1} illustrates those relationships for three
-hypothetical protest actions. \(\theta_{(z)}\) is the
-\emph{z}-standardized tendency for a participant to be more accepting of
-various actions so that \(\beta_i\) determines the probability that the
-average respondent (\(\theta_{(z)} = 0\)) considers an action an
-acceptable means of protest. Action 2 (\(\beta_2 = 1.28\)) is more
-acceptable than Actions 1 and 2 (\(\{ \beta_2, \beta_3 \} = 0\)) as the
-average respondent is more likely to consider the former (\(\Pr = .78\))
-than the latter (\(\Pr = .50\)) acceptable means of protest. Action 3
-(\(\alpha_3 = 3.60\)) is more discriminating than Actions 1 and 2
-(\(\{ \alpha_1, \alpha_2 \} = 1\)) as the slope for the relationship
-between how accepting a respondent is and how likely they are to
-consider an action acceptable is steeper for the former action. This
-means that knowing whether someone considers Action 3 acceptable
-provides more information about them than knowing whether they consider
-Actions 1 or 2 acceptable.
+where \(\theta_j\) estimates how \emph{accepting} respondent \(j\) is of
+various actions, \(\beta_i\) estimates how \emph{acceptable} action
+\(i\) is, and \(\alpha_i\) estimates how \emph{discriminating} action
+\(i\) is.
 
@@
 \includegraphics[scale=1]{../Scale Development/figures/figure-1.png}
-\caption*{\textit{Note.} $\theta_{(z)}$ follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).}
+\caption*{\textit{Note.} $\Pr ( y_\textit{ij} = 1 )$ is the probability that that respondent $j$ considers action $i$ an acceptable means of protest.
+$\theta_{(z)}$ is the $z$-standardized tendency for a participant to be more accepting of various actions.
+It follows a standard normal distribution and is shown as both $z$-scores (bottom) and percentiles (top).
+$\beta_i$ and $\alpha_i$ capture, respectively, how acceptable and discriminating an action is.}
 \label{fig:f1}
 \end{figure}
+
+Figure \ref{fig:f1} illustrates those relationships for three
+hypothetical protest actions. \(\theta_{(z)}\) is the
+\emph{z}-standardized tendency for a participant to be more accepting of
+various actions. It follows a normal distribution so that
+\(\theta_{(z)} = 1.28\) means that a respondent is more accepting than
+90\% of respondents. \(\beta_i\) determines the probability that the
+average respondent (\(\theta_{(z)} = 0\)) considers an action an
+acceptable means of protest. Action 2 (\(\beta_2 = 1.28\)) is more
+acceptable than Actions 1 and 3 (\(\{ \beta_1, \beta_3 \} = 0\)) as the
+average respondent is more likely to consider the former (\(\Pr = .78\))
+than the latter (\(\Pr = .50\)) acceptable means of protest.
+\(\alpha_i\) determines how much the probability that a respondent
+considers an action acceptable depends on how accepting that respondent
+is. Action 3 (\(\alpha_3 = 3.60\)) is more discriminating than Actions 1
+and 2 (\(\{ \alpha_1, \alpha_2 \} = 1\)) as the slope for the
+relationship between how accepting a respondent is and how likely they
+are to consider an action acceptable is steeper for the former action.
+This means that knowing whether someone considers Action 3 acceptable
+provides more information about them than knowing whether they consider
+Actions 1 or 2 acceptable.
 
@@
 \(\theta_{(z)}\) is \(z\)-standardized, correspond to Cohen's \(d\)
-effect sizes.
+effect sizes. As \(\theta_{(z)}\) determines the probability of
+respondents judging different protest actions acceptable, it can be
+converted to the predicted proportion \(p\) and number \(n\) of actions
+a respondent considers acceptable. Figure \ref{fig:f3} (A) illustrates
+these relationships. Figures \ref{fig:f3} (B), \ref{fig:f4},
+\ref{fig:f5}, and \ref{fig:f6} show the predicted proportion of actions
+a participant would consider acceptable as an intuitive and easily
+interpretable effect size.
 
@@
 
+\begin{figure*}[!t]
+\centering
+\caption{Examples of protest actions rated in Study 2}
+\includegraphics[scale=1]{../Scale Development/figures/figure-2}
+\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause.
+Darker bars mark actions that were included in the final scale.}
+\label{fig:f2}
+\end{figure*}
+
 In Study 1, we compiled protest actions from participants' responses and
@@
 actions that varied in how acceptable we would expect them to be. For
-details, see Supplemental Online Material (SOM).
+details, see Supplemental Online Material A (SOM-A).
 
@@
 143 participants for our analyses.
-
-\begin{figure*}[!t]
-\centering
-\caption{Examples of protest actions rated in Study 2}
-\includegraphics[scale=1]{../Scale Development/figures/figure-2}
-\caption*{\textit{Note.} Percentages reflect the proportion of participants who thought that an action was \textit{sometimes}, \textit{often}, or \textit{always} an acceptable means to a advance a cause.
-Darker bars mark actions that were included in the final scale.}
-\label{fig:f2}
-\end{figure*}
 
@@
 unacceptable (e.g., political violence; Figure \ref{fig:f2}). For
-detailed ratings and results, see SOM.
+detailed ratings and results, see SOM-A and Tables S1--S3.
 
@@
 observers judge collective action by lower-status (i.e.~working-class)
-and higher-status (middle-class) group members.\footnote{For Experiments
-  1 and 2, we had preregistered the additional hypotheses that all
-  participants would judge the same actions as more acceptable when the
-  protesters were from the higher-status group rather than the
-  lower-status group (Experiment 1) and when the goal of the protest was
-  to support rather than challenge the system (Experiment 2). These
-  hypotheses reflected the assumption that the motivation to justify the
-  system was, to some extent, universal. With hindsight, we no longer
-  think that this assumption is tenable and, therefore, omit the
-  hypotheses based on it from the revised manuscript. We still report
-  the same preregistered analyses and their results.} In
-non-preregistered analyses, we tested whether participants would judge
-the same protest actions as more or less acceptable depending on their
-own ideological positions in terms of political orientation,
-system-justifying beliefs, and social dominance orientation.\footnote{In
-  Experiment 1, we explored a broader range of operationalizations of
-  political ideology, including social dominance orientation, than in
-  the other experiments.} In this way, Experiment 1 provided a first
-test of the hypothesized class- and ideology-based double standards in
-judging collective action.
+and higher-status (middle-class) group members. In non-preregistered
+analyses, we tested whether participants would judge the same protest
+actions as more or less acceptable depending on their own ideological
+positions in terms of political orientation, system-justifying beliefs,
+and social dominance orientation.\footnote{In Experiment 1, we explored
+  a broader range of operationalizations of political ideology,
+  including social dominance orientation, than in the other experiments.}
+In this way, Experiment 1 provided a first test of the hypothesized
+class- and ideology-based double standards in judging collective action.
 
@@
 (\url{https://osf.io/24wrx/?view_only=8560eb60286149498f8da48bc9f9d99b}).
+For Experiments 1 and 2, we had preregistered the additional hypothesis
+that all participants would judge the same actions to be more acceptable
+when the protesters were from the higher-status group rather than the
+lower-status group (Experiment 1) and when the goal of the protest was
+to support rather than challenge the system (Experiment 2). This
+hypothesis reflected the assumption that the motivation to justify the
+system was, to some extent, universal. With hindsight, we no longer
+think that this assumption is tenable and, therefore, do not discuss the
+hypothesis based on it. We still follow the same preregistered analysis
+plan and report all results related to our original hypotheses.
+
 We made all materials, data, and analysis scripts available online
@@
 what working-class and middle-class/professional jobs are (for details,
-see SOM). Participants then answered, among other questions, whether
+see SOM-B). Participants then answered, among other questions, whether
 they considered their current job---or the jobs they had had in the past
@@
 to update prior distributions in light of the observed data to produce
-posterior distributions. In contrast to \(p\)-values and confidence
-intervals, posterior distributions have a straight-forward
-interpretation as stating how plausible different values of the model
-parameters are given the observed data.
+posterior distributions. Unlike \(p\)-values and confidence intervals,
+posterior distributions have a straight-forward interpretation as
+stating how plausible different values of the model parameters are given
+the observed data. We report point estimates, based on the median of
+posterior samples, and uncertainty intervals, based on the quantiles of
+posterior samples, that enclose the 95\% most plausible estimates.
 
@@
 informative prior distributions (Gelman et al., 2017) to all model
-parameters.\footnote{Our model assigned \(\text{Half-Cauchy} (0, 3)\)
-  prior distributions to the standard deviations of \(\alpha_i\),
-  \(\beta_i\), and \(\theta_j\) and \(\text{Normal} (0, 3)\) prior
-  distributions to all other model parameters.} We report point
-estimates, based on the median of posterior samples, and uncertainty
-intervals, based on the quantiles of posterior samples, that enclose the
-95\% most plausible estimates.
+parameters: \(\text{Half-Cauchy} (0, 3)\) prior distributions to the
+standard deviations of \(\alpha_i\), \(\beta_i\), and \(\theta_j\) and
+\(\text{Normal} (0, 3)\) prior distributions to all other model
+parameters.
+
+We report differences between conditions in terms of differences in the
+estimated tendency for a participant to consider more protest actions
+acceptable (\(\theta_{(z)}\)) which, as \(\theta_{(z)}\) is
+\(z\)-standardized, correspond to Cohen's \(d\) effect sizes. Likewise,
+we report standardized regression coefficients (\(\beta_{xy}\)) that
+correspond to the estimated change in the \(z\)-standardized tendency to
+consider more protest actions acceptable for each \(+1\textit{SD}\)
+change in the predictor variable. All figures show the predicted
+proportion of protest actions participants would consider acceptable in
+each condition based on the estimated models.
 
@@
 
-Figure \ref{fig:f3} shows (B) the \(z\)-standardized propensity to
-consider more controversial protest actions acceptable as a function of
+Figure \ref{fig:f3} shows (B) how participants' tendency to consider
+more controversial protest actions acceptable varied as a function of
 the three ideological orientation variables. We found that participants
@@
 (Gelman et al., 2012). Our models assigned weakly informative prior
-distributions to all model parameters.\footnote{Our model assigned
-  \(\text{LKJ} (2)\) prior distributions to the Cholesky-transformed
-  correlation matrices for varying effects and
-  \((\text{Half-})\text{Cauchy} (0, 5)\) prior distributions to all
-  other model parameters.} Figure \ref{fig:f4} shows results from three
+distributions: \(\text{LKJ} (2)\) prior distributions to the
+Cholesky-transformed correlation matrices for varying effects and
+\((\text{Half-})\text{Cauchy} (0, 5)\) prior distributions to all other
+model parameters. Figure \ref{fig:f4} shows results from three
 preregistered models we estimated to test our hypotheses.
@@
 predictor categories as well as how much of this change occurred between
-each of the four pairs of adjacent predictor categories.\footnote{Our
-  model assigned a Dirichlet prior, \(\alpha = {1, 1, 1, 1}\), to the
+each of the four pairs of adjacent predictor categories. This model
+assigned a Dirichlet prior, \(\alpha = {1, 1, 1, 1}\), to the
 proportions of the overall change that was expected to occur between
-  each of the four pairs of predictor categories.} Participants who
-supported the protesters' cause did not, on average, consider the same
-protest actions to be more acceptable than participants who opposed the
-protesters' cause (\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our
-results did not support the alternative hypothesis that ideology-based
-double standards can be reduced to support for, or opposition to, the
-cause of a protest.
+each of the four pairs of predictor categories. We found that
+participants who supported the protesters' cause did not, on average,
+consider the same protest actions to be more acceptable than
+participants who opposed the protesters' cause
+(\(\beta_{y} = 0.14, [-0.12, 0.40]\)). Ergo, our results did not support
+the alternative hypothesis that ideology-based double standards can be
+reduced to support for, or opposition to, the cause of a protest.
 
@@
 gender-related system-justifying beliefs (Jost \& Kay, 2005), and the
-updated moral foundations questionnaire (Atari et al., 2022, for
-details, see SOM). Within the latter questionnaire, we embedded three
+updated moral foundations questionnaire (Atari et al., 2023, for
+details, see SOM-C). Within the latter questionnaire, we embedded three
 further attention checks (e.g., ``To show that you are paying attention
@@
 than members of relatively disadvantaged groups. That is, our findings
-showed that participants' judgements of protest actions depended on
-their group memberships---but not in the directions implied by the
+showed that participants' judgments of protest actions depended on their
+group memberships---but not in the directions implied by the
 hypothesized group-based double standards.
@@
 
-In the remainder of this General Discussion, we discuss limiting
-conditions on our findings as well as theoretical and practical
-implications for understanding the often divided response to collective
-action.
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-Our research applied item response theory to judgments about collective
-action to investigate whether where people draw the line between
-acceptable and unacceptable means of protest depends on who the
-protesters are and what they are protesting. In so doing, we moved the
-distinction between normative and non-normative collective action from
-the realm of the researcher's intuition into the realm of scientific
-investigation, allowing us to test for double standards in judging
-collective action.
-
-Despite its strengths, our research has several limitations that
-constrain the generalizability of our findings. First, our research
-examined the hypothesized relationships for three political
-causes---protecting workers' rights, defunding the police, restricting
-abortion---and thus provided only limited evidence that our findings
-generalize beyond those causes. Relying on few stimuli to establish an
-effect threatens the validity and replicability of research findings and
-constrains the generalizability of psychological research (Yarkoni,
-2022). Future research should sample a wider range of causes to address
-this pervasive but often ignored problem (Judd et al., 2012). This is
-particularly important when studying collective action as recent
-theorizing (Jost et al., 2017) and research (Osborne et al., 2019)
-highlighted the importance of differentiating between progressive and
-conservative causes. Second, our research was based on samples from two
-Western, Educated, Industrialized, Rich, and Democratic (WEIRD, Henrich
-et al., 2010) contexts---the United Kingdom and the United
-States---which limits the cross-cultural generalizability of our
-findings.
+In the remainder of this General Discussion, we discuss theoretical and
+practical implications for understanding the often divided response to
+collective action as well as limitations that constraint the generality
+of our findings.
 
@@
 transgressive protest actions when taken by their most-liked, rather
-than their least-liked, political group. An identity-based account,
-however, implies that both conservatives and progressives would show
-ingroup bias and, therefore, cannot explain the observed asymmetry in
-the two groups' judgements.
+than their least-liked, political group. Indeed, researchers argued that
+opinion-based identities are more proximate and important predictors of
+collective action than social categories such as gender and race (Bliuc
+et al., 2007). Supporting this argument, meta-analyses found that
+politicized identification---that is, identifying with a social movement
+or activist group---has a stronger association with collective action
+than identification with other social categories (Agostini \& Zomeren,
+2021; Zomeren et al., 2008). An identity-based account, however, implies
+that both conservatives and progressives would show ingroup bias and
+cannot, on its own, explain the observed asymmetry in the two groups'
+judgments.
 
@@
 
-Most broadly, our findings showed that, for many protest actions (e.g.,
-blocking roads), there is no consensus on whether they are acceptable
-means to advance a cause. This confirms our contention that the
-distinction between normative and non-normative collective action should
-be situated in the eye of the beholder. By developing an instrument for
-measuring where people draw the line between acceptable and unacceptable
-forms of collective action, we provide a paradigm that we hope will
-stimulate research on when and why people differ in making this
-distinction.
+Our findings showed that, for many protest actions (e.g., blocking
+roads), there is no consensus on whether they are acceptable means to
+advance a cause. This confirms our contention that the distinction
+between normative and non-normative collective action should be situated
+in the eye of the beholder. By developing an instrument for measuring
+where people draw the line between acceptable and unacceptable forms of
+collective action, we provide a paradigm that we hope will stimulate
+research on when and why people differ in making this distinction.
 
@@
 
+\hypertarget{constraints-on-generality}{%
+\subsection{Constraints on Generality}\label{constraints-on-generality}}
+
+Our research applied item response theory to judgments about collective
+action to investigate whether where people draw the line between
+acceptable and unacceptable means of protest depends on who the
+protesters are and what they are protesting. In so doing, we moved the
+distinction between normative and non-normative collective action from
+the realm of the researcher's intuition into the realm of scientific
+investigation, allowing us to test for double standards in judging
+collective action.
+
+Despite its strengths, our research has several limitations that
+constrain the generality of our findings. First, our research examined
+the hypothesized relationships for three political causes---protecting
+workers' rights, defunding the police, restricting abortion---and thus
+provided only limited evidence that our findings generalize beyond those
+causes. Relying on few stimuli to establish an effect threatens the
+validity and replicability of research findings and constrains the
+generalizability of psychological research (Yarkoni, 2022). Future
+research should sample a wider range of causes to address this pervasive
+but often ignored problem (Judd et al., 2012). This is particularly
+important when studying collective action as recent theorizing (Jost et
+al., 2017) and research (Osborne et al., 2019) highlighted the
+importance of differentiating between progressive and conservative
+causes. Second, our research was based on samples from two Western,
+Educated, Industrialized, Rich, and Democratic (WEIRD, Henrich et al.,
+2010) contexts---the United Kingdom and the United States---which limits
+the cross-cultural generalizability of our findings. Both contexts are
+marked by severe political polarization which limits how well our
+findings generalize to less polarized contexts. Likewise, both contexts
+are liberal democracies which limits how well our findings generalize to
+more repressive contexts in which even peaceful protest might violate
+laws and social norms.
+
 \hypertarget{conclusion}{%
@@
 acceptable when the cause of the protest aligns with their own
-ideological position.
+ideological position. In other words, we found partisan double standards
+in judging collective action.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2022}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-agostini_toward_2021}{}}%
+Agostini, M., \& Zomeren, M. van. (2021). Toward a comprehensive and
+potentially cross-cultural model of why people engage in collective
+action: {A} quantitative research synthesis of four motivations and
+structural constraints. \emph{Psychological Bulletin}, \emph{147}(7),
+667--700. \url{https://doi.org/10.1037/bul0000256}
+
+\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2023}{}}%
 Atari, M., Haidt, J., Graham, J., Koleva, S., Stevens, S. T., \&
-Dehghani, M. (2022). Morality beyond the {WEIRD}: How the nomological
-network of morality varies across cultures. \emph{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/q6c9r}
+Dehghani, M. (2023). Morality beyond the {WEIRD}: {How} the nomological
+network of morality varies across cultures. \emph{Journal of Personality
+and Social Psychology}, \emph{125}(5), 1157--1188.
+\url{https://doi.org/10.1037/pspp0000470}
 
@@
 \url{https://perma.cc/YG99-EUYX}
+
+\leavevmode\vadjust pre{\hypertarget{ref-van_zomeren_toward_2008}{}}%
+Zomeren, M. van, Postmes, T., \& Spears, R. (2008). Toward an
+integrative social identity model of collective action: {A} quantitative
+research synthesis of three socio-psychological perspectives.
+\emph{Psychological Bulletin}, \emph{134}(4), 504--535.
+\url{https://doi.org/10.1037/0033-2909.134.4.504}
 

### 2024-11-05 Documents/supplemental-materials.qmd
2nd revision

@@
 
-We measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
+We measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2023] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*).
 We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").
@@
 
-# References
+\section*{References}
 

### 2024-11-05 Documents/supplemental-materials.tex
2nd revision

@@
 \usepackage[letterpaper, margin = 1in]{geometry}
-\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{2}
+\renewcommand{\thesection}{\Alph{section}} 
 \usepackage{titlesec}  
@@
 We measured moral concerns with the 36-item moral foundations
-questionnaire (MFQ-2, Atari et al., 2022) which assesses to what extent
+questionnaire (MFQ-2, Atari et al., 2023) which assesses to what extent
 participants endorse concerns about care, equality, proportionality,
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2022}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2023}{}}%
 Atari, M., Haidt, J., Graham, J., Koleva, S., Stevens, S. T., \&
-Dehghani, M. (2022). Morality beyond the {WEIRD}: How the nomological
-network of morality varies across cultures. \emph{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/q6c9r}
+Dehghani, M. (2023). Morality beyond the {WEIRD}: {How} the nomological
+network of morality varies across cultures. \emph{Journal of Personality
+and Social Psychology}, \emph{125}(5), 1157--1188.
+\url{https://doi.org/10.1037/pspp0000470}
 

### 2024-11-05 Experiment 1/figures/figure-3.R

@@ -86,7 +86,7 @@ rm(list = ls())
     facet_grid(observer ~ .) +
     theme_grey(base_size = 10) +
     labs(
-      x = expression(italic(p)),
+      x = expression(Pr),
       y = "Protesters"
     ) +
     theme(
@@ -161,45 +161,30 @@ rm(list = ls())
     )
   
   # Make figure
-  f_3b <- ggplot(d_explor, aes(x = x, y = z)) +
+  f_3b <- ggplot(d_explor, aes(x = x, y = p)) +
     stat_lineribbon(
       .width = c(.99),
       size = 0.455,
       alpha = 1.0,
       fill = "white"
     ) +
-    # geom_hline(
-    #   data = d_explor %>% 
-    #     filter(x == 0) %>% 
-    #     group_by(outcome) %>% 
-    #     summarise(z = median(z)),
-    #   aes(yintercept = z),
-    # size = 0.455,
-    #   colour = "grey20",
-    #   linetype = "dashed"
-    # ) +
-    geom_hline(
-      yintercept = 0,
-      size = 0.455,
-      colour = "grey20",
-      linetype = "dashed"
-    ) +
     stat_lineribbon(
       .width = c(.99, .95, .8, .5),
       size = 0.455,
       alpha = 0.6
     ) +
-    geom_vline(xintercept = c(-2.5, 2.5), size = 0.455) +
-    geom_hline(yintercept = c(-1 - 2 * 0.05, 1 + 2 * 0.05), size = 0.455) +
     scale_x_continuous(minor_breaks = NULL, expand = c(0, 0)) +
-    scale_y_continuous(minor_breaks = NULL) + 
-    scale_fill_brewer(type = "seq", palette = "Greys", direction = -1) +
-    coord_cartesian(ylim = c(-1, 1)) +
+    scale_y_continuous(
+      minor_breaks = NULL,
+      labels = ~numform::f_num(., digits = 2)
+    ) + 
+    scale_fill_brewer(type = "seq", palette = "Greys", direction = 1) +
+    # coord_cartesian(ylim = c(-1, 1)) +
     facet_grid(. ~ outcome, switch = "x") +
     labs(
       tag = "B",
       x = expression(italic(z)),
-      y = expression(theta[(italic(z))])
+      y = expression(Pr)
     ) +
     theme_grey(base_size = 10) +
     theme(
@@ -210,6 +195,7 @@ rm(list = ls())
       axis.title.x = element_blank(),
[…]

### 2024-11-05 Experiment 2/figures/figure-4.R

@@ -170,7 +170,7 @@ rm(list = ls())
   # Make plot
   f_4a_p <- m1_post %>% 
     mutate(ingroup = if_else(Participant == Protesters, "Ingroup", "Outgroup")) %>% 
-    ggplot(., aes(x = z, y = fct_rev(factor(order)))) + 
+    ggplot(., aes(x = p, y = fct_rev(factor(order)))) + 
     stat_pointinterval(
       aes(shape = ingroup),
       fill = "white",
@@ -179,10 +179,13 @@ rm(list = ls())
     ) +
     geom_hline(yintercept = c(0.5, 8.5), size = 0.455) +
     geom_hline(yintercept = c(4.5), size = 3*0.455, colour = "white") +
-    scale_x_continuous(minor_breaks = NULL) +
+    scale_x_continuous(
+      minor_breaks = NULL,
+      labels = ~numform::f_num(., digits = 2)
+    ) +
     scale_y_discrete(expand = c(0, 0)) +
     scale_shape_manual(values = c("Ingroup" = 21, "Outgroup" = 19)) +
-    coord_cartesian(xlim = c(-0.5, 0.5), ylim = c(0.5, 8.5)) +
+    # coord_cartesian(xlim = c(-0.5, 0.5), ylim = c(0.5, 8.5)) +
     theme_grey(base_size = 10) +
     theme(
       axis.text.x = element_text(colour = "black"),
@@ -207,7 +210,7 @@ rm(list = ls())
       plot.margin = margin(2, 0, 2, 2)
     ) +
     labs(
-      x = expression(theta[(italic(z))])
+      x = expression(Pr)
     )
     
 
@@ -223,50 +226,57 @@ rm(list = ls())
   m2_post_ii <- m2_ps %>% spread_draws(alpha[ii], beta[ii]) %>% ungroup()
   m2_post_jj <- m2_ps %>% spread_draws(theta[jj]) %>% ungroup()
   m2_post_kk <- m2_ps %>% spread_draws(delta[kk], b_sjb_kk[kk]) %>% ungroup()
-
-  # Make posterior predictions
-  m2_post <- dl %>% 
-    left_join(
-      tibble(
+  
+  # Calculate effect sizes (z)
+  m2_post <- dl |> 
+    distinct(jj, kk, x_sjb) |> 
+    left_join(m2_post_jj, by = join_by(jj)) |> 
[…]

### 2024-11-05 Experiment 2/figures/figure-5.R

@@ -79,7 +79,8 @@ rm(list = ls())
     rename(Participant = observers, Protesters = protesters, Cause = protest)
   
   
-# Predict -----------------------------------------------------------------
+
+# Predict (M2) ------------------------------------------------------------
 
   # Import results
   m2_fit <- read_rds("Experiment 2/results/m2_fit.rds")
@@ -92,26 +93,53 @@ rm(list = ls())
   m2_post_jj <- m2_ps %>% spread_draws(theta[jj]) %>% ungroup()
   m2_post_kk <- m2_ps %>% spread_draws(delta[kk], b_sjb_kk[kk]) %>% ungroup()
   
-  # Calculate effect sizes
-  m2_post <- dl %>% 
-    distinct(jj, kk, x_sjb) %>% 
-    left_join(m2_post_kk, by = c("kk")) %>% 
-    left_join(m2_post_jj, by = c("jj", ".draw", ".chain", ".iteration")) %>% 
-    group_by(.draw) %>% 
+  # Calculate effect sizes (z)
+  m2_post <- dl |> 
+    distinct(jj, kk, x_sjb) |> 
+    left_join(m2_post_jj, by = join_by(jj)) |> 
+    left_join(m2_post_kk, by = join_by(kk, .chain, .iteration, .draw)) |>
+    group_by(.draw) |> 
     summarize(
       mean = mean(theta + delta + b_sjb_kk * x_sjb),
       sd = sd(theta + delta + b_sjb_kk * x_sjb)
-    ) %>% 
+    ) |> 
     crossing(
-      kk = 1:8, 
+      kk = 1:8,
       x_sjb = c(-1, 1)
-    ) %>% 
-    left_join(m2_post_kk, by = c(".draw", "kk")) %>% 
+    ) |> 
+    left_join(m2_post_kk, by = join_by(.draw, kk)) |> 
     mutate(
       z = ((delta + b_sjb_kk * x_sjb) - mean)/sd
-    ) %>% 
-    left_join(dl %>% distinct(kk, order, Participant, Protesters, Cause), by = "kk") %>% 
-    select(.draw, kk, x_sjb, z, order, Participant, Protesters, Cause)
+    ) |> 
+    select(.draw, kk, x_sjb, z)
+  
+  # Calculate effect sizes (p, n)
+  m2_post <- dl |> 
+    distinct(ii) |> 
+    left_join(m2_post_ii, by = join_by(ii)) |>
+    crossing(
[…]

### 2024-11-05 Experiment 3/figures/figure-6.R

@@ -67,9 +67,9 @@ rm(list = ls())
   
   # Make plot
   f_6_p <- d_fig %>% 
-    ggplot(., aes(x = z, y = order)) +
+    ggplot(., aes(x = p, y = order)) +
     geom_ribbon(
-      data = d_fig %>% group_by(order, participant_ideology, protesters_cause) %>% median_qi(z),
+      data = d_fig %>% group_by(order, participant_ideology, protesters_cause) %>% median_qi(p),
       aes(group = interaction(protesters_cause, participant_ideology), xmin = .lower, xmax = .upper, fill = participant_ideology),
       alpha = 0.4
     ) +
@@ -82,12 +82,15 @@ rm(list = ls())
     ) +
     geom_hline(yintercept = c(0.5, 8.5), size = 0.455) +
     geom_hline(yintercept = c(4.5), size = 3*0.455, colour = "white") +
-    scale_x_continuous(minor_breaks = NULL) +
+    scale_x_continuous(
+      minor_breaks = NULL,
+      labels = ~numform::f_num(., digits = 2)
+    ) +
     scale_y_reverse(expand = c(0, 0)) +
     scale_shape_manual(values = c("Conservative" = 21, "Liberal" = 19)) +
     scale_colour_manual(values = c("Conservative" = "#DC267F", "Liberal" = "#648FFF")) +
     scale_fill_manual(values = c("Conservative" = "#DC267F", "Liberal" = "#648FFF")) +
-    coord_cartesian(xlim = c(-0.65, 1.15)) +
+    # coord_cartesian(xlim = c(-0.65, 1.15)) +
     facet_grid(. ~ "") +
     theme_grey(base_size = 10) +
     theme(
@@ -103,14 +106,14 @@ rm(list = ls())
       plot.margin = margin(2, 0, 2, 2)
     ) +
     labs(
-      x = expression(theta[(italic(z))])
+      x = expression(Pr)
     )
   
 
 # Combine -----------------------------------------------------------------
   
   # Assemble subfigures
-  f_6_t + f_6_p + plot_layout(widths = c(2, 1))
+  f_6_t + f_6_p + plot_layout(widths = c(3, 2))
 
 
 # Export ------------------------------------------------------------------
